---
title: "Kim data analysis"
author: "Jan Paul van Meenen"
date: "`r Sys.Date()`"
output: html_document
---

#Packages
```{r}
data_management_packages_to_install <- c("data.table", "tidyverse", "ggplot2", "remotes", "Platypus", "msa","cowplot")

packages_to_install <- c(data_management_packages_to_install)

# Check and install packages using a for loop
for (package in packages_to_install) {
  if (!requireNamespace(package, quietly = TRUE)) {
    install.packages(package)
  }
  library(package, character.only = TRUE)
}
```

#Load in Kim data
```{r}
load("data/vdj_Kim_sc_reclonotyped.RData") #vdj
```

## quick check
```{r}
head(vdj)
colnames(vdj)
```
Each row corresponds to a cell, sample_id links to the person, timepoint referes to the timepoint that sample was taken, tissue (=blood/lymphnode)

# Recreate plots
## cellcount and frequency
### cellcount
```{r, fig.width=14}
VDJ_cellcount <- vdj %>% 
  group_by(sample_id) %>%
  summarise(number_of_cells = n()) %>% 
  arrange(desc(number_of_cells))

cellcount_order <- VDJ_cellcount$sample_id

VDJ_cellcount_plot <- VDJ_cellcount %>%
  mutate(sample_id = factor(sample_id, levels = cellcount_order)) %>%
  ggplot(aes(x=sample_id, y = number_of_cells, fill=sample_id)) +
  geom_bar(position = "stack", stat="identity") +
  theme_minimal() + 
  ylab("Number of cells") +
  xlab("") +
  scale_y_continuous(limits = c(0,30000)) +
  scale_fill_manual(values = rep(c("darkorchid4", "darkorchid1", "mediumorchid1", "mediumpurple3"), length.out = length(unique(vdj$sample_id)))) +
  theme(legend.position = "none",
        panel.grid = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.text.y = element_text(size=20), 
        axis.text.x = element_text(size=20),
        plot.margin = margin(0,0,0,0) 
        ) 
  
VDJ_cellcount_plot
```

### frequency
Similar to S1
```{r}
VDJ_clonal_barplot
```
```{r, fig.width=14}
output_VDJ_barplots <- VDJ_clonal_barplot(VDJ=vdj, group.by="sample_id", counts.to.use = "clonotype_id")

output_VDJ_bardata_combined <-  output_VDJ_barplots %>%
  lapply(., function(out) out$data) %>%
  do.call(rbind, .) 

cl = setNames(output_VDJ_bardata_combined$manual_colors, output_VDJ_bardata_combined$Var1)

VDJ_barplot <- output_VDJ_bardata_combined %>%
  mutate(sample = factor(sample, levels = cellcount_order)) %>%
  group_by(sample) %>%
  arrange(desc(Freq)) %>%
  ggplot(aes(x = sample, y=Freq, fill = factor(Var1, levels = unique(Var1)))) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_manual(values = cl) +
  theme_minimal() +
  ylab("Percentage of cells") +
  xlab("") +
  theme(legend.position = "none",
        panel.grid = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.text.y = element_text(size=20), 
        axis.text.x = element_text(size=20),
        plot.margin = margin(0,0,0,0) 
        ) 
VDJ_barplot
```

```{r, fig.width = 14}
plot_grid(VDJ_cellcount_plot + theme(axis.text.x = element_blank()), 
          VDJ_barplot, 
          ncol = 1, 
          align = "hv",
          rel_heights = c(1:2)
          )
```


## clonal rank
### isotype
```{r}
VDJ_clonal_expansion
```

```{r}
VDJ_clonal_expansion_out_isotype <- VDJ_clonal_expansion(VDJ = vdj, subtypes = TRUE, celltype = "Bcells", color.by = "isotype")
VDJ_clonal_expansion_out_isotype
VDJ_clonal_expansion_out_isotype_combined <- VDJ_clonal_expansion_out_isotype[[2]] %>%
  do.call(rbind, .) %>%
  rownames_to_column(var = "row_id") %>% 
  mutate(sample = gsub("\\..*$", "", row_id))

colnames(VDJ_clonal_expansion_out_isotype_combined)
```

```{r, fig.width=14}
VDJ_clonal_expansion_out_isotype_combined %>%
  mutate(Color = gsub("[0-9]", "", Isotype)) %>%
  ggplot(aes(fill = Color, y=Counts, x=ClonalRank)) +
  geom_bar(stat="identity", width=0.6, color="black") + 
  theme_bw() + 
  scale_fill_manual(values = c("IGHG" = "green4", "IGHM" = "black", "IGHA" = "red3", "IGHD"="blue", "IGHE"="purple", "Unknown"="gray")) + 
  theme_classic() + 
  #title(sample) + 
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size = 12)) + 
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0.5)) + 
  labs(x = "Clonal rank", y = "Number of cells") +
  facet_wrap(~sample, scales = "free_y", nrow = 1) +
  geom_text(aes(label = sample, x=25, y=Inf, vjust=1.5), check_overlap = TRUE) + 
  guides(
    fill = guide_legend(title = "Isotype")
         ) + 
  theme(
    strip.placement = "inside",
    strip.text = element_blank(),
    strip.background = element_blank()
  )
```

#### subtype
```{r, fig.width=14}
VDJ_clonal_expansion_out_isotype_combined %>%
  mutate(Color = Isotype) %>%
  ggplot(aes(fill = Color, y=Counts, x=ClonalRank)) +
  geom_bar(stat="identity", width=0.6, color="black") + 
  theme_bw() + 
  #scale_fill_manual(values = c("IGHG" = "green4", "IGHM" = "black", "IGHA" = "red3", "IGHD"="blue", "IGHE"="purple", "Unknown"="gray")) + 
  scale_fill_manual(values = c("IGHG1" = "green","IGHG2" = "green3", "IGHG3"="green4", "IGHG4"="darkgreen", "IGHM" = "black", "IGHA1" = "red", "IGHA2"= "red4", "IGHD"="blue", "IGHE"="purple", "Unknown"="gray")) +
  theme_classic() + 
  #title(sample) + 
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size = 12)) + 
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0.5)) + 
  labs(x = "Clonal rank", y = "Number of cells") +
  geom_text(aes(label = sample, x=25, y=Inf, vjust=1.5), check_overlap = TRUE) + 
  guides(
    fill = guide_legend(title = "Isotype")
         ) + 
  theme(
    strip.placement = "inside",
    strip.text = element_blank(),
    strip.background = element_blank()
  )
```


### tissue
```{r}
VDJ_clonal_expansion_out_tissue <- VDJ_clonal_expansion(VDJ = vdj, subtypes = TRUE, celltype = "Bcells", color.by = "tissue")
VDJ_clonal_expansion_out_tissue
VDJ_clonal_expansion_out_tissue_combined <- VDJ_clonal_expansion_out_tissue[[2]] %>%
  do.call(rbind, .) %>%
  rownames_to_column(var = "row_id") %>% 
  mutate(sample = gsub("\\..*$", "", row_id))

colnames(VDJ_clonal_expansion_out_tissue_combined)
```

```{r, fig.width=14}
## Color palette from file
ggplotColours <- function(n, h = c(0, 360) + 15){
    if ((diff(h) %% 360) < 1) h[2] <- h[2] - 360/n
    grDevices::hcl(h = (seq(h[1], h[2], length = n)), c = 100, l = 65)
  }

VDJ_clonal_expansion_out_tissue_combined %>%
  #mutate(MainIsotype = gsub("[0-9]", "", Isotype)) %>%
  ggplot(aes(fill = Color, y=Counts, x=ClonalRank)) +
  geom_bar(stat="identity", width=0.6, color="black") + 
  theme_bw() + 
  scale_fill_manual(values = ggplotColours(n = length(unique(VDJ_clonal_expansion_out_tissue_combined$Color)))) +
  theme_classic() + 
  #title(sample) + 
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size = 12)) + 
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0.5)) + 
  labs(x = "Clonal rank", y = "Number of cells") +
  facet_wrap(~sample, scales = "free_y", nrow = 1) +
  geom_text(aes(label = sample, x=25, y=Inf, vjust=1.5), check_overlap = TRUE) + 
  guides(
    fill = guide_legend(title = "Tissue")
         ) + 
  theme(
    strip.placement = "inside",
    strip.text = element_blank(),
    strip.background = element_blank()
  ) -> plot_VDJ_clonal_expansion_out_tissue_combined

VDJ_clonal_expansion_out_tissue_combined
```

### timepoint
```{r}
VDJ_clonal_expansion_out_timepoint <- VDJ_clonal_expansion(
  VDJ = vdj %>%
  mutate(timepoint = case_when(
    timepoint == "d28+d35" ~ "d28d35",
    TRUE ~ timepoint
    )
  ),
  subtypes = TRUE, celltype = "Bcells", color.by = "timepoint")
VDJ_clonal_expansion_out_timepoint

VDJ_clonal_expansion_out_timepoint_combined <- VDJ_clonal_expansion_out_timepoint[[2]] %>%
  do.call(rbind, .) %>%
  rownames_to_column(var = "row_id") %>% 
  mutate(sample = gsub("\\..*$", "", row_id))

colnames(VDJ_clonal_expansion_out_timepoint_combined)
```

```{r, fig.width = 14}
VDJ_clonal_expansion_out_timepoint_combined %>%
  mutate(Color = as.character(Color)) %>%
  #mutate(MainIsotype = gsub("[0-9]", "", Isotype)) %>%
  ggplot(aes(fill = Color, y=Counts, x=ClonalRank)) +
  geom_bar(stat="identity", width=0.6, color="black") + 
  theme_bw() + 
  scale_fill_manual(values = ggplotColours(n = length(unique(VDJ_clonal_expansion_out_timepoint_combined$Color)))) +
  theme_classic() + 
  #title(sample) + 
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size = 12)) + 
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0.5)) + 
  labs(x = "Clonal rank", y = "Number of cells") +
  facet_wrap(~sample, scales = "free_y" #,
             ) +
             #nrow = 1) +
  geom_text(aes(label = sample, x=25, y=Inf, vjust=1.5), check_overlap = TRUE) + 
  guides(
    fill = guide_legend(title = "Timepoint")
         ) + 
  theme(
    strip.placement = "inside",
    strip.text = element_blank(),
    strip.background = element_blank()
  ) -> plot_VDJ_clonal_expansion_out_timepoint_combined

plot_VDJ_clonal_expansion_out_timepoint_combined
```

```{r, fig,width=14}
VDJ_clonal_expansion_out_timepoint[[2]]$`368-13` %>%
  mutate(Color = as.character(Color)) %>%
  ggplot(aes(fill = Color, y=Counts, x=ClonalRank)) +
  geom_bar(stat="identity", width=0.6, color="black") + 
  theme_bw() + 
  scale_fill_manual(values = ggplotColours(n = length(unique(VDJ_clonal_expansion_out_timepoint_combined$Color)))) +
  theme_classic() + 
  #title(sample) + 
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size = 12)) + 
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0.5)) + 
  labs(x = "Clonal rank", y = "Number of cells") +
  #facet_wrap(~sample, scales = "free_y", nrow = 1) +
  #geom_text(aes(label = sample, x=25, y=Inf, vjust=1.5), check_overlap = TRUE) + 
  guides(
    fill = guide_legend(title = "Timepoint")
         ) + 
  theme(
    strip.placement = "inside",
    strip.text = element_blank(),
    strip.background = element_blank()
  )
  
```


### v-gene
```{r}
length(unique(vdj$VDJ_vgene))
unique(vdj$VDJ_vgene)
```


```{r}
VDJ_clonal_expansion_out_vgene <- VDJ_clonal_expansion(VDJ = vdj, subtypes = TRUE, celltype = "Bcells", color.by = "VDJ_vgene")
VDJ_clonal_expansion_out_vgene
VDJ_clonal_expansion_out_vgene_combined <- VDJ_clonal_expansion_out_vgene[[2]] %>%
  do.call(rbind, .) %>%
  rownames_to_column(var = "row_id") %>% 
  mutate(sample = gsub("\\..*$", "", row_id))

colnames(VDJ_clonal_expansion_out_vgene_combined)
```

```{r, fig.width=14}
VDJ_clonal_expansion_out_vgene_combined %>%
  mutate(Color = sub("^([^0-9]+[0-9]).*", "\\1", Color)) %>%
  #select(Color) %>% unique()
  ggplot(aes(fill = Color, y=Counts, x=ClonalRank)) +
  geom_bar(stat="identity", width=0.6, color="black") + 
  theme_bw() + 
  scale_fill_manual(values = ggplotColours(
    n = length(
      VDJ_clonal_expansion_out_vgene_combined %>% 
        mutate(Color = sub("^([^0-9]+[0-9]).*", "\\1", Color)) %>%
        select(Color) %>% unique() %>% pull()
    )
  )) +
  theme_classic() + 
  #title(sample) + 
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size = 12)) + 
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0.5)) + 
  labs(x = "Clonal rank", y = "Number of cells") +
  facet_wrap(~sample, scales = "free_y", nrow = 1) +
  #facet_grid(cols = vars(sample), scales = "free_y", space = "free_y", axis = "all_y") +
  geom_text(aes(label = sample, x=25, y=Inf, vjust=1.5), check_overlap = TRUE) + 
  guides(
    fill = guide_legend(title = "V-gene")
  ) + 
  theme(
    strip.placement = "inside",
    strip.text = element_blank(),
    strip.background = element_blank()
  ) -> plot_VDJ_clonal_expansion_out_vgene_combined

plot_VDJ_clonal_expansion_out_vgene_combined
```

## together
```{}
VDJ_clonal_expansion_out_combined_colorby <- bind_rows(
  VDJ_clonal_expansion_out_isotype_combined %>%
    mutate(color_by = "isotype",
           Color = Isotype),
  VDJ_clonal_expansion_out_tissue_combined %>%
    mutate(color_by = "tissue"),
  VDJ_clonal_expansion_out_timepoint_combined %>%
    mutate(color_by = "timepoint")
  )
```

```{, fig.width = 20}
VDJ_clonal_expansion_out_combined_colorby %>%
  ggplot(aes(fill = Color, y=Counts, x=ClonalRank)) +
  geom_bar(stat="identity", width=0.6, color="black") + 
  theme_bw() + 
  #scale_fill_manual(values = ggplotColours(
  #  n = length(
  #    VDJ_clonal_expansion_out_vgene_combined %>% 
  #      mutate(Color = sub("^([^0-9]+[0-9]).*", "\\1", Color)) %>%
  #      select(Color) %>% unique() %>% pull()
  #  )
  #)) +
  theme_classic() + 
  #title(sample) + 
  theme(plot.title = element_text(hjust = 0.5), text = element_text(size = 12)) + 
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0.5)) + 
  labs(x = "Clonal rank", y = "Number of cells") +
  facet_wrap(sample~color_by, scales = "free_y", nrow=3) +
  geom_text(aes(label = sample, x=25, y=Inf, vjust=1.5), check_overlap = TRUE) + 
  #guides(
  #  fill = guide_legend(title = "V-gene")
  #) + 
  theme(
    strip.placement = "inside",
    strip.text = element_blank(),
    strip.background = element_blank()
  )
```

```{, fig.width = 20}
plot_VDJ_clonal_expansion_out_isotype_combined
plot_VDJ_clonal_expansion_out_tissue_combined
plot_VDJ_clonal_expansion_out_timepoint_combined

plot_grid(plotlist = plot_VDJ_clonal_expansion_out_isotype_combined, plot_VDJ_clonal_expansion_out_tissue_combined, plot_VDJ_clonal_expansion_out_timepoint_combined, nrow=3, ncol=1)
```

# Antibody forest
## packages
```{r}
bioc <- c("msa")

for (j in bioc)
{
  print(j)
  if (!require(j, character.only = TRUE)){
    BiocManager::install(j, dependencies = TRUE)
  }
  library(j, character.only = TRUE)
}
```



sourcing .R files
```{r}
# Load required libraries
library(jsonlite)  # For processing JSON data
library(httr)      # For making HTTP requests

# Define the GitHub repository details
repo_owner <- "alexyermanos"
repo_name <- "AntibodyForests"
folder_path <- "R"

# Construct the GitHub API URL to fetch folder content
api_url <- paste0("https://api.github.com/repos/", repo_owner, "/", repo_name, "/contents/", folder_path)

# Fetch the folder content using the GitHub API
response <- GET(api_url)

# Check if the request was successful
if (status_code(response) == 200) {
  
  # Parse the JSON response
  files_info <- content(response, "text", encoding = "UTF-8")
  files_info <- fromJSON(files_info)
  
  # Extract the raw URLs of all `.R` files
  r_files <- files_info[grepl("\\.R$", files_info$name), "download_url"]
  
  # Source each R file
  for (file_url in r_files) {
    message("Sourcing file: ", file_url)
    source(url(file_url))
  }
  
  message("All files sourced successfully!")
  
} else {
  stop("Failed to fetch folder content. HTTP status: ", status_code(response))
}
```

```{r}
head(vdj)
colnames(vdj)
```

```{r}
Af_build
```


```{r, message=FALSE}
AF_vdj_default <- Af_build(VDJ = vdj,
                           sequence.columns = c("VDJ_sequence_aa_trimmed"),
                           germline.columns = c("VDJ_germline_aa_trimmed")
                           )


```

```{r}
saveRDS(AF_vdj_default, file="AF_vdj_default.rds")
```

```{r}
head(AF_vdj_default)
```


## plot
```{r}
Af_plot_tree(AntibodyForests_object = AF_vdj_default,
             sample = "368-01a",
             clonotype = "clonotype1",
             label.by = "name",
             color.by = "isotype",
             edge.label = "lv",
             node.size = "expansion",
             node.size.scale = c(10, 20),
             main.title = "Lineage tree - Default algorithm",
             sub.title = "368-01a - clonotype1"
)
```


```{r}
names(AF_vdj_default$`368-01a`$clonotype1$igraph)
```


```{r, fig.width=14}
for (sample in names(AF_vdj_default)) {
    #print(sample)
    for (clonotype in head(names(AF_vdj_default[[sample]]),5)) {
      Af_plot_tree(AntibodyForests_object = AF_vdj_default,
                   sample = paste0(name),
                   clonotype = paste0(clonotype),
                   label.by = "name",
                   color.by = "isotype",
                   edge.label = "lv",
                   node.size = "expansion",
                   node.size.scale = c(10, 20),
                   sub.title = "Lineage tree - Default algorithm",
                   main.title = paste0(sample," - ",clonotype)
      )
    }
  }
```

