---
title: "Streptococcus pneumoniae data analysis"
author: "Jan Paul van Meenen"
date: "`r Sys.Date()`"
output: html_document
---

#Packages
```{r}
source("requirements/packages.R")
```

#Functions
```{r}
function_files <- list.files(path = "requirements/functions", pattern = "\\.R$",full.names = TRUE)

for (function_file in function_files) {
  source(function_file)
}
```

# Alphafold 3 structures
Here we predict the Alphafold 3 structures based on the recombinant sequences from the protein panel, to more easily use the alphafold server we create a .json file with all of the sequences.  
```{r}
# Define input and output paths
fasta_file <- "data/fasta/sp_recombinant_protein_sequences.fasta"
output_json_file <- "output/AlphaFold3/sp_recombinant_protein_sequences.json"

json_data <- generate_json(fasta_file, suffix = "recombinant protein")
write_json_file(json_data, output_json_file)
```

# Convert AlphaFold3 .cif to .pdb files
```{r}
AlphaFold3_zip_files <- list.files(path = "output/AlphaFold3/", pattern = "\\.zip$")

for (zip_file in AlphaFold3_zip_files) {
  unzip(zipfile = paste0("output/AlphaFold3/",zip_file), exdir = "output/AlphaFold3/individual_files/")
}
```

```{r}
AlphaFold3_folders <- list.dirs(path = "output/AlphaFold3/individual_files", full.names = FALSE, recursive = FALSE)

last(AlphaFold3_folders)
```


## Convert cif files with highest average b-value per protein
```{r}
for (AlphaFold3_folder in AlphaFold3_folders) {
  folder_path <- paste0("output/AlphaFold3/individual_files/", AlphaFold3_folder)
  cif_file_paths <- list.files(path = folder_path, pattern = "\\.cif$", full.names=TRUE)
  
  atom_df_cif <- NULL
  
  for (cif_file in cif_file_paths) {
    file_name <- sub("\\.cif$", "", basename(cif_file))
    
    protein_name <- gsub("fold_(.*?)_recombinant_protein.*", "\\1", file_name)
    fold_number <- gsub(".*_(\\d+)$", "\\1", file_name)
    
    cif_object <- read.cif.custom(cif_file)
    
    cif_df <- cif_object$atom %>%
      mutate(fold_number = fold_number)
    
    atom_df_cif <- bind_rows(atom_df_cif, cif_df)
  }
  
  highest_fold <- atom_df_cif %>%
    group_by(fold_number) %>%
    summarize(mean_b = mean(b)) %>%
    arrange(desc(mean_b)) %>%
    head(1) %>%
    pull(fold_number)
  
  highest_cif_file <- cif_file_paths[grepl(paste0("_", highest_fold, "\\.cif$"), cif_file_paths)]
  
  print(paste("Selected CIF file: ", highest_cif_file))
  
  # Read the CIF object of the selected file
  highest_cif_object <- read.cif.custom(highest_cif_file)
  
  write.pdb(highest_cif_object, file=paste0("output/AlphaFold3/generated_pdb/",protein_name,"_fold_",highest_fold,".pdb"))
}
```


## combine all into one df
```{r}
data_AF3_recombinant <- NULL

for (AlphaFold3_folder in AlphaFold3_folders) {
  folder_path <- paste0("output/AlphaFold3/individual_files/", AlphaFold3_folder)
  cif_file_paths <- list.files(path = folder_path, pattern = "\\.cif$", full.names=TRUE)
  
  for (cif_file in cif_file_paths) {
    file_name <- sub("\\.cif$", "", basename(cif_file))
    
    print(file_name)
    
    protein_name <- gsub("fold_(.*?)_recombinant_protein.*", "\\1", file_name)
    fold_number <- gsub(".*_(\\d+)$", "\\1", file_name)
    
    cif_object <- read.cif.custom(cif_file)
    
    cif_df <- cif_object$atom %>%
      mutate(protein = protein_name,
             fold_number = fold_number)
    
    data_AF3_recombinant <- bind_rows(data_AF3_recombinant, cif_df)
  }
}

data_AF3_recombinant
```

```{r}
write.csv(data_AF3_recombinant,
          "output/AlphaFold3/data_AF3_recombinant.csv", 
          row.names = FALSE)
```


### highest only
```{r}
recombinant_pdb_files_list <- list.files(path="output/AlphaFold3/generated_pdb/",pattern=".pdb", )
recombinant_pdb_files_list

data_AF3_recombinant_highest <- NULL

for (pdb_file in recombinant_pdb_files_list) {
  protein_name <- gsub("(.*?)_fold.*", "\\1", pdb_file)
  print(protein_name)

  pdb_object <- read.pdb(file =  paste0("output/AlphaFold3/generated_pdb/",pdb_file))
  
  atom_df_subset <- pdb_object$atom %>%
    mutate(protein = protein_name)
  
  data_AF3_recombinant_highest <- rbind(data_AF3_recombinant_highest, atom_df_subset)
}

data_AF3_recombinant_highest
```


## plot per protein
```{r, fig.width=14, warning=FALSE}
for (protein in unique(data_AF3_recombinant$protein)) {
  
  atom_subset <- data_AF3_recombinant %>% filter(protein == protein)

  plot <- atom_subset %>%
    ggplot(aes(x=fold_number, y=b)) +
    geom_violin(fill = "steelblue", alpha = 0.7, color = "black") +  # Violin plot
    geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + # Boxplot inside violin
    facet_wrap(~fold_number, nrow = 1) +
    labs(title = paste0(protein_name,"\npLDDT Score Distribution"), x = "fold number", y = "pLDDT") +
    scale_y_continuous(limits = c(NA, 100)) +
    theme_minimal() +
    theme(text = element_text(size = 14))
  
  print(plot)
}
```

## plot average pLDDTs of all proteins for highest average fold.
```{r}
data_AF3_recombinant_highest %>%
  group_by(protein) %>%
  summarize(average_b = mean(b)) %>%
  ggplot(aes(x="", y=average_b)) +
  geom_violin(fill = "steelblue", alpha = 0.7, color = "black") +  # Violin plot
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + # Boxplot inside violin
  #facet_wrap(~fold_number, nrow = 1) +
  scale_y_continuous(limits = c(NA, 100)) +
  labs(title = paste0("average pLDDT score distribution"), x = "", y = "pLDDT") +
  theme_minimal() +
  theme(text = element_text(size = 14))
  
```

```{r}
data_AF3_recombinant_highest %>%
  filter(elety == "CA") %>%
  group_by(protein) %>%
  summarize(average_b = mean(b)) %>%
  ggplot(aes(x="", y=average_b)) +
  geom_violin(fill = "steelblue", alpha = 0.7, color = "black") +  # Violin plot
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + # Boxplot inside violin
  scale_y_continuous(limits = c(NA, 100)) +
  labs(title = paste0("average pLDDT score distribution CA only"), x = "", y = "pLDDT") +
  theme_minimal() +
  theme(text = element_text(size = 14))
```

## join folds to Protein class
```{r, fig.width=14}
data_AF3_recombinant_highest_with_class <- data_AF3_recombinant_highest %>%
  filter(elety == "CA") %>%
  group_by(protein) %>%
  summarize(average_b = mean(b)) %>%
  arrange(desc(average_b)) %>%
  full_join(data_proteinpanel %>%
                         mutate(
                           clean_proteinname = gsub(" ","",tolower(gsub("\\(.*\\)", "", `Proteins (TIGR4)`)))
                           ) %>%
                         select(`Protein class`, clean_proteinname), 
                       by = c("protein" = "clean_proteinname"), keep = TRUE) 

group_order_median <- data_AF3_recombinant_highest_with_class %>% 
  group_by(`Protein class`) %>%
  summarize(group_average_b = median(average_b)) %>%
  arrange(desc(group_average_b)) %>%
  pull(`Protein class`)

data_AF3_recombinant_highest_with_class %>%
  ggplot(aes(x=average_b, y=reorder(protein, average_b), fill = factor(`Protein class`, levels = group_order_median))) +
  geom_col() +
  geom_vline(xintercept=80) +
  labs(title = "Protein Ranking by Average pLDDT Score",
       x = "Average pLDDT Score",
       y = "Protein",
       fill = "Protein Class (median sorted)") +
  theme_minimal() 

group_order_mean <- data_AF3_recombinant_highest_with_class %>% 
  group_by(`Protein class`) %>%
  summarize(group_average_b = mean(average_b)) %>%
  arrange(desc(group_average_b)) %>%
  pull(`Protein class`)

data_AF3_recombinant_highest_with_class %>%
  ggplot(aes(x=average_b, y=reorder(protein, average_b), fill = factor(`Protein class`, levels = group_order_mean))) +
  geom_col() +
  geom_vline(xintercept=80) +
  labs(title = "Protein Ranking by Average pLDDT Score",
       x = "Average pLDDT Score",
       y = "Protein",
       fill = "Protein Class (mean sorted)") +
  theme_minimal()
```

```{r, fig.width=14}
data_AF3_recombinant_highest_with_class %>%
  ggplot(aes(x = average_b, 
             fill = factor(`Protein class`, levels = group_order_mean),
             color = factor(`Protein class`, levels = group_order_mean))) +
  geom_density(alpha = 0.3) +
  #scale_fill_viridis_d(option = "plasma") + 
  labs(title = "Distribution of Average pLDDT Score by Protein Class",
       x = "Average pLDDT Score",
       y = "Density",
       fill = "Protein Class",
       color = "Protein Class",
       linetype = "Quantity") +
  scale_x_continuous(limits = c(NA, 100)) +
  #facet_wrap(~`Protein class`, scales = "free_y") +
  
  geom_vline(data = data_AF3_recombinant_highest_with_class %>%
               group_by(`Protein class`) %>%
               summarise(mean_b = median(average_b, na.rm = TRUE)),
             aes(xintercept = mean_b, color=factor(`Protein class`, levels = group_order_mean), linetype = "Median"),
              size= 1) +
  geom_vline(data = data_AF3_recombinant_highest_with_class %>%
               group_by(`Protein class`) %>%
               summarise(mean_b = mean(average_b, na.rm = TRUE)),
             aes(xintercept = mean_b, color=factor(`Protein class`, levels = group_order_mean), linetype = "Mean"),
              size= 1) +
  scale_linetype_manual(values = c("Mean" = "dashed", "Median" = "solid")) +
  
  guides(fill = guide_legend(override.aes = list(alpha = 0.5), order=1),  # Overlapping legends
         color = guide_legend(override.aes = list(alpha = 0.5), order=1),  # Overlapping legends
         linetype = guide_legend(order = 2)) +
  theme_minimal() 
```

```{r, fig.width=14}
data_AF3_recombinant_highest_with_class %>%
  ggplot(aes(x = average_b, 
             fill = factor(`Protein class`, levels = group_order_mean),
             color = factor(`Protein class`, levels = group_order_mean)
             )
         ) +
  geom_histogram(alpha = 0.5, bins=25) +
  #scale_fill_viridis_d(option = "plasma") + 
  labs(title = "Distribution of Average pLDDT Score by Protein Class",
       x = "Average pLDDT Score",
       y = "Density",
       fill = "Protein Class",
       color = "Protein Class",
       linetype = "Quantity") +
  scale_x_continuous(limits = c(NA, 100)) +
  #facet_wrap(~`Protein class`, scales = "free_y") +
  
  #geom_vline(data = data_AF3_recombinant_highest_with_class %>%
  #             group_by(`Protein class`) %>%
  #             summarise(mean_b = median(average_b, na.rm = TRUE)),
  #           aes(xintercept = mean_b, color=factor(`Protein class`, levels = group_order_mean), linetype = "Median"),
  #            size= 1) +
  #geom_vline(data = data_AF3_recombinant_highest_with_class %>%
  #             group_by(`Protein class`) %>%
  #             summarise(mean_b = mean(average_b, na.rm = TRUE)),
  #           aes(xintercept = mean_b, color=factor(`Protein class`, levels = group_order_mean), linetype = "Mean"),
  #            size= 1) +
  #scale_linetype_manual(values = c("Mean" = "dashed", "Median" = "solid")) +
  
  guides(fill = guide_legend(override.aes = list(alpha = 0.5), order=1),  # Overlapping legends
         color = guide_legend(override.aes = list(alpha = 0.5), order=1),  # Overlapping legends
         linetype = guide_legend(order = 2)) +
  theme_minimal() 
```

