---
title: "Streptococcus pneumoniae data analysis"
author: "Jan Paul van Meenen"
date: "`r Sys.Date()`"
output: html_document
---

#Packages
```{r}
data_management_packages_to_install <- c("data.table", "tidyverse", "ggplot2", "readxl")

packages_to_install <- c(data_management_packages_to_install)

# Check and install packages using a for loop
for (package in packages_to_install) {
  if (!requireNamespace(package, quietly = TRUE)) {
    install.packages(package)
  }
  library(package, character.only = TRUE)
}
```

#Data Protein panel
##loading
```{r}
data_proteinpanel <- rbind(
    read_xlsx("data/20231124 Pneumococcal protein panel (with accession number)_SH_WM_FV.xlsx", range = "B3:M31"),
    read_xlsx("data/20231124 Pneumococcal protein panel (with accession number)_SH_WM_FV.xlsx", range = "N3:Y30")
    ) %>%
  mutate(
    "Protein class" = case_when(
      `No.` < 24 ~ "Lipoproteins",
      `No.` >= 24 & `No.` < 28 ~ "Cytoplasmic P.",
      `No.` == 28 ~ "Membrane associated P.",
      `No.` >= 29 & `No.` < 38 ~ "Choline-binding P.", 
      `No.` >= 38 & `No.` < 46 ~ "Sortase-anchored P.",
      `No.` == 46 ~ "NCSP",
      TRUE ~ "Others"),
    "Proteins (TIGR4)" = na_if(`Proteins (TIGR4)`, "-"),
    "Proteins (D39)" = na_if(`Proteins (D39)`, "-")
  )

data_proteinpanel
```
```{r}
colnames(data_proteinpanel)
```


```{r}
data_proteinpanel %>%
  select("Proteins (TIGR4)","AA sequence of recombinant protein")
```

# Create fasta    
```{r}
fasta_content <- apply(data_proteinpanel, 1, function(row) {
  paste0(">", row["Proteins (TIGR4)"]," ", row["Protein class"], "\n", row["AA sequence of recombinant protein"])
})

writeLines(fasta_content, "sp_recombinant_protein_sequences.fasta")
```

```{r}
cat(fasta_content)
```

#Data structure
##loading
```{r}
data_structural <- read_xlsx("data/Pneumoproteins_PDB_structuralData.xlsx") %>%
  rename_at(vars(3, 4), ~ c("Proteins (TIGR4) name", "Proteins (TIGR4)")) %>%
  mutate(
    "Protein class" = case_when(
      `No.` < 24 ~ "Lipoproteins",
      `No.` >= 24 & `No.` < 33 ~ "CPBs", 
      `No.` >= 33 & `No.` < 41 ~ "Sortase-anchored P.",
      `No.` >= 41 & `No.` < 45 ~ "Cytoplasmic P.",
      `No.` == 45 ~ "Membrane associated P.",
      `No.` == 46 ~ "NCSP",
      TRUE ~ "Others"),
    "Proteins (TIGR4)" = na_if(`Proteins (TIGR4)`, "-"),
    "Proteins (D39)" = na_if(`Proteins (D39)`, "-")
  )

data_structural  
```

```{r}
structural_packages_to_install <- c("bio3d","NGLVieweR")

packages_to_install <- c(structural_packages_to_install)

# Check and install packages using a for loop
for (package in packages_to_install) {
  if (!requireNamespace(package, quietly = TRUE)) {
    install.packages(package)
  }
  library(package, character.only = TRUE)
}
```

```{r}
data_structural_with_pdb <- data_structural %>%
  mutate(pdb_objects = map(
    `PDB entry`,  # Column containing PDB entry IDs or file names
    ~ if (!is.na(.x)) {
      # Split the PDB entries if multiple IDs are present
      pdb_list <- strsplit(.x, ",\\s*")[[1]]
      
      # Map over each PDB entry to read the corresponding PDB file
      setNames(
        map(pdb_list, function(pdb_entry) {
          tryCatch(
            {
              # Read the PDB file using read.pdb
              pdb_obj <- read.pdb(pdb_entry)
              pdb_obj$pdb_entry <- pdb_entry
              
              write.pdb(pdb_obj, file = paste0("output/pdb/",pdb_entry, ".pdb"))
              
              return(pdb_obj)
            },
            error = function(e) {
              message(paste("Error reading PDB file:", pdb_entry))  # Optionally print a message on error
              return(NA)  # Return NA if an error occurs
            }
          )
        }),
        pdb_list  # Set the names of the list elements to be the PDB IDs
      )
    } else {
      NA  # For rows where `PDB entry` is NA, return NA
    }
    
    
  )) 
  
```


```{r}
data_structural_with_pdb %>%
  filter(!is.na(`PDB entry`)#,
         #`Protein class` != "Others"
         ) %>%
  pull(pdb_objects) %>%
  map(~ {
    map(.x, function(pdb_object) {
      # Check if pdb_object is a list and contains the necessary components
      if (is.list(pdb_object) && !is.null(pdb_object$pdb_entry) && !is.null(pdb_object$seqres)) {
        # Extract sequence information
        fasta_header <- pdb_object$pdb_entry
        fasta_sequence <- paste(aa321(pdb_object$seqres), collapse = "")
        
        # If both header and sequence are non-empty, return FASTA formatted string
        if (nchar(fasta_header) > 0 && nchar(fasta_sequence) > 0) {
          paste0(">", fasta_header, "\n", fasta_sequence)
        } else {
          # Print a message if either header or sequence is missing
          message("Skipping empty PDB entry: ", pdb_object$pdb_entry)
          return(NULL)  # Skip this entry
        }
      } else {
        return(NULL)  # Return NA if invalid
      }
    })
  }) %>%
  unlist() %>%
  writeLines("pdb_protein_sequences.fasta")
```

### split into parts
Bepi-2.0 only allows for 50 proteins at a time.
```{r}
fasta_lines <- readLines("pdb_protein_sequences.fasta")

# Group lines into proteins based on `>...` headers
protein_entries <- split(fasta_lines, cumsum(grepl("^>", fasta_lines)))

# Split protein entries into chunks of 50
protein_chunks <- split(protein_entries, ceiling(seq_along(protein_entries) / 50))

for (i in seq_along(protein_chunks)) {
  # Combine lines for each chunk
  chunk_lines <- unlist(protein_chunks[[i]])
  
  # Write the chunk to a new file
  writeLines(chunk_lines, paste0("pdb_protein_sequences_part_", i, ".fasta"))
}

```


# BepiPred-2.0 results:
https://services.healthtech.dtu.dk/services/BepiPred-2.0/ 
```{r}
data_BepiPred_2 <- rbind(
    read_csv("output/BepiPred2.0/BepiPred_part1_summary_67868578000F647FB45B7803.csv"),
    read_csv("output/BepiPred2.0/BepiPred_part2_summary_67868580000F64B5DAEDC816.csv")
    ) 

data_BepiPred_2 %>%
  filter(Entry == "3CX3") %>%
  mutate(AminoAcid = aa123(AminoAcid))
```

## 3CX3 test
### fixing alignment
```{r}
pdb_3CX3 <- read.pdb("3CX3", rm.alt = FALSE)

aa321(pdb_3CX3$seqres)
#pdb_3CX3$atom

pdbseq(pdb_3CX3)
```


### checking alignment
```{r, fig.width=14}
pdb_3CX3$atom %>%
  ggplot(aes(x=resno, y=chain, col = resid)) +
  geom_vline(xintercept=34) +
  geom_vline(xintercept=50) +
  geom_vline(xintercept=100) +
  geom_point() +
  geom_point(data = data_BepiPred_2 %>% filter(Entry == "3CX3"),
             aes(x=Position+27, col=aa123(AminoAcid), y="FASTA aligned")
             ) +
  geom_point(data = data_BepiPred_2 %>% filter(Entry == "3CX3"),
             aes(x=Position, col=aa123(AminoAcid), y="FASTA")
             ) +
  scale_x_continuous(limits = c(NA,150)) +
  #scale_y_discrete(expand = c(0,0)) +
  theme_bw() 
```

```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("Biostrings")
BiocManager::install("pwalign")
#install.packages("pwalign")
library("pwalign")
```

```{r}
chr_list <- pdbseq(pdb_3CX3)
chr_positions <- names(chr_list)
chr_amino_acids_1 <- unname(chr_list)
chr_amino_acids_3 <- aa123(chr_amino_acids_1)

#n <- length(chr_amino_acids_3)

chr_subset <- data.frame(Position = as.numeric(chr_positions),
                         AminoAcid = chr_amino_acids_3) %>%
  distinct(Position, .keep_all = TRUE)

BepiPred_subset <- data_BepiPred_2 %>%
  filter(Entry == "3CX3") 

pattern_str <- chr_subset %>% 
  mutate(AminoAcid = aa321(AminoAcid)) %>%
  #replace_na(list(AminoAcid = "-")) %>%
  pull(AminoAcid) %>%
  str_c(collapse = "")

bepi_str <- BepiPred_subset %>% 
  pull(AminoAcid) %>%
  str_c(collapse = "")

alignment <- pairwiseAlignment(pattern_str, bepi_str)

# Extract the aligned sequences
aligned_pattern <- alignedPattern(alignment)

#shift <- 0
#while (substr(aligned_pattern, shift + 1, shift + 1) == "-") { #count the leading gaps
#  shift <- shift + 1
#}
shift <- str_locate(as.character(aligned_pattern), "[^\\-]")[1]

```

```{r}
starting_resno <- chr_subset$Position[1]

required_adjustment <- starting_resno - shift

BepiPred_subset <- BepiPred_subset %>%
  mutate(resno = Position + required_adjustment) %>%
  select(Entry, Position, resno, AminoAcid, EpitopeProbability)
```




### 3D plots
```{r}
highlight_data <- BepiPred_subset %>%
  filter(resno >= min(chr_subset$Position), resno <= max(chr_subset$Position)) %>%
  mutate(
    EpiColor = ifelse(
      EpitopeProbability <= 0.5,
      "white",  # Assign white for probabilities below 0.6
      colorRampPalette(c("#f0f8ff", "darkblue","black"))(10)[
        pmin(pmax(round((EpitopeProbability - 0.5) / 0.05) + 1, 1), 10)
      ]
    )
  ) 

# Create the selection string as a space-separated string 
selections <- paste0(highlight_data$resno, ":A")

selection_str_base <- paste(selections[highlight_data$EpiColor == "white"], collapse = " ") #non-epitopes
selection_str <- paste(selections[highlight_data$EpiColor != "white"], collapse = " ") #epitopes


```

```{r}
viewer_3CX3 <- NGLVieweR("output/pdb/3CX3.pdb") %>%
  addRepresentation("cartoon",
    param = list(colorScheme = "residueindex", 
                     sele = ":A")
  ) %>%
  addRepresentation(
    "surface",  # Add surface representation once
    param = list(
      sele = selection_str_base,
      colorValue = "white"
    )
  ) %>%
 stageParameters(backgroundColor = "white") %>%
   addRepresentation("label",
    param = list(
      sele = selection_str,
      labelType = "format",
      labelFormat = "[%(resname)s]%(resno)s",
      labelGrouping = "residue", 
      color = "black",
      fontFamiliy = "sans-serif",
      xOffset = 0,
      yOffset = 0,
      zOffset = 20,
      fixedSize = TRUE,
      radiusType = 1,
      radiusSize = 1.5, # Label size
      showBackground = FALSE
      # backgroundColor="black",
      # backgroundOpacity=0.5
    )
  ) 

for (i in seq_len(nrow(highlight_data))) {
  residue <- highlight_data$resno[i]  # Residue position
  color <- highlight_data$EpiColor[i]  # Color based on EpitopeProbability
  
  if (color != "white") { # Add colored surface representation for each residue
  viewer_3CX3 <- viewer_3CX3 %>%
    addRepresentation(
      "surface",
      param = list(
        sele = paste0(residue, ":A"),  # Select the specific residue (change chain if necessary)
        colorValue = color  # Color based on the probability
      )
    ) }
}

viewer_3CX3
```


```{r}
library("htmlwidgets")
library("htmltools")

saveWidget(viewer_3CX3, file = "output/snapshot/3CXC.html", title ="3CX3", selfcontained = TRUE)
```

### Heatmap
```{r}
pdb_entry <- "3CX3"

pdb_object <- read.pdb(pdb_entry,rm.alt=FALSE)
  
required_adjustment <- map.seqresno.to.resno(pdb_object)
  #print(required_adjustment)
  
resmap <- create.resmap.bepipred(pdb_object$atom, 
                                           data_BepiPred_2 %>% filter(Entry == pdb_entry),
                                           required_adjustment)

resmap

plot_bepipred <- resmap %>%
  ggplot(aes(x=resno)) +
  geom_tile(aes(y=0.1,fill = EpitopeProbability), height = 0.1) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0.5, limits = c(0, 1)) +  
  theme_minimal() + 
  theme(
    axis.text.y = element_blank(),   
    axis.title.y = element_blank(),  
    axis.ticks.y = element_blank()
  ) +
  labs(fill = "BepiPred-2.0 prediction")

plot_bepipred
```

## Generalized
```{r}
PDB_list <- data_structural_with_pdb %>%
  pull(`PDB entry`) %>% 
  na.omit() %>% 
  str_split(", ") %>% 
  unlist() %>% 
  unique() 

PDB_list
```

```{r}
pdb_files_list <- sub("\\.pdb$", "",
                      list.files(path = "output/pdb/", 
                             pattern = "\\.pdb$")
                      )

for (pdb_entry in pdb_files_list) {
  print(pdb_entry)

}
```

### function map.seqresno.to.resno
#### function map.seqposno.to.resno
```{r}
map.seqposno.to.resno <- function(sequence_AA1, pdb_object, verbose=FALSE, type="overlap") {
  pattern <- pdbseq(pdb_object)
  unique_pattern <- pattern[!duplicated(names(pattern))]
  pattern_resno <- as.numeric(names(unique_pattern))
  pattern_seq <- str_c(
    unname(unique_pattern),
    collapse = ""
    )
  
  subject_seq <- sequence_AA1
  
  alignment <- pairwiseAlignment(pattern_seq, subject_seq, type=type)

  # Extract the aligned sequences
  aligned_pattern <- alignedPattern(alignment)
  #first_10_residues_pattern <- substr(as.character(aligned_pattern), 1, 10)
  residues_up_to_gap <- str_extract(as.character(aligned_pattern), "^[^-]*")
  
  shift <- str_locate(subject_seq, residues_up_to_gap)[1]
  
  if(is.na(shift)) {
    warning("Missing sequence residue compared to pdb structure")
    aligned_subject <- alignedSubject(alignment)
    residues_up_to_gap <- str_extract(as.character(aligned_subject), "^[^-]*")
    shift <- str_locate(subject_seq, residues_up_to_gap)[1]
  }
  
  required_adjustment <- pattern_resno[1] - shift
  
  #verbose
  if(verbose == TRUE) {
    print(deparse(substitute(pdb_object)))
    print(aligned_pattern)
    print(alignedSubject(alignment)) 
    print(residues_up_to_gap)
    print(shift)
  }
  
  return(c(required_adjustment))
}

```

#### function map.seqresno.to.resno
```{r}
map.seqresno.to.resno <- function(pdb_object, verbose=FALSE, type="overlap") {
  subject_seq <- str_c(
    aa321(pdb_object$seqres),
    collapse = ""
    )
  
  required_adjustment <- map.seqposno.to.resno(subject_seq, pdb_object, verbose=verbose, type=type)
  
  return(c(required_adjustment))
}
```

```{r}
pdb_entry <- "3E0M"

pdb_object <- read.pdb(pdb_entry,rm.alt=FALSE)
pdbseq(pdb_object)
aa321(pdb_object$seqres)
  
required_adjustment <- map.seqresno.to.resno(pdb_object, verbose = TRUE, type="overlap")
print(required_adjustment)
  
resmap <- create.resmap.bepipred(pdb_object$atom, 
                                           data_BepiPred_2 %>% filter(Entry == pdb_entry),
                                           required_adjustment)

resmap
```


### function prediction.viewer
```{r, function prediction.viewer}
prediction.viewer <- function(pdb_entry, directory, predictions_df=NULL) {
  file_path <- paste0(directory,pdb_entry,".pdb")
  view <- NGLVieweR(file_path) %>%
  stageParameters(backgroundColor = "white", zoomSpeed = 1) %>%
  addRepresentation("cartoon",
    param = list(colorScheme = "residueindex", 
                     sele = ":A")
  ) 
  
  if(!is.null(predictions_df)) {
    for (row in 1:nrow(predictions_df)) {
      #print(predictions_df)
      #print(predictions_df$resno_string[row])
      #print(predictions_df$color[row])
      view <- view %>%
        addRepresentation(
          "surface",  # Add surface representation once
          param = list(
            sele = predictions_df$resno_string[row],
            colorValue = predictions_df$color[row]
      )
    )
    } 
    
  }
  
  return(view)
}

```

```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  prediction.viewer(pdb_entry, directory="output/pdb/")

}
```

### function create.resmap.bepipred
```{r}
create.resmap.bepipred <- function(pdb_atom_df, bepipred_epitope_prediction_df,required_adjustment) {
  resmap_df <- pdb_atom_df %>% 
    select(resid,resno) %>%
    unique() %>%
    filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
    left_join(bepipred_epitope_prediction_df %>%
                #select(Position,AminoAcid,EpitopeProbability) %>%
                mutate(resno = Position + required_adjustment),
              by = c("resno")
              )
  return(resmap_df)
}
```

## BepiPred-2.0 with resno
```{r}
data_BepiPred_2_with_resno <- NULL

for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  
  pdb_object <- read.pdb(pdb_entry,rm.alt=FALSE)
  
  required_adjustment <- map.seqresno.to.resno(pdb_object)
  #print(required_adjustment)
  
  resmap <- create.resmap.bepipred(pdb_object$atom, 
                                           data_BepiPred_2 %>% filter(Entry == pdb_entry),
                                           required_adjustment)
  
  if (is.null(data_BepiPred_2_with_resno)) {
    data_BepiPred_2_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_BepiPred_2_with_resno <- rbind(data_BepiPred_2_with_resno, resmap)
  }
}

data_BepiPred_2_with_resno
```

```{r}
data_BepiPred_2 %>%
  filter(Entry == "3E0M")
```


```{r}
pdbs_path = "output/pdb/"

for (pdb_entry in pdb_files_list) {
  file_path = paste0(pdbs_path,pdb_entry,".pdb")
  print(file_path)
  #if(file.exists(file_path)) {
  #  pdb_object <- read.pdb(file_path, rm.alt = FALSE)
  #} else {
  #print(paste0("file missing: ", pdb_entry))
  #}
  pdb_object <- read.pdb(pdb_entry,rm.alt=FALSE)
  
  required_adjustment <- map.seqresno.to.resno(pdb_object)
  #print(required_adjustment)
  
  resmap <- create.resmap.bepipred(pdb_object$atom, 
                                   data_BepiPred_2 %>% filter(Entry == pdb_entry),
                                   required_adjustment)
  print(resmap)
  
  color_map <- data.frame(
    ProbabilityBin = levels(cut(seq(0.0, 1, by = 0.05), breaks = seq(0, 1, by = 0.05), include.lowest = TRUE)),
    color = colorRampPalette(c("darkblue","white","red"))(20)
    ) %>%
    rbind(c(NA,"black"))
  
  #print(color_map)
  
  predictions_df <- resmap %>%
    mutate(ProbabilityBin = cut(EpitopeProbability, 
                              breaks = seq(0, 1, by = 0.05), 
                              include.lowest = TRUE)
    ) %>%
    group_by(ProbabilityBin) %>% 
    summarize(resno_list = list(paste0(resno,":A")), .groups = "drop") %>%
    mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " "))) %>%
    left_join(color_map)
  
  #print(predictions_df) 
  
  view <- prediction.viewer(pdb_entry, pdbs_path, predictions_df = predictions_df)
  
  saveWidget(view, file = paste0("output/snapshot/",pdb_entry,".html"), title=pdb_entry, selfcontained = TRUE)
}
```




```{r}
data_BepiPred_2 %>% filter(Entry == "8QLK")
```

```{r}
levels(cut(seq(0.05, 1, by = 0.05), breaks = seq(0, 1, by = 0.05), include.lowest = TRUE))
```


```{r}
library(viridisLite)
my_palette <- inferno(20)
my_palette <- colorRampPalette(c("blue","white","red"))(20)
barplot(rep(1, length(my_palette)), col = my_palette, border = NA, axes=FALSE, space = 0, main = "Visualization of colorRampPalette")
```

```{r}
# Create a mapping of bins to colors
color_map <- data.frame(
  ProbabilityBin = levels(cut(seq(0.05, 1, by = 0.05), breaks = seq(0, 1, by = 0.05), include.lowest = TRUE)),
  color = inferno(20)
)

color_map
```


```{r}
map.seqresno.to.resno(pdb_3CX3)

map.seqresno.to.resno(read.pdb(paste0(pdbs_path,"3CX3",".pdb"),rm.alt = FALSE))
```

## Ggplot
```{r}
library("cowplot")
```


```{r, fig.width=14}
plot_bepipred <- data_BepiPred_2 %>%
  filter(Entry == "3CX3") %>%
  ggplot(aes(x=Position)) +
  geom_tile(aes(y=0.1,fill = EpitopeProbability), height = 0.1) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0.5, limits = c(0, 1)) +  
  theme_minimal() + 
  theme(
    axis.text.y = element_blank(),   
    axis.title.y = element_blank(),  
    axis.ticks.y = element_blank()
  ) +
  labs(fill = "BepiPred-2.0 prediction")

plot_surface <- data_BepiPred_2 %>%
  filter(Entry == "3CX3") %>%
  ggplot(aes(x=Position)) +
  geom_tile(aes(y=0.1,fill = RelativeSurfaceAccessilibity), height = 0.1) +
  theme_minimal() + 
  theme(
    axis.text.y = element_blank(),   
    axis.title.y = element_blank(),  
    axis.ticks.y = element_blank()
  )

plot_grid(
  plot_bepipred + theme(
    axis.text.x = element_blank(),   
    axis.title.x = element_blank(),  
    axis.ticks.x = element_blank()
  ), 
  plot_surface, align = "vh", ncol = 1)
```

# BepiPred-3.0 results
https://services.healthtech.dtu.dk/services/BepiPred-3.0/
```{r}
data_BepiPred_3 <- rbind(
    read_csv("output/BepiPred3.0/raw_output_part1.csv"),
    read_csv("output/BepiPred3.0/raw_output_part2.csv")
    ) %>%
  #mutate(Residue = aa123(Residue)) %>%
  group_by(Accession) %>%
  mutate(Position = row_number())
```

```{r}
data_BepiPred_3 %>%
  filter(Accession == "3E0M") 
```

## BepiPred-3.0 with resno
```{r}
data_BepiPred_3_with_resno <- NULL

for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb(pdb_entry,rm.alt=FALSE)
  
  required_adjustment <- map.seqresno.to.resno(pdb_object)
  #print(required_adjustment)
  
  resmap <- create.resmap.bepipred(pdb_object$atom, 
                                           data_BepiPred_3 %>% filter(Accession == pdb_entry),
                                           required_adjustment)
  
  if (is.null(data_BepiPred_3_with_resno)) {
    data_BepiPred_3_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_BepiPred_3_with_resno <- rbind(data_BepiPred_3_with_resno, resmap)
  }
}

data_BepiPred_3_with_resno
```

```{r}
data_BepiPred_3_with_resno %>%
  pull(`BepiPred-3.0 score`) %>%
  summary()
```


# DiscoTope-3.0
https://services.healthtech.dtu.dk/services/DiscoTope-3.0/
```{r}
library(purrr)
library(readr)

csv_files <- list.files("output/DiscoTope", pattern = "*.csv", full.names = TRUE)

data_DiscoTope <- NULL

for (csv_file in csv_files) {
  temp_data <- read_csv(csv_file) 
  
  if (is.null(data_DiscoTope)) {
    data_DiscoTope <- temp_data
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_DiscoTope <- rbind(data_DiscoTope, temp_data)
  }
}

data_DiscoTope
```

```{r}
data_DiscoTope %>%
  filter(pdb == "3CX3")
```

## Test 3CX3
```{r}
data_DiscoTope <- data_DiscoTope %>%
  mutate(pdb = sub("_.*", "", pdb)) %>%
  rename("resno" = res_id)

data_DiscoTope %>%
  filter(pdb == "3CX3")

data_BepiPred_2 %>%
  filter(Entry == "3CX3")
```


```{r}
plot_discotope <- data_DiscoTope %>%
  filter(pdb == "3CX3") %>%
  ggplot(aes(x=resno, y=calibrated_score, fill=calibrated_score)) +
  geom_col(width = 1) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +  
  theme_minimal() + 
  theme(
    axis.text.y = element_blank(),   
    axis.title.y = element_blank(),  
    axis.ticks.y = element_blank()
  ) +
  labs(fill = "DiscoTope-3.0 prediction")

plot_discotope
```

# Epigraph
http://epigraph.kaist.ac.kr/
```{r}
epigraph_zip_files <- list.files(path = "output/EpiGraph/", pattern = "\\.zip$")

for (zip_file in epigraph_zip_files) {
  unzip(zipfile = paste0("output/EpiGraph/",zip_file), exdir = "output/EpiGraph")
}

epigraph_csv_files <- list.files(path = "output/EpiGraph/", pattern = "\\.csv$")
epigraph_csv_files
```

```{r}
data_Epigraph <- NULL

for (csv_file in epigraph_csv_files) {
  temp_data <- read_csv(paste0("output/EpiGraph/",csv_file)) 
  
  if (is.null(data_Epigraph)) {
    data_Epigraph <- temp_data
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_Epigraph <- rbind(data_Epigraph, temp_data)
  }
}

data_Epigraph <- data_Epigraph %>% 
  separate(Residue, into = c("chain", "resid", "resno"), sep = ":", convert = TRUE) %>%
  mutate(PDB = toupper(PDB))

data_Epigraph

write.csv(data_Epigraph,"output/EpiGraph/data_Epigraph.csv", row.names=FALSE)
```

```{r}
data_Epigraph %>%
  count(chain) 
```


```{r}
data_Epigraph_subset <- data_Epigraph %>% filter(PDB == "3ZPP")

plot_epigraph <- data_Epigraph_subset %>% 
    ggplot(aes(x=resno, y=Score, fill=Score)) +
    geom_col(width = 1,) +
    scale_fill_gradient2(low = "blue", mid = "gray95", high = "red") +  
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "EpiGraph prediction")

plot_epigraph
```

#GraphBepi
```{r}

```



# Plot all together
```{r, warning=FALSE, fig.width=14, fig.height=10}
for (pdb_entry in pdb_files_list) {
  #print(pdb_entry)
  
  data_BepiPred_2_with_resno_subset <- data_BepiPred_2_with_resno %>% filter(Entry == pdb_entry)
  #print(data_BepiPred_2_with_resno_subset)
  
  plot_bepipred_2_RSA <- data_BepiPred_2_with_resno_subset %>% 
    ggplot(aes(x=resno, y=RelativeSurfaceAccessilibity, fill=RelativeSurfaceAccessilibity)) +
    geom_col(width = 1) +  
    scale_fill_gradient2(low = "darkcyan", mid = "gray95", high = "deeppink4", midpoint = 0.5, limits = c(0, 1)) +  
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "Relative Surface Accessilibity\n(NetsurfP)")
  
  
  plot_bepipred_2 <- data_BepiPred_2_with_resno_subset %>% 
    ggplot(aes(x=resno, y=EpitopeProbability, fill=EpitopeProbability)) +
    geom_col(width = 1) +  
    scale_fill_gradient2(low = "blue", mid = "gray95", high = "red", midpoint = 0.5, limits = c(0, 1)) +  
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "BepiPred-2.0 prediction")
  #print(plot_bepipred_2)
  
  data_BepiPred_3_with_resno_subset <- data_BepiPred_3_with_resno %>% filter(Accession == pdb_entry)
  median_BepiPred_3_subset_score <- median(data_BepiPred_3_with_resno_subset$`BepiPred-3.0 score`)
  
  plot_bepipred_3 <- data_BepiPred_3_with_resno_subset %>%
    ggplot(aes(x=resno, y=`BepiPred-3.0 score`, fill=`BepiPred-3.0 score`)) +
    geom_col(width = 1) +
    scale_fill_gradient2(low = "blue", mid = "gray95", high = "red", 
                         midpoint = median_BepiPred_3_subset_score#, 
                         #limits = c(0, NA)
                         ) +
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "BepiPred-3.0 prediction")
  
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  #print(data_DiscoTope_subset)
  plot_discotope <- data_DiscoTope_subset %>% 
    ggplot(aes(x=resno, y=calibrated_score, fill=calibrated_score)) +
    geom_col(width = 1) +
    scale_fill_gradientn(colors = c("blue", "gray95", "#FFC0C0", "#FF7070", "red"), 
                         values = scales::rescale(c(-1, 0, 0.4, 0.9, 1.5), to = c(0, 1)),
                         limits = c(-1, 1.5),
                         oob = scales::squish
                         ) +  
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "DiscoTope-3.0 prediction")
  #print(plot_discotope)
  
  plot_discotope_RSA <- data_DiscoTope_subset %>% 
    ggplot(aes(x=resno, y=rsa, fill=rsa)) +
    geom_col(width = 1) +
    scale_fill_gradient2(low = "darkcyan", mid = "gray95", high = "deeppink4", midpoint = 0.5, limits = c(0, 1)) +  
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "Relative surface accessibility\n(Shrake-Rupley)")
  #print(plot_discotope)
  
  limit_lower <- data_DiscoTope_subset %>% pull(resno) %>% min(na.rm = TRUE)
  limit_upper <- data_DiscoTope_subset %>% pull(resno) %>% max(na.rm = TRUE)
  
  data_Epigraph_subset <- data_Epigraph %>% filter(PDB == pdb_entry)
  
  plot_epigraph <- data_Epigraph_subset %>% 
    ggplot(aes(x=resno, y=Score, fill=Score)) +
    geom_col(width = 1,) +
    scale_fill_gradient2(low = "blue", mid = "gray95", high = "red") +  
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "EpiGraph prediction")
  
  #Make the full figures
  figure <- plot_grid(
    plot_bepipred_2_RSA + scale_x_continuous(limits = c(limit_lower,limit_upper)) +
      theme(
        axis.ticks = element_blank(),  
        axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ) +
      labs(title = paste0(pdb_entry)), 
    
    plot_bepipred_2 + scale_x_continuous(limits = c(limit_lower,limit_upper)) +
      theme(
        axis.ticks = element_blank(),  
        axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ), 
    
    plot_bepipred_3 + scale_x_continuous(limits = c(limit_lower,limit_upper)) +
      theme(
        axis.ticks = element_blank(),  
        axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ),
    
    plot_discotope + scale_x_continuous(limits = c(limit_lower,limit_upper)) + 
      theme(
        axis.ticks = element_blank(),  
        axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ),
    
    plot_discotope_RSA + scale_x_continuous(limits = c(limit_lower,limit_upper)) + 
      theme(
        axis.ticks = element_blank(),  
        axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ),
    
    plot_epigraph + scale_x_continuous(limits = c(limit_lower,limit_upper)) + 
      theme(
        axis.ticks = element_blank(),  
        #axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        #axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        #axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ),
    
    align = "v", ncol = 1)
  
  print(figure)
  
  ggsave(filename = paste0("output/figures/tracks/", pdb_entry, ".png"), 
       plot = figure, 
       width = 14, height = 10, dpi = 300,
       bg="white")

}
```

## As heatmap
```{r}
toupper(AMINO_ACID_CODE)
```


```{r}
data_BepiPred_2_with_resno %>%
  select(Entry,resno,resid,EpitopeProbability,RelativeSurfaceAccessilibity) %>%
  rename("PDB" = Entry,
         "BepiPred-2.0 probability" = EpitopeProbability,
         "RSA NetsurfP" = RelativeSurfaceAccessilibity) %>%
  filter(resid %in% toupper(AMINO_ACID_CODE))
```

```{r}
data_BepiPred_3_with_resno %>%
  select(Accession, resno, resid, `BepiPred-3.0 score`,`BepiPred-3.0 linear epitope score`) %>%
  rename("PDB" = Accession) %>%
  filter(resid %in% toupper(AMINO_ACID_CODE))
```

```{r}
data_DiscoTope %>%
  select(pdb, chain, resno, residue, `DiscoTope-3.0_score`, calibrated_score, rsa) %>%
  distinct(pdb, resno, .keep_all = TRUE) %>%
  mutate(residue = aa123(residue)) %>%
  rename("PDB" = pdb,
         "resid" = residue,
         "RSA Shrake-Rupley" = rsa,
         "DiscoTope-3.0_calibrated_score" = calibrated_score)
```

```{r}
data_Epigraph %>%
  select(PDB, chain, resno, resid, Score) %>%
  distinct(PDB, resno, .keep_all = TRUE) %>%
  rename("Epigraph_score" = Score)
```

```{r}
data_predictions <- left_join(
  data_BepiPred_2_with_resno %>%
  select(Entry,resno,resid,EpitopeProbability,RelativeSurfaceAccessilibity) %>%
  rename("PDB" = Entry,
         "BepiPred-2.0 probability" = EpitopeProbability,
         "RSA NetsurfP" = RelativeSurfaceAccessilibity) %>%
  filter(resid %in% toupper(AMINO_ACID_CODE))
  ,
  
  data_BepiPred_3_with_resno %>%
    select(Accession, resno, resid, `BepiPred-3.0 score`,`BepiPred-3.0 linear epitope score`) %>%
    rename("PDB" = Accession) %>%
    filter(resid %in% toupper(AMINO_ACID_CODE))
) %>%
  left_join(
    data_DiscoTope %>%
      select(pdb, chain, resno, residue, `DiscoTope-3.0_score`, calibrated_score, rsa) %>%
      distinct(pdb, resno, .keep_all = TRUE) %>%
      mutate(residue = aa123(residue)) %>%
      rename("PDB" = pdb,
             "resid" = residue,
             "RSA Shrake-Rupley" = rsa,
             "DiscoTope-3.0_calibrated_score" = calibrated_score)
    
  ) %>%
  left_join(
    data_Epigraph %>%
      select(PDB, chain, resno, resid, Score) %>%
      distinct(PDB, resno, .keep_all = TRUE) %>%
      rename("Epigraph_score" = Score),
    by = c("PDB","resno","resid")
  )

data_predictions
```

```{r}
library(scales)

data_predictions_normalized <- data_predictions %>%
  group_by(PDB) %>%
  mutate(
    across(
      c(`BepiPred-3.0 score` , `BepiPred-3.0 linear epitope score`, 
        `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, Epigraph_score), 
      rescale
    )
  ) %>%
  ungroup() %>%
  na.omit()

data_predictions_normalized
```

```{r}
data_Epigraph %>%
      select(PDB, chain, resno, resid, Score) %>%
      distinct(PDB, resno, .keep_all = TRUE) %>%
      rename("Epigraph_score" = Score) %>%
  filter(PDB == "6YAG")
```

```{r}
data_predictions %>%
  filter(PDB == "6YAG")

data_predictions_normalized %>%
  filter(PDB == "6YAG")
```


```{r, fig.width = 14}
library(pheatmap)

data_test_wide <- data_predictions_normalized %>%
  filter(PDB == "6YAG") %>%
  select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
        `DiscoTope-3.0_calibrated_score`, Epigraph_score) %>%
  column_to_rownames("resno") %>%
  na.omit() %>%
  as.data.frame()

pheatmap(t(data_test_wide), 
         cluster_rows = TRUE,         
         cluster_cols = FALSE,         
         main = "6YAG", 
         color = colorRampPalette(c("blue", "white", "red"))(50),
         labels_col = "resno",
         clustering_distance_rows = "correlation") 
```

```{r, fig.width = 14}
for (pdb_entry in unique(data_predictions_normalized$PDB)) {
  data_predictions_normalised_subset_wide <- data_predictions_normalized %>%
  filter(PDB == pdb_entry) %>%
  distinct(resno, .keep_all = TRUE) %>%
  select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
        `DiscoTope-3.0_calibrated_score`, Epigraph_score) %>%
  column_to_rownames("resno") %>%
  na.omit() %>%
  as.data.frame()

  figure <- pheatmap(t(data_predictions_normalised_subset_wide), 
         cluster_rows = TRUE,         
         cluster_cols = FALSE,         
         main = paste0(pdb_entry), 
         color = colorRampPalette(c("blue", "white", "red"))(50),
         labels_col = "resno",
         clustering_distance_rows = "correlation") 
  
  print(figure)
  
  ggsave(filename = paste0("output/figures/heatmap/", pdb_entry, "_heatmap.png"), 
       plot = figure, 
       width = 14, height = 10, dpi = 300,
       bg="white")
}
```


## 3D
### bepipred-3.0
```{r}
pdbs_path = "output/pdb/"

for (pdb_entry in pdb_files_list) {
  file_path = paste0(pdbs_path,pdb_entry,".pdb")
  print(file_path)

  pdb_object <- read.pdb(pdb_entry,rm.alt=FALSE)
  
  data_BepiPred_3_with_resno_subset <- data_BepiPred_3_with_resno %>% filter(Accession == pdb_entry)
  min_score <- min(data_BepiPred_3_with_resno_subset$`BepiPred-3.0 score`, na.rm = TRUE)
  median_score <- median(data_BepiPred_3_with_resno_subset$`BepiPred-3.0 score`, na.rm = TRUE)
  max_score <- max(data_BepiPred_3_with_resno_subset$`BepiPred-3.0 score`, na.rm = TRUE)
  
  
  low_to_median <- colorRampPalette(c("blue", "gray95"))(10)
  median_to_high <- colorRampPalette(c("gray95", "red"))(10)
  
  breaks <- seq(min_score, max_score, length.out = 21)  # 20 intervals
  colors <- c(low_to_median, median_to_high)  # 20 colors
  
  color_map <- data.frame(
    ProbabilityBin = cut(seq(min_score, max_score, length.out = 20), 
                         breaks = breaks, include.lowest = TRUE),
    color = colors
  ) %>%
  rbind(c(NA,"black"))
  
  predictions_df <- data_BepiPred_3_with_resno_subset %>%
    mutate(ProbabilityBin = cut(`BepiPred-3.0 score`, 
                                breaks = breaks, include.lowest = TRUE)
    ) %>%
    group_by(ProbabilityBin) %>% 
    summarize(resno_list = list(paste0(resno,":A")), .groups = "drop") %>%
    mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " "))) %>%
    left_join(color_map)
  
  view <- prediction.viewer(pdb_entry, pdbs_path, predictions_df = predictions_df)
  
  #print(view)
  
  saveWidget(view, file = paste0("output/BepiPred3.0/snapshot/",pdb_entry,"_bepipred3.html"), title=pdb_entry, selfcontained = TRUE)
}
```


### discotope-3.0
```{r}
pdbs_path = "output/pdb/"

for (pdb_entry in pdb_files_list) {
  file_path = paste0(pdbs_path,pdb_entry,".pdb")
  print(file_path)

  pdb_object <- read.pdb(pdb_entry,rm.alt=FALSE)
  
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  min_score <- min(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  median_score <- median(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  max_score <- max(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  
  
  low_to_median <- colorRampPalette(c("blue", "gray95"))(10)
  median_to_high <- colorRampPalette(c("gray95", "red"))(10)
  
  breaks <- seq(min_score, max_score, length.out = 21)  # 20 intervals
  colors <- c(low_to_median, median_to_high)  # 20 colors
  
  color_map <- data.frame(
    ProbabilityBin = cut(seq(min_score, max_score, length.out = 20), 
                         breaks = breaks, include.lowest = TRUE),
    color = colors
  ) %>%
  rbind(c(NA,"black"))
  
  predictions_df <- data_DiscoTope_subset %>%
    mutate(ProbabilityBin = cut(calibrated_score, 
                                breaks = breaks, include.lowest = TRUE)
    ) %>%
    group_by(ProbabilityBin) %>% 
    summarize(resno_list = list(paste0(resno,":A")), .groups = "drop") %>%
    mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " "))) %>%
    left_join(color_map)
  
  view <- prediction.viewer(pdb_entry, pdbs_path, predictions_df = predictions_df)
  
  #print(view)
  
  saveWidget(view, file = paste0("output/DiscoTope/snapshot/",pdb_entry,"_discotope3.html"), title=pdb_entry, selfcontained = TRUE)
}
```

# Variability
## AmiA
```{,don't work }
install.packages("seqinr")
library("seqinr")
library("read.gb")

BiocManager::install("genbankr", dependencies=TRUE)
install.packages("genbankr", repos = "https://bioconductor.org/packages/3.17/bioc")
library("genbankr")

gb_data_AmiA <- read.gb("data/6. AmiA.gb")
```


```{r}
gb_lines_AmiA <- readLines("data/6. AmiA.gb")
head(gb_lines_AmiA, 20)
```

```{r}
origin_indices <- grep("^ORIGIN", gb_lines_AmiA)
end_indices <- grep("^//", gb_lines_AmiA)
header_prefix <- "AmiA"

fasta_content <- list()

for(i in 1:length(origin_indices)) {
  start_index <- origin_indices[i] + 2
  end_index <- end_indices[i] - 1
  
  protein_sequence_raw <- gb_lines_AmiA[start_index:end_index] 
  
  protein_sequence <- protein_sequence_raw %>%
    gsub("[0-9]+", "", .) %>%
    gsub("\\s+", "", .) %>%
    paste(collapse = "")
  
  header_line <- gb_lines_AmiA[grep("^TITLE", gb_lines_AmiA)[i]]
  header <- sub("^TITLE\\s+", "", header_line)

  fasta_content[[i]] <- paste0(">",header_prefix,"_", header, "\n", protein_sequence)
}

writeLines(unlist(fasta_content), "output/AmiA_sequences.fasta")
```

### multiple sequence alignment
```{r}
AmiA_sequences <- readAAStringSet("output/AmiA_sequences.fasta")
AmiA_sequences
```

```{r}
library("msa")
```


```{r}
AmiA_alignment <- msa(AmiA_sequences)
```

```{r}
AmiA_alignment
```


```{r}
print(AmiA_alignment, show="complete")
```

#### ggmsa
```{}
if (!requireNamespace("devtools", quietly=TRUE))
    install.packages("devtools")
devtools::install_github("YuLab-SMU/ggmsa")

library(ggmsa)
```

```{}
AmiA_sequences %>% 
  ggmsa()
```

#### ggseqlogo
```{r}
install.packages("ggseqlogo")  
library(ggseqlogo)
```

```{r, fig.width= 50}
#AmiA_alignment_matrix

seq_vector <- apply(AmiA_alignment_matrix, 1, paste, collapse ="")

names(seq_vector) <- paste0("Seq_", seq_along(seq_vector))

ggseqlogo(seq_vector, method = "bits") + 
  theme_minimal() +
  ggtitle("Sequence Logo of AmiA Alignment")
```


### alignment matrix
```{r}
AmiA_alignment_matrix <- as.matrix(AmiA_alignment)

AmiA_variability <- apply(AmiA_alignment_matrix, 2, function(col) {
  length(unique(col))
})

AmiA_variability

highly_variable_positions <- which(AmiA_variability == 2)

highly_variable_positions
length(highly_variable_positions)
```

```{r}
dim(AmiA_alignment_matrix)
```

```{}
#apply(AmiA_alignment_matrix, 2, function(col) {
#  names(sort(table(col), decreasing = TRUE))
#})

#apply(AmiA_alignment_matrix, 2, function(col) {
#  sort(table(col), decreasing = TRUE)
#})
```

### data_variability
```{r}
consensus_seq <- unlist(strsplit(msaConsensusSequence(AmiA_alignment),""))

variability_counts <- sapply(seq_len(ncol(AmiA_alignment_matrix)), function(i) {
  sum(AmiA_alignment_matrix[, i] != consensus_seq[i] & AmiA_alignment_matrix[, i] != "-")
})

#mutation_counts

total_sequences <- nrow(AmiA_alignment)

variability_frequencies <- variability_counts / total_sequences

data_variability_AmiA <- data.frame(
  Position = 1:length(variability_counts),
  Consensus_residue = consensus_seq,
  Variability_counts = variability_counts,
  Variability_frequency = variability_frequencies
)

data_variability_AmiA

```

```{r}
data_variability_AmiA %>%
  ggplot(aes(x=Position, y=Variability_counts)) +
  geom_col(width = 1) +
  theme_minimal()
```


### consensus matrix
```{r}
AmiA_consensus_matrix <- consensusMatrix(AmiA_alignment)
AmiA_consensus_matrix
```


```{r}
library("tinytex")

tinytex::tlmgr_install("pdflatex")
library("pdflatex")
```


```{r}
msaPrettyPrint(AmiA_alignment, output="asis", y=c(164, 213), showNames="none", showLogo="none", askForOverwrite=FALSE)
```

## (WIP) matching AmiA consensus to to epitope predictions
AmiA corresponds to pdb entries 8A42, 8QM0
```{r}
data.frame(Consensus_Residue = unlist(strsplit(msaConsensusSequence(AmiA_alignment), "")),
           Position = seq_along(unlist(strsplit(msaConsensusSequence(AmiA_alignment), ""))))
```


```{r}
pdb_object <- read.pdb("8A42")

required_adjustment <- map.seqposno.to.resno(sequence_AA1 = msaConsensusSequence(AmiA_alignment), pdb_object = pdb_object, verbose = TRUE)
required_adjustment
```

```{r}
pdb_object$atom %>%
  filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
  ggplot(aes(x=resno, y=chain, col = resid)) +
  geom_vline(xintercept=34) +
  geom_vline(xintercept=50) +
  geom_vline(xintercept=100) +
  geom_point() +
  geom_point(data = data_variability_AmiA ,
             aes(x=Position, col=aa123(Consensus_residue), y="FASTA")
             ) +
  scale_x_continuous(limits = c(NA,50)) +
  #scale_y_discrete(expand = c(0,0)) +
  theme_bw() 
```


#### mutations_data_with_resno
```{r}
data_variability_AmiA

pdb_entry <- "8A42"

pdb_object <- read.pdb(pdb_entry,rm.alt=FALSE)

required_adjustment <- map.seqposno.to.resno(sequence_AA1 = msaConsensusSequence(AmiA_alignment), pdb_object = pdb_object)

data_variability_AmiA_with_resno_8A42 <- create.resmap.bepipred(pdb_object$atom, 
                                 data_variability_AmiA,
                                 required_adjustment)

data_variability_AmiA_with_resno_8A42


```

```{r}
data_predictions_with_variability <- data_predictions %>%
  filter(PDB == "8A42") %>%
  left_join(data_variability_AmiA_with_resno_8A42 %>% select(resid, resno, Variability_counts, Variability_frequency),
            by = c("resno", "resid"))

data_predictions_with_variability
```

```{r}
names(data_predictions)
```


```{r, fig.width = 14}
data_predictions_with_variability %>%
  pivot_longer(cols = c(
    `BepiPred-2.0 probability`, `RSA NetsurfP`, `BepiPred-3.0 score`,
    `BepiPred-3.0 linear epitope score`, `DiscoTope-3.0_score`,
    `DiscoTope-3.0_calibrated_score`, `RSA Shrake-Rupley`, `Epigraph_score`, Variability_frequency
  ), names_to = "Score_Type", values_to = "Score_Value") %>%
  ggplot(aes(x=resno)) +
  geom_col(aes(y=Score_Value)) +
  
  geom_text(aes(label = Score_Type, x = (min(resno) + (max(resno) - min(resno))/2)), color = "black",y = Inf, vjust = 1.5, size = 4, family = "Arial", check_overlap = TRUE) +
  
  facet_wrap(~Score_Type, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(strip.text = element_blank())
```

```{r}
data_predictions_with_variability_normalized <- data_predictions_with_variability %>%
  group_by(PDB) %>%
  mutate(
    across(
      c(`BepiPred-3.0 score` , `BepiPred-3.0 linear epitope score`, 
        `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, Epigraph_score, Variability_frequency), 
      rescale
    )
  ) %>%
  ungroup() %>%
  na.omit()

data_predictions_with_variability_normalized
```

```{r}
data_test_wide <- data_predictions_with_variability_normalized %>%
  select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
        `DiscoTope-3.0_calibrated_score`, Epigraph_score, Variability_frequency) %>%
  column_to_rownames("resno") %>%
  na.omit() %>%
  as.data.frame()

pheatmap(t(data_test_wide), 
         cluster_rows = TRUE,         
         cluster_cols = FALSE,         
         main = "8A42", 
         color = colorRampPalette(c("blue", "white", "red"))(50),
         labels_col = "resno",
         clustering_distance_rows = "correlation") 
```

