---
title: "Streptococcus pneumoniae epitope prediction tools"
author: "Jan Paul van Meenen"
date: "`r Sys.Date()`"
output: html_document
---

#Packages
```{r}
source("requirements/packages.R")
```

#Functions
```{r}
function_files <- list.files(path = "requirements/functions", pattern = "\\.R$",full.names = TRUE)

for (function_file in function_files) {
  source(function_file)
}
```

# BepiPred-2.0
## split fasta into parts
BepiPred-2.0 only allows for 50 proteins at a time.
```{r}
fasta_lines <- readLines("data/fasta/pdb_protein_sequences.fasta")

# Group lines into proteins based on `>...` headers
protein_entries <- split(fasta_lines, cumsum(grepl("^>", fasta_lines)))

# Split protein entries into chunks of 50
protein_chunks <- split(protein_entries, ceiling(seq_along(protein_entries) / 50))

for (i in seq_along(protein_chunks)) {
  # Combine lines for each chunk
  chunk_lines <- unlist(protein_chunks[[i]])
  
  # Write the chunk to a new file
  writeLines(chunk_lines, paste0("data/fasta/pdb_protein_sequences_part_", i, ".fasta"))
}

```

https://services.healthtech.dtu.dk/services/BepiPred-2.0/ 
```{r}
data_BepiPred_2 <- rbind(
    read_csv("output/BepiPred2.0/BepiPred_part1_summary_67C1B456000FBFBBD072AA59.csv"),
    read_csv("output/BepiPred2.0/BepiPred_part2_summary_67C1B4C8000FC0D59FA5D3FB.csv"),
    read_csv("output/BepiPred2.0/BepiPred_part3_summary_681B08A3001AD0BAAA45311C.csv")
    ) 
```

## Generalized
### get valid all pdb entires and files
```{r}
pdbs_path = "data/pdb/"

pdb_files_list <- list.files(pdbs_path, pattern = "\\.pdb$") %>% sub("\\.pdb$", "", .)

pdb_files_list
```

## BepiPred-2.0 with resno
```{r}
data_BepiPred_2_with_resno <- NULL

for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  required_adjustment <- map.seqresno.to.resno(pdb_object)
  #print(required_adjustment)
  
  resmap <- create.resmap(pdb_object$atom, 
                                           data_BepiPred_2 %>% filter(Entry == pdb_entry),
                                           required_adjustment)
  
  if (is.null(data_BepiPred_2_with_resno)) {
    data_BepiPred_2_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_BepiPred_2_with_resno <- rbind(data_BepiPred_2_with_resno, resmap)
  }
}

data_BepiPred_2_with_resno
```

## Store results as csv
```{r}
write.csv(data_BepiPred_2_with_resno,
          "output/BepiPred2.0/data_BepiPred2.csv", 
          row.names = FALSE)
```



# BepiPred-3.0 results
https://services.healthtech.dtu.dk/services/BepiPred-3.0/

```{r}
data_BepiPred_3 <- rbind(
    read_csv("output/BepiPred3.0/raw_output_part1_0228.csv"),
    read_csv("output/BepiPred3.0/raw_output_part2_0228.csv"),
    read_csv("output/BepiPred3.0/raw_output_part3.csv")
    ) %>%
  #mutate(Residue = aa123(Residue)) %>%
  group_by(Accession) %>%
  mutate(Position = row_number())
```

## BepiPred-3.0 with resno
```{r}
data_BepiPred_3_with_resno <- NULL

for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  required_adjustment <- map.seqresno.to.resno(pdb_object)
  #print(required_adjustment)
  
  resmap <- create.resmap(pdb_object$atom, 
                                           data_BepiPred_3 %>% filter(Accession == pdb_entry),
                                           required_adjustment)
  
  if (is.null(data_BepiPred_3_with_resno)) {
    data_BepiPred_3_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_BepiPred_3_with_resno <- rbind(data_BepiPred_3_with_resno, resmap)
  }
}

data_BepiPred_3_with_resno
```

## Store results as csv
```{r}
write.csv(data_BepiPred_3_with_resno,
          "output/BepiPred3.0/data_BepiPred3.csv", 
          row.names = FALSE)
```

# DiscoTope-3.0
https://services.healthtech.dtu.dk/services/DiscoTope-3.0/
```{r}
csv_files <- list.files("output/DiscoTope/individual_files/", pattern = "*.csv", full.names = TRUE)

data_DiscoTope <- NULL

for (csv_file in csv_files) {
  print(csv_file)
  
  pdb_id <- str_extract(basename(csv_file), "^[^_]+") 
  
  temp_data <- read_csv(csv_file, show_col_types = FALSE)
  
  temp_data <- temp_data %>%
    mutate(pdb = pdb_id,
           resno = res_id)

  if (is.null(data_DiscoTope)) {
    data_DiscoTope <- temp_data
  } else {
    data_DiscoTope <- rbind(data_DiscoTope, temp_data)
  }
}

data_DiscoTope
```

```{r}
data_DiscoTope %>% pull(pdb) %>% unique()
```
## Store results as csv
```{r}
write.csv(data_DiscoTope,
          "output/DiscoTope/data_DiscoTope.csv", 
          row.names = FALSE)
```

# Epigraph
http://epigraph.kaist.ac.kr/
```{r}
epigraph_zip_files <- list.files(path = "output/EpiGraph/individual_files/", pattern = "\\.zip$")

for (zip_file in epigraph_zip_files) {
  unzip(zipfile = paste0("output/EpiGraph/individual_files/",zip_file), exdir = "output/EpiGraph/individual_files/")
}

epigraph_csv_files <- list.files(path = "output/EpiGraph/individual_files/", pattern = "\\.csv$")
epigraph_csv_files
```

```{r}
data_Epigraph <- NULL

for (csv_file in epigraph_csv_files) {
  temp_data <- read_csv(paste0("output/EpiGraph/individual_files/",csv_file), show_col_types = FALSE) 
  print(csv_file)
  
  if (is.null(data_Epigraph)) {
    data_Epigraph <- temp_data
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_Epigraph <- rbind(data_Epigraph, temp_data)
  }
}

data_Epigraph <- data_Epigraph %>% 
  separate(Residue, into = c("chain", "resid", "resno"), sep = ":", convert = TRUE) %>%
  mutate(PDB = toupper(PDB))

data_Epigraph
```

```{r}
data_Epigraph %>% pull(PDB) %>% unique()
```


```{r}
data_Epigraph_subset <- data_Epigraph %>% filter(PDB == "2BIB")

plot_epigraph <- data_Epigraph_subset %>% 
    ggplot(aes(x=resno, y=Score, fill=Score)) +
    geom_col(width = 1,) +
    scale_fill_gradient2(low = "blue", mid = "gray95", high = "red") +  
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "EpiGraph prediction")

plot_epigraph
```

## Store results as csv
```{r}
write.csv(data_Epigraph,
          "output/EpiGraph/data_Epigraph.csv", 
          row.names=FALSE)
```

#(WIP) GraphBepi
Lookup the longest chain per protein
```{r}
data_DiscoTope %>%
  filter(residue %in% names(AMINO_ACID_CODE)[1:20]) %>%
  select(pdb, chain, length) %>%
  pivot_wider(names_from = chain, values_from = length) %>%
  mutate(across(where(is.list), ~sapply(., function(x) ifelse(is.null(x), NA, length(x)))))
```

Visually check all chains per protein
```{r}
for (pdb_entry in unique(data_DiscoTope$pdb)) {
  print(pdb_entry)
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  
  plot <- data_DiscoTope_subset %>%
    ggplot(aes(x=res_id, colour = residue)) +
    geom_point(aes(y=0)) +
    facet_wrap(~chain,ncol = 1) +
    labs(title=paste0(pdb_entry)) +
    theme_minimal()
  
  
  print(plot)
}
```

## read in the results
```{r}
graphbepi_txt_files <- list.files(path = "output/GraphBepi/individual_files/", pattern = "\\.txt$")

graphbepi_txt_files
```

```{r}
data_GraphBepi <- NULL

for (txt_file in graphbepi_txt_files) {
  print(txt_file)
  PDB <- sub("_.*", "", txt_file)
  chain <- sub("^[^_]+_([A-Za-z0-9])_.*", "\\1", txt_file)

  temp_data <- read.table(paste0("output/GraphBepi/individual_files/",txt_file), header = TRUE, skip=1)
  
  temp_data <- temp_data %>% 
    mutate(
      PDB = PDB,
      chain = chain,
      resid = aa123(AA),
      `GraphBepi probability` = Prob,
      `GraphBepi prediction` = Pred,
     Position = row_number()
    ) %>%
    select(PDB, chain, resid, Position, `GraphBepi probability`, `GraphBepi prediction`)
  
  if (is.null(data_GraphBepi)) {
    data_GraphBepi <- temp_data
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_GraphBepi <- rbind(data_GraphBepi, temp_data)
  }
}

data_GraphBepi
```

```{r}
data_GraphBepi_with_resno <- NULL

for (pdb_entry in unique(data_GraphBepi$PDB)) {
  print(pdb_entry)
  
  pdb_file <- paste0(pdbs_path,pdb_entry,".pdb") 
  
  if (file.exists(pdb_file)) {
    print(pdb_file)
    pdb_object <- read.pdb(file = pdb_file, rm.alt = FALSE)
  } else {
    pdb_object <- read.pdb(pdb_entry, rm.alt = FALSE)
  }
  
  graphbepi_subset <- data_GraphBepi %>% filter(PDB == pdb_entry)
  graphbepi_sequence <- graphbepi_subset %>% pull(resid) %>% aa321() %>% str_c(collapse = "")
  graphbepi_chain <- graphbepi_subset %>% pull(chain) %>% unique()
  
  required_adjustment <- map.seqposno.to.resno(sequence_AA1 = graphbepi_sequence, pdb_object)
  print(required_adjustment)
  
  resmap <- create.resmap(pdb_object$atom %>% filter(chain == graphbepi_chain), 
                                           graphbepi_subset,
                                           required_adjustment, join_by = c("resno","chain","resid")
                                   )
  
  if (is.null(data_GraphBepi_with_resno)) {
    data_GraphBepi_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_GraphBepi_with_resno <- rbind(data_GraphBepi_with_resno, resmap)
  }
}

data_GraphBepi_with_resno
```

# InterPro
https://www.ebi.ac.uk/interpro/search/sequence/
InterPro gives domain/topology prediction of proteins
```{r}
interpro_files <- list.files(path = "output/InterPro/", pattern = "\\.tsv$", full.names = TRUE)

interpro_files

interpro_colnames <- c(
  "Protein_accession", "MD5", "Sequence_length", "Analysis",
  "Signature_accession", "Signature_description", "Start", "End",
  "Score", "Status", "Date", "InterPro_accession", "InterPro_description",
  "GO_annotations", "Pathways"
)

interpro_coltypes <- cols(
  Protein_accession = col_character(),
  MD5 = col_character(),
  Sequence_length = col_integer(),
  Analysis = col_character(),
  Signature_accession = col_character(),
  Signature_description = col_character(),
  Start = col_integer(),
  End = col_integer(),
  Score = col_double(),  
  Status = col_logical(),
  Date = col_date(format = "%d-%m-%Y"),
  InterPro_accession = col_character(),
  InterPro_description = col_character(),
  GO_annotations = col_character(),
  Pathways = col_character()
)

# Load the file with correct column types and headers
data_interpro <- read_tsv(
  interpro_files,
  col_names = interpro_colnames,
  col_types = interpro_coltypes
) %>%
  setNames(interpro_colnames)
```

```{r}
data_interpro %>% filter(Protein_accession == "4QQQ") %>%
  select(Analysis, Signature_description, InterPro_accession, Start, End, Score) %>%
  arrange(-log10(Score))
```

```{r}
data_interpro %>% 
  #filter(!(Signature_description == "-" & InterPro_description == "-")) %>% 
  pull(Analysis) %>%
  table() %>%
  as.data.frame() %>%
  arrange(desc(Freq))

data_interpro %>% 
  filter(!(Signature_description == "-" & InterPro_description == "-")) %>% 
  pull(Analysis) %>%
  table() %>%
  as.data.frame() %>%
  arrange(desc(Freq))
```

```{r}
data_interpro %>%
  filter(Analysis != "-") %>%  # Exclude empty entries
  mutate(present = TRUE) %>%   # Mark presence explicitly
  distinct(Protein_accession, Analysis, present, Score) %>%  # Avoid duplicates
  pivot_wider(names_from = Analysis, values_from = Score, values_fill = NA)
```



```{r, fig.width=14}
data_interpro %>%
  filter(Protein_accession == "2BIB") %>%
  mutate(id = row_number()) %>%
  ggplot() +
  geom_segment(
    aes(
      x = Start, xend = End,
      y = Signature_description, yend = Signature_description,
      color = InterPro_description,
    ),
    size=8
  ) +
  theme_bw() +
  labs(
    x = "Amino acid position",
    y = "Signature description"
  )

data_interpro %>%
  filter(Protein_accession == "2BIB") %>%
  mutate(id = row_number()) %>%
  ggplot() +
  geom_segment(
    aes(
      x = Start, xend = End,
      y = InterPro_description, yend = InterPro_description,
      color = Signature_description,
    ),
    size=8
  ) +
  theme_bw() +
  labs(
    x = "Amino acid position",
    y = "InterPro description"
  )

data_interpro %>%
  filter(Protein_accession == "2BIB") %>%
  mutate(id = row_number()) %>%
  ggplot() +
  geom_segment(
    aes(
      x = Start, xend = End,
      y = Analysis, yend = Analysis,
      color = Signature_description,
    ),
    size=8
  ) +
  theme_bw() +
  labs(
    x = "Amino acid position",
    y = "Analysis method"
  )

data_interpro %>%
  filter(Protein_accession == "2BIB") %>%
  mutate(id = row_number()) %>%
  ggplot() +
  geom_segment(
    aes(
      x = Start, xend = End,
      y = Analysis, yend = Analysis,
      color = InterPro_description,
    ),
    size=8
  ) +
  theme_bw() +
  labs(
    x = "Amino acid position",
    y = "Analysis method"
  )

data_interpro %>%
  filter(Protein_accession == "2BIB") %>%
  mutate(id = row_number(),
         Score = as.numeric(Score)) %>%
  ggplot() +
  geom_segment(
    aes(
      x = Start, xend = End,
      y = -log10(Score), yend = -log10(Score),
      color = interaction(Analysis,InterPro_description),
    ),
    size=8
  ) +
  theme_bw() +
  labs(
    x = "Amino acid position",
    y = "Significance"
  )
```

```{r, fig.width=14}
data_interpro %>%
  filter(Analysis == "SUPERFAMILY") %>%
  arrange(Protein_accession) %>%
  ggplot() +
  geom_segment(
    aes(
      x = Start, xend = End,
      y = Protein_accession, yend = Protein_accession,
      color = Signature_description,
    ),
    size=4
  ) +
  theme_bw() +
  labs(
    x = "Amino acid position",
    y = "Analysis method"
  )

data_interpro %>%
  arrange(Protein_accession) %>%
  ggplot() +
  geom_segment(
    aes(
      x = Start, xend = End,
      y = Protein_accession, yend = Protein_accession,
      color = InterPro_description,
    ),
    size=4
  ) +
  theme_bw() +
  labs(
    x = "Amino acid position",
    y = "Analysis method"
  )
```

```{r fig.width=14}
for (protein in sort(unique(data_interpro$Protein_accession))) {
  print(protein)
  
  data_interpro_subset <- data_interpro %>%
    filter(Protein_accession == protein) %>%
    pivot_longer(
      cols = c(Signature_description, InterPro_description),
      names_to = "Description_type",
      values_to = "Description"
    )
  
  plot <- data_interpro_subset %>%
  ggplot() +
  geom_segment(
    aes(
      x = Start, xend = End,
      y = Analysis, yend = Analysis,
      color = Description,
    ),
    size=8
  ) +
  facet_wrap(~ Description_type) +
  theme_bw() +
  labs(
    title = paste0("Domains in ", protein),
    x = "Amino acid position",
    y = "Analysis method"
  )
  
  print(plot)
}
```

```{r, fig.width=14}
for (protein in sort(unique(data_interpro$Protein_accession))) {
  
  data_interpro_subset <- data_interpro %>%
    filter(Protein_accession == protein) %>%
    pivot_longer(
      cols = c(Signature_description, InterPro_description),
      names_to = "Description_type",
      values_to = "Description"
    ) %>%
    filter(!is.na(Description) & Description != "-") %>%
    mutate(
      combination = interaction(Description_type, Analysis)
    )

  # Use numeric y_pos for small vertical shifts
  plot <- ggplot(data_interpro_subset) +
    geom_segment(
      aes(x = Start, xend = End,
          y = -log10(Score), yend = -log10(Score),
          group = combination,
          color = Description),
      size = 5
    ) +
    #scale_y_continuous(
    #  breaks = unique(as.numeric(factor(data_interpro_subset$Analysis))),
    #  labels = levels(factor(data_interpro_subset$Analysis))
    #) +
    labs(
      title = paste0("Protein: ", protein),
      x = "Amino acid position",
      y = "E-value"
    ) +
    theme_minimal() 

  print(plot)
}
```

```{r, fig.width=14, warning=FALSE}
for (protein in sort(unique(data_interpro$Protein_accession))) {
  print(protein)
  
  data_interpro_subset <- data_interpro %>%
    filter(Protein_accession == protein,
           !is.na(Score)) %>%
    mutate(e_value = -log10(Score),
           Description = case_when(
             InterPro_description != "-" ~ InterPro_description,
             TRUE ~ Signature_description
             )
           ) %>%
    # Calculate the max e_value for each Analysis group and reorder Analysis factor
    group_by(Analysis) %>%
    mutate(max_e_value = max(e_value, na.rm = TRUE)) %>%
    ungroup() %>%
    mutate(Analysis = factor(Analysis, levels = unique(Analysis[order(max_e_value)]))) %>%
    select(-max_e_value)  # Remove max_e_value column
  
  plot <- data_interpro_subset %>%
    filter(!is.na(Description) & Description != "-") %>%
    ggplot() +
    geom_segment(
      aes(
        x = Start, xend = End,
        y = Analysis, yend = Analysis,
        color = Description,
        alpha = e_value
      ),
      size = 8
    ) +
    scale_alpha_continuous() +
    theme_bw() +
    labs(
      title = paste0("Domains in ", protein),
      x = "Amino acid position",
      y = "Analysis"
    )
  
  print(plot)
}

```


## cleaned
We want to build up annotation, using the most significant (based on E-value) non-overlapping domain hits

```{r}
data_interpro %>%
  #filter(Protein_accession == "2BIB") %>%
  group_by(Protein_accession) %>%
  mutate(Description = case_when(
             InterPro_description != "-" ~ InterPro_description,
             !is.na(InterPro_description) ~ InterPro_description,
             TRUE ~ Signature_description
             )
         ) %>%
  ungroup() %>%
  filter(!is.na(Description), Description != "-") %>%
  arrange(Protein_accession, Score, Description) 
```

```{r}
keep_non_overlapping <- function(df) {
  df <- df %>%
    filter(!is.na(Score),
           Score != 0) %>%
    mutate(e_value = -log10(Score)) %>%
    arrange(desc(e_value))  # Most significant first
  
  kept <- list()
  occupied_ranges <- IRanges::IRanges()  # Efficient range management
  
  for (i in seq_len(nrow(df))) {
    this_range <- IRanges::IRanges(start = df$Start[i], end = df$End[i])
    
    # Check for overlap with any previously kept range
    if (length(occupied_ranges) == 0 || length(IRanges::findOverlaps(this_range, occupied_ranges)) == 0) {
      occupied_ranges <- c(occupied_ranges, this_range)
      kept[[length(kept) + 1]] <- df[i, ]
    }
  }
  
  bind_rows(kept)
}
```

```{r}
data_interpro %>%
  filter(Protein_accession == "2BIB",
         !is.na(Score)) %>%
  mutate(e_value = -log10(Score)) %>%
  arrange(desc(e_value))
```


```{r}
filtered_hits <- data_interpro %>%
  filter(Protein_accession == "3E0M") %>%
  group_by(Protein_accession) %>%
  mutate(Description = case_when(
             InterPro_description != "-" ~ InterPro_description,
             !is.na(InterPro_description) ~ InterPro_description,
             TRUE ~ Signature_description
             )
         ) %>%
  ungroup() %>%
  filter(!is.na(Description), Description != "-") %>%
  arrange(Protein_accession, Score, Description) %>%
  group_by(Protein_accession) %>%
  group_modify(~keep_non_overlapping(.x)) %>%
  ungroup()

filtered_hits
```

```{r}
data_interpro_filtered <- data_interpro %>%
  group_by(Protein_accession) %>%
  mutate(Description = case_when(
             InterPro_description == "-" ~ Signature_description,
             is.na(InterPro_description) ~ Signature_description,
             TRUE ~ InterPro_description
             )
         ) %>%
  filter(!is.na(Description), Description != "-") %>%
  arrange(Protein_accession, Score) %>%
  group_modify(~ keep_non_overlapping(.x)) %>%
  select(Protein_accession, Sequence_length, Analysis, )
  ungroup()

data_interpro_filtered
```

