---
title: "Streptococcus pneumoniae epitope prediction tools"
author: "Jan Paul van Meenen"
date: "`r Sys.Date()`"
output: html_document
---

#Packages
```{r}
source("requirements/packages.R")
```

#Functions
```{r}
function_files <- list.files(path = "requirements/functions", pattern = "\\.R$",full.names = TRUE)

for (function_file in function_files) {
  source(function_file)
}
```

# BepiPred-2.0
## split fasta into parts
BepiPred-2.0 only allows for 50 proteins at a time.
```{r}
fasta_lines <- readLines("data/fasta/pdb_protein_sequences.fasta")

# Group lines into proteins based on `>...` headers
protein_entries <- split(fasta_lines, cumsum(grepl("^>", fasta_lines)))

# Split protein entries into chunks of 50
protein_chunks <- split(protein_entries, ceiling(seq_along(protein_entries) / 50))

for (i in seq_along(protein_chunks)) {
  # Combine lines for each chunk
  chunk_lines <- unlist(protein_chunks[[i]])
  
  # Write the chunk to a new file
  writeLines(chunk_lines, paste0("data/fasta/pdb_protein_sequences_part_", i, ".fasta"))
}

```

https://services.healthtech.dtu.dk/services/BepiPred-2.0/ 
```{r}
data_BepiPred_2 <- rbind(
    read_csv("output/BepiPred2.0/BepiPred_part1_summary_67C1B456000FBFBBD072AA59.csv"),
    read_csv("output/BepiPred2.0/BepiPred_part2_summary_67C1B4C8000FC0D59FA5D3FB.csv")
    ) 
```

## Generalized
### get valid all pdb entires and files
```{r}
pdbs_path = "data/pdb/"

pdb_files_list <- list.files(pdbs_path, pattern = "\\.pdb$") %>% sub("\\.pdb$", "", .)

pdb_files_list
```

## BepiPred-2.0 with resno
```{r}
data_BepiPred_2_with_resno <- NULL

for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  required_adjustment <- map.seqresno.to.resno(pdb_object)
  #print(required_adjustment)
  
  resmap <- create.resmap(pdb_object$atom, 
                                           data_BepiPred_2 %>% filter(Entry == pdb_entry),
                                           required_adjustment)
  
  if (is.null(data_BepiPred_2_with_resno)) {
    data_BepiPred_2_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_BepiPred_2_with_resno <- rbind(data_BepiPred_2_with_resno, resmap)
  }
}

data_BepiPred_2_with_resno
```

## Store results as csv
```{r}
write.csv(data_BepiPred_2_with_resno,
          "output/BepiPred2.0/data_BepiPred2.csv", 
          row.names = FALSE)
```



# BepiPred-3.0 results
https://services.healthtech.dtu.dk/services/BepiPred-3.0/
```{r}
data_BepiPred_3 <- rbind(
    read_csv("output/BepiPred3.0/raw_output_part1_0228.csv"),
    read_csv("output/BepiPred3.0/raw_output_part2_0228.csv")
    ) %>%
  #mutate(Residue = aa123(Residue)) %>%
  group_by(Accession) %>%
  mutate(Position = row_number())
```

## BepiPred-3.0 with resno
```{r}
data_BepiPred_3_with_resno <- NULL

for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  required_adjustment <- map.seqresno.to.resno(pdb_object)
  #print(required_adjustment)
  
  resmap <- create.resmap(pdb_object$atom, 
                                           data_BepiPred_3 %>% filter(Accession == pdb_entry),
                                           required_adjustment)
  
  if (is.null(data_BepiPred_3_with_resno)) {
    data_BepiPred_3_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_BepiPred_3_with_resno <- rbind(data_BepiPred_3_with_resno, resmap)
  }
}

data_BepiPred_3_with_resno
```

## Store results as csv
```{r}
write.csv(data_BepiPred_3_with_resno,
          "output/BepiPred3.0/data_BepiPred3.csv", 
          row.names = FALSE)
```

# DiscoTope-3.0
https://services.healthtech.dtu.dk/services/DiscoTope-3.0/
```{r}
csv_files <- list.files("output/DiscoTope", pattern = "*.csv", full.names = TRUE)

data_DiscoTope <- NULL

for (csv_file in csv_files) {
  temp_data <- read_csv(csv_file, show_col_types = FALSE) 
  
  if (is.null(data_DiscoTope)) {
    data_DiscoTope <- temp_data
  } else {
    data_DiscoTope <- rbind(data_DiscoTope, temp_data)
  }
}

data_DiscoTope <- data_DiscoTope %>%
  mutate(pdb = gsub("_.*", "", pdb),
         resno = res_id)

data_DiscoTope
```

## Store results as csv
```{r}
write.csv(data_DiscoTope,
          "output/DiscoTope/data_DiscoTope.csv", 
          row.names = FALSE)
```

# Epigraph
http://epigraph.kaist.ac.kr/
```{r}
epigraph_zip_files <- list.files(path = "output/EpiGraph/individual_files/", pattern = "\\.zip$")

for (zip_file in epigraph_zip_files) {
  unzip(zipfile = paste0("output/EpiGraph/individual_files/",zip_file), exdir = "output/EpiGraph/individual_files/")
}

epigraph_csv_files <- list.files(path = "output/EpiGraph/individual_files/", pattern = "\\.csv$")
epigraph_csv_files
```

```{r}
data_Epigraph <- NULL

for (csv_file in epigraph_csv_files) {
  temp_data <- read_csv(paste0("output/EpiGraph/individual_files/",csv_file), show_col_types = FALSE) 
  print(csv_file)
  
  if (is.null(data_Epigraph)) {
    data_Epigraph <- temp_data
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_Epigraph <- rbind(data_Epigraph, temp_data)
  }
}

data_Epigraph <- data_Epigraph %>% 
  separate(Residue, into = c("chain", "resid", "resno"), sep = ":", convert = TRUE) %>%
  mutate(PDB = toupper(PDB))

data_Epigraph
```

```{r}
data_Epigraph_subset <- data_Epigraph %>% filter(PDB == "3ZPP")

plot_epigraph <- data_Epigraph_subset %>% 
    ggplot(aes(x=resno, y=Score, fill=Score)) +
    geom_col(width = 1,) +
    scale_fill_gradient2(low = "blue", mid = "gray95", high = "red") +  
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "EpiGraph prediction")

plot_epigraph
```

## Store results as csv
```{r}
write.csv(data_Epigraph,
          "output/EpiGraph/data_Epigraph.csv", 
          row.names=FALSE)
```

#(WIP) GraphBepi
Lookup the longest chain per protein
```{r}
data_DiscoTope %>%
  filter(residue %in% names(AMINO_ACID_CODE)[1:20]) %>%
  select(pdb, chain, length) %>%
  pivot_wider(names_from = chain, values_from = length) %>%
  mutate(across(where(is.list), ~sapply(., function(x) ifelse(is.null(x), NA, length(x)))))
```

Visually check all chains per protein
```{r}
for (pdb_entry in unique(data_DiscoTope$pdb)) {
  print(pdb_entry)
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  
  plot <- data_DiscoTope_subset %>%
    ggplot(aes(x=res_id, colour = residue)) +
    geom_point(aes(y=0)) +
    facet_wrap(~chain,ncol = 1) +
    labs(title=paste0(pdb_entry)) +
    theme_minimal()
  
  
  print(plot)
}
```

## read in the results
```{r}
graphbepi_txt_files <- list.files(path = "output/GraphBepi/individual_files/", pattern = "\\.txt$")

graphbepi_txt_files
```

```{r}
data_GraphBepi <- NULL

for (txt_file in graphbepi_txt_files) {
  print(txt_file)
  PDB <- sub("_.*", "", txt_file)
  chain <- sub("^[^_]+_([A-Za-z0-9])_.*", "\\1", txt_file)

  temp_data <- read.table(paste0("output/GraphBepi/individual_files/",txt_file), header = TRUE, skip=1)
  
  temp_data <- temp_data %>% 
    mutate(
      PDB = PDB,
      chain = chain,
      resid = aa123(AA),
      `GraphBepi probability` = Prob,
      `GraphBepi prediction` = Pred,
     Position = row_number()
    ) %>%
    select(PDB, chain, resid, Position, `GraphBepi probability`, `GraphBepi prediction`)
  
  if (is.null(data_GraphBepi)) {
    data_GraphBepi <- temp_data
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_GraphBepi <- rbind(data_GraphBepi, temp_data)
  }
}

data_GraphBepi
```

```{r}
data_GraphBepi_with_resno <- NULL

for (pdb_entry in unique(data_GraphBepi$PDB)) {
  print(pdb_entry)
  
  pdb_file <- paste0(pdbs_path,pdb_entry,".pdb") 
  
  if (file.exists(pdb_file)) {
    print(pdb_file)
    pdb_object <- read.pdb(file = pdb_file, rm.alt = FALSE)
  } else {
    pdb_object <- read.pdb(pdb_entry, rm.alt = FALSE)
  }
  
  graphbepi_subset <- data_GraphBepi %>% filter(PDB == pdb_entry)
  graphbepi_sequence <- graphbepi_subset %>% pull(resid) %>% aa321() %>% str_c(collapse = "")
  graphbepi_chain <- graphbepi_subset %>% pull(chain) %>% unique()
  
  required_adjustment <- map.seqposno.to.resno(sequence_AA1 = graphbepi_sequence, pdb_object)
  print(required_adjustment)
  
  resmap <- create.resmap(pdb_object$atom %>% filter(chain == graphbepi_chain), 
                                           graphbepi_subset,
                                           required_adjustment, join_by = c("resno","chain","resid")
                                   )
  
  if (is.null(data_GraphBepi_with_resno)) {
    data_GraphBepi_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_GraphBepi_with_resno <- rbind(data_GraphBepi_with_resno, resmap)
  }
}

data_GraphBepi_with_resno
```
