---
title: "Streptococcus pneumoniae epitope prediction tools"
author: "Jan Paul van Meenen"
date: "`r Sys.Date()`"
output: html_document
---

#Packages
```{r}
source("requirements/packages.R")
```

#Functions
```{r}
function_files <- list.files(path = "requirements/functions", pattern = "\\.R$",full.names = TRUE)

for (function_file in function_files) {
  source(function_file)
}
```

# BepiPred-2.0
## split fasta into parts
BepiPred-2.0 only allows for 50 proteins at a time.
```{r}
fasta_lines <- readLines("data/fasta/pdb_protein_sequences.fasta")

# Group lines into proteins based on `>...` headers
protein_entries <- split(fasta_lines, cumsum(grepl("^>", fasta_lines)))

# Split protein entries into chunks of 50
protein_chunks <- split(protein_entries, ceiling(seq_along(protein_entries) / 50))

for (i in seq_along(protein_chunks)) {
  # Combine lines for each chunk
  chunk_lines <- unlist(protein_chunks[[i]])
  
  # Write the chunk to a new file
  writeLines(chunk_lines, paste0("data/fasta/pdb_protein_sequences_part_", i, ".fasta"))
}

```

https://services.healthtech.dtu.dk/services/BepiPred-2.0/ 
```{r}
data_BepiPred_2_raw <- rbind(
    read_csv("output/BepiPred2.0/BepiPred_part1_summary_684033E6001F9BFB3C151E32.csv"),
    read_csv("output/BepiPred2.0/BepiPred_part2_summary_684034B0001F9E18B0E5074D.csv"),
    read_csv("output/BepiPred2.0/BepiPred_vws_summary_68358292003BC30E33A4DFA1.csv")
    ) 

data_BepiPred_2 <- data_BepiPred_2_raw %>%
  mutate(
    PDB = case_when(
      Entry %in% data_namesclasspdb$PDB ~ Entry,
      TRUE ~ NA
    ),
    protein = case_when(
     !is.na(PDB) ~ 
        data_namesclasspdb$protein[match(PDB, data_namesclasspdb$PDB)],
      !is.na(match(tolower(Entry), tolower(data_namesclasspdb$protein))) ~
        data_namesclasspdb$protein[match(tolower(Entry), tolower(data_namesclasspdb$protein))],
      !is.na(match(tolower(Entry), tolower(data_namesclasspdb$TIGR4_name))) ~
       data_namesclasspdb$protein[match(tolower(Entry), tolower(data_namesclasspdb$TIGR4_name))],
      TRUE ~ data_namesclasspdb$TIGR4_name[match(tolower(Entry), tolower(data_namesclasspdb$TIGR4_name))]
    )
  ) %>%
  left_join(
    data_namesclasspdb, by = c("protein","PDB")
  )
data_BepiPred_2
```


## Generalized
### get valid all pdb entires and files
```{r}
pdb_files_list <- list.files(pdbs_path, pattern = "\\.pdb$") %>% sub("\\.pdb$", "", .)

AF3_pdb_files_list <- list.files(AF3_pdbs_path, pattern = "\\.pdb$") %>% sub("\\.pdb$", "", .)
```

## BepiPred-2.0 with resno
```{r}
data_BepiPred_2_with_resno <- NULL

for (i in seq_len(nrow(data_namesclasspdb))) {
  protein_indexed <- data_namesclasspdb$protein[i]
  pdb_entry <- data_namesclasspdb$PDB[i]
  TIGR4_entry <- data_namesclasspdb$TIGR4_name[i]
  
  if (!is.na(pdb_entry)) {
    
    print(paste("Processing", protein_indexed, "-", pdb_entry))
    
    pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
    
    data_BepiPred_2_subset <- data_BepiPred_2 %>% 
      filter(Entry == pdb_entry)
    
  } else {
    match_idx <- grep(paste0("^", tolower(protein_indexed), "_model_\\d+_fold_\\d+"), tolower(AF3_pdb_files_list))
    
    if (length(match_idx) == 0) {
      warning(paste("No model found for protein: ", protein_indexed, "retrying using TIGR4 name \n"))
      
      match_idx <- grep(paste0("^", tolower(TIGR4_entry), "_model_\\d+_fold_\\d+"), tolower(AF3_pdb_files_list))
      if (length(match_idx) == 0) {
        warning(paste("No model found for protein: ", protein_indexed, "or TIGR4 name: ", TIGR4_entry, "\n"))
        next
      }
    }
    
    print(paste("Processing", protein_indexed, "-", AF3_pdb_files_list[match_idx]))
    
    pdb_object <- read.pdb.custom(pdb_entry = AF3_pdb_files_list[match_idx], pdb_path = AF3_pdbs_path)
    
    data_BepiPred_2_subset <- data_BepiPred_2 %>% 
      mutate(Entry = tolower(Entry)) %>%
      filter(
        is.na(PDB),
        Entry == tolower(protein_indexed) | Entry == tolower(TIGR4_entry)
      )
  
  }
  
  bepipred2_seqres <- data_BepiPred_2_subset %>% 
    pull(AminoAcid) %>%
    str_c(collapse = "") %>%
    AAString()
  
  ca_atoms_reconstructed <- reconstruct.ca_atoms(pdb_object)
  pdb_seqres <- ca_atoms_reconstructed %>%
    pull(seqres_resid) %>%
    aa321() %>%
    str_c(collapse = "") %>%
    AAString()
  
  alignment <- pairwiseAlignment(
    pattern = pdb_seqres,
    subject = bepipred2_seqres
  )
  
  aligned_seqres <- as.character(alignedPattern(alignment))
  aligned_bepipred2_seqres <- as.character(alignedSubject(alignment))
  
  alignment_df <- tibble(
    seqres_resid = strsplit(aligned_seqres, "")[[1]],
    bepipred2_resid = strsplit(aligned_bepipred2_seqres, "")[[1]],
  ) %>%
    mutate(
      seqresno = row_number(),
      Position = case_when(
        bepipred2_resid != "-" ~ cumsum(bepipred2_resid != "-"),
        TRUE ~ NA
      ),
      seqres_resid = aa123(seqres_resid)
    ) %>%
    left_join(
      ca_atoms_reconstructed,
      by = c("seqresno", "seqres_resid")
      )
  
  resmap <- alignment_df %>%
    left_join(
      data_BepiPred_2_subset,
      by = c(
        "Position",
        "bepipred2_resid" = "AminoAcid"
             )
    ) 
  
  if (is.null(data_BepiPred_2_with_resno)) {
    data_BepiPred_2_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_BepiPred_2_with_resno <- bind_rows(data_BepiPred_2_with_resno, resmap)
  }
}

data_BepiPred_2_with_resno
```

## Store results as csv
```{r}
write.csv(data_BepiPred_2_with_resno,
          "output/BepiPred2.0/data_BepiPred2.csv", 
          row.names = FALSE)
```

# BepiPred-3.0 results
https://services.healthtech.dtu.dk/services/BepiPred-3.0/

```{r}
data_BepiPred_3_raw <- rbind(
    read_csv("output/BepiPred3.0/raw_output_part1.csv"),
    read_csv("output/BepiPred3.0/raw_output_part2.csv"),
    read_csv("output/BepiPred3.0/raw_output_vws.csv")
    )

data_BepiPred_3 <- data_BepiPred_3_raw %>%
  mutate(Residue = aa123(Residue)) %>%
  group_by(Accession) %>%
  mutate(Position = row_number()) %>%
  ungroup() %>%
  mutate(
    PDB = case_when(
      Accession %in% data_namesclasspdb$PDB ~ Accession,
      TRUE ~ NA
    ),
    protein = case_when(
     !is.na(PDB) ~ 
        data_namesclasspdb$protein[match(PDB, data_namesclasspdb$PDB)],
      !is.na(match(tolower(Accession), tolower(data_namesclasspdb$protein))) ~
        data_namesclasspdb$protein[match(tolower(Accession), tolower(data_namesclasspdb$protein))],
      !is.na(match(tolower(Accession), tolower(data_namesclasspdb$TIGR4_name))) ~
       data_namesclasspdb$protein[match(tolower(Accession), tolower(data_namesclasspdb$TIGR4_name))],
      TRUE ~ data_namesclasspdb$TIGR4_name[match(tolower(Accession), tolower(data_namesclasspdb$TIGR4_name))]
    )
  ) %>%
  left_join(
    data_namesclasspdb, by = c("protein","PDB")
  )

data_BepiPred_3
```

```{r}
data_BepiPred_3 %>% filter(PDB == "3CX3")
```


## BepiPred-3.0 with resno
```{r}
data_BepiPred_3_with_resno <- NULL

for (i in seq_len(nrow(data_namesclasspdb))) {
  protein_indexed <- data_namesclasspdb$protein[i]
  pdb_entry <- data_namesclasspdb$PDB[i]
  TIGR4_entry <- data_namesclasspdb$TIGR4_name[i]
  
  if (!is.na(pdb_entry)) {
    
    print(paste("Processing", protein_indexed, "-", pdb_entry))
    
    pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
    
    data_BepiPred_3_subset <- data_BepiPred_3 %>% 
      filter(Accession == pdb_entry)
                          
  } else {
    match_idx <- grep(paste0("^", tolower(protein_indexed), "_model_\\d+_fold_\\d+"), tolower(AF3_pdb_files_list))
    
    if (length(match_idx) == 0) {
      warning(paste("No model found for protein: ", protein_indexed, "retrying using TIGR4 name \n"))
      
      match_idx <- grep(paste0("^", tolower(TIGR4_entry), "_model_\\d+_fold_\\d+"), tolower(AF3_pdb_files_list))
      if (length(match_idx) == 0) {
        warning(paste("No model found for protein: ", protein_indexed, "or TIGR4 name: ", TIGR4_entry, "\n"))
        next
      }
    }
    
    print(paste("Processing", protein_indexed, "-", AF3_pdb_files_list[match_idx]))
    
    pdb_object <- read.pdb.custom(pdb_entry = AF3_pdb_files_list[match_idx], pdb_path = AF3_pdbs_path)
    
    data_BepiPred_3_subset <- data_BepiPred_3 %>% 
      mutate(Accession = tolower(Accession)) %>%
      filter(
        is.na(PDB),
        Accession == tolower(protein_indexed) | Accession == tolower(TIGR4_entry)
      )
    
  }
  
  bepipred3_seqres <- data_BepiPred_3_subset %>% 
    pull(Residue) %>%
    aa321() %>%
    str_c(collapse = "") %>%
    AAString()
  
  ca_atoms_reconstructed <- reconstruct.ca_atoms(pdb_object)
  pdb_seqres <- ca_atoms_reconstructed %>%
    pull(seqres_resid) %>%
    aa321() %>%
    str_c(collapse = "") %>%
    AAString()
  
  alignment <- pairwiseAlignment(
    pattern = pdb_seqres,
    subject = bepipred3_seqres
  )
  
  aligned_seqres <- as.character(alignedPattern(alignment))
  aligned_bepipred2_seqres <- as.character(alignedSubject(alignment))
  
  alignment_df <- tibble(
    seqres_resid = strsplit(aligned_seqres, "")[[1]],
    bepipred3_resid = strsplit(aligned_bepipred2_seqres, "")[[1]],
  ) %>%
    mutate(
      seqresno = row_number(),
      Position = case_when(
        bepipred3_resid != "-" ~ cumsum(bepipred3_resid != "-"),
        TRUE ~ NA
      ),
      seqres_resid = aa123(seqres_resid)
    ) %>%
    left_join(
      ca_atoms_reconstructed,
      by = c("seqresno", "seqres_resid")
      )
  
  resmap <- alignment_df %>%
    left_join(
      data_BepiPred_3_subset %>% mutate(Residue = aa321(Residue)),
      by = c(
        "Position",
        "bepipred3_resid" = "Residue"
             )
    ) 
  
  if (is.null(data_BepiPred_3_with_resno)) {
    data_BepiPred_3_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_BepiPred_3_with_resno <- bind_rows(data_BepiPred_3_with_resno, resmap)
  }
}

data_BepiPred_3_with_resno
```


## Store results as csv
```{r}
write.csv(data_BepiPred_3_with_resno,
          "output/BepiPred3.0/data_BepiPred3.csv", 
          row.names = FALSE)
```

# DiscoTope-3.0
https://services.healthtech.dtu.dk/services/DiscoTope-3.0/
```{r}
csv_files <- list.files("output/DiscoTope/individual_files/", pattern = "*.csv", full.names = TRUE)

data_DiscoTope_raw <- NULL

for (csv_file in csv_files) {
  #print(csv_file)
  
  file_id <- str_extract(basename(csv_file), "^(.*)(?=_[A-Z]_discotope3\\.csv)")
  
  temp_data <- read_csv(csv_file, show_col_types = FALSE) #%>%
   # mutate(file_id = file_id)

  if (is.null(data_DiscoTope_raw)) {
    data_DiscoTope_raw <- temp_data
  } else {
    data_DiscoTope_raw <- rbind(data_DiscoTope_raw, temp_data)
  }
}

data_DiscoTope_raw
```

```{r}
data_DiscoTope <- data_DiscoTope_raw %>%
  mutate(
    resno = res_id,
    
    resid = aa123(residue),
    
    file_id = sub("_[^_]+$", "", pdb),
    
    PDB = case_when(
      file_id %in% data_namesclasspdb$PDB ~ file_id,
      TRUE ~ NA
    ),
    
    model_id = case_when(
      !is.na(PDB) ~ NA,
      TRUE ~ str_extract(file_id, "^(.*?)(?=_model_)")
    ),
    
    protein = case_when(
      !is.na(PDB) ~ 
        data_namesclasspdb$protein[match(PDB, data_namesclasspdb$PDB)],
      !is.na(match(tolower(model_id), tolower(data_namesclasspdb$protein))) ~
        data_namesclasspdb$protein[match(tolower(model_id), tolower(data_namesclasspdb$protein))],
      !is.na(match(tolower(model_id), tolower(data_namesclasspdb$TIGR4_name))) ~
        data_namesclasspdb$protein[match(tolower(model_id), tolower(data_namesclasspdb$TIGR4_name))],
      TRUE ~ data_namesclasspdb$TIGR4_name[match(tolower(model_id), tolower(data_namesclasspdb$TIGR4_name))]
    )
  ) %>%
  select(-pdb, -res_id, -file_id, -model_id) %>%
  left_join(
    data_namesclasspdb, by = c("protein", "PDB")
  )

data_DiscoTope
```

## Store results as csv
```{r}
write.csv(data_DiscoTope,
          "output/DiscoTope/data_DiscoTope.csv", 
          row.names = FALSE)
```

# Epigraph
http://epigraph.kaist.ac.kr/
```{r}
epigraph_zip_files <- list.files(path = "output/EpiGraph/individual_files/", pattern = "\\.zip$")

for (zip_file in epigraph_zip_files) {
  unzip(zipfile = paste0("output/EpiGraph/individual_files/",zip_file), exdir = "output/EpiGraph/individual_files/")
}

epigraph_csv_files <- list.files(path = "output/EpiGraph/individual_files/", pattern = "\\.csv$")
epigraph_csv_files
```

```{r}
data_Epigraph_raw <- NULL

for (csv_file in epigraph_csv_files) {
  temp_data <- read_csv(paste0("output/EpiGraph/individual_files/",csv_file), show_col_types = FALSE) 
  print(csv_file)
  
  if (is.null(data_Epigraph_raw)) {
    data_Epigraph_raw <- temp_data
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_Epigraph_raw <- rbind(data_Epigraph_raw, temp_data)
  }
}

data_Epigraph <- data_Epigraph_raw %>% 
  separate(Residue, into = c("chain", "resid", "resno"), sep = ":", convert = TRUE) %>%
  mutate(file_id = PDB) %>%
  select(-PDB) %>%
  mutate(
    PDB = case_when(
      toupper(file_id) %in% data_namesclasspdb$PDB ~ toupper(file_id),
      TRUE ~ NA
    ),
    
    model_id = case_when(
      !is.na(PDB) ~ NA,
      TRUE ~ str_extract(file_id, "^(.*?)(?=_model_)")
    ),
    
    protein = case_when(
      !is.na(PDB) ~ 
        data_namesclasspdb$protein[match(PDB, data_namesclasspdb$PDB)],
      !is.na(match(tolower(model_id), tolower(data_namesclasspdb$protein))) ~
        data_namesclasspdb$protein[match(tolower(model_id), tolower(data_namesclasspdb$protein))],
      !is.na(match(tolower(model_id), tolower(data_namesclasspdb$TIGR4_name))) ~
        data_namesclasspdb$protein[match(tolower(model_id), tolower(data_namesclasspdb$TIGR4_name))],
      TRUE ~ data_namesclasspdb$TIGR4_name[match(tolower(model_id), tolower(data_namesclasspdb$TIGR4_name))]
    )
  ) %>%
  select(-file_id, -model_id) %>%
  left_join(
    data_namesclasspdb, by = c("protein","PDB")
  )
  
data_Epigraph %>% filter(is.na(PDB))
```

```{r}
data_Epigraph %>% filter(protein == "GlpO")
```


## Store results as csv
```{r}
write.csv(data_Epigraph,
          "output/EpiGraph/data_Epigraph.csv", 
          row.names=FALSE)
```

#(WIP) GraphBepi
Lookup the longest chain per protein
```{r}
data_DiscoTope %>%
  filter(residue %in% names(AMINO_ACID_CODE)[1:20]) %>%
  select(pdb, chain, length) %>%
  pivot_wider(names_from = chain, values_from = length) %>%
  mutate(across(where(is.list), ~sapply(., function(x) ifelse(is.null(x), NA, length(x)))))
```

Visually check all chains per protein
```{r}
for (pdb_entry in unique(data_DiscoTope$pdb)) {
  print(pdb_entry)
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  
  plot <- data_DiscoTope_subset %>%
    ggplot(aes(x=res_id, colour = residue)) +
    geom_point(aes(y=0)) +
    facet_wrap(~chain,ncol = 1) +
    labs(title=paste0(pdb_entry)) +
    theme_minimal()
  
  
  print(plot)
}
```

## read in the results
```{r}
graphbepi_txt_files <- list.files(path = "output/GraphBepi/individual_files/", pattern = "\\.txt$")

graphbepi_txt_files
```

```{r}
data_GraphBepi <- NULL

for (txt_file in graphbepi_txt_files) {
  print(txt_file)
  PDB <- sub("_.*", "", txt_file)
  chain <- sub("^[^_]+_([A-Za-z0-9])_.*", "\\1", txt_file)

  temp_data <- read.table(paste0("output/GraphBepi/individual_files/",txt_file), header = TRUE, skip=1)
  
  temp_data <- temp_data %>% 
    mutate(
      PDB = PDB,
      chain = chain,
      resid = aa123(AA),
      `GraphBepi probability` = Prob,
      `GraphBepi prediction` = Pred,
     Position = row_number()
    ) %>%
    select(PDB, chain, resid, Position, `GraphBepi probability`, `GraphBepi prediction`)
  
  if (is.null(data_GraphBepi)) {
    data_GraphBepi <- temp_data
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_GraphBepi <- rbind(data_GraphBepi, temp_data)
  }
}

data_GraphBepi
```

```{r}
data_GraphBepi_with_resno <- NULL

for (pdb_entry in unique(data_GraphBepi$PDB)) {
  print(pdb_entry)
  
  pdb_file <- paste0(pdbs_path,pdb_entry,".pdb") 
  
  if (file.exists(pdb_file)) {
    print(pdb_file)
    pdb_object <- read.pdb(file = pdb_file, rm.alt = FALSE)
  } else {
    pdb_object <- read.pdb(pdb_entry, rm.alt = FALSE)
  }
  
  graphbepi_subset <- data_GraphBepi %>% filter(PDB == pdb_entry)
  graphbepi_sequence <- graphbepi_subset %>% pull(resid) %>% aa321() %>% str_c(collapse = "")
  graphbepi_chain <- graphbepi_subset %>% pull(chain) %>% unique()
  
  required_adjustment <- map.seqposno.to.resno(sequence_AA1 = graphbepi_sequence, pdb_object)
  print(required_adjustment)
  
  resmap <- create.resmap(pdb_object$atom %>% filter(chain == graphbepi_chain), 
                                           graphbepi_subset,
                                           required_adjustment, join_by = c("resno","chain","resid")
                                   )
  
  if (is.null(data_GraphBepi_with_resno)) {
    data_GraphBepi_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_GraphBepi_with_resno <- rbind(data_GraphBepi_with_resno, resmap)
  }
}

data_GraphBepi_with_resno
```

# InterPro
https://www.ebi.ac.uk/interpro/search/sequence/
InterPro gives domain/topology prediction of proteins
```{r}
interpro_files <- list.files(path = "output/InterPro/", pattern = "\\.tsv$", full.names = TRUE)

interpro_files

interpro_colnames <- c(
  "Protein_accession", "MD5", "Sequence_length", "Analysis",
  "Signature_accession", "Signature_description", "Start", "End",
  "Score", "Status", "Date", "InterPro_accession", "InterPro_description",
  "GO_annotations", "Pathways"
)

interpro_coltypes <- cols(
  Protein_accession = col_character(),
  MD5 = col_character(),
  Sequence_length = col_integer(),
  Analysis = col_character(),
  Signature_accession = col_character(),
  Signature_description = col_character(),
  Start = col_integer(),
  End = col_integer(),
  Score = col_double(),  
  Status = col_logical(),
  Date = col_date(format = "%d-%m-%Y"),
  InterPro_accession = col_character(),
  InterPro_description = col_character(),
  GO_annotations = col_character(),
  Pathways = col_character()
)

# Load the file with correct column types and headers
data_interpro_raw <- read_tsv(
  interpro_files,
  col_names = interpro_colnames,
  col_types = interpro_coltypes
) %>%
  setNames(interpro_colnames)
```

```{r}
map_namesclasspdb_AF3 <- read.csv("data/map_nameclass_pdbAF3.csv")
```

```{r}
data_interpro <- data_interpro_raw %>% mutate(
  structure = case_when(
    Protein_accession %in% map_namesclasspdb_AF3$PDB ~ Protein_accession,
    tolower(Protein_accession) %in% tolower(map_namesclasspdb_AF3$protein) ~ 
      map_namesclasspdb_AF3$AF3[
        match(tolower(Protein_accession), tolower(map_namesclasspdb_AF3$protein))
      ],
    Protein_accession %in% map_namesclasspdb_AF3$TIGR4_name ~ {
        idx <- match(Protein_accession, map_namesclasspdb_AF3$TIGR4_name)
        ifelse(!is.na(map_namesclasspdb_AF3$PDB[idx]),
               map_namesclasspdb_AF3$PDB[idx],
               map_namesclasspdb_AF3$AF3[idx])
      },
    TRUE ~ NA_character_
  )
)
```


```{r}
data_interpro %>% select(Protein_accession, structure) %>% unique()
```


```{r}
data_interpro %>% filter(structure == "alia_model_4_fold_0.pdb") %>%
  select(Analysis, Signature_description, InterPro_accession, Start, End, Score) %>%
  arrange(-log10(Score))
```

```{r}
data_interpro %>% 
  #filter(!(Signature_description == "-" & InterPro_description == "-")) %>% 
  pull(Analysis) %>%
  table() %>%
  as.data.frame() %>%
  arrange(desc(Freq))

data_interpro %>% 
  filter(!(Signature_description == "-" & InterPro_description == "-")) %>% 
  pull(Analysis) %>%
  table() %>%
  as.data.frame() %>%
  arrange(desc(Freq))
```

```{r}
data_interpro %>%
  filter(Analysis != "-") %>%  # Exclude empty entries
  mutate(present = TRUE) %>%   # Mark presence explicitly
  distinct(Protein_accession, Analysis, present, Score) %>%  # Avoid duplicates
  pivot_wider(names_from = Analysis, values_from = Score, values_fill = NA)
```

```{r}
data_interpro %>% pull(structure) %>% unique()
```


```{r, fig.width=14, warning=FALSE}
charweight = 0.3

for (protein in sort(unique(data_interpro$structure))) {
  print(protein)
  
  data_interpro_subset <- data_interpro %>%
    filter(structure == protein,
           !is.na(Score)) %>%
    mutate(e_value = -log10(Score),
           Description = case_when(
             InterPro_description != "-" ~ InterPro_description,
             TRUE ~ Signature_description
             )
           ) %>%
    filter(!is.na(Description) & Description != "-") %>%
    # Calculate the max e_value for each Analysis group and reorder Analysis factor
    group_by(Analysis) %>%
    mutate(max_e_value = max(e_value, na.rm = TRUE)) %>%
    
    ungroup() %>%
    mutate(Analysis = factor(Analysis, levels = unique(Analysis[order(max_e_value)]))) %>%
    select(-max_e_value)  # Remove max_e_value column
  
  plot <- data_interpro_subset %>%
    group_by(Analysis,Description) %>%
    mutate(
      x_midpoint = Start + (End - Start)/2,
      min_start = min(Start),
      max_end = max(End),
      group_midpoint = min_start + (max_end - min_start)/2,
      size_calc = (End - Start)/(nchar(Description)*charweight),
      group_size_calc = (max_end - min_start)/(nchar(Description)*charweight) 
        
    ) %>%
    ggplot() +
    geom_segment(
      aes(
        x = Start, xend = End,
        y = Analysis, yend = Analysis,
        color = Description,
        alpha = e_value
      ),
      size = 8
    ) +
    geom_text(
      aes(
        x = ifelse(
          size_calc >= 5,
          x_midpoint,
          group_midpoint
        ),
        y = Analysis,
        color = Description, 
        label = Description,
        #size = ifelse(
        #  size_calc >= 5,
        #  pmin(size_calc, 0.1),
        #  pmin(group_size_calc, 0.1)
        #)
        size = ifelse(
          size_calc >= 5,
          pmin(size_calc/100, 0.1),
          pmin(pmax(group_size_calc/100,0.05), 0.1)
        )
      ),
      color = "black",
      show.legend = FALSE
    ) +
    scale_alpha_continuous() +
    theme_bw() +
    labs(
      title = paste0("Domains in ", protein),
      x = "Amino acid position",
      y = "Analysis"
    )
  
  print(plot)
}

```


## cleaned
We want to build up annotation, using the most significant (based on E-value) non-overlapping domain hits

```{r}
Analysis_remove_selection <- c(
  "CDD"
  )

data_interpro %>%
  filter(!Analysis %in% Analysis_remove_selection) %>%
  group_by(Protein_accession) %>%
  mutate(Description = case_when(
             InterPro_description != "-" ~ InterPro_description,
             !is.na(InterPro_description) ~ InterPro_description,
             TRUE ~ Signature_description
             )
         ) %>%
  ungroup() %>%
  filter(!is.na(Description), Description != "-") %>%
  arrange(Protein_accession, Score, Description) 
```

```{r}
keep_non_overlapping <- function(df) {
  df <- df %>%
    filter(!is.na(Score),
           Score != 0) %>%
    mutate(e_value = -log10(Score)) %>%
    arrange(desc(e_value))  # Most significant first
  
  kept <- list()
  occupied_ranges <- IRanges::IRanges()  # Efficient range management
  
  for (i in seq_len(nrow(df))) {
    this_range <- IRanges::IRanges(start = df$Start[i], end = df$End[i])
    
    # Check for overlap with any previously kept range
    if (length(occupied_ranges) == 0 || length(IRanges::findOverlaps(this_range, occupied_ranges)) == 0) {
      occupied_ranges <- c(occupied_ranges, this_range)
      kept[[length(kept) + 1]] <- df[i, ]
    }
  }
  
  bind_rows(kept)
}
```

```{r}
data_interpro %>%
  filter(Protein_accession == "trxB",
         !is.na(Score)) %>%
  mutate(e_value = -log10(Score)) %>%
  arrange(desc(e_value))
```


```{r}
get_shortest <- function(x) {
  sapply(x, function(val) {
    if (is.na(val) || val == "-") return(NA_character_)
    parts <- str_split(val, ",")[[1]]
    parts <- str_trim(parts)
    parts[which.min(nchar(parts))]
  })
}

filtered_hits <- data_interpro %>%
  group_by(Protein_accession) %>%
  mutate(
    Description = case_when(
      # case 1: only InterPro is valid
      (!is.na(InterPro_description) & InterPro_description != "-") &
        (is.na(Signature_description) | Signature_description == "-") ~ 
        get_shortest(InterPro_description),
      
      # case 2: only Signature is valid
      (!is.na(Signature_description) & Signature_description != "-") &
        (is.na(InterPro_description) | InterPro_description == "-") ~ 
        get_shortest(Signature_description),
      
      # case 3: both valid → take the shortest of the two shortest splits
      (!is.na(InterPro_description) & InterPro_description != "-") &
        (!is.na(Signature_description) & Signature_description != "-") ~ {
          interpro_short  <- get_shortest(InterPro_description)
          signature_short <- get_shortest(Signature_description)
          ifelse(nchar(interpro_short) <= nchar(signature_short), 
                 interpro_short, 
                 signature_short)
        },
      
      # fallback
      TRUE ~ NA_character_
    )
  ) %>%
  ungroup() %>%
  filter(!is.na(Description), Description != "-") %>%
  arrange(Protein_accession, Score, Description) %>%
  group_by(Protein_accession) %>%
  group_modify(~keep_non_overlapping(.x)) %>%
  ungroup()


filtered_hits
```

```{r}
data_interpro_filtered <- data_interpro %>%
  group_by(Protein_accession) %>%
  mutate(Description = case_when(
    InterPro_description != "-" ~ InterPro_description,
    !is.na(InterPro_description) ~ InterPro_description,
    TRUE ~ Signature_description
  ),
  Description = sapply(str_split(Description, ","), function(x) {
    x_trimmed <- str_trim(x)              # remove extra spaces
    x_shortest <- x_trimmed[which.min(nchar(x_trimmed))] # keep shortest
    x_shortest
  })
  ) %>%
  filter(
    !is.na(Description), 
    Description != "-",
    !Analysis %in% Analysis_remove_selection
  ) %>%
  arrange(structure, Score) %>%
  group_modify(~ keep_non_overlapping(.x)) %>%
  select(structure, Sequence_length, Start, End, Description, InterPro_description, Signature_description, Analysis, e_value, Score) %>%
  ungroup()

data_interpro_filtered
```

```{r}
for (protein in sort(unique(data_interpro_filtered$Protein_accession))) {
  print(protein)
  
  data_interpro_subset <- data_interpro_filtered %>%
    filter(Protein_accession == protein,
           !is.na(Score)) %>%
    mutate(
      e_value = -log10(Score)
           ) %>%
    filter(!is.na(Description) & Description != "-") %>%
    # Calculate the max e_value for each Analysis group and reorder Analysis factor
    group_by(Analysis) %>%
    mutate(max_e_value = max(e_value, na.rm = TRUE)) %>%
    
    ungroup() %>%
    mutate(Analysis = factor(Analysis, levels = unique(Analysis[order(max_e_value)]))) %>%
    select(-max_e_value)  # Remove max_e_value column
  
  plot <- data_interpro_subset %>%
    group_by(Analysis,Description) %>%
    mutate(
      x_midpoint = Start + (End - Start)/2,
      min_start = min(Start),
      max_end = max(End),
      group_midpoint = min_start + (max_end - min_start)/2,
      size_calc = (End - Start)/(nchar(Description)*charweight),
      group_size_calc = (max_end - min_start)/(nchar(Description)*charweight) 
        
    ) %>%
    ggplot() +
    geom_segment(
      aes(
        x = Start, xend = End,
        y = Analysis, yend = Analysis,
        color = Description,
        alpha = e_value
      ),
      size = 8
    ) +
    geom_text(
      aes(
        x = ifelse(
          size_calc >= 5,
          x_midpoint,
          group_midpoint
        ),
        y = Analysis,
        color = Description, 
        label = Description,
        #size = ifelse(
        #  size_calc >= 5,
        #  pmin(size_calc, 0.1),
        #  pmin(group_size_calc, 0.1)
        #)
        size = ifelse(
          size_calc >= 5,
          pmin(size_calc/100, 0.1),
          pmin(pmax(group_size_calc/100,0.05), 0.1)
        )
      ),
      color = "black",
      show.legend = FALSE
    ) +
    scale_alpha_continuous() +
    theme_bw() +
    labs(
      title = paste0("Domains in ", protein),
      x = "Amino acid position",
      y = "Analysis"
    )
  
  print(plot)
}
```

