---
title: "Streptococcus pneumoniae epitope prediction tools"
author: "Jan Paul van Meenen"
date: "`r Sys.Date()`"
output: html_document
---

#Packages
```{r}
data_management_packages_to_install <- c("data.table", "tidyverse" ,"readxl","purrr","readr","scales","msa")
data_visualisation_packages_to_install <- c("ggplot2","cowplot","pheatmap","viridisLite","ggseqlogo")
structural_packages_to_install <- c("bio3d","NGLVieweR", "htmlwidgets","htmltools")

packages_to_install <- c(data_management_packages_to_install, 
                         data_visualisation_packages_to_install,
                         structural_packages_to_install)

# Check and install packages using a for loop
for (package in packages_to_install) {
  if (!requireNamespace(package, quietly = TRUE)) {
    install.packages(package)
  }
  library(package, character.only = TRUE)
}

BiocManager_packages_to_install <- c("Biostrings","pwalign")

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

for (package in BiocManager_packages_to_install) {
  if (!requireNamespace(package, quietly = TRUE)) {
    BiocManager::install(package)
  }
  library(package, character.only = TRUE)
}
```

#Functions
```{r}
source("functions/mrp_pc_functions.R")
```

# BepiPred-2.0
## split fasta into parts
Bepi-2.0 only allows for 50 proteins at a time.
```{r}
fasta_lines <- readLines("data/fasta/pdb_protein_sequences.fasta")

# Group lines into proteins based on `>...` headers
protein_entries <- split(fasta_lines, cumsum(grepl("^>", fasta_lines)))

# Split protein entries into chunks of 50
protein_chunks <- split(protein_entries, ceiling(seq_along(protein_entries) / 50))

for (i in seq_along(protein_chunks)) {
  # Combine lines for each chunk
  chunk_lines <- unlist(protein_chunks[[i]])
  
  # Write the chunk to a new file
  writeLines(chunk_lines, paste0("data/fasta/pdb_protein_sequences_part_", i, ".fasta"))
}

```
https://services.healthtech.dtu.dk/services/BepiPred-2.0/ 
```{r}
data_BepiPred_2 <- rbind(
    read_csv("output/BepiPred2.0/BepiPred_part1_summary_67868578000F647FB45B7803.csv"),
    read_csv("output/BepiPred2.0/BepiPred_part2_summary_67868580000F64B5DAEDC816.csv")
    ) 

data_BepiPred_2 %>%
  filter(Entry == "3CX3") %>%
  mutate(AminoAcid = aa123(AminoAcid))
```

## Generalized
### get valid all pdb entires and files
```{r}
PDB_list <- data_structural_with_pdb %>%
  pull(`PDB entry`) %>% 
  na.omit() %>% 
  str_split(", ") %>% 
  unlist() %>% 
  unique() 

PDB_list
```

```{r}
pdbs_path = "output/pdb/"

pdb_files_list <- sub("\\.pdb$", "",
                      list.files(path = pdbs_path, 
                             pattern = "\\.pdb$")
                      )

for (pdb_entry in pdb_files_list) {
  print(pdb_entry)

}
```

## BepiPred-2.0 with resno
```{r}
data_BepiPred_2_with_resno <- NULL

for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_file <- paste0(pdbs_path,pdb_entry,".pdb")
  
  if (file.exists(pdb_file)) {
    pdb_object <- read.pdb(file = pdb_file, rm.alt = FALSE)
    print(pdb_file)
  } else {
    pdb_object <- read.pdb(pdb_entry, rm.alt = FALSE)
  }
  
  required_adjustment <- map.seqresno.to.resno(pdb_object)
  #print(required_adjustment)
  
  resmap <- create.resmap.bepipred(pdb_object$atom, 
                                           data_BepiPred_2 %>% filter(Entry == pdb_entry),
                                           required_adjustment)
  
  if (is.null(data_BepiPred_2_with_resno)) {
    data_BepiPred_2_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_BepiPred_2_with_resno <- rbind(data_BepiPred_2_with_resno, resmap)
  }
}

data_BepiPred_2_with_resno
```

```{r}
pdbs_path = "output/pdb/"

for (pdb_entry in pdb_files_list) {
  file_path = paste0(pdbs_path,pdb_entry,".pdb")
  print(file_path)
  #if(file.exists(file_path)) {
  #  pdb_object <- read.pdb(file_path, rm.alt = FALSE)
  #} else {
  #print(paste0("file missing: ", pdb_entry))
  #}
  pdb_object <- read.pdb(pdb_entry,rm.alt=FALSE)
  
  required_adjustment <- map.seqresno.to.resno(pdb_object)
  #print(required_adjustment)
  
  resmap <- create.resmap.bepipred(pdb_object$atom, 
                                   data_BepiPred_2 %>% filter(Entry == pdb_entry),
                                   required_adjustment)
  print(resmap)
  
  color_map <- data.frame(
    ProbabilityBin = levels(cut(seq(0.0, 1, by = 0.05), breaks = seq(0, 1, by = 0.05), include.lowest = TRUE)),
    color = colorRampPalette(c("darkblue","white","red"))(20)
    ) %>%
    rbind(c(NA,"black"))
  
  #print(color_map)
  
  predictions_df <- resmap %>%
    mutate(ProbabilityBin = cut(EpitopeProbability, 
                              breaks = seq(0, 1, by = 0.05), 
                              include.lowest = TRUE)
    ) %>%
    group_by(ProbabilityBin) %>% 
    summarize(resno_list = list(paste0(resno,":A")), .groups = "drop") %>%
    mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " "))) %>%
    left_join(color_map)
  
  #print(predictions_df) 
  
  view <- prediction.viewer(pdb_entry, pdbs_path, predictions_df = predictions_df)
  
  saveWidget(view, file = paste0("output/snapshot/",pdb_entry,".html"), title=pdb_entry, selfcontained = TRUE)
}
```


```{r}
levels(cut(seq(0.05, 1, by = 0.05), breaks = seq(0, 1, by = 0.05), include.lowest = TRUE))
```


```{r}
my_palette <- inferno(20)
my_palette <- colorRampPalette(c("blue","white","red"))(20)
barplot(rep(1, length(my_palette)), col = my_palette, border = NA, axes=FALSE, space = 0, main = "Visualization of colorRampPalette")
```

```{r}
# Create a mapping of bins to colors
color_map <- data.frame(
  ProbabilityBin = levels(cut(seq(0.05, 1, by = 0.05), breaks = seq(0, 1, by = 0.05), include.lowest = TRUE)),
  color = inferno(20)
)

color_map
```

## Ggplot


```{r, fig.width=14}
plot_bepipred <- data_BepiPred_2 %>%
  filter(Entry == "3CX3") %>%
  ggplot(aes(x=Position)) +
  geom_tile(aes(y=0.1,fill = EpitopeProbability), height = 0.1) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0.5, limits = c(0, 1)) +  
  theme_minimal() + 
  theme(
    axis.text.y = element_blank(),   
    axis.title.y = element_blank(),  
    axis.ticks.y = element_blank()
  ) +
  labs(fill = "BepiPred-2.0 prediction")

plot_surface <- data_BepiPred_2 %>%
  filter(Entry == "3CX3") %>%
  ggplot(aes(x=Position)) +
  geom_tile(aes(y=0.1,fill = RelativeSurfaceAccessilibity), height = 0.1) +
  theme_minimal() + 
  theme(
    axis.text.y = element_blank(),   
    axis.title.y = element_blank(),  
    axis.ticks.y = element_blank()
  )

plot_grid(
  plot_bepipred + theme(
    axis.text.x = element_blank(),   
    axis.title.x = element_blank(),  
    axis.ticks.x = element_blank()
  ), 
  plot_surface, align = "vh", ncol = 1)
```

# BepiPred-3.0 results
https://services.healthtech.dtu.dk/services/BepiPred-3.0/
```{r}
data_BepiPred_3 <- rbind(
    read_csv("output/BepiPred3.0/raw_output_part1.csv"),
    read_csv("output/BepiPred3.0/raw_output_part2.csv")
    ) %>%
  #mutate(Residue = aa123(Residue)) %>%
  group_by(Accession) %>%
  mutate(Position = row_number())
```

## BepiPred-3.0 with resno
```{r}
data_BepiPred_3_with_resno <- NULL

for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_file <- paste0(pdbs_path,pdb_entry,".pdb") 
  
  if (file.exists(pdb_file)) {
    print(pdb_file)
    pdb_object <- read.pdb(file = pdb_file, rm.alt = FALSE)
  } else {
    pdb_object <- read.pdb(pdb_entry, rm.alt = FALSE)
  }
  
  required_adjustment <- map.seqresno.to.resno(pdb_object)
  #print(required_adjustment)
  
  resmap <- create.resmap.bepipred(pdb_object$atom, 
                                           data_BepiPred_3 %>% filter(Accession == pdb_entry),
                                           required_adjustment)
  
  if (is.null(data_BepiPred_3_with_resno)) {
    data_BepiPred_3_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_BepiPred_3_with_resno <- rbind(data_BepiPred_3_with_resno, resmap)
  }
}

data_BepiPred_3_with_resno
```

```{r}
data_BepiPred_3_with_resno %>%
  pull(`BepiPred-3.0 score`) %>%
  summary()
```


# DiscoTope-3.0
https://services.healthtech.dtu.dk/services/DiscoTope-3.0/
```{r}
csv_files <- list.files("output/DiscoTope", pattern = "*.csv", full.names = TRUE)

data_DiscoTope <- NULL

for (csv_file in csv_files) {
  temp_data <- read_csv(csv_file) 
  
  if (is.null(data_DiscoTope)) {
    data_DiscoTope <- temp_data
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_DiscoTope <- rbind(data_DiscoTope, temp_data)
  }
}

data_DiscoTope
```

```{r}
data_DiscoTope %>%
  filter(pdb == "3CX3")
```

## Test 3CX3
```{r}
data_DiscoTope <- data_DiscoTope %>%
  mutate(pdb = sub("_.*", "", pdb)) %>%
  rename(res_id ="resno" )

data_DiscoTope %>%
  filter(pdb == "3CX3")

data_BepiPred_2 %>%
  filter(Entry == "3CX3")
```


```{r}
plot_discotope <- data_DiscoTope %>%
  filter(pdb == "3CX3") %>%
  ggplot(aes(x=resno, y=calibrated_score, fill=calibrated_score)) +
  geom_col(width = 1) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +  
  theme_minimal() + 
  theme(
    axis.text.y = element_blank(),   
    axis.title.y = element_blank(),  
    axis.ticks.y = element_blank()
  ) +
  labs(fill = "DiscoTope-3.0 prediction")

plot_discotope
```

# Epigraph
http://epigraph.kaist.ac.kr/
```{r}
epigraph_zip_files <- list.files(path = "output/EpiGraph/individual_files/", pattern = "\\.zip$")

for (zip_file in epigraph_zip_files) {
  unzip(zipfile = paste0("output/EpiGraph/individual_files/",zip_file), exdir = "output/EpiGraph/individual_files/")
}

epigraph_csv_files <- list.files(path = "output/EpiGraph/individual_files/", pattern = "\\.csv$")
epigraph_csv_files
```

```{r}
data_Epigraph <- NULL

for (csv_file in epigraph_csv_files) {
  temp_data <- read_csv(paste0("output/EpiGraph/individual_files/",csv_file)) 
  print(csv_file)
  
  if (is.null(data_Epigraph)) {
    data_Epigraph <- temp_data
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_Epigraph <- rbind(data_Epigraph, temp_data)
  }
}

data_Epigraph <- data_Epigraph %>% 
  separate(Residue, into = c("chain", "resid", "resno"), sep = ":", convert = TRUE) %>%
  mutate(PDB = toupper(PDB))

data_Epigraph

write.csv(data_Epigraph,"output/EpiGraph/data_Epigraph.csv", row.names=FALSE)
```

```{r}
data_Epigraph_subset <- data_Epigraph %>% filter(PDB == "3ZPP")

plot_epigraph <- data_Epigraph_subset %>% 
    ggplot(aes(x=resno, y=Score, fill=Score)) +
    geom_col(width = 1,) +
    scale_fill_gradient2(low = "blue", mid = "gray95", high = "red") +  
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "EpiGraph prediction")

plot_epigraph
```

#GraphBepi
```{r}
data_DiscoTope %>%
  filter(residue %in% names(AMINO_ACID_CODE)) %>%
  select(pdb, chain, length) %>%
  pivot_wider(names_from = chain, values_from = length) %>%
  mutate(across(where(is.list), ~sapply(., function(x) ifelse(is.null(x), NA, length(x)))))
```

```{r}
for (pdb_entry in unique(data_DiscoTope$pdb)) {
  print(pdb_entry)
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  
  
  plot <- data_DiscoTope_subset %>%
    ggplot(aes(x=resno, colour = residue)) +
    geom_point(aes(y=1)) +
    facet_wrap(~chain,ncol = 1) +
    labs(title=paste0(pdb_entry)) +
    theme_minimal()
  
  
  print(plot)
}
```


# Plot all together
```{r, warning=FALSE, fig.width=14, fig.height=10}
for (pdb_entry in pdb_files_list) {
  #print(pdb_entry)
  
  data_BepiPred_2_with_resno_subset <- data_BepiPred_2_with_resno %>% filter(Entry == pdb_entry)
  #print(data_BepiPred_2_with_resno_subset)
  
  plot_bepipred_2_RSA <- data_BepiPred_2_with_resno_subset %>% 
    ggplot(aes(x=resno, y=RelativeSurfaceAccessilibity, fill=RelativeSurfaceAccessilibity)) +
    geom_col(width = 1) +  
    scale_fill_gradient2(low = "darkcyan", mid = "gray95", high = "deeppink4", midpoint = 0.5, limits = c(0, 1)) +  
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "Relative Surface Accessilibity\n(NetsurfP)")
  
  
  plot_bepipred_2 <- data_BepiPred_2_with_resno_subset %>% 
    ggplot(aes(x=resno, y=EpitopeProbability, fill=EpitopeProbability)) +
    geom_col(width = 1) +  
    scale_fill_gradient2(low = "blue", mid = "gray95", high = "red", midpoint = 0.5, limits = c(0, 1)) +  
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "BepiPred-2.0 prediction")
  #print(plot_bepipred_2)
  
  data_BepiPred_3_with_resno_subset <- data_BepiPred_3_with_resno %>% filter(Accession == pdb_entry)
  median_BepiPred_3_subset_score <- median(data_BepiPred_3_with_resno_subset$`BepiPred-3.0 score`)
  
  plot_bepipred_3 <- data_BepiPred_3_with_resno_subset %>%
    ggplot(aes(x=resno, y=`BepiPred-3.0 score`, fill=`BepiPred-3.0 score`)) +
    geom_col(width = 1) +
    scale_fill_gradient2(low = "blue", mid = "gray95", high = "red", 
                         midpoint = median_BepiPred_3_subset_score#, 
                         #limits = c(0, NA)
                         ) +
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "BepiPred-3.0 prediction")
  
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  #print(data_DiscoTope_subset)
  plot_discotope <- data_DiscoTope_subset %>% 
    ggplot(aes(x=resno, y=calibrated_score, fill=calibrated_score)) +
    geom_col(width = 1) +
    scale_fill_gradientn(colors = c("blue", "gray95", "#FFC0C0", "#FF7070", "red"), 
                         values = scales::rescale(c(-1, 0, 0.4, 0.9, 1.5), to = c(0, 1)),
                         limits = c(-1, 1.5),
                         oob = scales::squish
                         ) +  
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "DiscoTope-3.0 prediction")
  #print(plot_discotope)
  
  plot_discotope_RSA <- data_DiscoTope_subset %>% 
    ggplot(aes(x=resno, y=rsa, fill=rsa)) +
    geom_col(width = 1) +
    scale_fill_gradient2(low = "darkcyan", mid = "gray95", high = "deeppink4", midpoint = 0.5, limits = c(0, 1)) +  
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "Relative surface accessibility\n(Shrake-Rupley)")
  #print(plot_discotope)
  
  limit_lower <- data_DiscoTope_subset %>% pull(resno) %>% min(na.rm = TRUE)
  limit_upper <- data_DiscoTope_subset %>% pull(resno) %>% max(na.rm = TRUE)
  
  data_Epigraph_subset <- data_Epigraph %>% filter(PDB == pdb_entry)
  
  plot_epigraph <- data_Epigraph_subset %>% 
    ggplot(aes(x=resno, y=Score, fill=Score)) +
    geom_col(width = 1,) +
    scale_fill_gradient2(low = "blue", mid = "gray95", high = "red") +  
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "EpiGraph prediction")
  
  #Make the full figures
  figure <- plot_grid(
    plot_bepipred_2_RSA + scale_x_continuous(limits = c(limit_lower,limit_upper)) +
      theme(
        axis.ticks = element_blank(),  
        axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ) +
      labs(title = paste0(pdb_entry)), 
    
    plot_bepipred_2 + scale_x_continuous(limits = c(limit_lower,limit_upper)) +
      theme(
        axis.ticks = element_blank(),  
        axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ), 
    
    plot_bepipred_3 + scale_x_continuous(limits = c(limit_lower,limit_upper)) +
      theme(
        axis.ticks = element_blank(),  
        axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ),
    
    plot_discotope + scale_x_continuous(limits = c(limit_lower,limit_upper)) + 
      theme(
        axis.ticks = element_blank(),  
        axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ),
    
    plot_discotope_RSA + scale_x_continuous(limits = c(limit_lower,limit_upper)) + 
      theme(
        axis.ticks = element_blank(),  
        axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ),
    
    plot_epigraph + scale_x_continuous(limits = c(limit_lower,limit_upper)) + 
      theme(
        axis.ticks = element_blank(),  
        #axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        #axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        #axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ),
    
    align = "v", ncol = 1)
  
  print(figure)
  
  ggsave(filename = paste0("output/figures/tracks/", pdb_entry, ".png"), 
       plot = figure, 
       width = 14, height = 10, dpi = 300,
       bg="white")

}
```

## create joined predictions df
```{r}
data_predictions_sequencebased <- full_join(
  data_BepiPred_2_with_resno %>%
    mutate(PDB = Entry,
           `BepiPred-2.0 probability` = EpitopeProbability,
           `RSA NetsurfP` = RelativeSurfaceAccessilibity) %>%
    select(PDB, chain, resno,resid,`BepiPred-2.0 probability`,`RSA NetsurfP`) %>%
    filter(resid %in% toupper(AMINO_ACID_CODE))
  ,
  
  data_BepiPred_3_with_resno %>%
    mutate(PDB = Accession) %>%
    select(PDB, chain, resno, resid, `BepiPred-3.0 score`,`BepiPred-3.0 linear epitope score`) %>%
    filter(resid %in% toupper(AMINO_ACID_CODE))
  ,
  by = c("PDB","chain","resno","resid")
)

data_predictions_sequencebased

data_predictions_sequencebased %>%
  filter(if_any(everything(), is.na))
```

```{r}
data_predictions_structurebased <- full_join(
  data_Epigraph %>%
    mutate(Epigraph_score = Score) %>%
    select(PDB, chain, resno, resid, Epigraph_score)
  ,
  data_DiscoTope %>%
    mutate(PDB = pdb,
           resid = aa123(residue),
           `RSA Shrake-Rupley` = rsa,
           `DiscoTope-3.0_calibrated_score` = calibrated_score
    ) %>%
    select(PDB, chain, resno, resid, `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, `RSA Shrake-Rupley`)
  , 
  by = c("PDB","resno","chain","resid")
)
data_predictions_structurebased

data_predictions_structurebased %>%
  filter(if_any(everything(), is.na))

data_predictions_structurebased <- data_predictions_structurebased %>%
  na.omit()
  
data_predictions_structurebased %>%
  filter(if_any(everything(), is.na))
```


```{r}
data_predictions <- full_join(
  data_predictions_sequencebased,
  data_predictions_structurebased,
  by = c("PDB", "chain", "resno","resid")
  )

data_predictions %>%
  filter(if_any(everything(), is.na))

data_predictions <- data_predictions %>%
  na.omit() 

data_predictions
```

## normalize
```{r}
data_predictions_normalized <- data_predictions %>%
  group_by(PDB, chain) %>%
  mutate(
    across(
      c(#`BepiPred-2.0 probability`,
        `BepiPred-3.0 score` , `BepiPred-3.0 linear epitope score`, 
        `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, Epigraph_score), 
      rescale
    )
  ) %>%
  ungroup()

data_predictions_normalized
```

```{r, fig.width = 14}
data_test_wide <- data_predictions_normalized %>%
  filter(PDB == "6YAG", chain == "A") %>%
  #distinct(resno, .keep_all = TRUE) %>%
  select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
        `DiscoTope-3.0_calibrated_score`, Epigraph_score) %>%
  column_to_rownames("resno") %>%
  as.data.frame()

pheatmap(t(data_test_wide), 
         cluster_rows = TRUE,         
         cluster_cols = FALSE,         
         main = "6YAG", 
         color = colorRampPalette(c("blue", "gray95", "red"))(50),
         labels_col = "resno",
         clustering_distance_rows = "correlation") 
```

```{r, fig.width = 14}
for (pdb_entry in unique(data_predictions_normalized$PDB)) {
  data_predictions_normalised_subset <- data_predictions_normalized %>%
    filter(PDB == pdb_entry) 
  
  for (chain_index in unique(data_predictions_normalised_subset$chain)) {
    
    data_predictions_normalised_subset_wide <- data_predictions_normalised_subset %>%
      filter(chain == chain_index) %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score) %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    figure <- pheatmap(t(data_predictions_normalised_subset_wide), 
                       cluster_rows = TRUE,         
                       cluster_cols = FALSE,         
                       main = paste0(pdb_entry, " chain ", chain_index), 
                       color = colorRampPalette(c("blue", "gray95", "red"))(50),
                       labels_col = "resno",
                       clustering_distance_rows = "correlation") 
    
    print(figure)
    
    ggsave(filename = paste0("output/figures/heatmap/", pdb_entry, " chain ", chain_index, "_heatmap.png"), 
           plot = figure, 
           width = 14, height = 10, dpi = 300,
           bg="white")
  }
}
```


## 3D
### bepipred-3.0
```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_file <- paste0(pdbs_path,pdb_entry,".pdb") 
  
  if (file.exists(pdb_file)) {
    print(pdb_file)
    pdb_object <- read.pdb(file = pdb_file, rm.alt = FALSE)
  } else {
    pdb_object <- read.pdb(pdb_entry, rm.alt = FALSE)
  }
  
  data_BepiPred_3_with_resno_subset <- data_BepiPred_3_with_resno %>% filter(Accession == pdb_entry) %>% na.omit()
  
  for (chain_index in unique(data_BepiPred_3_with_resno_subset$chain)) {
    data_BepiPred_3_with_resno_subset_chain <- data_BepiPred_3_with_resno_subset %>% filter(chain == chain_index)
    
    min_score <- min(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, na.rm = TRUE)
    median_score <- median(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, na.rm = TRUE)
    max_score <- max(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, na.rm = TRUE)
    
    
    low_to_median <- colorRampPalette(c("blue", "gray95"))(10)
    median_to_high <- colorRampPalette(c("gray95", "red"))(10)
    
    breaks <- seq(min_score, max_score, length.out = 21)  # 20 intervals
    colors <- c(low_to_median, median_to_high)  # 20 colors
    
    color_map <- data.frame(
      ProbabilityBin = cut(seq(min_score, max_score, length.out = 20), 
                           breaks = breaks, include.lowest = TRUE),
      color = colors
    ) %>%
      rbind(c(NA,"black"))
    
    predictions_df <- data_BepiPred_3_with_resno_subset_chain %>%
      mutate(ProbabilityBin = cut(`BepiPred-3.0 score`, 
                                  breaks = breaks, include.lowest = TRUE)
      ) %>%
      group_by(ProbabilityBin) %>% 
      summarize(resno_list = list(paste0(resno,":",chain_index)), .groups = "drop") %>%
      mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " "))) %>%
      left_join(color_map, by = "ProbabilityBin")
    
    view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index, predictions_df = predictions_df)
    
    #print(view)
    
    saveWidget(view, file = paste0("output/BepiPred3.0/snapshot/",pdb_entry, " chain ", chain_index,"_bepipred3.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```


### discotope-3.0
```{r}
pdbs_path = "output/pdb/"

for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_file <- paste0(pdbs_path,pdb_entry,".pdb") 
  
  if (file.exists(pdb_file)) {
    print(pdb_file)
    pdb_object <- read.pdb(file = pdb_file, rm.alt = FALSE)
  } else {
    pdb_object <- read.pdb(pdb_entry, rm.alt = FALSE)
  }
  
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  
  for (chain_index in unique(data_DiscoTope_subset$chain)) {
  data_DiscoTope_subset_chain <- data_DiscoTope_subset %>% filter(chain == chain_index)
  
  min_score <- min(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  median_score <- median(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  max_score <- max(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  
  
  low_to_median <- colorRampPalette(c("blue", "gray95"))(10)
  median_to_high <- colorRampPalette(c("gray95", "red"))(10)
  
  breaks <- seq(min_score, max_score, length.out = 21)  # 20 intervals
  colors <- c(low_to_median, median_to_high)  # 20 colors
  
  color_map <- data.frame(
    ProbabilityBin = cut(seq(min_score, max_score, length.out = 20), 
                         breaks = breaks, include.lowest = TRUE),
    color = colors
  ) %>%
  rbind(c(NA,"black"))
  
  predictions_df <- data_DiscoTope_subset %>%
    mutate(ProbabilityBin = cut(calibrated_score, 
                                breaks = breaks, include.lowest = TRUE)
    ) %>%
    group_by(ProbabilityBin) %>% 
    summarize(resno_list = list(paste0(resno,":",chain_index)), .groups = "drop") %>%
    mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " "))) %>%
    left_join(color_map, by = "ProbabilityBin")
  
  view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index,predictions_df = predictions_df)
  
  #print(view)
  
  saveWidget(view, file = paste0("output/DiscoTope/snapshot/",pdb_entry, " chain ", chain_index,"_discotope3.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```

# Variability
## AmiA
```{r}
gb_lines_AmiA <- readLines("data/6. AmiA.gb")
head(gb_lines_AmiA, 20)
```

```{r}
origin_indices <- grep("^ORIGIN", gb_lines_AmiA)
end_indices <- grep("^//", gb_lines_AmiA)
header_prefix <- "AmiA"

fasta_content <- list()

for(i in 1:length(origin_indices)) {
  start_index <- origin_indices[i] + 2
  end_index <- end_indices[i] - 1
  
  protein_sequence_raw <- gb_lines_AmiA[start_index:end_index] 
  
  protein_sequence <- protein_sequence_raw %>%
    gsub("[0-9]+", "", .) %>%
    gsub("\\s+", "", .) %>%
    paste(collapse = "")
  
  header_line <- gb_lines_AmiA[grep("^TITLE", gb_lines_AmiA)[i]]
  header <- sub("^TITLE\\s+", "", header_line)

  fasta_content[[i]] <- paste0(">",header_prefix,"_", header, "\n", protein_sequence)
}

writeLines(unlist(fasta_content), "data/fasta//AmiA_sequences.fasta")
```

### multiple sequence alignment
```{r}
AmiA_sequences <- readAAStringSet("data/fasta//AmiA_sequences.fasta")
AmiA_sequences
```

```{r}
AmiA_alignment <- msa(AmiA_sequences)
```

```{r}
AmiA_alignment
```


```{r}
print(AmiA_alignment, show="complete")
```

#### ggmsa

### alignment matrix
```{r}
AmiA_alignment_matrix <- as.matrix(AmiA_alignment)

AmiA_variability <- apply(AmiA_alignment_matrix, 2, function(col) {
  length(unique(col))
})

AmiA_variability

highly_variable_positions <- which(AmiA_variability == 2)

highly_variable_positions
length(highly_variable_positions)
```

```{r}
dim(AmiA_alignment_matrix)
```

#### ggseqlogo
```{r, fig.width= 50}
#AmiA_alignment_matrix

seq_vector <- apply(AmiA_alignment_matrix, 1, paste, collapse ="")

names(seq_vector) <- paste0("Seq_", seq_along(seq_vector))

ggseqlogo(seq_vector, method = "bits") + 
  theme_minimal() +
  ggtitle("Sequence Logo of AmiA Alignment")
```

### data_variability
```{r}
consensus_seq <- unlist(strsplit(msaConsensusSequence(AmiA_alignment),""))

variability_counts <- sapply(seq_len(ncol(AmiA_alignment_matrix)), function(i) {
  sum(AmiA_alignment_matrix[, i] != consensus_seq[i] & AmiA_alignment_matrix[, i] != "-")
})

#mutation_counts

total_sequences <- nrow(AmiA_alignment)

variability_frequencies <- variability_counts / total_sequences

data_variability_AmiA <- data.frame(
  Position = 1:length(variability_counts),
  Consensus_residue = consensus_seq,
  Variability_counts = variability_counts,
  Variability_frequency = variability_frequencies
)

data_variability_AmiA

```

```{r}
data_variability_AmiA %>%
  ggplot(aes(x=Position, y=Variability_counts)) +
  geom_col(width = 1) +
  theme_minimal()
```


### consensus matrix
```{r}
AmiA_consensus_matrix <- consensusMatrix(AmiA_alignment)
AmiA_consensus_matrix
```

```{r}
msaPrettyPrint(AmiA_alignment, output="asis", y=c(164, 213), showNames="none", showLogo="none", askForOverwrite=FALSE)
```

## (WIP) matching AmiA consensus to to epitope predictions
AmiA corresponds to pdb entries 8A42, 8QM0
```{r}
data.frame(Consensus_Residue = unlist(strsplit(msaConsensusSequence(AmiA_alignment), "")),
           Position = seq_along(unlist(strsplit(msaConsensusSequence(AmiA_alignment), ""))))
```


```{r}
pdb_object <- read.pdb("8A42")

required_adjustment <- map.seqposno.to.resno(sequence_AA1 = msaConsensusSequence(AmiA_alignment), pdb_object = pdb_object, verbose = TRUE)
required_adjustment
```

```{r}
pdb_object$atom %>%
  filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
  ggplot(aes(x=resno, y=chain, col = resid)) +
  geom_vline(xintercept=34) +
  geom_vline(xintercept=50) +
  geom_vline(xintercept=100) +
  geom_point() +
  geom_point(data = data_variability_AmiA ,
             aes(x=Position, col=aa123(Consensus_residue), y="FASTA")
             ) +
  scale_x_continuous(limits = c(NA,50)) +
  #scale_y_discrete(expand = c(0,0)) +
  theme_bw() 
```


#### mutations_data_with_resno
```{r}
data_variability_AmiA

pdb_entry <- "8A42"

pdb_object <- read.pdb(pdb_entry,rm.alt=FALSE)

required_adjustment <- map.seqposno.to.resno(sequence_AA1 = msaConsensusSequence(AmiA_alignment), pdb_object = pdb_object)

data_variability_AmiA_with_resno_8A42 <- create.resmap.bepipred(pdb_object$atom, 
                                 data_variability_AmiA,
                                 required_adjustment)

data_variability_AmiA_with_resno_8A42


```

```{r}
data_predictions_with_variability <- data_predictions %>%
  filter(PDB == "8A42") %>%
  left_join(data_variability_AmiA_with_resno_8A42 %>% select(resid, resno, Variability_counts, Variability_frequency),
            by = c("resno", "resid"))

data_predictions_with_variability
```

```{r}
names(data_predictions)
```


```{r, fig.width = 14}
data_predictions_with_variability %>%
  pivot_longer(cols = c(
    `BepiPred-2.0 probability`, `RSA NetsurfP`, `BepiPred-3.0 score`,
    `BepiPred-3.0 linear epitope score`, `DiscoTope-3.0_score`,
    `DiscoTope-3.0_calibrated_score`, `RSA Shrake-Rupley`, `Epigraph_score`, Variability_frequency
  ), names_to = "Score_Type", values_to = "Score_Value") %>%
  ggplot(aes(x=resno)) +
  geom_col(aes(y=Score_Value)) +
  
  geom_text(aes(label = Score_Type, x = (min(resno) + (max(resno) - min(resno))/2)), color = "black",y = Inf, vjust = 1.5, size = 4, family = "Arial", check_overlap = TRUE) +
  
  facet_wrap(~Score_Type, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(strip.text = element_blank())
```

```{r}
data_predictions_with_variability_normalized <- data_predictions_with_variability %>%
  group_by(PDB) %>%
  mutate(
    across(
      c(`BepiPred-3.0 score` , `BepiPred-3.0 linear epitope score`, 
        `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, Epigraph_score, Variability_frequency), 
      rescale
    )
  ) %>%
  ungroup() %>%
  na.omit()

data_predictions_with_variability_normalized
```

```{r}
data_test_wide <- data_predictions_with_variability_normalized %>%
  select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
        `DiscoTope-3.0_calibrated_score`, Epigraph_score, Variability_frequency) %>%
  column_to_rownames("resno") %>%
  na.omit() %>%
  as.data.frame()

pheatmap(t(data_test_wide), 
         cluster_rows = TRUE,         
         cluster_cols = FALSE,         
         main = "8A42", 
         color = colorRampPalette(c("blue", "white", "red"))(50),
         labels_col = "resno",
         clustering_distance_rows = "correlation") 
```

# Alphafold 3 structures
Here we predict the alphafold 3 structures based on the recombinant sequences
```{r}
library(jsonlite)

# Function to parse FASTA file and extract headers and sequences correctly
parse_fasta <- function(fasta_file) {
  if (!file.exists(fasta_file)) {
    stop("Error: FASTA file does not exist at the specified path: ", fasta_file)
  }
  
  fasta_lines <- readLines(fasta_file)  
  sequences <- list()
  
  current_name <- NULL
  current_sequence <- ""
  
  for (line in fasta_lines) {
    if (startsWith(line, ">")) {  
      if (!is.null(current_name)) {
        
        sequences <- append(sequences, list(list(name = current_name, sequence = current_sequence)))
      }
      current_name <- substr(line, 2, nchar(line))  
      current_sequence <- ""  
    } else {
      current_sequence <- paste0(current_sequence, line)  
    }
  }
  

  if (!is.null(current_name)) {
    sequences <- append(sequences, list(list(name = current_name, sequence = current_sequence)))
  }
  
  return(sequences)
}

# Function to generate the correct JSON structure
generate_json <- function(fasta_file) {
  sequences <- parse_fasta(fasta_file)
  
  json_data <- lapply(sequences, function(seq_data) {

    clean_name <- sub("([^(\\s]+)(\\s|\\().*", "\\1", seq_data$name)
    clean_name <- gsub("\\s+", "", clean_name)
    clean_name <- paste(clean_name, "recombinant protein")
    
    list(
      name = as.character(clean_name),  
      modelSeeds = list(),  
      sequences = list(
        list(
          proteinChain = list(
            sequence = as.character(seq_data$sequence),  
            count = 1 
          )
        )
      )
    )
  })
  
  return(json_data)
}


# Function to write the JSON data to a file
write_json_file <- function(json_data, output_file) {
  output_dir <- dirname(output_file)
  
  if (!dir.exists(output_dir)) {
    dir.create(output_dir, recursive = TRUE)
  }
  
  write_json(json_data, output_file, pretty = TRUE, auto_unbox = TRUE) #auto_unbox is very important as otherwise the Alphafoldserver will not except the .json
}



# Define input and output paths
fasta_file <- "data/fasta/sp_recombinant_protein_sequences.fasta"
output_json_file <- "output/AlphaFold3/sp_recombinant_protein_sequences.json"

json_data <- generate_json(fasta_file)
write_json_file(json_data, output_json_file)
```

