---
title: "Streptococcus pneumoniae epitope prediction tools"
author: "Jan Paul van Meenen"
date: "`r Sys.Date()`"
output: html_document
---

#Packages
```{r}
data_management_packages_to_install <- c("data.table", "tidyverse" ,"readxl","purrr","readr","scales","msa")
data_visualisation_packages_to_install <- c("ggplot2","cowplot","pheatmap","viridisLite","ggseqlogo")
structural_packages_to_install <- c("bio3d","NGLVieweR", "htmlwidgets","htmltools")

packages_to_install <- c(data_management_packages_to_install, 
                         data_visualisation_packages_to_install,
                         structural_packages_to_install)

# Check and install packages using a for loop
for (package in packages_to_install) {
  if (!requireNamespace(package, quietly = TRUE)) {
    install.packages(package)
  }
  library(package, character.only = TRUE)
}

BiocManager_packages_to_install <- c("Biostrings","pwalign")

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

for (package in BiocManager_packages_to_install) {
  if (!requireNamespace(package, quietly = TRUE)) {
    BiocManager::install(package)
  }
  library(package, character.only = TRUE)
}
```

#Functions
```{r}
function_files <- list.files(path = "functions", pattern = "\\.R$",full.names = TRUE)

for (function_file in function_files) {
  source(function_file)
}
```

#File paths
```{r}
pdbs_path = "data/pdb/"
```


# BepiPred-2.0
## split fasta into parts
Bepi-2.0 only allows for 50 proteins at a time.
```{r}
fasta_lines <- readLines("data/fasta/pdb_protein_sequences.fasta")

# Group lines into proteins based on `>...` headers
protein_entries <- split(fasta_lines, cumsum(grepl("^>", fasta_lines)))

# Split protein entries into chunks of 50
protein_chunks <- split(protein_entries, ceiling(seq_along(protein_entries) / 50))

for (i in seq_along(protein_chunks)) {
  # Combine lines for each chunk
  chunk_lines <- unlist(protein_chunks[[i]])
  
  # Write the chunk to a new file
  writeLines(chunk_lines, paste0("data/fasta/pdb_protein_sequences_part_", i, ".fasta"))
}

```

https://services.healthtech.dtu.dk/services/BepiPred-2.0/ 
```{r}
data_BepiPred_2 <- rbind(
    read_csv("output/BepiPred2.0/BepiPred_part1_summary_67C1B456000FBFBBD072AA59.csv"),
    read_csv("output/BepiPred2.0/BepiPred_part2_summary_67C1B4C8000FC0D59FA5D3FB.csv")
    ) 

data_BepiPred_2 %>%
  filter(Entry == "3CX3") %>%
  mutate(AminoAcid = aa123(AminoAcid))
```

## Generalized
### get valid all pdb entires and files
```{r}
pdb_files_list <- list.files(pdbs_path, pattern = "\\.pdb$") %>% sub("\\.pdb$", "", .)

pdb_files_list
```

## BepiPred-2.0 with resno
```{r}
data_BepiPred_2_with_resno <- NULL

for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  required_adjustment <- map.seqresno.to.resno(pdb_object)
  #print(required_adjustment)
  
  resmap <- create.resmap(pdb_object$atom, 
                                           data_BepiPred_2 %>% filter(Entry == pdb_entry),
                                           required_adjustment)
  
  if (is.null(data_BepiPred_2_with_resno)) {
    data_BepiPred_2_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_BepiPred_2_with_resno <- rbind(data_BepiPred_2_with_resno, resmap)
  }
}

data_BepiPred_2_with_resno
```




# BepiPred-3.0 results
https://services.healthtech.dtu.dk/services/BepiPred-3.0/
```{r}
data_BepiPred_3 <- rbind(
    read_csv("output/BepiPred3.0/raw_output_part1_0228.csv"),
    read_csv("output/BepiPred3.0/raw_output_part2_0228.csv")
    ) %>%
  #mutate(Residue = aa123(Residue)) %>%
  group_by(Accession) %>%
  mutate(Position = row_number())
```

## BepiPred-3.0 with resno
```{r}
data_BepiPred_3_with_resno <- NULL

for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  required_adjustment <- map.seqresno.to.resno(pdb_object)
  #print(required_adjustment)
  
  resmap <- create.resmap(pdb_object$atom, 
                                           data_BepiPred_3 %>% filter(Accession == pdb_entry),
                                           required_adjustment)
  
  if (is.null(data_BepiPred_3_with_resno)) {
    data_BepiPred_3_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_BepiPred_3_with_resno <- rbind(data_BepiPred_3_with_resno, resmap)
  }
}

data_BepiPred_3_with_resno
```

# DiscoTope-3.0
https://services.healthtech.dtu.dk/services/DiscoTope-3.0/
```{r}
csv_files <- list.files("output/DiscoTope", pattern = "*.csv", full.names = TRUE)

data_DiscoTope <- NULL

for (csv_file in csv_files) {
  temp_data <- read_csv(csv_file, show_col_types = FALSE) 
  
  if (is.null(data_DiscoTope)) {
    data_DiscoTope <- temp_data
  } else {
    data_DiscoTope <- rbind(data_DiscoTope, temp_data)
  }
}

data_DiscoTope <- data_DiscoTope %>%
  mutate(pdb = gsub("_.*", "", pdb),
         resno = res_id)

data_DiscoTope
```

# Epigraph
http://epigraph.kaist.ac.kr/
```{r}
epigraph_zip_files <- list.files(path = "output/EpiGraph/individual_files/", pattern = "\\.zip$")

for (zip_file in epigraph_zip_files) {
  unzip(zipfile = paste0("output/EpiGraph/individual_files/",zip_file), exdir = "output/EpiGraph/individual_files/")
}

epigraph_csv_files <- list.files(path = "output/EpiGraph/individual_files/", pattern = "\\.csv$")
epigraph_csv_files
```

```{r}
data_Epigraph <- NULL

for (csv_file in epigraph_csv_files) {
  temp_data <- read_csv(paste0("output/EpiGraph/individual_files/",csv_file), show_col_types = FALSE) 
  print(csv_file)
  
  if (is.null(data_Epigraph)) {
    data_Epigraph <- temp_data
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_Epigraph <- rbind(data_Epigraph, temp_data)
  }
}

data_Epigraph <- data_Epigraph %>% 
  separate(Residue, into = c("chain", "resid", "resno"), sep = ":", convert = TRUE) %>%
  mutate(PDB = toupper(PDB))

data_Epigraph

write.csv(data_Epigraph,"output/EpiGraph/data_Epigraph.csv", row.names=FALSE)
```

```{r}
data_Epigraph_subset <- data_Epigraph %>% filter(PDB == "3ZPP")

plot_epigraph <- data_Epigraph_subset %>% 
    ggplot(aes(x=resno, y=Score, fill=Score)) +
    geom_col(width = 1,) +
    scale_fill_gradient2(low = "blue", mid = "gray95", high = "red") +  
    theme_minimal() + 
    theme(
      axis.text.y = element_blank(),   
      axis.title.y = element_blank(),  
      axis.ticks.y = element_blank()
    ) +
    labs(fill = "EpiGraph prediction")

plot_epigraph
```

#GraphBepi
Lookup the longest chain per protein
```{r}
data_DiscoTope %>%
  filter(residue %in% names(AMINO_ACID_CODE)[1:20]) %>%
  select(pdb, chain, length) %>%
  pivot_wider(names_from = chain, values_from = length) %>%
  mutate(across(where(is.list), ~sapply(., function(x) ifelse(is.null(x), NA, length(x)))))
```

Visually check all chains per protein
```{r}
for (pdb_entry in unique(data_DiscoTope$pdb)) {
  print(pdb_entry)
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  
  plot <- data_DiscoTope_subset %>%
    ggplot(aes(x=res_id, colour = residue)) +
    geom_point(aes(y=0)) +
    facet_wrap(~chain,ncol = 1) +
    labs(title=paste0(pdb_entry)) +
    theme_minimal()
  
  
  print(plot)
}
```

## read in the results
```{r}
graphbepi_txt_files <- list.files(path = "output/GraphBepi/individual_files/", pattern = "\\.txt$")

graphbepi_txt_files
```

```{r}
data_GraphBepi <- NULL

for (txt_file in graphbepi_txt_files) {
  print(txt_file)
  PDB <- sub("_.*", "", txt_file)
  chain <- sub("^[^_]+_([A-Za-z0-9])_.*", "\\1", txt_file)

  temp_data <- read.table(paste0("output/GraphBepi/individual_files/",txt_file), header = TRUE, skip=1)
  
  temp_data <- temp_data %>% 
    mutate(
      PDB = PDB,
      chain = chain,
      resid = aa123(AA),
      `GraphBepi probability` = Prob,
      `GraphBepi prediction` = Pred,
     Position = row_number()
    ) %>%
    select(PDB, chain, resid, Position, `GraphBepi probability`, `GraphBepi prediction`)
  
  if (is.null(data_GraphBepi)) {
    data_GraphBepi <- temp_data
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_GraphBepi <- rbind(data_GraphBepi, temp_data)
  }
}

data_GraphBepi
```

```{r}
data_GraphBepi_with_resno <- NULL

for (pdb_entry in unique(data_GraphBepi$PDB)) {
  print(pdb_entry)
  
  pdb_file <- paste0(pdbs_path,pdb_entry,".pdb") 
  
  if (file.exists(pdb_file)) {
    print(pdb_file)
    pdb_object <- read.pdb(file = pdb_file, rm.alt = FALSE)
  } else {
    pdb_object <- read.pdb(pdb_entry, rm.alt = FALSE)
  }
  
  graphbepi_subset <- data_GraphBepi %>% filter(PDB == pdb_entry)
  graphbepi_sequence <- graphbepi_subset %>% pull(resid) %>% aa321() %>% str_c(collapse = "")
  graphbepi_chain <- graphbepi_subset %>% pull(chain) %>% unique()
  
  required_adjustment <- map.seqposno.to.resno(sequence_AA1 = graphbepi_sequence, pdb_object)
  print(required_adjustment)
  
  resmap <- create.resmap(pdb_object$atom %>% filter(chain == graphbepi_chain), 
                                           graphbepi_subset,
                                           required_adjustment, join_by = c("resno","chain","resid")
                                   )
  
  if (is.null(data_GraphBepi_with_resno)) {
    data_GraphBepi_with_resno <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_GraphBepi_with_resno <- rbind(data_GraphBepi_with_resno, resmap)
  }
}

data_GraphBepi_with_resno

```


## all predictions, one chain
```{}
data_predictions_1chain <- data_GraphBepi_with_resno %>% 
  
  select(!Position) %>%
  inner_join(data_predictions,
            by=c("PDB","chain","resno","resid"))

data_predictions_1chain
```

```{}
for (pdb_entry in unique(data_predictions_1chain$PDB)) {
  data_test_wide <- data_predictions_1chain %>%
    filter(PDB == pdb_entry) %>%
    group_by(PDB, chain) %>%
    mutate(
      across(
        c(`BepiPred-2.0 probability`,
          `BepiPred-3.0 score` , `BepiPred-3.0 linear epitope score`, 
          `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, Epigraph_score, `GraphBepi probability`), 
        rescale
      )
    ) %>%
    ungroup() %>%
    select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
           `DiscoTope-3.0_calibrated_score`, Epigraph_score, `GraphBepi probability`) %>%
    column_to_rownames("resno") %>%
    as.data.frame()
  
  pheatmap(t(data_test_wide), 
           cluster_rows = TRUE,         
           cluster_cols = FALSE,         
           main = pdb_entry, 
           color = colorRampPalette(c("blue", "gray95", "red"))(50),
           labels_col = "resno",
           clustering_distance_rows = "correlation")  
}
```

### PCA subset
```{}
data_predictions_1chain

score_columns <- c("GraphBepi probability","Epigraph_score" ,"DiscoTope-3.0_calibrated_score",
                   "BepiPred-2.0 probability", 
                   "BepiPred-3.0 score")

score_columns
```

```{}
scaled_scores <- data_predictions_1chain %>%
  select(all_of(score_columns)) %>%
  scale()
  
pca_result <- prcomp(t(scaled_scores),center = TRUE, scale. = TRUE)

pca_scores <- as.data.frame(pca_result$x) %>%
  rownames_to_column(var = "score_type")
```

```{}
library(ggrepel)

pca_scores <- as.data.frame(pca_result$x) %>%
  rownames_to_column(var = "score_type")

# Scatter plot of PC1 vs PC2
ggplot(pca_scores, aes(x = PC1, y = PC2, label = score_type)) +
  geom_point(alpha = 0.7, color = "blue") +
  geom_text_repel() +
  labs(title = "PCA of concatenated epitope prediction scores across different tools", x = "PC1", y = "PC2") +
  theme_minimal()
```



# Plot all together
## create joined predictions df
```{r}
data_predictions_sequencebased <- full_join(
  data_BepiPred_2_with_resno %>%
    mutate(PDB = Entry,
           `BepiPred-2.0 probability` = EpitopeProbability,
           `RSA NetsurfP` = RelativeSurfaceAccessilibity) %>%
    select(PDB, chain, resno,resid,`BepiPred-2.0 probability`,`RSA NetsurfP`) %>%
    filter(resid %in% toupper(AMINO_ACID_CODE))
  ,
  
  data_BepiPred_3_with_resno %>%
    mutate(PDB = Accession) %>%
    select(PDB, chain, resno, resid, `BepiPred-3.0 score`,`BepiPred-3.0 linear epitope score`) %>%
    filter(resid %in% toupper(AMINO_ACID_CODE))
  ,
  by = c("PDB","chain","resno","resid")
)

data_predictions_sequencebased

data_predictions_sequencebased %>%
  filter(if_any(everything(), is.na))
```

```{r}
data_predictions_structurebased <- full_join(
  data_Epigraph %>%
    mutate(Epigraph_score = Score) %>%
    select(PDB, chain, resno, resid, Epigraph_score)
  ,
  data_DiscoTope %>%
    mutate(PDB = pdb,
           resid = aa123(residue),
           `RSA Shrake-Rupley` = rsa,
           `DiscoTope-3.0_calibrated_score` = calibrated_score
    ) %>%
    select(PDB, chain, resno, resid, `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, `RSA Shrake-Rupley`)
  , 
  by = c("PDB","resno","chain","resid")
)
data_predictions_structurebased

data_predictions_structurebased %>%
  filter(if_any(everything(), is.na))

data_predictions_structurebased <- data_predictions_structurebased %>%
  na.omit()
  
data_predictions_structurebased %>%
  filter(if_any(everything(), is.na))
```


```{r}
data_predictions <- full_join(
  data_predictions_sequencebased,
  data_predictions_structurebased,
  by = c("PDB", "chain", "resno","resid")
  )

data_predictions %>%
  filter(if_any(everything(), is.na))

data_predictions <- data_predictions %>%
  na.omit() 

data_predictions
```

### get the protein names 
```{r}
data_predictions <- data_predictions %>%
  left_join(
    data_structural %>%
      separate_rows(`PDB entry`, sep = ", ") %>%
      mutate(PDB = `PDB entry`,
             proteinclass = `Protein class`,
             proteinname = `Proteins (TIGR4) name`) %>%
      select(proteinclass, proteinname, PDB)
  )

data_predictions
```

## normalize
```{r}
data_predictions_normalized <- data_predictions %>%
  group_by(PDB, chain) %>%
  mutate(
    across(
      c(#`BepiPred-2.0 probability`,
        `BepiPred-3.0 score` , `BepiPred-3.0 linear epitope score`, 
        `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, Epigraph_score), 
      rescale
    )
  ) %>%
  ungroup()

data_predictions_normalized
```

###track
```{r}
track_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score"
)

for (pdb_entry in unique(data_predictions_normalized$PDB)) {
  data_predictions_subset <- data_predictions %>%
    filter(PDB == pdb_entry) 
  
  proteinname <- unique(data_predictions_subset$proteinname)
  
  for (chain_index in unique(data_predictions_subset$chain)) {
    
    data_predictions_subset_long <- data_predictions_subset %>%
      filter(chain == chain_index) %>%
      pivot_longer(cols = track_list, 
                   names_to = "Score_Type", values_to = "Score_Value") %>%
      group_by(Score_Type) %>%
      mutate(normalized_score = case_when(
        Score_Type == "BepiPred-2.0 probability" ~ Score_Value,
        TRUE ~ rescale(Score_Value, to = c(0, 1))
        )
      )
    
    figure <- data_predictions_subset_long %>%
      ggplot(aes(x=resno, y=Score_Value, fill = normalized_score)) +
      geom_col() +
      
      facet_wrap(~Score_Type, ncol = 1, scales = "free_y") +
      
      scale_fill_gradient2(low = "blue", mid = "gray95", high = "red", 
                           midpoint = 0.5) +  # Now uses median per facet
      
      geom_text(aes(label = Score_Type, 
                    x = (min(resno) + (max(resno) - min(resno)) / 2)), 
                color = "black", y = Inf, vjust = 1.5, size = 4, family = "Arial", 
                check_overlap = TRUE) +
      
      labs(title = paste0(pdb_entry, " chain ", chain_index, " (",proteinname,")")) +
      
      theme_minimal() +
      theme(
        strip.text = element_blank(),
        axis.ticks = element_blank(),  
        axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      )
    
    print(figure)
    
    ggsave(filename = paste0("output/figures/tracks/", pdb_entry, "_", chain_index, "_track.png"), 
           plot = figure, 
           width = 14, height = 10, dpi = 300,
           bg="white")
  }
}
```


###heatmap
```{r, fig.width = 14}
for (pdb_entry in unique(data_predictions_normalized$PDB)) {
  data_predictions_normalised_subset <- data_predictions_normalized %>%
    filter(PDB == pdb_entry) 
  
  proteinname <- unique(data_predictions_normalised_subset$proteinname)
  
  for (chain_index in unique(data_predictions_normalised_subset$chain)) {
    
    data_predictions_normalised_subset_wide <- data_predictions_normalised_subset %>%
      filter(chain == chain_index) %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score) %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    figure <- pheatmap(t(data_predictions_normalised_subset_wide), 
                       cluster_rows = TRUE,         
                       cluster_cols = FALSE,         
                       main = paste0(pdb_entry, " chain ", chain_index, " (",proteinname,")"), 
                       color = colorRampPalette(c("blue", "gray95", "red"))(50),
                       labels_col = "resno",
                       clustering_distance_rows = "correlation") 
    
    print(figure)
    
    ggsave(filename = paste0("output/figures/heatmap/", pdb_entry, "_", chain_index, "_heatmap.png"), 
           plot = figure, 
           width = 14, height = 10, dpi = 300,
           bg="white")
  }
}
```


## 3D
### bepipred-3.0
```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_BepiPred_3_with_resno_subset <- data_BepiPred_3_with_resno %>% filter(Accession == pdb_entry) %>% na.omit()
  
  for (chain_index in unique(data_BepiPred_3_with_resno_subset$chain)) {
    data_BepiPred_3_with_resno_subset_chain <- data_BepiPred_3_with_resno_subset %>% filter(chain == chain_index)
    
    min_score <- min(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, na.rm = TRUE)
    median_score <- median(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, na.rm = TRUE)
    max_score <- max(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, na.rm = TRUE)
    
    
    low_to_median <- colorRampPalette(c("blue", "gray95"))(10)
    median_to_high <- colorRampPalette(c("gray95", "red"))(10)
    
    breaks <- seq(min_score, max_score, length.out = 21)  # 20 intervals
    colors <- c(low_to_median, median_to_high)  # 20 colors
    
    color_map <- data.frame(
      ProbabilityBin = cut(seq(min_score, max_score, length.out = 20), 
                           breaks = breaks, include.lowest = TRUE),
      color = colors
    ) %>%
      rbind(c(NA,"black"))
    
    predictions_df <- data_BepiPred_3_with_resno_subset_chain %>%
      mutate(ProbabilityBin = cut(`BepiPred-3.0 score`, 
                                  breaks = breaks, include.lowest = TRUE)
      ) %>%
      group_by(ProbabilityBin) %>% 
      summarize(resno_list = list(paste0(resno,":",chain_index)), .groups = "drop") %>%
      mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " "))) %>%
      left_join(color_map, by = "ProbabilityBin")
    
    view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index, sele_color_df = predictions_df)
    
    #print(view)
    
    saveWidget(view, file = paste0("output/BepiPred3.0/snapshot/score/",pdb_entry, "_", chain_index,"_bepipred3.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```


### discotope-3.0
```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  
  for (chain_index in unique(data_DiscoTope_subset$chain)) {
  data_DiscoTope_subset_chain <- data_DiscoTope_subset %>% filter(chain == chain_index)
  
  min_score <- min(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  median_score <- median(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  max_score <- max(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  
  
  low_to_median <- colorRampPalette(c("blue", "gray95"))(10)
  median_to_high <- colorRampPalette(c("gray95", "red"))(10)
  
  breaks <- seq(min_score, max_score, length.out = 21)  # 20 intervals
  colors <- c(low_to_median, median_to_high)  # 20 colors
  
  color_map <- data.frame(
    ProbabilityBin = cut(seq(min_score, max_score, length.out = 20), 
                         breaks = breaks, include.lowest = TRUE),
    color = colors
  ) %>%
  rbind(c(NA,"black"))
  
  predictions_df <- data_DiscoTope_subset %>%
    mutate(ProbabilityBin = cut(calibrated_score, 
                                breaks = breaks, include.lowest = TRUE)
    ) %>%
    group_by(ProbabilityBin) %>% 
    summarize(resno_list = list(paste0(resno,":",chain_index)), .groups = "drop") %>%
    mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " "))) %>%
    left_join(color_map, by = "ProbabilityBin")
  
  view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index,sele_color_df = predictions_df)
  
  #print(view)
  
  saveWidget(view, file = paste0("output/DiscoTope/snapshot/score/",pdb_entry, "_", chain_index,"_discotope3.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```

### colour epitopes by resno
Colouring predicted epitopes by resno could help identify discontinous epitopes
```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_BepiPred_3_with_resno_subset <- data_BepiPred_3_with_resno %>% filter(Accession == pdb_entry) %>% na.omit()
  
  quantile_threshold <- 0.85  #0.8 == top 20%
  
  for (chain_index in unique(data_BepiPred_3_with_resno_subset$chain)) {
    data_BepiPred_3_with_resno_subset_chain <- data_BepiPred_3_with_resno_subset %>% filter(chain == chain_index)
    
    score_threshold <- quantile(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, 
                                probs = quantile_threshold, na.rm = TRUE)
    
    predictions_df <- data_BepiPred_3_with_resno_subset_chain %>%
      mutate(color = ifelse(`BepiPred-3.0 score` >= score_threshold, "residueindex", "white")) %>%
      group_by(color) %>% 
      summarize(resno_list = list(paste0(resno, ":", chain_index)), .groups = "drop") %>%
      mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " ")))
    
    view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index, sele_color_df = predictions_df)
    
    #print(view)
    
    saveWidget(view, file = paste0("output/BepiPred3.0/snapshot/conformational/",pdb_entry, "_", chain_index,"_resid_bepipred3.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```

```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  
  quantile_threshold <- 0.85  #0.8 == top 20%
  
  for (chain_index in unique(data_DiscoTope_subset$chain)) {
    data_DiscoTope_subset_chain <- data_DiscoTope_subset %>% filter(chain == chain_index)
    
    score_threshold <- quantile(data_DiscoTope_subset_chain$`DiscoTope-3.0_score`, 
                                probs = quantile_threshold, na.rm = TRUE)
    
    predictions_df <- data_DiscoTope_subset_chain %>%
      mutate(color = ifelse(`DiscoTope-3.0_score` >= score_threshold, "residueindex", "white")) %>%
      group_by(color) %>% 
      summarize(resno_list = list(paste0(resno, ":", chain_index)), .groups = "drop") %>%
      mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " ")))
    
    view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index, sele_color_df = predictions_df)
    
    print(view)
    
    saveWidget(view, file = paste0("output/DiscoTope/snapshot/conformational/",pdb_entry, "_", chain_index,"_resid_discotope.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```


## Longest chain
### Getting the longest chain
```{r}
data_predictions_longestchain <- inner_join(
  data_predictions_structurebased %>%
    add_count(PDB, chain, name = "chain_length") %>%  # Add chain length count
    arrange(PDB,desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
    group_by(PDB) %>%
    filter(chain_length == max(chain_length)) %>%  
    filter(chain == min(chain)) %>% 
    ungroup() %>%  
    select(-chain_length),
  data_predictions_sequencebased,
  by = c("PDB", "chain", "resno","resid")
  ) %>%
  
  left_join(
    data_structural %>%
      separate_rows(`PDB entry`, sep = ", ") %>%
      mutate(PDB = `PDB entry`,
             proteinclass = `Protein class`,
             proteinname = `Proteins (TIGR4) name`) %>%
      select(proteinclass, proteinname, PDB)
  )

data_predictions_longestchain
```

### PCA
```{r}
score_columns <- c("Epigraph_score", "DiscoTope-3.0_calibrated_score",
                   "BepiPred-2.0 probability", 
                   "BepiPred-3.0 score")

score_columns
```
```{r}
scaled_scores <- data_predictions_longestchain %>%
  select(all_of(score_columns)) %>%
  scale()
  
pca_result <- prcomp(t(scaled_scores),center = TRUE, scale. = TRUE)
```


```{r}
explained_variance <- data.frame(
  PC = paste0("PC", 1:length(pca_result$sdev)),
  Variance = (pca_result$sdev^2) / sum(pca_result$sdev^2) * 100
)

# Scree plot
ggplot(explained_variance, aes(x = PC, y = Variance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Scree Plot", x = "Principal Components", y = "Variance Explained (%)") +
  theme_minimal()
```

```{r}
library(ggrepel)

pca_scores <- as.data.frame(pca_result$x) %>%
  rownames_to_column(var = "score_type")

# Scatter plot of PC1 vs PC2
ggplot(pca_scores, aes(x = PC1, y = PC2, label = score_type)) +
  geom_point(alpha = 0.7, color = "blue") +
  geom_text_repel() +
  labs(title = "PCA of concatenated epitope prediction scores across different tools", x = "PC1", y = "PC2") +
  theme_minimal()
```


# Variability
```{r}
gbs_path <- "data/gb/"

gb_list <- sub(
  "\\.gb$", "", sub(
    "^\\d+\\.\\s*", "", list.files(gbs_path, pattern = "\\.gb$")
  )
)

gb_list
```
```{r}
for (gb in gb_list) {
  gb_file_path <- list.files(gbs_path,pattern=paste0(gb,"\\.gb$"), full.names = TRUE)
  gb2fasta(gb_file_path,output_dir="data/fasta/msa/")
}

```

### multiple sequence alignment
```{r}
AmiA_sequences <- readAAStringSet("data/fasta/msa/AmiA_msa_sequences.fasta")
AmiA_sequences
```

```{r}
AmiA_alignment <- msa(AmiA_sequences)
```

```{r}
AmiA_alignment
```


```{r}
print(AmiA_alignment, show="complete")
```

### alignment matrix
```{r}
AmiA_alignment_matrix <- as.matrix(AmiA_alignment)

AmiA_variability <- apply(AmiA_alignment_matrix, 2, function(col) {
  length(unique(col))
})

AmiA_variability

highly_variable_positions <- which(AmiA_variability == 2)

highly_variable_positions
length(highly_variable_positions)
```

```{r}
dim(AmiA_alignment_matrix)
```

#### ggseqlogo
```{r, fig.width= 50}
#AmiA_alignment_matrix

seq_vector <- apply(AmiA_alignment_matrix, 1, paste, collapse ="")

names(seq_vector) <- paste0("Seq_", seq_along(seq_vector))

ggseqlogo(seq_vector, method = "bits") + 
  theme_minimal() +
  ggtitle("Sequence Logo of AmiA Alignment")
```

### data_variability
```{r}
consensus_seq <- unlist(strsplit(msaConsensusSequence(AmiA_alignment),""))

variability_counts <- sapply(seq_len(ncol(AmiA_alignment_matrix)), function(i) {
  sum(AmiA_alignment_matrix[, i] != consensus_seq[i] & AmiA_alignment_matrix[, i] != "-")
})

#mutation_counts

total_sequences <- nrow(AmiA_alignment)

variability_frequencies <- variability_counts / total_sequences

data_variability_AmiA <- data.frame(
  Position = 1:length(variability_counts),
  Consensus_residue = consensus_seq,
  Variability_counts = variability_counts,
  Variability_frequency = variability_frequencies
)

data_variability_AmiA

```

```{r}
data_variability_AmiA %>%
  ggplot(aes(x=Position, y=Variability_counts)) +
  geom_col(width = 1) +
  theme_minimal()
```


### consensus matrix
```{r}
AmiA_consensus_matrix <- consensusMatrix(AmiA_alignment)
AmiA_consensus_matrix
```

```{r}
msaPrettyPrint(AmiA_alignment, output="asis", y=c(164, 213), showNames="none", showLogo="none", askForOverwrite=FALSE)
```

## (WIP) matching AmiA consensus to to epitope predictions
AmiA corresponds to pdb entries 8A42, 8QM0
```{r}
data.frame(Consensus_Residue = unlist(strsplit(msaConsensusSequence(AmiA_alignment), "")),
           Position = seq_along(unlist(strsplit(msaConsensusSequence(AmiA_alignment), ""))))
```


```{r}
pdb_object <- read.pdb("8A42")

required_adjustment <- map.seqposno.to.resno(sequence_AA1 = msaConsensusSequence(AmiA_alignment), pdb_object = pdb_object, verbose = TRUE)
required_adjustment
```

```{r}
pdb_object$atom %>%
  filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
  ggplot(aes(x=resno, y=chain, col = resid)) +
  geom_vline(xintercept=34) +
  geom_vline(xintercept=50) +
  geom_vline(xintercept=100) +
  geom_point() +
  geom_point(data = data_variability_AmiA ,
             aes(x=Position, col=aa123(Consensus_residue), y="FASTA")
             ) +
  scale_x_continuous(limits = c(NA,50)) +
  #scale_y_discrete(expand = c(0,0)) +
  theme_bw() 
```


#### mutations_data_with_resno
```{r}
short_name <- "AmiA"

protein_name_pdb_entries <- data_structural %>%
  filter(str_detect(`Proteins (TIGR4) name`, short_name)) %>%
  pull(`PDB entry`) %>%
  str_split(pattern = ", ") %>%
  unlist()

protein_name_pdb_entries
```


```{r}
data_variability_AmiA

data_variability_AmiA_with_pdb <- NULL

for (pdb_entry in protein_name_pdb_entries) {
  print(pdb_entry)
  
  pdb_file <- paste0(pdbs_path,pdb_entry,".pdb") 
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  required_adjustment <- map.seqposno.to.resno(sequence_AA1 = msaConsensusSequence(AmiA_alignment), pdb_object = pdb_object)
  
  resmap <- create.resmap(pdb_object$atom, 
                                 data_variability_AmiA,
                                 required_adjustment) %>%
    mutate(PDB = pdb_entry)
  
  if (is.null(data_variability_AmiA_with_pdb)) {
    data_variability_AmiA_with_pdb <- resmap
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_variability_AmiA_with_pdb <- rbind(data_variability_AmiA_with_pdb, resmap)
  }
  
}

data_variability_AmiA_with_pdb

```

```{r}
data_predictions_with_variability <- data_predictions_longestchain %>%
  right_join(data_variability_AmiA_with_pdb %>% select(PDB, chain, resno,resid, Variability_counts, Variability_frequency),
            by = c("PDB","chain","resno", "resid"))

data_predictions_with_variability
```

```{r, fig.width = 14}
pivot_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score", 
  "Variability_frequency"
)

for (pdb_entry in unique(data_predictions_with_variability$PDB)) {
  data_subset <- data_predictions_with_variability %>% filter(PDB == pdb_entry) %>%
    select(proteinname,PDB,chain,resno,pivot_list) %>%
  pivot_longer(cols = pivot_list, 
  names_to = "Score_Type", values_to = "Score_Value") %>%
    group_by(Score_Type) %>%
    mutate(normalized_score = case_when(
        Score_Type == "BepiPred-2.0 probability" ~ Score_Value,
        TRUE ~ rescale(Score_Value, to = c(0, 1))
        )
    )
  
  plot <- data_subset %>%
    ggplot(aes(x=resno, y=normalized_score, fill = normalized_score)) +
    geom_col() +
    
    facet_wrap(~Score_Type, ncol = 1, scales = "free_y") +
    
    scale_fill_gradient2(low = "blue", mid = "gray95", high = "red", 
                         midpoint = 0.5) +  # Now uses median per facet
    
    geom_text(aes(label = Score_Type, 
                  x = (min(resno) + (max(resno) - min(resno)) / 2)), 
              color = "black", y = Inf, vjust = 1.5, size = 4, family = "Arial", 
              check_overlap = TRUE) +
    
    labs(title = paste0(pdb_entry, " chain ", data_subset$chain[1], " (",data_subset$proteinname[1],")")) +
    
    theme_minimal() +
    theme(
      strip.text = element_blank(),
      axis.ticks = element_blank(),  
      axis.ticks.x = element_blank(),  
      axis.ticks.y = element_blank(),
      axis.text.x = element_blank(),  
      axis.text.y = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
  
  print(plot)
}
```

```{r}
data_predictions_with_variability_normalized <- data_predictions_with_variability %>%
  group_by(PDB) %>%
  mutate(
    across(
      c(Variability_frequency,
        `BepiPred-3.0 score` , `BepiPred-3.0 linear epitope score`, 
        `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, Epigraph_score), 
      rescale
    )
  ) %>%
  ungroup() %>%
  na.omit()

data_predictions_with_variability_normalized
```

```{r}
data_test_wide <- data_predictions_with_variability_normalized %>%
  filter(PDB == "8A42", chain == "A") %>%
  select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
        `DiscoTope-3.0_calibrated_score`, Epigraph_score, Variability_frequency) %>%
  column_to_rownames("resno") %>%
  as.data.frame()

pheatmap(t(data_test_wide), 
         cluster_rows = TRUE,         
         cluster_cols = FALSE,         
         main = "8A42", 
         color = colorRampPalette(c("blue", "white", "red"))(50),
         labels_col = "resno",
         clustering_distance_rows = "correlation") 
```

# Convert AlphaFold3 .cif to .pdb files

```{r}
AlphaFold3_zip_files <- list.files(path = "output/AlphaFold3/", pattern = "\\.zip$")

for (zip_file in AlphaFold3_zip_files) {
  unzip(zipfile = paste0("output/AlphaFold3/",zip_file), exdir = "output/AlphaFold3/individual_files/")
}
```

```{r}
AlphaFold3_folders <- list.dirs(path = "output/AlphaFold3/individual_files", full.names = FALSE, recursive = FALSE)

AlphaFold3_folders
```

## plot per protein
```{r, fig.width=14, warning=FALSE}
for (AlphaFold3_folder in AlphaFold3_folders) {
  folder_path <- paste0("output/AlphaFold3/individual_files/", AlphaFold3_folder)
  cif_file_paths <- list.files(path = folder_path, pattern = "\\.cif$", full.names=TRUE)
  
  atom_df_cif <- NULL
  
  for (cif_file in cif_file_paths) {
    #print(cif_file)
    file_name <- sub("\\.cif$", "", basename(cif_file))
    
    protein_name <- gsub(".*_(.*?)_.*", "\\1", file_name)
    fold_number <- gsub(".*_(\\d+)$", "\\1", file_name)
    
    cif_object <- read.cif(cif_file)
    
    cif_df <- cif_object$atom %>%
      mutate(fold_number = fold_number)
    
    atom_df_cif <- bind_rows(atom_df_cif, cif_df)
  }
  
  atom_df_cif %<>%
    group_by(fold_number) %>%
    mutate(mean_b = mean(b))
  
  #print(atom_df_cif)
  
  plot <- atom_df_cif %>%
    ggplot(aes(x=fold_number, y=b)) +
    geom_violin(fill = "steelblue", alpha = 0.7, color = "black") +  # Violin plot
    geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + # Boxplot inside violin
    #facet_wrap(~fold_number, nrow = 1) +
    labs(title = paste0(protein_name,"\npLDDT Score Distribution"), x = "fold number", y = "pLDDT") +
    theme_minimal() +
    theme(text = element_text(size = 14))
  
  print(plot)
}
```

## plot all proteins with average
```{r}
atom_df_cif_wide <- NULL

for (AlphaFold3_folder in AlphaFold3_folders[2]) {
  folder_path <- paste0("output/AlphaFold3/individual_files/", AlphaFold3_folder)
  cif_file_paths <- list.files(path = folder_path, pattern = "\\.cif$", full.names=TRUE)
  
  atom_df_cif_subset <- NULL
  atom_df_cif_subset_wide <- NULL 
  
  for (cif_file in cif_file_paths) {
    print(cif_file)
    file_name <- sub("\\.cif$", "", basename(cif_file))
    
    protein_name <- gsub(".*_(.*?)_.*", "\\1", file_name)
    fold_number <- gsub(".*_(\\d+)$", "\\1", file_name)
    
    cif_object <- read.cif(cif_file)
    
    cif_df <- cif_object$atom %>%
      mutate(protein = protein_name,
             fold_number = fold_number)
    
    atom_df_cif_subset <- bind_rows(atom_df_cif_subset, cif_df)
  }
  
  # Apply pivot_wider AFTER accumulating all folds for a single protein
  atom_df_cif_subset_wide <- atom_df_cif_subset %>%
    pivot_wider(names_from = fold_number, values_from = b, names_prefix = "Fold_") %>%
    mutate(Fold_average = rowMeans(select(., starts_with("Fold_")), na.rm = TRUE))
  
  # Append to atom_df_cif_wide after ensuring the structure is correct
  atom_df_cif_wide <- bind_rows(atom_df_cif_wide, atom_df_cif_subset_wide)
}

# Check final structure
print(atom_df_cif_wide)

plot <- atom_df_cif_wide %>%
    ggplot(aes(x=protein, y=b)) +
    geom_violin(fill = "steelblue", alpha = 0.7, color = "black") +  # Violin plot
    geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + # Boxplot inside violin
    facet_wrap(~protein, nrow = 1) +
    labs(title = paste0(protein_name,"\npLDDT Score Distribution"), x = "fold number", y = "pLDDT") +
    theme_minimal() +
    theme(text = element_text(size = 14))
  
  print(plot)
```

## read.cif
```{r}
for (AlphaFold3_folder in AlphaFold3_folders) {
  folder_path <- paste0("output/AlphaFold3/individual_files/", AlphaFold3_folder)
  cif_file_paths <- list.files(path = folder_path, pattern = "\\.cif$", full.names=TRUE)
  
  atom_df_cif <- NULL
  
  for (cif_file in cif_file_paths) {
    file_name <- sub("\\.cif$", "", basename(cif_file))
    
    protein_name <- gsub(".*_(.*?)_.*", "\\1", file_name)
    fold_number <- gsub(".*_(\\d+)$", "\\1", file_name)
    
    cif_object <- read.cif.custom(cif_file)
    
    cif_df <- cif_object$atom %>%
      mutate(fold_number = fold_number)
    
    atom_df_cif <- bind_rows(atom_df_cif, cif_df)
  }
  
  highest_fold <- atom_df_cif %>%
    group_by(fold_number) %>%
    summarize(mean_b = mean(b)) %>%
    arrange(desc(mean_b)) %>%
    head(1) %>%
    pull(fold_number)
  
  highest_cif_file <- cif_file_paths[grepl(paste0("_", highest_fold, "\\.cif$"), cif_file_paths)]
  
  print(paste("Selected CIF file: ", highest_cif_file))
  
  # Read the CIF object of the selected file
  highest_cif_object <- read.cif.custom(highest_cif_file)
  
  write.pdb(highest_cif_object, file=paste0("output/AlphaFold3/generated_pdb/",protein_name,"_fold_",highest_fold,".pdb"))
}
```

