#Packages
```{r}
source("requirements/packages.R")
```

#Functions
```{r}
function_files <- list.files(path = "requirements/functions", pattern = "\\.R$",full.names = TRUE)

for (function_file in function_files) {
  source(function_file)
}
```

# Variability
## fas
```{r}
fasfiles_path <- "data/bioedit/fas/"
fasfiles <- list.files(path = fasfiles_path, pattern = "\\.aln.fas", full.names = TRUE)
#fasfiles

aln_protein_names <- sapply(basename(fasfiles), function(x) {
  if (grepl("^\\d+ SP_\\d{4}", x)) {
    # Extract "SP_xxxx" pattern
    sub(".*(SP_\\d{4}).*", "\\1", x)
  } else {
    # Extract first word before underscore
    sub("^\\d+ ([^_]+).*", "\\1", x)
  }
})

aln_protein_names
```

###redoing MSA 
Takes a lot of time
```{}
data_variability <- NULL

for (fasfile in fasfiles) {
  print(basename(fasfile))
  
  protein_name <- sapply(basename(fasfile), function(x) {
    if (grepl("^\\d+ SP_\\d{4}", x)) {
      # Extract "SP_xxxx" pattern
      sub(".*(SP_\\d{4}).*", "\\1", x)
    } else {
      # Extract first word before underscore
      sub("^\\d+ ([^_]+).*", "\\1", x)
    }
  })
  
  tryCatch({
    # Attempt to read the file
    sequences <- readAAStringSet(fasfile)
    print(sequences)
    
    # Perform the alignment
    alignment <- msa(sequences)
    consensus_seq <- unlist(strsplit(msaConsensusSequence(alignment), ""))
    print(consensus_seq)
    
    alignment_matrix <- as.matrix(alignment)
    
    variability <- apply(alignment_matrix, 2, function(col) {
      length(unique(col))
    })
    
    variability_counts <- sapply(seq_len(ncol(alignment_matrix)), function(i) {
      sum(alignment_matrix[, i] != consensus_seq[i] & alignment_matrix[, i] != "-")
    })
    
    total_sequences <- nrow(alignment)
    variability_frequencies <- variability_counts / total_sequences
    
    data_variability_subset <- data.frame(
      protein = protein_name,
      position = 1:total_sequences,
      consensus_res = consensus_seq,
      var_counts = variability_counts,
      Var_freq = variability_frequencies,
      n_seq = total_sequences
    )
    
    data_variability <- rbind(data_variability, data_variability_subset)
    
  }, error = function(e) {
    warning(paste("Skipping file due to error:", fasfile, "\nError message:", e$message))
  })
}

data_variability
```


###skipping MSA
Please note that the input DNA sequences are already properly aligned and without ambiguous residues, i.e. just "ATCG". The chunk below shouldn't be used if the DNA sequences aren't aligned yet: in that case first translating and then running the MSA would be appropriate.
In our case, however, as the DNA sequences are already algned. We can keep the alignment by converting the gaps "-" to ambiguous nucleotides N, which translate to ambigous residue X and then transforming these back to gaps "-".

```{r}
data_variability <- NULL

for (fasfile in fasfiles) {
  print(basename(fasfile))
  
  protein_name <- sapply(basename(fasfile), function(x) {
    if (grepl("^\\d+ SP_\\d{4}", x)) {
      # Extract "SP_xxxx" pattern
      sub(".*(SP_\\d{4}).*", "\\1", x)
    } else {
      # Extract first word before underscore
      sub("^\\d+ ([^_]+).*", "\\1", x)
    }
  })
  
  tryCatch({
    # Attempt to read the file
    DNA_sequences <- readDNAStringSet(fasfile)
    #print(DNA_sequences)
    
    
    masked_DNA_sequences <- chartr("-", "N", DNA_sequences)
    #print(masked_DNA_sequences)
    
    masked_sequences <- translate(masked_DNA_sequences, no.init.codon = TRUE, if.fuzzy.codon = "X")
    #print(masked_sequences)
    
    sequences <- chartr("X", "-", masked_sequences)
    
    # Convert sequences to a character matrix
    alignment_matrix <- as.matrix(sequences)
    #print(alignment_matrix)
    
    # Determine consensus sequence manually
    consensus_seq <- apply(alignment_matrix, 2, function(col) {
      residues <- col[col != "-"]
      unique_residues <- unique(col[col != "-"])  # Ignore gaps
      most_frequent <- names(sort(table(residues), decreasing = TRUE))[1]
      if (is.na(most_frequent)) "-" else most_frequent
    })
    
    variability <- apply(alignment_matrix, 2, function(col) {
      length(unique(col[col != "-"]))
    })
    
    variability_counts <- sapply(seq_len(ncol(alignment_matrix)), function(i) {
      sum(alignment_matrix[, i] != consensus_seq[i] & alignment_matrix[, i] != "-")
    })
    
    total_sequences <- length(sequences)
    sequence_length <- length(consensus_seq)
    variability_frequencies <- variability_counts / total_sequences
    
    #print(paste("Length of protein_name:", length(protein_name)))
    #print(paste("Length of consensus_seq:", length(consensus_seq)))
    #print(paste("Length of variability_counts:", length(variability_counts)))
    #print(paste("Length of variability_frequencies:", length(variability_frequencies)))
    #print(paste("Total number of sequences:", total_sequences))
    
    data_variability_subset <- data.frame(
      protein = rep(protein_name,sequence_length),
      position = 1:sequence_length,
      consensus_res = consensus_seq,
      var_counts = variability_counts,
      var_freq = variability_frequencies,
      n_seq = total_sequences,
      seq_length = sequence_length
    )
    
    data_variability <- rbind(data_variability, data_variability_subset)
    
  }, error = function(e) {
    warning(paste("Skipping file due to error:", fasfile, "\nError message:", e$message))
  })
}

data_variability
```

#missing proteins
```{r}
setdiff(aln_protein_names, unique(data_variability$protein))
```


```{r}
write.csv(data_variability,
          "output/data_variability.csv", 
          row.names=FALSE)
```

# plot
```{r}
for (proteinindex in unique(data_variability$protein)) {
  data_subset <- data_variability %>% filter(protein == proteinindex)
  
  plot <- data_subset %>%
    ggplot(aes(x=position, y=var_counts, fill=var_freq)) +
    geom_col() +
    #scale_x_continuous(limits = c(1, 20), breaks = 1:20) +
    labs(title = paste0(proteinindex,"\nnumber of seqs: ",unique(data_subset$n_seq))) 
    
    print(plot)
}

```

```{r}
data_variability <- NULL

for (fasfile in fasfiles) {
  print(basename(fasfile))
  
  protein_name <- sapply(basename(fasfile), function(x) {
    if (grepl("^\\d+ SP_\\d{4}", x)) {
      # Extract "SP_xxxx" pattern
      sub(".*(SP_\\d{4}).*", "\\1", x)
    } else {
      # Extract first word before underscore
      sub("^\\d+ ([^_]+).*", "\\1", x)
    }
  })
  
  tryCatch({
    # Attempt to read the file
    DNA_sequences <- readDNAStringSet(fasfile)
    #print(DNA_sequences)
    
    
    masked_DNA_sequences <- chartr("-", "N", DNA_sequences)
    #print(masked_DNA_sequences)
    
    masked_sequences <- translate(masked_DNA_sequences, no.init.codon = TRUE, if.fuzzy.codon = "X")
    #print(masked_sequences)
    
    sequences <- chartr("X", "-", masked_sequences)
    
    # Convert sequences to a character matrix
    alignment_matrix <- as.matrix(sequences)
    #print(alignment_matrix)
    
    # Determine consensus sequence manually
    consensus_seq <- apply(alignment_matrix, 2, function(col) {
      residues <- col[col != "-"]
      unique_residues <- unique(col[col != "-"])  # Ignore gaps
      most_frequent <- names(sort(table(residues), decreasing = TRUE))[1]
      if (is.na(most_frequent)) "-" else most_frequent
    })
    
    variability <- apply(alignment_matrix, 2, function(col) {
      length(unique(col[col != "-"]))
    })
    
    variability_counts <- sapply(seq_len(ncol(alignment_matrix)), function(i) {
      sum(alignment_matrix[, i] != consensus_seq[i] & alignment_matrix[, i] != "-")
    })
    
    total_sequences <- length(sequences)
    sequence_length <- length(consensus_seq)
    variability_frequencies <- variability_counts / total_sequences
    
    #print(paste("Length of protein_name:", length(protein_name)))
    #print(paste("Length of consensus_seq:", length(consensus_seq)))
    #print(paste("Length of variability_counts:", length(variability_counts)))
    #print(paste("Length of variability_frequencies:", length(variability_frequencies)))
    #print(paste("Total number of sequences:", total_sequences))
    
    data_variability_subset <- data.frame(
      protein = rep(protein_name,sequence_length),
      position = 1:sequence_length,
      consensus_res = consensus_seq,
      var_counts = variability_counts,
      var_freq = variability_frequencies,
      n_seq = total_sequences,
      seq_length = sequence_length
    )
    
    data_variability <- rbind(data_variability, data_variability_subset)
    
  }, error = function(e) {
    warning(paste("Skipping file due to error:", fasfile, "\nError message:", e$message))
  })
}

data_variability
```

```{r}

for (proteinindex in unique(data_variability$protein)) {
  data_subset <- data_variability %>% filter(protein == proteinindex)
  
  plot <- data_subset %>%
    ggplot(aes(x=position, y=var_counts, fill=var_freq)) +
    geom_col() +
    scale_x_continuous(limits = c(1, 20), breaks = 1:20) +
    labs(title = paste0(proteinindex,"\nnumber of seqs: ",unique(data_subset$n_seq))) 
    
    print(plot)
}

seq_vector <- apply(alignment_matrix, 1, paste, collapse ="")

names(seq_vector) <- paste0("Seq_", seq_along(seq_vector))

seq_vector <- substr(seq_vector, 1, 10)

ggseqlogo(seq_vector, method = "bits", ) + 
  theme_minimal()
```

