#Packages
```{r}
source("requirements/packages.R")
```

#Functions
```{r}
function_files <- list.files(path = "requirements/functions", pattern = "\\.R$",full.names = TRUE)

for (function_file in function_files) {
  source(function_file)
}
```

# Variability data
## fas
```{r}
fasfiles <- list.files(path = DNA_alnfiles_path, pattern = "\\.aln.fas", full.names = TRUE)
#fasfiles

aln_protein_names <- sapply(basename(fasfiles), function(x) {
  if (grepl("^\\d+[ _]SP_\\d{4}", x)) {
    # number + (space or underscore) + SP_xxxx
    sub(".*(SP_\\d{4}).*", "\\1", x)
  } else {
    # number + (space or underscore) + Name
    sub("^\\d+[ _]([^_]+).*", "\\1", x)
  }
})

aln_protein_names
```

## Translation & realignment

1. convert DNA to AA, removing any gaps in the aligned DNA sequences 
2. redo the alignments on AA level using DECIPHER

```{r, warning=FALSE}
data_variability <- NULL

for (fasfile in fasfiles) {
  print(basename(fasfile))
  
  protein_name <- sapply(basename(fasfile), function(x) {
    if (grepl("^\\d+[ _]SP_\\d{4}", x)) {
      # number + (space or underscore) + SP_xxxx
      sub(".*(SP_\\d{4}).*", "\\1", x)
    } else {
      # number + (space or underscore) + Name
      sub("^\\d+[ _]([^_]+).*", "\\1", x)
    }
  })
  
  output_file_name <- sub("\\.aln\\.fas$", ".AA.aln.fas", basename(fasfile))
  
  output_file_path <- file.path(AA_alnfiles_path, output_file_name)
  
  tryCatch({
    # Attempt to read the file
    dna_seqs <- readDNAStringSet(fasfile) 
    #print(dna_seqs)
    
    dna_ungapped <- RemoveGaps(dna_seqs)
    
    aa_seqs <- translate(dna_ungapped)

    #print(aa_seqs)
    
    # Perform the alignment
    alignment <- AlignSeqs(aa_seqs,
                          gapOpening = c(-18, -10),
                          gapExtension = -3,
                          terminalGap = -0.1)
    
    aligned_aa_seqs <- as(alignment, "AAStringSet")
    #print(aligned_aa_seqs)
    
    writeXStringSet(aligned_aa_seqs, filepath = output_file_path, format = "fasta")
    
  }, error = function(e) {
    warning(paste("Skipping file due to error:", fasfile, "\nError message:", e$message))
  })
  
}

```

```{r}
AA_fasfiles <- list.files(path = AA_alnfiles_path, pattern = "\\.AA.aln.fas", full.names = TRUE)

aln_protein_names <- sapply(basename(AA_fasfiles), function(x) {
  if (grepl("^\\d+[ _]SP_\\d{4}", x)) {
    # number + (space or underscore) + SP_xxxx
    sub(".*(SP_\\d{4}).*", "\\1", x)
  } else {
    # number + (space or underscore) + Name
    sub("^\\d+[ _]([^_]+).*", "\\1", x)
  }
})

aln_protein_names
```




## Map to PDB's
```{r}
map_namesclasspdb_aln <- data_namesclasspdb %>%
  left_join(
    tibble(
      protein = unname(aln_protein_names),
      matched_filename = names(aln_protein_names)
    ) %>%
  filter(
    !matched_filename %in% c(
      "20 NgtS_old_group_1787.AA.aln.fas",
      "28 MsrAB2_new_group_213.AA.aln.fas", 
      "28 MsrAB2_old_group_213.AA.aln.fas",
      "28 MsrAB2_msrAB.AA.aln.fas")
    )
  ) 

map_namesclasspdb_aln
```

```{r}
map_namesclasspdb_aln_AF3 <- left_join(
  map_namesclasspdb_AF3,
  map_namesclasspdb_aln 
)

map_namesclasspdb_aln_AF3
```



```{r}
map_namesclasspdb_aln_AF3 %>% filter(
  protein %in% 
    setdiff(map_namesclasspdb_aln$protein %>% unique(), data_DiscoTope$protein %>% unique()) 
)
```

###test
```{r}
msa_msrAB1 <- readAAStringSet(filepath = "data/bioedit/fas/AA/msrAB.AA.aln.fas")
msa_msrAB <- readAAStringSet(filepath = "data/bioedit/fas/AA/msrAB1.AA.aln.fas")

msrAB_profiles <- AlignProfiles(
    pattern = msa_msrAB1, 
    subject = msa_msrAB,
    #gapOpening = c(-15,-10), 
    #gapExtension = -4, 
    terminalGap = -0.1#,
  )

fasta_3E0M <- readAAStringSet(filepath = "data/fasta/pdb_protein_sequences.fasta")["3E0M"]

aligned_profiles <- AlignProfiles(
    pattern = fasta_3E0M, 
    subject = msrAB_profiles,
    #gapOpening = c(-15,-10), 
    #gapExtension = -4, 
    terminalGap = -0.1#,
  )

BrowseSeqs(aligned_profiles)
```

### realign
```{}
data_variability_with_resno <- NULL
map_aln_resno <- NULL
aligned_resno <- NULL

# only do this for complete cases
for (protein_name in unique(na.omit(map_namesclasspdb_aln)$protein)) {
  cat("\nProcessing protein:", protein_name, "\n")
  
  # Match the filename for the current protein
  filename <- map_namesclasspdb_aln %>%
    filter(protein == protein_name) %>%
    pull(matched_filename) %>% unique()
  
  if (length(filename) == 0) {
    warning("No matching .fas file for protein: ", protein_name)
    next
  }
  
  # Load existing alignment and remove gaps
  aa_file <- file.path(AA_alnfiles_path, filename)
  
  aa_seqs <- RemoveGaps(readAAStringSet(aa_file))
  
  aa_profile <- readAAStringSet(aa_file)
  
  # Get PDB entries corresponding to this protein
  pdb_entries <- map_namesclasspdb_aln %>% filter(protein == protein_name) %>% pull(PDB)
  
  # Collect all sequences from PDB entries
  pdb_strings <- c()
  
  for (pdb_entry in pdb_entries) {
    print(pdb_entry)
    pdb_object <- tryCatch({
      read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
    }, error = function(e) {
      warning("Could not read PDB: ", pdb_entry, " — skipping. Reason: ", conditionMessage(e))
      return(NULL)
    })
    
    if (is.null(pdb_object)) next
    
    longest_chain <- pdb_object$atom %>%
      filter(type == "ATOM",
             elety == "CA",
             alt %in% c(NA, "A"),
             resid %in% toupper(AMINO_ACID_CODE)) %>%
      add_count(chain, name = "chain_length") %>%
      arrange(desc(chain_length), chain) %>%
      filter(chain_length == max(chain_length)) %>%
      slice_min(chain, with_ties = FALSE) %>%
      pull(chain)

    
    if (length(longest_chain) != 1) {
      warning("Ambiguous chain in PDB:", pdb_entry)
      next
    }
    
    ca_atoms <- pdb_object$atom %>%
      filter(type == "ATOM",
             elety == "CA",
             chain == longest_chain,
             alt %in% c(NA, "A"),
             resid %in% toupper(AMINO_ACID_CODE))
    
    new_seq <- str_c(aa321(ca_atoms$resid), collapse = "")
    names(new_seq) <- pdb_entry
    pdb_strings <- c(pdb_strings, new_seq)
    
    map_aln_resno_subset <- ca_atoms %>% 
      mutate(
        PDB = pdb_entry, 
        protein = unique(na.omit(
          map_namesclasspdb_aln$protein[map_namesclasspdb_aln$PDB == pdb_entry]
          )),
        proteinclass = unique(na.omit(
          map_namesclasspdb_aln$proteinclass[map_namesclasspdb_aln$PDB == pdb_entry]
          )),
        resid_1 = aa321(resid),
        posno = row_number()
      ) %>% 
      select(proteinclass,protein,PDB,chain,resno,resid, posno, resid_1) 
    
    if (is.null(map_aln_resno)) {
      map_aln_resno <- map_aln_resno_subset
    } else {
      # Otherwise, rbind the current data to the combined data frame
      map_aln_resno <- rbind(
        map_aln_resno, 
        map_aln_resno_subset
      )
    }
  }
  
  if (length(pdb_strings) == 0) {
    warning("No alignable PDBs for protein: ", protein_name, " — skipping.")
    next
  }
  
  pdb_stringset <- AAStringSet(pdb_strings, use.names = TRUE)
  
  ifelse(
    length(pdb_stringset) > 1, 
    pdb_profile <- AlignSeqs(pdb_stringset, terminalGap = -0.1),
    pdb_profile <- pdb_stringset
    )

  aligned_profiles <- AlignProfiles(
    pattern = pdb_profile, 
    subject = aa_profile,
    #gapOpening = c(-15,-10), 
    #gapExtension = -4, 
    terminalGap = -0.1#,
  )
  
  combined_seqs <- c(pdb_stringset, aa_seqs)
  
  aligned_combined <- AlignSeqs(
    combined_seqs,  
    #gapOpening = c(-30,-10), 
    #gapExtension = -3, 
    terminalGap = -0.1
  )
  
  aligned_new_seqs <- aligned_combined[names(pdb_strings)]
  
  aligned_pdb_df  <- aligned_new_seqs %>%
    as.character() %>%
    enframe(name = "PDB", value = "aligned_resid_seq") %>%
    mutate(aligned_resid = strsplit(aligned_resid_seq, "")) %>%
    select(-aligned_resid_seq) %>%
    unnest(aligned_resid) %>%
    group_by(PDB) %>%
    mutate(aligned_posno = row_number(),
           nogap_posno = cumsum(aligned_resid != "-")) %>%
    ungroup()
  
  aligned_no_pdb <- aligned_combined[!names(aligned_combined) %in% names(pdb_strings)]
  
  data_variability_subset <- calculate_variability(aligned_no_pdb) 
  
  aligned_resno_df <- full_join(
    map_aln_resno,
    aligned_pdb_df,
    by = c("PDB", "posno" = "nogap_posno", "resid_1" = "aligned_resid")
  )
  
  aligned_resno_subset <- full_join(
    aligned_resno_df,
    data_variability_subset,
    by = c("aligned_posno" = "position")
  )
  
  if (is.null(aligned_resno)) {
    aligned_resno <- aligned_resno_subset
  } else {
    aligned_resno <- rbind(aligned_resno, aligned_resno_subset)
  }
  
}

aligned_resno
```


```{}
#check missing ones
sort(unique(map_namesclasspdb_aln$protein))
sort(unique(aligned_resno$protein))

setdiff(
  sort(unique(map_namesclasspdb_aln$protein)),
  sort(unique(aligned_resno$protein))
)
```


```{}
aligned_pdb_df  <- aligned_new_seqs %>%
  as.character() %>%
  enframe(name = "PDB", value = "aligned_resid_seq") %>%
  mutate(aligned_resid = strsplit(aligned_resid_seq, "")) %>%
  select(-aligned_resid_seq) %>%
  unnest(aligned_resid) %>%
  group_by(PDB) %>%
  mutate(aligned_posno = row_number(),
         nogap_posno = cumsum(aligned_resid != "-")) %>%
  ungroup()

aligned_pdb_df
```

```{}
full_join(
  map_aln_resno,
  aligned_pdb_df,
  by = c("PDB", "posno" = "nogap_posno", "resid_1" = "aligned_resid")
)
```

## check out bad matches
```{r}
data_variability_with_resno <- NULL
map_aln_resno <- NULL

# only do this for complete cases
for (protein_name in unique(na.omit(map_namesclasspdb_aln$protein))[11]) {
cat("\nProcessing protein:", protein_name, "\n")
  
  # Match the filename for the current protein
  filename <- map_namesclasspdb_aln %>%
    filter(protein == protein_name) %>%
    pull(matched_filename) %>% unique()
  
  if (length(filename) == 0) {
    warning("No matching .fas file for protein: ", protein_name)
    next
  }
  
  # Load existing alignment and remove gaps
  aa_file <- file.path(AA_alnfiles_path, filename)
  
  aa_seqs <- RemoveGaps(readAAStringSet(aa_file))
  
  aa_profile <- readAAStringSet(aa_file)
  
  # Get PDB entries corresponding to this protein
  pdb_entries <- map_namesclasspdb_aln %>% filter(protein == protein_name) %>% pull(PDB)
  
  # Collect all sequences from PDB entries
  pdb_strings <- c()
  
  for (pdb_entry in pdb_entries) {
    print(pdb_entry)
    pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
    
    ca_atoms <- reconstruct.ca_atoms(pdb_object)
    
    if(!is.null(pdb_object$seqres)) { #use seqres
      pdb_fasta_sequence <- pdb_object$seqres %>% enframe(name = "chain", value = "resid") %>%
        add_count(chain, name = "chain_length") %>%
        arrange(desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
        filter(chain_length == max(chain_length)) %>%
        filter(chain == min(chain)) %>%
        pull(resid) %>%
        aa321() %>%
        str_c(collapse = "")
      
    } else { # Use atom data
      message("No seqres, using atom data")
      
      pdb_fasta_sequence <- pdb_object$atom %>%
        filter(type == "ATOM",
               elety == "CA",
               alt %in% c(NA, "A"),
               resid %in% toupper(AMINO_ACID_CODE)[1:20]) %>%
        add_count(chain, name = "chain_length") %>%
        arrange(desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
        filter(chain_length == max(chain_length)) %>%
        filter(chain == min(chain)) %>%
        pull(resid) %>%
        aa321() %>%
        str_c(collapse = "")
      
    }
    
    names(pdb_fasta_sequence) <- pdb_entry
    pdb_strings <- c(pdb_strings, pdb_fasta_sequence)
    
    map_aln_resno_subset <- ca_atoms %>% 
      mutate(
        PDB = pdb_entry, 
        protein = unique(na.omit(
          map_namesclasspdb_aln$protein[map_namesclasspdb_aln$PDB == pdb_entry]
          )),
        proteinclass = unique(na.omit(
          map_namesclasspdb_aln$proteinclass[map_namesclasspdb_aln$PDB == pdb_entry]
          ))#,
        #resid_1 = aa321(resid),
        #posno = row_number()
      ) %>% 
      select(proteinclass,protein,PDB,chain,resno,resid, seqresno, seqres_resid, x,y,z,b) 
    
    if (is.null(map_aln_resno)) {
      map_aln_resno <- map_aln_resno_subset
    } else {
      # Otherwise, rbind the current data to the combined data frame
      map_aln_resno <- rbind(
        map_aln_resno, 
        map_aln_resno_subset
      )
    }
  }
  
  pdb_stringset <- AAStringSet(pdb_strings, use.names = TRUE)

  # Combine and align
  combined_seqs <- c(pdb_stringset, aa_seqs)
  aligned_combined <- AlignSeqs(
    combined_seqs,  
    #gapOpening = c(-30,-10), 
    #gapExtension = -3, 
    terminalGap = -0.1
    )

  aligned_combined <- AlignSeqs(combined_seqs, terminalGap = -0.1)
  
}
```

```{r}
BrowseSeqs(aligned_combined)
```


### completely redo alignment
```{r}
data_variability_with_resno <- NULL
map_aln_resno <- NULL
aligned_resno <- NULL

# only do this for complete cases
for (protein_name in sort(unique(na.omit(map_namesclasspdb_aln_AF3$protein)))) {
  cat("\nProcessing protein:", protein_name, "\n")
  
  map_aln_resno <- NULL
  
  # Match the filename for the current protein
  filename <- map_namesclasspdb_aln_AF3 %>%
    filter(protein == protein_name) %>%
    pull(matched_filename) %>% unique() %>% na.omit() 
  
  if (length(filename) == 0) {
    warning("No matching .fas file for protein: ", protein_name)
    next
  }
  
  # Load existing alignment and remove gaps
  aa_file <- file.path(AA_alnfiles_path, filename)
  aa_seqs <- RemoveGaps(readAAStringSet(aa_file))
  
  pdb_entries <- map_namesclasspdb_aln_AF3 %>% filter(protein == protein_name) %>% pull(PDB) %>% na.omit()
  
  # Get PDB entries corresponding to this protein
  if(length(pdb_entries) > 0) {
    # Collect all sequences from PDB entries
    pdb_strings <- c()
    
    for (pdb_entry in pdb_entries) {
      print(pdb_entry)
      pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
      
      ca_atoms <- reconstruct.ca_atoms(pdb_object)
      
      if(!is.null(pdb_object$seqres)) { #use seqres
        pdb_fasta_sequence <- pdb_object$seqres %>% enframe(name = "chain", value = "resid") %>%
          add_count(chain, name = "chain_length") %>%
          arrange(desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
          filter(chain_length == max(chain_length)) %>%
          filter(chain == min(chain)) %>%
          pull(resid) %>%
          aa321() %>%
          str_c(collapse = "")
        
      } else { # Use atom data
        message("No seqres, using atom data")
        
        pdb_fasta_sequence <- pdb_object$atom %>%
          filter(type == "ATOM",
                 elety == "CA",
                 alt %in% c(NA, "A"),
                 resid %in% toupper(AMINO_ACID_CODE)[1:20]) %>%
          add_count(chain, name = "chain_length") %>%
          arrange(desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
          filter(chain_length == max(chain_length)) %>%
          filter(chain == min(chain)) %>%
          pull(resid) %>%
          aa321() %>%
          str_c(collapse = "")
      }
      
      names(pdb_fasta_sequence) <- pdb_entry
      pdb_strings <- c(pdb_strings, pdb_fasta_sequence)
      
      map_aln_resno_subset <- ca_atoms %>% 
        mutate(
          structure = pdb_entry,
          protein = unique(na.omit(
            map_namesclasspdb_aln_AF3$protein[map_namesclasspdb_aln_AF3$PDB == pdb_entry]
          )),
          proteinclass = unique(na.omit(
            map_namesclasspdb_aln_AF3$proteinclass[map_namesclasspdb_aln_AF3$PDB == pdb_entry]
          ))#,
          #resid_1 = aa321(resid),
          #posno = row_number()
        ) %>% 
        select(proteinclass,protein,structure,chain,resno,resid, seqresno, seqres_resid, x, y, z, b) 
      
      if (is.null(map_aln_resno)) {
        map_aln_resno <- map_aln_resno_subset
      } else {
        # Otherwise, rbind the current data to the combined data frame
        map_aln_resno <- rbind(
          map_aln_resno, 
          map_aln_resno_subset
        )
      }
    } 
  
  pdb_stringset <- AAStringSet(pdb_strings, use.names = TRUE)
  
  # Combine and align
  combined_seqs <- c(pdb_stringset, aa_seqs)
  aligned_combined <- AlignSeqs(
    combined_seqs,  
    #gapOpening = c(-30,-10), 
    #gapExtension = -3, 
    terminalGap = -0.1
    )
  
  } else {
  warning(paste(protein_name, " has no PDBs. Attempting AF3 fallback..."))

  # Get potential match names: protein + TIGR4
  potential_names <- map_namesclasspdb_aln_AF3 %>%
    filter(protein == protein_name) %>%
    select(protein, TIGR4_name) %>%
    unlist(use.names = FALSE) %>%
    unique() %>%
    na.omit()
  
  match_idx <- grep(
    pattern = paste0("^(", paste0(tolower(potential_names), collapse = "|"), ")_model_\\d+_fold_\\d+"),
    x = tolower(AF3_pdb_files_list)
  )
  
  if (length(match_idx) == 0) {
    warning("No AF3 model found for protein: ", protein_name)
    next
  }
  
  used_pdb_id <- AF3_pdb_files_list[match_idx[1]]
  message("Using AF3 model: ", used_pdb_id)

  pdb_object <- read.pdb.custom(pdb_entry = tools::file_path_sans_ext(used_pdb_id), pdb_path = AF3_pdbs_path)
  
  ca_atoms <- reconstruct.ca_atoms(pdb_object)

  # Reuse seqres or atom logic
  if (!is.null(pdb_object$seqres)) {
    pdb_fasta_sequence <- pdb_object$seqres %>%
      enframe(name = "chain", value = "resid") %>%
      add_count(chain, name = "chain_length") %>%
      arrange(desc(chain_length), chain) %>%
      filter(chain_length == max(chain_length)) %>%
      filter(chain == min(chain)) %>%
      pull(resid) %>%
      aa321() %>%
      str_c(collapse = "")
  } else {
    message("No seqres, using atom data")
    pdb_fasta_sequence <- pdb_object$atom %>%
      filter(type == "ATOM",
             elety == "CA",
             alt %in% c(NA, "A"),
             resid %in% toupper(AMINO_ACID_CODE)[1:20]) %>%
      add_count(chain, name = "chain_length") %>%
      arrange(desc(chain_length), chain) %>%
      filter(chain_length == max(chain_length)) %>%
      filter(chain == min(chain)) %>%
      pull(resid) %>%
      aa321() %>%
      str_c(collapse = "")
  }
  
  names(pdb_fasta_sequence) <- used_pdb_id
  pdb_strings <- c(pdb_fasta_sequence)  # AF3 version only has one structure
  
  map_aln_resno_subset <- ca_atoms %>%
    mutate(
      structure = used_pdb_id,
      protein = protein_name,
      proteinclass = unique(na.omit(
        map_namesclasspdb_aln$proteinclass[map_namesclasspdb_aln$protein == protein_name]
      ))
    ) %>%
    select(proteinclass, protein, structure, chain, resno, resid, seqresno, seqres_resid, x, y, z, b)

  if (is.null(map_aln_resno)) {
    map_aln_resno <- map_aln_resno_subset
  } else {
    map_aln_resno <- rbind(map_aln_resno, map_aln_resno_subset)
  }

  pdb_stringset <- AAStringSet(pdb_strings, use.names = TRUE)
  combined_seqs <- c(pdb_stringset, aa_seqs)
  aligned_combined <- AlignSeqs(combined_seqs, terminalGap = -0.1)
}

  
  aligned_combined <- AlignSeqs(combined_seqs, terminalGap = -0.1)
  
  aligned_new_seqs <- aligned_combined[names(pdb_strings)]
  
  aligned_pdb_df  <- aligned_new_seqs %>%
    as.character() %>%
    enframe(name = "structure", value = "seq") %>%  # Turn it into a tibble
    rowwise() %>%
    mutate(
      protein = protein_name,
      proteinclass = unique(na.omit(
        map_namesclasspdb_aln$proteinclass[map_namesclasspdb_aln$protein == protein_name]
      )),
      aligned_seq_resid = strsplit(seq, "")
    ) %>%
    select(-seq) %>%
    unnest(aligned_seq_resid) %>%
    group_by(structure) %>%
    mutate(aligned_posno = row_number(),
           nogap_posno = case_when(
             aligned_seq_resid == "-" ~ NA,
             aligned_seq_resid != "-" ~ cumsum(aligned_seq_resid != "-")
                                   ),
           aligned_seq_resid = aa123(aligned_seq_resid)) %>%
    ungroup()
  
  aligned_no_pdb <- aligned_combined[!names(aligned_combined) %in% names(pdb_strings)]
  
  data_variability_subset <- calculate_variability(aligned_no_pdb) 
  
  aligned_resno_df <- full_join(
    map_aln_resno,
    aligned_pdb_df,
    by = c("proteinclass","protein","structure", "seqresno" = "nogap_posno")
  ) %>%
    mutate(
      PDB = case_when(
        structure %in% map_namesclasspdb_aln_AF3$PDB ~ structure,
        TRUE ~ NA
      ),
      AF3 = case_when(
        structure %in% map_namesclasspdb_aln_AF3$AF3 ~ structure,
        TRUE ~ NA
      )
    )
  
  aligned_resno_subset <- full_join(
    aligned_resno_df,
    data_variability_subset,
    by = c("aligned_posno" = "position")
  )
  
  if (is.null(aligned_resno)) {
    aligned_resno <- aligned_resno_subset
  } else {
    aligned_resno <- rbind(aligned_resno, aligned_resno_subset)
  }
  
}

aligned_resno
```

```{r}
aligned_resno %>% filter(protein == "AliA")
```


```{r}
write.csv(
  aligned_resno,
  "output/data_variability.csv", 
  row.names=FALSE
)
```

#### calculate mismatch
```{r}
data_variability <- read.csv("output/data_variability.csv")
```


```{r}
mismatch <- data_variability %>%
  select(proteinclass, protein, PDB, chain, resno, resid, seqres_resid,consensus_res, residue_tables) %>%
  na.omit(seqres_resid) %>%
  mutate(mapcheck = consensus_res == aa321(seqres_resid),
         mapcheck_full = mapply(function(x, y) {
      if (is.na(y) || y == "") {
        NA
      } else {
        grepl(x, y)
      }
    }, aa321(seqres_resid), residue_tables)
  ) %>%
  group_by(PDB, protein) %>%
  summarize(
    match = sum(mapcheck, na.rm = TRUE),
    mismatch = sum(!mapcheck, na.rm = TRUE),
    matchpercentage = match / (match + mismatch),
    match_relaxed = sum(mapcheck_full, na.rm = TRUE),
    mismatch_relaxed = sum(!mapcheck_full, na.rm = TRUE),
    matchpercentage_relaxed = match_relaxed / (match_relaxed + mismatch_relaxed),
    .groups = "drop"
  ) %>%
  arrange(matchpercentage_relaxed, matchpercentage, PDB)
mismatch
```

## Calculate variability
```{r}
AA_fasfiles

AA_alnfiles_path
```

```{r, warning=FALSE}
data_variability <- NULL

for (fasfile in na.omit(unique(map_namesclasspdb_aln$matched_filename))){
  print(basename(fasfile))
  
  protein_name <- if (grepl("^\\d+[ _]SP_\\d{4}", basename(fasfile))) {
    sub(".*(SP_\\d{4}).*", "\\1", basename(fasfile))
  } else {
    sub("^\\d+[ _]([^_]+).*", "\\1", basename(fasfile))
  }
  
  print(protein_name)
  
  tryCatch({
    aligned_aa_seqs <- readAAStringSet(paste0(AA_alnfiles_path,"/",fasfile))
    #print(aligned_aa_seqs)
    
    data_variability_subset <- calculate_variability(aligned_aa_seqs)
    
    data_variability_subset$protein <- protein_name
    
    data_variability <- rbind(data_variability, data_variability_subset)
    
  }, error = function(e) {
    warning(paste("Skipping file due to error:", fasfile, "\nError message:", e$message))
  })
  
}

data_variability <- data_variability %>% 
  left_join(map_namesclasspdb_aln, by = "protein") %>%
  select(-matched_filename)

data_variability
```


```{r}
data_variability %>% filter(is.na(proteinclass))
```


#missing proteins
```{r}
#proteins without variability
setdiff(data_namesclasspdb$protein, aln_protein_names)
```


```{r}
proteins_without_structure <- setdiff(aln_protein_names, unique(data_variability$protein))
proteins_without_structure

saveRDS(proteins_without_structure, file = "data/variable_proteins_without_structure.rds")
```


# load variability
```{r}
data_variability <- read.csv("output/data_variability.csv")

data_variability
```

```{r}
data_variability %>% filter(PDB == "1W9R")
```


#shannon evenness
```{r}
data_variability_shannon <- data_variability %>%
  filter(str_detect(residue_tables, "-")) %>%
  rowwise() %>%
  mutate(
    # Calculate entropy first
    shannon_entropy = {
      entries <- str_split(residue_tables, ";\\s*")[[1]]

      counts <- sapply(entries, function(entry) {
        parts <- str_split(entry, ":")[[1]]
        residue <- parts[1]
        count <- as.numeric(parts[2])
        if (residue %in% c("*", "-")) return(NA)
        count
      }, USE.NAMES = FALSE)

      counts <- counts[!is.na(counts)]

      total <- sum(counts, na.rm = TRUE)

      if (total == 0 || length(counts) == 0) {
        NA_real_
      } else {
        probs <- counts / total
        -sum(probs * log(probs))
      }
    },
    # Then calculate evenness using entropy & number of residues
    shannon_evenness = {
      entries <- str_split(residue_tables, ";\\s*")[[1]]

      counts <- sapply(entries, function(entry) {
        parts <- str_split(entry, ":")[[1]]
        residue <- parts[1]
        if (residue %in% c("*", "-")) return(NA)
        1L  # just count unique valid residues
      }, USE.NAMES = FALSE)

      k <- sum(!is.na(counts))  # number of residue types

      if (is.na(shannon_entropy) || k <= 1) {
        1
      } else {
        shannon_entropy / log(k)
      }
    }
  ) %>%
  ungroup()

data_variability_shannon %>%
  select(seqresno,residue_tables, var_counts, var_freq, indel_freq, shannon_entropy, shannon_evenness)
```


# plot
```{r}
for (proteinindex in unique(data_variability$protein)[1]) {
  data_subset <- data_variability %>% filter(protein == proteinindex)
  
  plot <- data_subset %>%
    ggplot(aes(x=seqresno, y=var_counts)) +
    geom_col() +
    #scale_x_continuous(limits = c(1, 20), breaks = 1:20) +
    labs(title = paste0(proteinindex,"\nnumber of seqs: ",unique(data_subset$n_seq))) 
    
    print(plot)
}

```

```{r}
for (proteinindex in unique(data_variability$protein)[1]) {
  data_subset <- data_variability %>% filter(protein == proteinindex)
  
  data_long <- data_subset %>%
    gather(key = "count_type", value = "count_value", var_counts, indel_counts) %>%
    mutate(
      count_type = case_when(
        count_type == "var_counts" ~ "Variation",  
        indel_type == "insertion" ~ "Insertion", 
        indel_type == "deletion" ~ "Deletion",    
        TRUE ~ "Other"                           
      ),
      count_type = factor(count_type, levels = rev(c("Deletion", "Insertion", "Variation")))
    )
  
  plot <- data_long %>%
    ggplot(aes(x = seqresno, y = count_value, fill = count_type)) +
    geom_bar(stat = "identity", position = "stack") +
    labs(title = paste0(proteinindex,"\nnumber of seqs: ",unique(data_subset$n_seq))) +
    scale_fill_manual(values = c("Variation" = "black", "Insertion" = "skyblue", "Deletion" = "maroon")) +
    scale_y_continuous(limits = c(0, max(data_long$n_seq)),
                       expand = c(0, 0) ) +
    scale_x_discrete(
      limits = c(0, max(data_long$seq_length)),
      expand = c(0, 0)  
    ) +
    
    theme_bw()
  
  print(plot)
}
```


```{r}
for (proteinindex in unique(data_variability$protein)[1:5]) {
  data_subset <- data_variability %>% filter(protein == proteinindex)
  
  data_long <- data_subset %>%
    gather(key = "count_type", value = "count_value", var_counts, indel_counts) %>%
    mutate(
      count_type = case_when(
        count_type == "var_counts" ~ "Variation",  
        indel_type == "insertion" ~ "Insertion", 
        indel_type == "deletion" ~ "Deletion",    
        TRUE ~ "Other"                           
      ),
      count_type = factor(count_type, levels = rev(c("Deletion", "Insertion", "Variation"))),
      freq_value = count_value / n_seq
    )
  
  plot <- data_long %>%
    ggplot(aes(x = seqresno, y = freq_value, fill = count_type)) +
    geom_bar(stat = "identity", position = "stack") +
    labs(title = paste0(proteinindex,"\nnumber of seqs: ",unique(data_subset$n_seq))) +
    scale_fill_manual(values = c("Variation" = "black", "Insertion" = "skyblue", "Deletion" = "maroon")) +
    scale_y_continuous(limits = c(0, 0.5)) +
    theme_bw()
    
    print(plot)
}
```

```{r}
for (structureindex in unique(data_variability_shannon$structure)[1:5]) {
  data_subset <- data_variability_shannon %>% filter(PDB == structureindex)
  print(data_subset)
  
  plot <- data_subset %>%
    ggplot(aes(x = seqresno, y = shannon_evenness)) +
    geom_bar(stat = "identity", position = "stack") +
    labs(title = paste0(unique(data_subset$protein),"\nnumber of seqs: ",unique(data_subset$n_seq))) +
    #scale_fill_manual(values = c("Variation" = "black", "Insertion" = "skyblue", "Deletion" = "maroon")) +
    #scale_y_continuous(limits = c(0, 0.5)) +
    theme_bw()
    
    print(plot)
}
```


```{r, fig.width=14}
for (structureindex in unique(data_variability_shannon$structure)[1:5]) {
  data_subset <- data_variability_shannon %>% filter(structure == structureindex)
  
  data_long <- data_subset %>%
    gather(key = "count_type", value = "count_value", var_counts, indel_counts) %>%
    mutate(
      count_type = case_when(
        count_type == "var_counts" ~ "Variation",  
        indel_type == "insertion" ~ "Insertion", 
        indel_type == "deletion" ~ "Deletion",    
        TRUE ~ "Other"                           
      ),
      count_type = factor(count_type, levels = rev(c("Deletion", "Insertion", "Variation"))),
      freq_value = count_value / n_seq
    )
  
  plot_upper <- data_long %>%
    ggplot(aes(x = seqresno, y = freq_value, fill = count_type)) +
    geom_bar(stat = "identity", position = "stack") +
    labs(title = paste0(unique(data_subset$protein),"\nnumber of seqs: ",unique(data_subset$n_seq))) +
    scale_fill_manual(values = c("Variation" = "black", "Insertion" = "skyblue", "Deletion" = "maroon")) +
    scale_y_continuous(expand = c(0,0)) +
    scale_x_continuous(limits = c(0, max(data_long$seqresno)), expand = c(0,0)) +
    theme_bw() #+
    #theme(legend.position = "none")
  
  plot_lower <- data_subset %>%
    ggplot(aes(x = seqresno, y = shannon_entropy)) +
    geom_bar(stat = "identity", position = "stack") +
    #labs(title = paste0(unique(data_subset$protein),"\nnumber of seqs: ",unique(data_subset$n_seq))) +
    #scale_fill_manual(values = c("Variation" = "black", "Insertion" = "skyblue", "Deletion" = "maroon")) +
    scale_y_continuous(expand = c(0,0)) +
    scale_x_continuous(limits = c(0, max(data_subset$seqresno)), expand = c(0,0)) +
    theme_bw()
  
  plot <- plot_grid(
    plot_upper + 
    theme(
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.ticks.x = element_blank(),
      plot.margin = margin(t = 5, r = 5, b = 3, l = 5)
    )
  ,
  plot_lower + 
    theme(
      strip.text = element_blank(),
      strip.background = element_blank(),
      plot.margin = margin(t = 0, r = 5, b = 5, l = 5) 
    )
  ,
    ncol = 1, 
    align = "v",  # vertical alignment
    axis = "lr"   # align left and right axes
  )
    
  print(plot)
}
```

```{r, fig.width=14}
for (structureindex in unique(data_variability_shannon$structure)[5]) {
  data_subset <- data_variability_shannon %>% filter(structure == structureindex)
  
  
  plot_upper <- data_subset %>%
    ggplot(aes(x = seqresno, y = var_freq)) +
    geom_bar(stat = "identity", position = "stack") +
    labs(title = paste0(unique(data_subset$protein),"\nnumber of seqs: ",unique(data_subset$n_seq))) +
    scale_y_continuous(limits = c(0,0.5),expand = c(0,0)) +
    scale_x_continuous(limits = c(0, max(data_long$seqresno)), expand = c(0,0)) +
    labs(y = "Residue variability") +
    theme_bw() +
    theme(
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
  ) 
  
  plot_lower <- data_subset %>%
    ggplot(aes(x = seqresno, y = shannon_entropy)) +
    geom_bar(stat = "identity", position = "stack") +
    #labs(title = paste0(unique(data_subset$protein),"\nnumber of seqs: ",unique(data_subset$n_seq))) +
    #scale_fill_manual(values = c("Variation" = "black", "Insertion" = "skyblue", "Deletion" = "maroon")) +
    scale_y_continuous(limits = c(0,1),expand = c(0,0)) +
    scale_x_continuous(limits = c(0, max(data_subset$seqresno)), expand = c(0,0)) +
    labs(x = "",
         y = "Residue Shannon entropy") +
    theme_bw() +
    theme(
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
  )
  
  plot <- plot_grid(
    plot_upper + 
    theme(
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      axis.ticks.x = element_blank(),
      plot.margin = margin(t = 5, r = 5, b = 3, l = 5)
    )
  ,
  plot_lower + 
    theme(
      strip.text = element_blank(),
      strip.background = element_blank(),
      plot.margin = margin(t = 0, r = 5, b = 5, l = 5) 
    )
  ,
    ncol = 1, 
    align = "v",  # vertical alignment
    axis = "lr"   # align left and right axes
  )
    
  print(plot)
}
```

```{r}
data_subset %>% filter(resno > 267, resno <= 277) %>%
  select(seqresno, var_freq, shannon_entropy, residue_tables)
```


## shannon vs var
```{r}
data_variability_shannon %>%
  filter(var_freq < 0.05, shannon_entropy > 0.3)
```


```{r, fig.width=14}
data_variability_shannon %>%
  filter(consensus_res != "-") %>%
  ggplot(aes(x=var_freq,y=shannon_entropy)) +
  geom_point() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw() +
  labs(x = "Residue variability",
       y = "Residue Shannon entropy",
       title = "Correlation between Shannon entropy and variability") +
  theme(
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
  )

```


#### filter out all INDELS
```{r}
#without consensus_gaps
for (proteinindex in unique(data_variability$protein)) {
  data_subset <- data_variability %>% filter(protein == proteinindex,
                                             consensus_res != "-") %>%
    mutate(position_new = row_number())
  
  plot <- data_subset %>%
    ggplot(aes(x=position_new, y=var_counts, fill=var_freq)) +
    geom_col() +
    #scale_x_continuous(limits = c(1, 20), breaks = 1:20) +
    labs(title = paste0(proteinindex,"\nnumber of seqs: ",unique(data_subset$n_seq))) 
    
    print(plot)
}
```

#### show all gaps
```{r}
for (proteinindex in unique(data_variability$protein)) {
  data_subset <- data_variability %>% filter(protein == proteinindex)
  
  plot <- data_subset %>%
    ggplot(aes(x=position, y=indel_counts, fill=indel_type)) +
    geom_col() +
    #scale_x_continuous(limits = c(1, 20), breaks = 1:20) +
    labs(title = paste0(proteinindex,"\nnumber of seqs: ",unique(data_subset$n_seq))) 
    
    print(plot)
}
```



# summary statistics
```{r}
data_variability_with_seroprof <- data_variability %>%
  mutate(
    protein_adj = case_when(
      protein == "Pneumolysin" ~ "Ply",
      TRUE ~ protein
    ),
    seroprof_titers = factor(
      case_when(
        protein_adj %in% c("Ply", "Hic2", "PspA", "SP_2063", "PpmA", "PnrA", "LytC", "SP_0107", "PavB", "AmiA", "AliC", "AliB", "PsaA", "PhtD", "PcsB", "PrtA2", "PcpA", "PspC") ~ "High",
        protein_adj %in% c("PitB", "PccL", "PiaA", "PfbA", "LytA", "GlpO", "AliA", "PhpP", "PstS2", "SlrA", "CbpL", "CbpC", "CbpE", "DacB", "NanA") ~ "Intermediate",
        protein_adj %in% c("MetQ", "MalX", "Enolase", "AliD", "SP_0699", "SP_1069", "NgtS", "RrgB", "RrgA", "MsrAB2", "GshT", "AdcAII", "TrxB", "SP_2141", "SP_0060", "Etrx2", "SP_0191", "LytB", "PepO", "SP_1992", "Etrx1") ~ "Low",
        TRUE ~ NA_character_
      ),
      levels = c("High", "Intermediate", "Low")
    )
  )

data_variability_with_seroprof

```

```{r}
data_variability %>% filter(is.na(protein)) 
```


```{r}
data_variability %>% pull(protein) %>% unique()
```

```{r}
ordered_proteins <- data_variability_with_seroprof %>%
  select(seroprof_titers, proteinclass, protein, n_seq, seq_length) %>%
  distinct() %>%
  arrange(desc(n_seq)) %>%
  pull(protein) %>%
  unique()
```


```{r, fig.width=14}
upper_plot <- data_variability_with_seroprof %>%
  select(seroprof_titers, proteinclass,protein,n_seq,seq_length) %>%
  drop_na(protein) %>%
  distinct() %>%
  group_by(seroprof_titers) %>%
  arrange(desc(n_seq)) %>% 
  mutate(
    protein = factor(
      protein, 
      levels = ordered_proteins
    )
  ) %>%
  
  ggplot(aes(x=protein)) +
  
  geom_rect(
        aes(
            xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = seroprof_titers
        ),
        inherit.aes = FALSE, alpha = 0.3,
        data = data_variability_with_seroprof %>% distinct(seroprof_titers),
        show.legend = FALSE  # Exclude facet-based color fill from the legend
    ) +
  
  scale_fill_brewer(palette = "RdYlBu") +
  
  new_scale_fill() +
  
  geom_col(aes(y = n_seq, fill = proteinclass)) +
  
  facet_grid(
        ~seroprof_titers,
        scales = "free_x",
        space = "free_x",
        labeller = label_wrap_gen(width = 10)
    ) +
  
  labs(y = "Number of sequences", x = NULL) +
  
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, NA)) +
  
  theme_bw() +
  
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 16), # Adjust facet label text size
    panel.background = element_rect(fill = "white", colour = "white"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.spacing = unit(0, "lines"),
    axis.line = element_line(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

upper_plot
```

```{r, fig.width=14}
min_var_freq = 0.05

lower_plot <- data_variability_with_seroprof %>%
  drop_na(protein) %>%
  group_by(seroprof_titers,proteinclass, protein, seq_length) %>%
  distinct() %>%
  summarize(
    nonzero_var_count = sum(var_freq > min_var_freq),
    .groups = "drop"
  ) %>%
  filter(!is.na(seq_length)) %>%
  mutate(
    normalized_var = nonzero_var_count / seq_length,
    protein = factor(protein, levels = ordered_proteins)
  ) %>%
  
  ggplot(aes(x = protein, y = normalized_var)) +
  
  geom_rect(
    aes(
      xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = seroprof_titers
    ),
    inherit.aes = FALSE, alpha = 0.3,
    data = data_variability_with_seroprof %>% distinct(seroprof_titers),
    show.legend = FALSE  # Exclude facet-based color fill from the legend
  ) +
  
  scale_fill_brewer(palette = "RdYlBu") +
  
  new_scale_fill() +
  
  geom_col(aes(fill = proteinclass)) +
  
  facet_grid(
    ~seroprof_titers,
    scales = "free_x",
    space = "free_x",
    labeller = label_wrap_gen(width = 10)
  ) +
  
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, NA), breaks = seq(0, 1, by = 0.1) ) +
  
  theme_bw() +
  
  labs(y = "Normalized Variability", x = NULL) +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 16), # Adjust facet label text size
    panel.background = element_rect(fill = "white", colour = "white"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.spacing = unit(0, "lines"),
    axis.line = element_line(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

lower_plot
```

```{r, fig.width=14}
plot_grid(
  upper_plot + 
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      plot.margin = margin(t = 5, r = 5, b = 1.5, l = 5)
    )
  ,
  lower_plot + 
    theme(
      strip.text = element_blank(),
      strip.background = element_blank(),
      legend.position = "none",
      plot.margin = margin(t = 0, r = 5, b = 5, l = 5) 
    ),
  ncol = 1,
  align = "v",
  axis = "rl",
  rel_heights = c(1, 0.5)  # adjust height ratio as needed
)
```


