---
title: "Streptococcus pneumoniae epitope visualization"
author: "Jan Paul van Meenen"
date: "`r Sys.Date()`"
output: html_document
---

#Packages
```{r}
source("requirements/packages.R")
```

#Functions
```{r}
function_files <- list.files(path = "functions", pattern = "\\.R$",full.names = TRUE)

for (function_file in function_files) {
  source(function_file)
}
```

# Plot all together
## Load in from 01
```{r}
data_BepiPred_2_with_resno <- read.csv("output/BepiPred2.0/data_BepiPred2.csv", check.names = FALSE)
data_BepiPred_3_with_resno <- read.csv("output/BepiPred3.0/data_BepiPred3.csv", check.names = FALSE)
data_Epigraph <- read.csv("output/EpiGraph/data_Epigraph.csv", check.names = FALSE)
data_DiscoTope <- read.csv("output/DiscoTope/data_DiscoTope.csv", check.names = FALSE)
map_namesclasspdb_AF3 <- read.csv("data/map_nameclass_pdbAF3.csv", check.names = FALSE)
```


## create joined predictions df
```{r}
data_predictions_sequencebased <- full_join(
  data_BepiPred_2_with_resno %>%
    mutate(
      `BepiPred-2.0 probability` = EpitopeProbability,
      `RSA_NetsurfP` = RelativeSurfaceAccessilibity
    ) %>%
    select(
      proteinclass, protein, TIGR4_name, PDB, chain, resno, resid, x,y,z,b, seqresno, seqres_resid,
      `BepiPred-2.0 probability`,RSA_NetsurfP
    ) 
  ,
  
  data_BepiPred_3_with_resno %>%
    select(
      proteinclass, protein, TIGR4_name, PDB, chain, resno, resid, x,y,z,b, seqresno, seqres_resid, 
      `BepiPred-3.0 score`,`BepiPred-3.0 linear epitope score`
    ) 
  ,
  by = c("proteinclass","protein","TIGR4_name","PDB","chain","resno","resid","seqres_resid","seqresno","x","y","z","b")
)

data_predictions_sequencebased
```

```{r}
data_predictions_structurebased <- full_join(
  data_Epigraph %>%
    mutate(
      Epigraph_score = Score
      ) %>%
    select(proteinclass,protein,TIGR4_name,PDB, chain, resno, resid, Epigraph_score, RSA)
  ,
  data_DiscoTope %>%
    mutate(
           `RSA Shrake-Rupley` = rsa,
           `DiscoTope-3.0_calibrated_score` = calibrated_score
    ) %>%
    select(
      proteinclass,protein,TIGR4_name,PDB, chain, resno, resid, `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, `RSA Shrake-Rupley`
      )
  , 
  by = c("proteinclass","protein","TIGR4_name","PDB","chain","resno","resid")
)
data_predictions_structurebased
```


```{r}
data_predictions <- full_join(
  data_predictions_sequencebased,
  data_predictions_structurebased %>%
    group_by(proteinclass,protein,TIGR4_name,PDB) %>%
    add_count(chain, name = "chain_length") %>%
    arrange(desc(chain_length), chain) %>%
    filter(chain_length == max(chain_length)) %>%
    filter(chain == min(chain)),
  by = c("proteinclass","protein","TIGR4_name","PDB","chain","resno","resid")
  ) %>%
  left_join(map_namesclasspdb_AF3 %>% select(protein, PDB, AF3)) %>%
  mutate(structure = case_when(
    !is.na(PDB) ~ PDB,
    !is.na(AF3) ~ str_remove(AF3, "\\.pdb$"),
    TRUE ~ NA
  ))

data_predictions %>%
  filter(if_any(everything(), is.na))

#data_predictions <- data_predictions %>%
#  na.omit() 

data_predictions
```

##save
```{r}
write.csv(data_predictions,
          "output/data_predictions.csv", 
          row.names=FALSE) 
```


# Variability
matching variability consensus sequence to epitope predictions sequences
## variability_data_with_resno
```{r}
data_variability <- read.csv("output/data_variability_filtered.csv") #%>%
  #distinct(structure,chain,x,y,z, .keep_all = TRUE)

data_variability
```

```{r}
data_variability %>%
  filter(structure == "2BIB") %>%
  arrange(aligned_posno)
```


#!LOW n_seq MATCH FILTER
```{r}
data_variability %>% 
  select(proteinclass, protein, structure, n_seq) %>% 
  group_by(proteinclass, protein, structure) %>% 
  #mutate(mean_b = mean(b, na.rm = TRUE)) %>% select (-b) %>% 
  arrange(n_seq) %>% unique()
```

```{r}
low_n_seq <- c("PspC")

data_variability <- data_variability %>% filter(!protein %in% c(low_n_seq))
```

### check the missing ones
Currently working with PDB entries only, not all proteins have PDB's associated with them, so some are missing
```{}
sort(unique(data_predictions$protein))
sort(unique(data_variability$protein))

setdiff(
  tolower(unique(data_predictions$protein)),
  tolower(unique(data_variability$protein))
)

setdiff(
  tolower(unique(data_variability$protein)),
  tolower(unique(data_predictions$protein))
)
```

```{r}
data_predictions_with_variability <- full_join(
  data_predictions,
  data_variability %>% select(-chain),
  by = c("proteinclass","protein","structure","resno","resid", "seqresno", "seqres_resid", "x", "y", "z", "b", "PDB", "AF3")) %>%
  mutate(
    seroprof_titers = data_nameclasspdb_with_seroprof$seroprof_titers[
      match(protein, data_nameclasspdb_with_seroprof$protein)
    ]
  )

data_predictions_with_variability
```


```{r}
data_predictions_with_variability %>% filter(protein == "PavB")

data_predictions_with_variability %>% 
  group_by(proteinclass, protein) %>%
  summarize(
    has_pdb = any(!is.na(PDB)),
    has_AF3 = any(!is.na(AF3)),
    has_AF3_conf = (has_AF3 && (mean(b, na.rm=TRUE) >= 80)),
    has_var = any(!is.na(var_freq)),
    complete = (has_pdb &  has_var),
    complete_adj = ((has_pdb | has_AF3_conf) &  has_var),
    .groups = "drop"
  ) %>%
  filter(complete_adj == TRUE)
```


###save
```{r}
write.csv(data_predictions_with_variability,
          "output/data_predictions_with_variability.csv", 
          row.names=FALSE) 
```
