---
title: "Streptococcus pneumoniae epitope visualization"
author: "Jan Paul van Meenen"
date: "`r Sys.Date()`"
output: html_document
---

#Packages
```{r}
source("requirements/packages.R")
```

#Functions
```{r}
function_files <- list.files(path = "functions", pattern = "\\.R$",full.names = TRUE)

for (function_file in function_files) {
  source(function_file)
}
```

# Plot all together
## Load in from 01
```{r}
data_BepiPred_2_with_resno <- read.csv("output/BepiPred2.0/data_BepiPred2.csv", check.names = FALSE)
data_BepiPred_3_with_resno <- read.csv("output/BepiPred3.0/data_BepiPred3.csv", check.names = FALSE)
data_Epigraph <- read.csv("output/EpiGraph/data_Epigraph.csv", check.names = FALSE)
data_DiscoTope <- read.csv("output/DiscoTope/data_DiscoTope.csv", check.names = FALSE)
```


## create joined predictions df
```{r}
data_predictions_sequencebased <- full_join(
  data_BepiPred_2_with_resno %>%
    mutate(PDB = Entry,
           `BepiPred-2.0 probability` = EpitopeProbability,
           `RSA NetsurfP` = RelativeSurfaceAccessilibity) %>%
    select(PDB, chain, resno,resid,`BepiPred-2.0 probability`,`RSA NetsurfP`) %>%
    filter(resid %in% toupper(AMINO_ACID_CODE))
  ,
  
  data_BepiPred_3_with_resno %>%
    mutate(PDB = Accession) %>%
    select(PDB, chain, resno, resid, `BepiPred-3.0 score`,`BepiPred-3.0 linear epitope score`) %>%
    filter(resid %in% toupper(AMINO_ACID_CODE))
  ,
  by = c("PDB","chain","resno","resid")
)

data_predictions_sequencebased

data_predictions_sequencebased %>%
  filter(if_any(everything(), is.na))
```

```{r}
data_predictions_structurebased <- full_join(
  data_Epigraph %>%
    mutate(Epigraph_score = Score) %>%
    select(PDB, chain, resno, resid, Epigraph_score)
  ,
  data_DiscoTope %>%
    mutate(PDB = pdb,
           resid = aa123(residue),
           `RSA Shrake-Rupley` = rsa,
           `DiscoTope-3.0_calibrated_score` = calibrated_score
    ) %>%
    select(PDB, chain, resno, resid, `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, `RSA Shrake-Rupley`)
  , 
  by = c("PDB","resno","chain","resid")
)
data_predictions_structurebased

data_predictions_structurebased %>%
  filter(if_any(everything(), is.na))

data_predictions_structurebased <- data_predictions_structurebased %>%
  na.omit()
  
data_predictions_structurebased %>%
  filter(if_any(everything(), is.na))
```


```{r}
data_predictions <- full_join(
  data_predictions_sequencebased,
  data_predictions_structurebased,
  by = c("PDB", "chain", "resno","resid")
  )

data_predictions %>%
  filter(if_any(everything(), is.na))

data_predictions <- data_predictions %>%
  na.omit() 

data_predictions
```

### get the protein names 
names are not consisten betwen the structural and protein panel data, so they have to be cleaned
```{r}
data_namesclasspdb <- data_structural %>%
  separate_rows(`PDB entry`, sep = ", ") %>%
  mutate(PDB = `PDB entry`,
         structure_proteinname = `Proteins (TIGR4) name`,
         structure_proteinname_clean = ifelse(
           grepl("^SP_", gsub(" ", "", gsub("\\(.*", "", `Proteins (TIGR4) name`))), 
           sub("^(SP_\\d{4}).*", "\\1", gsub(" ", "", gsub("\\(.*", "", `Proteins (TIGR4) name`))),
           sub("_.*", "", gsub(" ", "", gsub("\\(.*", "", `Proteins (TIGR4) name`)))
         )
  ) %>%
  select(structure_proteinname_clean, PDB) %>%
  full_join(
    data_proteinpanel %>% 
      mutate(panel_proteinname = `Proteins (TIGR4)`,
             panel_proteinname_clean = gsub(" ","",gsub("\\(.*", "", panel_proteinname)),
             proteinclass = `Protein class`
      ) %>%
      select(proteinclass, panel_proteinname_clean), keep = TRUE,
    by = c("structure_proteinname_clean" = "panel_proteinname_clean")
  ) %>%
  mutate(proteinname = structure_proteinname_clean) %>%
  select(proteinclass, proteinname, PDB)

data_namesclasspdb
```


```{r}
data_predictions <- data_predictions %>%
  left_join(
    data_namesclasspdb
  ) 

data_predictions
```

## normalize
```{r}
data_predictions_normalized <- data_predictions %>%
  group_by(PDB, chain) %>%
  mutate(
    across(
      c(#`BepiPred-2.0 probability`,
        `BepiPred-3.0 score` , `BepiPred-3.0 linear epitope score`, 
        `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, Epigraph_score), 
      rescale
    )
  ) %>%
  ungroup()

data_predictions_normalized
```

###track
```{r}
track_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score"
)

for (pdb_entry in unique(data_predictions_normalized$PDB)) {
  data_predictions_subset <- data_predictions %>%
    filter(PDB == pdb_entry) 
  
  proteinname <- unique(data_predictions_subset$proteinname)
  
  for (chain_index in unique(data_predictions_subset$chain)) {
    
    data_predictions_subset_long <- data_predictions_subset %>%
      filter(chain == chain_index) %>%
      pivot_longer(cols = track_list, 
                   names_to = "Score_Type", values_to = "Score_Value") %>%
      group_by(Score_Type) %>%
      mutate(normalized_score = case_when(
        Score_Type == "BepiPred-2.0 probability" ~ Score_Value,
        TRUE ~ rescale(Score_Value, to = c(0, 1))
        )
      )
    
    figure <- data_predictions_subset_long %>%
      ggplot(aes(x=resno, y=Score_Value, fill = normalized_score)) +
      geom_col() +
      
      facet_wrap(~Score_Type, ncol = 1, scales = "free_y") +
      
      scale_fill_gradient2(low = "blue", mid = "gray95", high = "red", 
                           midpoint = 0.5) +  # Now uses median per facet
      
      geom_text(aes(label = Score_Type, 
                    x = (min(resno) + (max(resno) - min(resno)) / 2)), 
                color = "black", y = Inf, vjust = 1.5, size = 4, family = "Arial", 
                check_overlap = TRUE) +
      
      labs(title = paste0(pdb_entry, " chain ", chain_index, " (",proteinname,")")) +
      
      theme_minimal() +
      theme(
        strip.text = element_blank(),
        axis.ticks = element_blank(),  
        axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      )
    
    print(figure)
    
    ggsave(filename = paste0("output/figures/tracks/", pdb_entry, "_", chain_index, "_track.png"), 
           plot = figure, 
           width = 14, height = 10, dpi = 300,
           bg="white")
  }
}
```


###heatmap
```{r, fig.width = 14}
for (pdb_entry in unique(data_predictions_normalized$PDB)) {
  data_predictions_normalised_subset <- data_predictions_normalized %>%
    filter(PDB == pdb_entry) 
  
  proteinname <- unique(data_predictions_normalised_subset$proteinname)
  
  for (chain_index in unique(data_predictions_normalised_subset$chain)) {
    
    data_predictions_normalised_subset_wide <- data_predictions_normalised_subset %>%
      filter(chain == chain_index) %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score) %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    figure <- pheatmap(t(data_predictions_normalised_subset_wide), 
                       cluster_rows = TRUE,         
                       cluster_cols = FALSE,         
                       main = paste0(pdb_entry, " chain ", chain_index, " (",proteinname,")"), 
                       color = colorRampPalette(c("blue", "gray95", "red"))(50),
                       labels_col = "resno",
                       clustering_distance_rows = "correlation") 
    
    print(figure)
    
    ggsave(filename = paste0("output/figures/heatmap/", pdb_entry, "_", chain_index, "_heatmap.png"), 
           plot = figure, 
           width = 14, height = 10, dpi = 300,
           bg="white")
  }
}
```


## 3D
### bepipred-3.0
```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_BepiPred_3_with_resno_subset <- data_BepiPred_3_with_resno %>% filter(Accession == pdb_entry) %>% na.omit()
  
  for (chain_index in unique(data_BepiPred_3_with_resno_subset$chain)) {
    data_BepiPred_3_with_resno_subset_chain <- data_BepiPred_3_with_resno_subset %>% filter(chain == chain_index)
    
    min_score <- min(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, na.rm = TRUE)
    median_score <- median(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, na.rm = TRUE)
    max_score <- max(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, na.rm = TRUE)
    
    
    low_to_median <- colorRampPalette(c("blue", "gray95"))(10)
    median_to_high <- colorRampPalette(c("gray95", "red"))(10)
    
    breaks <- seq(min_score, max_score, length.out = 21)  # 20 intervals
    colors <- c(low_to_median, median_to_high)  # 20 colors
    
    color_map <- data.frame(
      ProbabilityBin = cut(seq(min_score, max_score, length.out = 20), 
                           breaks = breaks, include.lowest = TRUE),
      color = colors
    ) %>%
      rbind(c(NA,"black"))
    
    predictions_df <- data_BepiPred_3_with_resno_subset_chain %>%
      mutate(ProbabilityBin = cut(`BepiPred-3.0 score`, 
                                  breaks = breaks, include.lowest = TRUE)
      ) %>%
      group_by(ProbabilityBin) %>% 
      summarize(resno_list = list(paste0(resno,":",chain_index)), .groups = "drop") %>%
      mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " "))) %>%
      left_join(color_map, by = "ProbabilityBin")
    
    view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index, sele_color_df = predictions_df)
    
    #print(view)
    
    saveWidget(view, file = paste0("output/BepiPred3.0/snapshot/score/",pdb_entry, "_", chain_index,"_bepipred3.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```


### discotope-3.0
```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  
  for (chain_index in unique(data_DiscoTope_subset$chain)) {
  data_DiscoTope_subset_chain <- data_DiscoTope_subset %>% filter(chain == chain_index)
  
  min_score <- min(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  median_score <- median(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  max_score <- max(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  
  
  low_to_median <- colorRampPalette(c("blue", "gray95"))(10)
  median_to_high <- colorRampPalette(c("gray95", "red"))(10)
  
  breaks <- seq(min_score, max_score, length.out = 21)  # 20 intervals
  colors <- c(low_to_median, median_to_high)  # 20 colors
  
  color_map <- data.frame(
    ProbabilityBin = cut(seq(min_score, max_score, length.out = 20), 
                         breaks = breaks, include.lowest = TRUE),
    color = colors
  ) %>%
  rbind(c(NA,"black"))
  
  predictions_df <- data_DiscoTope_subset %>%
    mutate(ProbabilityBin = cut(calibrated_score, 
                                breaks = breaks, include.lowest = TRUE)
    ) %>%
    group_by(ProbabilityBin) %>% 
    summarize(resno_list = list(paste0(resno,":",chain_index)), .groups = "drop") %>%
    mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " "))) %>%
    left_join(color_map, by = "ProbabilityBin")
  
  view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index,sele_color_df = predictions_df)
  
  #print(view)
  
  saveWidget(view, file = paste0("output/DiscoTope/snapshot/score/",pdb_entry, "_", chain_index,"_discotope3.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```

### colour epitopes by resno
Colouring predicted epitopes by resno could help identify discontinous epitopes
```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_BepiPred_3_with_resno_subset <- data_BepiPred_3_with_resno %>% filter(Accession == pdb_entry) %>% na.omit()
  
  quantile_threshold <- 0.85  #0.8 == top 20%
  
  for (chain_index in unique(data_BepiPred_3_with_resno_subset$chain)) {
    data_BepiPred_3_with_resno_subset_chain <- data_BepiPred_3_with_resno_subset %>% filter(chain == chain_index)
    
    score_threshold <- quantile(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, 
                                probs = quantile_threshold, na.rm = TRUE)
    
    predictions_df <- data_BepiPred_3_with_resno_subset_chain %>%
      mutate(color = ifelse(`BepiPred-3.0 score` >= score_threshold, "residueindex", "white")) %>%
      group_by(color) %>% 
      summarize(resno_list = list(paste0(resno, ":", chain_index)), .groups = "drop") %>%
      mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " ")))
    
    view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index, sele_color_df = predictions_df)
    
    #print(view)
    
    saveWidget(view, file = paste0("output/BepiPred3.0/snapshot/conformational/",pdb_entry, "_", chain_index,"_resid_bepipred3.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```

```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  
  quantile_threshold <- 0.85  #0.8 == top 20%
  
  for (chain_index in unique(data_DiscoTope_subset$chain)) {
    data_DiscoTope_subset_chain <- data_DiscoTope_subset %>% filter(chain == chain_index)
    
    score_threshold <- quantile(data_DiscoTope_subset_chain$`DiscoTope-3.0_score`, 
                                probs = quantile_threshold, na.rm = TRUE)
    
    predictions_df <- data_DiscoTope_subset_chain %>%
      mutate(color = ifelse(`DiscoTope-3.0_score` >= score_threshold, "residueindex", "white")) %>%
      group_by(color) %>% 
      summarize(resno_list = list(paste0(resno, ":", chain_index)), .groups = "drop") %>%
      mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " ")))
    
    view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index, sele_color_df = predictions_df)
    
    print(view)
    
    saveWidget(view, file = paste0("output/DiscoTope/snapshot/conformational/",pdb_entry, "_", chain_index,"_resid_discotope.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```


## Longest chain
### Getting the longest chain
```{r}
data_predictions_longestchain <- inner_join(
  data_predictions_structurebased %>%
    add_count(PDB, chain, name = "chain_length") %>%  # Add chain length count
    arrange(PDB,desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
    group_by(PDB) %>%
    filter(chain_length == max(chain_length)) %>%  
    filter(chain == min(chain)) %>% 
    ungroup() %>%  
    select(-chain_length),
  data_predictions_sequencebased,
  by = c("PDB", "chain", "resno","resid")
  ) %>%
  
  left_join(
    data_namesclasspdb
  )

data_predictions_longestchain
```

### PCA
```{r}
score_columns <- c("Epigraph_score", "DiscoTope-3.0_calibrated_score",
                   "BepiPred-2.0 probability", 
                   "BepiPred-3.0 score")

score_columns
```

```{r}
scaled_scores <- data_predictions_longestchain %>%
  select(all_of(score_columns)) %>%
  scale()
  
pca_result <- prcomp(t(scaled_scores),center = TRUE, scale. = TRUE)
```


```{r}
explained_variance <- data.frame(
  PC = paste0("PC", 1:length(pca_result$sdev)),
  Variance = (pca_result$sdev^2) / sum(pca_result$sdev^2) * 100
)

# Scree plot
ggplot(explained_variance, aes(x = PC, y = Variance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Scree Plot", x = "Principal Components", y = "Variance Explained (%)") +
  theme_minimal()
```

```{r}
library(ggrepel)

pca_scores <- as.data.frame(pca_result$x) %>%
  rownames_to_column(var = "score_type")

# Scatter plot of PC1 vs PC2
ggplot(pca_scores, aes(x = PC1, y = PC2, label = score_type)) +
  geom_point(alpha = 0.7, color = "blue") +
  geom_text_repel() +
  labs(title = "PCA of concatenated epitope prediction scores across different tools", x = "PC1", y = "PC2") +
  theme_minimal()
```



# (WIP) matching AmiA consensus to to epitope predictions
## AmiA
AmiA corresponds to pdb entries 8A42, 8QM0
```{r}
data.frame(Consensus_Residue = unlist(strsplit(msaConsensusSequence(AmiA_alignment), "")),
           Position = seq_along(unlist(strsplit(msaConsensusSequence(AmiA_alignment), ""))))
```


```{r}
pdb_object <- read.pdb("8A42")

required_adjustment <- map.seqposno.to.resno(sequence_AA1 = msaConsensusSequence(AmiA_alignment), pdb_object = pdb_object, verbose = TRUE)
required_adjustment
```

```{r}
pdb_object$atom %>%
  filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
  ggplot(aes(x=resno, y=chain, col = resid)) +
  geom_vline(xintercept=34) +
  geom_vline(xintercept=50) +
  geom_vline(xintercept=100) +
  geom_point() +
  geom_point(data = data_variability_AmiA ,
             aes(x=Position, col=aa123(Consensus_residue), y="FASTA")
             ) +
  scale_x_continuous(limits = c(NA,50)) +
  #scale_y_discrete(expand = c(0,0)) +
  theme_bw() 
```


## mutations_data_with_resno
```{r}
data_variability <- read.csv("output/data_variability.csv")

data_variability <- data_variability %>%
  left_join(data_namesclasspdb,
            by = c("protein" = "proteinname"))

data_variability
```

### check the missing ones
Currently working with PDB entries only, not all proteins have PDB's associated with them, so some are missing
```{r}
sort(unique(data_predictions$proteinname))
sort(unique(data_variability$protein))

setdiff(
  tolower(unique(data_predictions$proteinname)),
  tolower(unique(data_variability$protein))
)
```



```{r}
data_variability_with_resno <- NULL

for (pdb_entry in unique(data_predictions$PDB)) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_variability_subset <- data_variability %>% filter(PDB == pdb_entry, consensus_res != "*") %>% 
    mutate(Position = position) %>% select(!position)
  
  consensus_sequence <- data_variability_subset %>%
    pull(consensus_res) %>% 
    str_c(collapse = "")
  
  required_adjustment <- map.seqposno.to.resno(sequence_AA1 = consensus_sequence, 
                                               pdb_object = pdb_object)
  
  #print(required_adjustment)

  resmap <- create.resmap(pdb_object$atom, 
                          data_variability_subset,
                          required_adjustment) 
  
  data_variability_with_resno <- rbind(data_variability_with_resno, resmap)
  
}

data_variability_with_resno 
data_variability_with_resno %>%
  select(PDB,resno,resid, consensus_res,var_counts,var_freq,n_seq,seq_length)
```

```{r}
data_predictions_with_variability <- inner_join(data_predictions_longestchain,
                                                data_variability_with_resno %>% 
                                                  select(PDB, chain, resno, resid,  consensus_res, var_counts,
                                                         var_freq, n_seq,seq_length),
                                                by = c("PDB", "chain","resno","resid"))

data_predictions_with_variability
```


## plot
```{r, fig.width = 14}
pivot_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score", 
  "var_freq"
)

for (pdb_entry in unique(data_predictions_with_variability$PDB)) {
  data_subset <- data_predictions_with_variability %>% filter(PDB == pdb_entry) %>%
    select(proteinname,PDB,chain,resno,pivot_list) %>%
  pivot_longer(cols = pivot_list, 
  names_to = "Score_Type", values_to = "Score_Value") %>%
    group_by(Score_Type) %>%
    mutate(normalized_score = case_when(
        Score_Type == "BepiPred-2.0 probability" ~ Score_Value,
        TRUE ~ rescale(Score_Value, to = c(0, 1))
        )
    )
  
  plot <- data_subset %>%
    ggplot(aes(x=resno, y=Score_Value, fill = normalized_score)) +
    geom_col() +
    
    facet_wrap(~Score_Type, ncol = 1, scales = "free_y") +
    
    scale_fill_gradient2(low = "blue", mid = "gray95", high = "red", 
                         midpoint = 0.5) +  # Now uses median per facet
    
    geom_text(aes(label = Score_Type, 
                  x = (min(resno) + (max(resno) - min(resno)) / 2)), 
              color = "black", y = Inf, vjust = 1.5, size = 4, family = "Arial", 
              check_overlap = TRUE) +
    
    labs(
      title = paste0(pdb_entry, " chain ", data_subset$chain[1], " (",data_subset$proteinname[1],")"), 
         x = "Residue number") +
    
    theme_minimal() +
    theme(
      strip.text = element_blank(),
      axis.ticks = element_blank(),  
      axis.ticks.x = element_blank(),  
      axis.ticks.y = element_blank(),
      #axis.text.x = element_blank(),  
      axis.text.y = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
  
  print(plot)
}
```

```{r}
data_predictions_with_variability %>% filter(PDB =="2L4O") 
data_predictions_longestchain %>% filter(PDB =="2L4O") 
data_variability_with_resno %>% filter(PDB =="2L4O") 
```


```{r}
data_predictions_with_variability_normalized <- data_predictions_with_variability %>%
  group_by(PDB) %>%
  mutate(
    across(
      c(`BepiPred-3.0 score` , `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq), 
      rescale
    )
  ) %>%
  ungroup() %>%
  na.omit()

data_predictions_with_variability_normalized
```

```{r}
for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  
  protein_name <- unique(data_subset$proteinname)
 
  for (chain_index in unique(data_subset$chain)) {
    data_subset_chain <- data_subset %>% filter(chain == chain_index)
    
    data_subset_wide <- data_subset_chain %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    #not all sequences have variability 
    if (var(data_subset_wide$var_freq, na.rm = TRUE) == 0) {
      data_subset_wide <- data_subset_wide %>% select(-var_freq)
    }

    plot <- pheatmap(t(data_subset_wide), 
         cluster_rows = TRUE,         
         cluster_cols = FALSE,         
         main = paste0(PDB_entry," chain ", chain_index, " (", protein_name, ")"), 
         color = colorRampPalette(c("blue", "white", "red"))(50),
         labels_col = "resno",
         clustering_distance_rows = "correlation") 
    
    print(plot)
  }
}
```






```{r}
data_variability %>% filter(PDB == "2L4O") %>%
  select(PDB, protein, position, consensus_res) 
```


```{r}
pdb_object$atom %>% 
  select(chain,resno,resid) %>% unique()
```


```{r}
resmap
```




