---
title: "Streptococcus pneumoniae epitope visualization"
author: "Jan Paul van Meenen"
date: "`r Sys.Date()`"
output: html_document
---

#Packages
```{r}
source("requirements/packages.R")
```

#Functions
```{r}
function_files <- list.files(path = "functions", pattern = "\\.R$",full.names = TRUE)

for (function_file in function_files) {
  source(function_file)
}
```

# Plot all together
## Load in from 01
```{r}
data_BepiPred_2_with_resno <- read.csv("output/BepiPred2.0/data_BepiPred2.csv", check.names = FALSE)
data_BepiPred_3_with_resno <- read.csv("output/BepiPred3.0/data_BepiPred3.csv", check.names = FALSE)
data_Epigraph <- read.csv("output/EpiGraph/data_Epigraph.csv", check.names = FALSE)
data_DiscoTope <- read.csv("output/DiscoTope/data_DiscoTope.csv", check.names = FALSE)
```


## create joined predictions df
```{r}
data_predictions_sequencebased <- full_join(
  data_BepiPred_2_with_resno %>%
    mutate(PDB = Entry,
           `BepiPred-2.0 probability` = EpitopeProbability,
           `RSA NetsurfP` = RelativeSurfaceAccessilibity) %>%
    select(PDB, chain, resno,resid,`BepiPred-2.0 probability`,`RSA NetsurfP`) %>%
    filter(resid %in% toupper(AMINO_ACID_CODE))
  ,
  
  data_BepiPred_3_with_resno %>%
    mutate(PDB = Accession,
           seqposno = Position) %>%
    select(PDB, chain, resno, seqposno, resid, `BepiPred-3.0 score`,`BepiPred-3.0 linear epitope score`) %>%
    filter(resid %in% toupper(AMINO_ACID_CODE))
  ,
  by = c("PDB","chain","resno","resid")
)

data_predictions_sequencebased

data_predictions_sequencebased %>%
  filter(if_any(everything(), is.na))
```


```{r}
data_predictions_structurebased <- full_join(
  data_Epigraph %>%
    mutate(Epigraph_score = Score) %>%
    select(PDB, chain, resno, resid, Epigraph_score)
  ,
  data_DiscoTope %>%
    mutate(PDB = pdb,
           resid = aa123(residue),
           `RSA Shrake-Rupley` = rsa,
           `DiscoTope-3.0_calibrated_score` = calibrated_score
    ) %>%
    select(PDB, chain, resno, resid, `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, `RSA Shrake-Rupley`)
  , 
  by = c("PDB","resno","chain","resid")
)
data_predictions_structurebased

data_predictions_structurebased %>%
  filter(if_any(everything(), is.na))

data_predictions_structurebased <- data_predictions_structurebased #%>%
  #na.omit()
  
#data_predictions_structurebased %>%
#  filter(if_any(everything(), is.na))
```


```{r}
data_predictions_structurebased %>% filter(PDB == "2BIB")
```

```{r}
data_predictions <- full_join(
  data_predictions_sequencebased,
  data_predictions_structurebased,
  by = c("PDB", "chain", "resno","resid")
  )

data_predictions %>%
  filter(if_any(everything(), is.na))

#data_predictions <- data_predictions %>%
#  na.omit() 

data_predictions
```

### get the protein names 
names are not consistent between the structural and protein panel data, so they have to be cleaned
```{r}
data_namesclasspdb <- data_structural %>%
  separate_rows(`PDB entry`, sep = ", ") %>%
  mutate(PDB = `PDB entry`,
         structure_proteinname = `Proteins (TIGR4) name`,
         structure_proteinname_clean = ifelse(
           grepl("^SP_", gsub(" ", "", gsub("\\(.*", "", `Proteins (TIGR4) name`))), 
           sub("^(SP_\\d{4}).*", "\\1", gsub(" ", "", gsub("\\(.*", "", `Proteins (TIGR4) name`))),
           sub("_.*", "", gsub(" ", "", gsub("\\(.*", "", `Proteins (TIGR4) name`)))
         )
  ) %>%
  select(structure_proteinname_clean, PDB) %>%
  full_join(
    data_proteinpanel %>% 
      mutate(panel_proteinname = `Proteins (TIGR4)`,
             panel_proteinname_clean = gsub(" ","",gsub("\\(.*", "", panel_proteinname)),
             proteinclass = `Protein class`
      ) %>%
      select(proteinclass, panel_proteinname_clean), keep = TRUE,
    by = c("structure_proteinname_clean" = "panel_proteinname_clean")
  ) %>%
  mutate(protein = structure_proteinname_clean) %>%
  select(proteinclass, protein, PDB)

data_namesclasspdb
```


```{r}
data_predictions <- data_predictions %>%
  left_join(
    data_namesclasspdb
  ) 

data_predictions
```

## normalize
```{r}
data_predictions_normalized <- data_predictions %>%
  group_by(PDB, chain) %>%
  mutate(
    across(
      c(#`BepiPred-2.0 probability`,
        `BepiPred-3.0 score` , `BepiPred-3.0 linear epitope score`, 
        `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, Epigraph_score), 
      rescale
    )
  ) %>%
  ungroup()

data_predictions_normalized
```

###track
```{, warning=FALSE}
track_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score"
)

for (pdb_entry in unique(data_predictions$PDB)) {
  data_predictions_subset <- data_predictions %>%
    filter(PDB == pdb_entry) 
  
  proteinname <- unique(data_predictions_subset$proteinname)
  
  for (chain_index in unique(data_predictions_subset$chain)) {
    
    data_predictions_subset_long <- data_predictions_subset %>%
      filter(chain == chain_index) %>%
      pivot_longer(cols = track_list, 
                   names_to = "Score_Type", values_to = "Score_Value") %>%
      group_by(Score_Type) %>%
      mutate(normalized_score = case_when(
        Score_Type == "BepiPred-2.0 probability" ~ Score_Value,
        TRUE ~ rescale(Score_Value, to = c(0, 1))
        )
      )
    
    figure <- data_predictions_subset_long %>%
      ggplot(aes(x=resno, y=Score_Value, fill = normalized_score)) +
      geom_col() +
      
      facet_wrap(~Score_Type, ncol = 1, scales = "free_y") +
      
      scale_fill_gradient2(low = "blue", mid = "gray95", high = "red", 
                           midpoint = 0.5) +  # Now uses median per facet
      
      geom_text(aes(label = Score_Type, 
                    x = (min(resno) + (max(resno) - min(resno)) / 2)), 
                color = "black", y = Inf, vjust = 1.5, size = 4, family = "Arial", 
                check_overlap = TRUE) +
      
      labs(title = paste0(pdb_entry, " chain ", chain_index, " (",proteinname,")")) +
      
      theme_minimal() +
      theme(
        strip.text = element_blank(),
        axis.ticks = element_blank(),  
        axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      )
    
    print(figure)
    
    ggsave(filename = paste0("output/figures/tracks/", pdb_entry, "_", chain_index, "_track.png"), 
           plot = figure, 
           width = 14, height = 10, dpi = 300,
           bg="white")
  }
}
```


###heatmap (pheatmap)
```{, fig.width = 14}
for (pdb_entry in unique(data_predictions_normalized$PDB)) {
  data_predictions_normalised_subset <- data_predictions_normalized %>% na.omit() %>%
    filter(PDB == pdb_entry) 
  
  proteinname <- unique(data_predictions_normalised_subset$proteinname)
  
  for (chain_index in unique(data_predictions_normalised_subset$chain)) {
    
    data_predictions_normalised_subset_wide <- data_predictions_normalised_subset %>%
      filter(chain == chain_index) %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score) %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    figure <- pheatmap(t(data_predictions_normalised_subset_wide), 
                       cluster_rows = TRUE,         
                       cluster_cols = FALSE,         
                       main = paste0(pdb_entry, " chain ", chain_index, " (",proteinname,")"), 
                       color = colorRampPalette(c("blue", "gray95", "red"))(50),
                       labels_col = "resno",
                       clustering_distance_rows = "correlation") 
    
    print(figure)
    
    ggsave(filename = paste0("output/figures/heatmap/", pdb_entry, "_", chain_index, "_heatmap.png"), 
           plot = figure, 
           width = 14, height = 10, dpi = 300,
           bg="white")
  }
}
```

```{, fig.width=14}
for (pdb_entry in unique(data_predictions_normalized$PDB)[2]) {
  data_predictions_normalised_subset <- data_predictions_normalized %>% na.omit() %>%
    filter(PDB == pdb_entry) 
  
  proteinname <- unique(data_predictions_normalised_subset$proteinname)
  
  for (chain_index in unique(data_predictions_normalised_subset$chain)) {
    
    data_predictions_normalised_subset_wide <- data_predictions_normalised_subset %>%
      filter(chain == chain_index) %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score) %>%
      distinct(resno, .keep_all = TRUE) %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    residues <- as.numeric(rownames(data_predictions_normalised_subset_wide))
    
    colorbar_df <- get_residueindex_colors_v2(residues)
    
    annotation_col <- data.frame(NGL_Color = residues)
    rownames(annotation_col) <- residues
    
    annotation_colors <- list(
      NGL_Color = setNames(colorbar_df$color, colorbar_df$resno)
    )
    
    figure <- pheatmap::pheatmap(t(data_predictions_normalised_subset_wide), 
                       cluster_rows = TRUE,         
                       cluster_cols = FALSE,         
                       main = paste0(pdb_entry, " chain ", chain_index, " (",proteinname,")"), 
                       color = colorRampPalette(c("blue", "gray95", "red"))(50),
                       labels_col = "resno",
                       clustering_distance_rows = "correlation",
                       annotation_col = annotation_col,
                       annotation_colors = annotation_colors,
                       annotation_legend = FALSE, 
                       annotation_names_col = FALSE
    ) 
    
    print(figure)
    
    ggsave(filename = paste0("output/figures/heatmap/", pdb_entry, "_", chain_index, "_heatmap.png"), 
           plot = figure, 
           width = 14, height = 10, dpi = 300,
           bg="white")
  }
}
```

### heatmap (complexheatmap)
```{r}
for (pdb_entry in unique(data_predictions_normalized$PDB)) {
  data_predictions_normalised_subset <- data_predictions_normalized %>% na.omit() %>%
    filter(PDB == pdb_entry) 
  
  proteinname <- unique(data_predictions_normalised_subset$proteinname)
  
  for (chain_index in unique(data_predictions_normalised_subset$chain)) {
    
    data_predictions_normalised_subset_wide <- data_predictions_normalised_subset %>%
      filter(chain == chain_index) %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score) %>%
      distinct(resno, .keep_all = TRUE) %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    residues <- as.numeric(rownames(data_predictions_normalised_subset_wide))
    
    colorbar_df <- get_residueindex_colors(residues, rev = FALSE)
    
    annotation_col <- data.frame(NGL_Color = factor(residues))
    
    annotation_colors <- list(
      NGL_Color = setNames(colorbar_df$color, residues)
    )
    
    figure <- ComplexHeatmap::pheatmap(
      t(data_predictions_normalised_subset_wide), 
      cluster_rows = TRUE,         
      cluster_cols = FALSE,         
      main = paste0(pdb_entry, " chain ", chain_index, " (", proteinname, ")"), 
      color = colorRampPalette(c("blue", "gray95", "red"))(50),
      #labels_col = rownames(data_predictions_normalised_subset_wide),
      clustering_distance_rows = "correlation",
      annotation_col = annotation_col,
      annotation_colors = annotation_colors,
      annotation_legend = FALSE, 
      annotation_names_col = FALSE
    )
    
    
    print(figure)
    
    #ggsave(filename = paste0("output/figures/heatmap/", pdb_entry, "_", chain_index, "_heatmap.png"), 
    #       plot = figure, 
    #       width = 14, height = 10, dpi = 300,
    #       bg="white")
  }
}
```


## 3D
### bepipred-3.0
```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_BepiPred_3_with_resno_subset <- data_BepiPred_3_with_resno %>% filter(Accession == pdb_entry) %>% na.omit()
  
  for (chain_index in unique(data_BepiPred_3_with_resno_subset$chain)) {
    data_BepiPred_3_with_resno_subset_chain <- data_BepiPred_3_with_resno_subset %>% filter(chain == chain_index)
    
    min_score <- min(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, na.rm = TRUE)
    median_score <- median(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, na.rm = TRUE)
    max_score <- max(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, na.rm = TRUE)
    
    
    low_to_median <- colorRampPalette(c("blue", "gray95"))(10)
    median_to_high <- colorRampPalette(c("gray95", "red"))(10)
    
    breaks <- seq(min_score, max_score, length.out = 21)  # 20 intervals
    colors <- c(low_to_median, median_to_high)  # 20 colors
    
    color_map <- data.frame(
      ProbabilityBin = cut(seq(min_score, max_score, length.out = 20), 
                           breaks = breaks, include.lowest = TRUE),
      color = colors
    ) %>%
      rbind(c(NA,"black"))
    
    predictions_df <- data_BepiPred_3_with_resno_subset_chain %>%
      mutate(ProbabilityBin = cut(`BepiPred-3.0 score`, 
                                  breaks = breaks, include.lowest = TRUE)
      ) %>%
      group_by(ProbabilityBin) %>% 
      summarize(resno_list = list(paste0(resno,":",chain_index)), .groups = "drop") %>%
      mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " "))) %>%
      left_join(color_map, by = "ProbabilityBin")
    
    view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index, sele_color_df = predictions_df)
    
    #print(view)
    
    saveWidget(view, file = paste0("output/BepiPred3.0/snapshot/score/",pdb_entry, "_", chain_index,"_bepipred3.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```


### discotope-3.0
```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  
  for (chain_index in unique(data_DiscoTope_subset$chain)) {
  data_DiscoTope_subset_chain <- data_DiscoTope_subset %>% filter(chain == chain_index)
  
  min_score <- min(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  median_score <- median(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  max_score <- max(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  
  
  low_to_median <- colorRampPalette(c("blue", "gray95"))(10)
  median_to_high <- colorRampPalette(c("gray95", "red"))(10)
  
  breaks <- seq(min_score, max_score, length.out = 21)  # 20 intervals
  colors <- c(low_to_median, median_to_high)  # 20 colors
  
  color_map <- data.frame(
    ProbabilityBin = cut(seq(min_score, max_score, length.out = 20), 
                         breaks = breaks, include.lowest = TRUE),
    color = colors
  ) %>%
  rbind(c(NA,"black"))
  
  predictions_df <- data_DiscoTope_subset %>%
    mutate(ProbabilityBin = cut(calibrated_score, 
                                breaks = breaks, include.lowest = TRUE)
    ) %>%
    group_by(ProbabilityBin) %>% 
    summarize(resno_list = list(paste0(resno,":",chain_index)), .groups = "drop") %>%
    mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " "))) %>%
    left_join(color_map, by = "ProbabilityBin")
  
  view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index,sele_color_df = predictions_df)
  
  #print(view)
  
  saveWidget(view, file = paste0("output/DiscoTope/snapshot/score/",pdb_entry, "_", chain_index,"_discotope3.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```

### colour epitopes by resno
Colouring predicted epitopes by resno could help identify discontinous epitopes
```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_BepiPred_3_with_resno_subset <- data_BepiPred_3_with_resno %>% filter(Accession == pdb_entry) %>% na.omit()
  
  quantile_threshold <- 0.85  #0.8 == top 20%
  
  for (chain_index in unique(data_BepiPred_3_with_resno_subset$chain)) {
    data_BepiPred_3_with_resno_subset_chain <- data_BepiPred_3_with_resno_subset %>% filter(chain == chain_index)
    
    score_threshold <- quantile(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, 
                                probs = quantile_threshold, na.rm = TRUE)
    
    predictions_df <- data_BepiPred_3_with_resno_subset_chain %>%
      mutate(color = ifelse(`BepiPred-3.0 score` >= score_threshold, "residueindex", "white")) %>%
      group_by(color) %>% 
      summarize(resno_list = list(paste0(resno, ":", chain_index)), .groups = "drop") %>%
      mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " ")))
    
    view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index, sele_color_df = predictions_df)
    
    #print(view)
    
    saveWidget(view, file = paste0("output/BepiPred3.0/snapshot/conformational/",pdb_entry, "_", chain_index,"_resid_bepipred3.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```

```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  
  quantile_threshold <- 0.85  #0.8 == top 20%
  
  for (chain_index in unique(data_DiscoTope_subset$chain)) {
    data_DiscoTope_subset_chain <- data_DiscoTope_subset %>% filter(chain == chain_index)
    
    score_threshold <- quantile(data_DiscoTope_subset_chain$`DiscoTope-3.0_score`, 
                                probs = quantile_threshold, na.rm = TRUE)
    
    predictions_df <- data_DiscoTope_subset_chain %>%
      mutate(color = ifelse(`DiscoTope-3.0_score` >= score_threshold, "residueindex", "white")) %>%
      group_by(color) %>% 
      summarize(resno_list = list(paste0(resno, ":", chain_index)), .groups = "drop") %>%
      mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " ")))
    
    view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index, sele_color_df = predictions_df)
    
    print(view)
    
    saveWidget(view, file = paste0("output/DiscoTope/snapshot/conformational/",pdb_entry, "_", chain_index,"_resid_discotope.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```


## Longest chain
### Getting the longest chain
```{r}
data_predictions_longestchain <- inner_join(
  data_predictions_structurebased %>%
    add_count(PDB, chain, name = "chain_length") %>%  # Add chain length count
    arrange(PDB,desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
    group_by(PDB) %>%
    filter(chain_length == max(chain_length)) %>%  
    filter(chain == min(chain)) %>% 
    ungroup() %>%  
    select(-chain_length),
  data_predictions_sequencebased,
  by = c("PDB", "chain", "resno","resid")
  ) %>%
  
  left_join(
    data_namesclasspdb
  )

data_predictions_longestchain
```

### PCA
```{r}
score_columns <- c("Epigraph_score", "DiscoTope-3.0_calibrated_score",
                   "BepiPred-2.0 probability", 
                   "BepiPred-3.0 score")

score_columns
```

```{r}
scaled_scores <- data_predictions_longestchain %>%
  select(all_of(score_columns)) %>%
  scale()
  
pca_result <- prcomp(t(scaled_scores),center = TRUE, scale. = TRUE)
```


```{r}
explained_variance <- data.frame(
  PC = paste0("PC", 1:length(pca_result$sdev)),
  Variance = (pca_result$sdev^2) / sum(pca_result$sdev^2) * 100
)

# Scree plot
ggplot(explained_variance, aes(x = PC, y = Variance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Scree Plot", x = "Principal Components", y = "Variance Explained (%)") +
  theme_minimal()
```

```{r}
library(ggrepel)

pca_scores <- as.data.frame(pca_result$x) %>%
  rownames_to_column(var = "score_type")

# Scatter plot of PC1 vs PC2
ggplot(pca_scores, aes(x = PC1, y = PC2, label = score_type)) +
  geom_point(alpha = 0.7, color = "blue") +
  geom_text_repel() +
  labs(title = "PCA of concatenated epitope prediction scores across different tools", x = "PC1", y = "PC2") +
  theme_minimal()
```



# Variability
matching variability consensus sequence to epitope predictions sequences
## variability_data_with_resno
```{r}
data_variability <- read.csv("output/data_variability.csv")

data_variability
```

### check the missing ones
Currently working with PDB entries only, not all proteins have PDB's associated with them, so some are missing
```{r}
sort(unique(data_predictions$protein))
sort(unique(data_variability$protein))

setdiff(
  tolower(unique(data_predictions$protein)),
  tolower(unique(data_variability$protein))
)
```

## msa pdbseq with protein
```{r}

```

FIXME: Currently we filter out all consensus_res with character "-", but it might be better to just align the pdb sequence to the variability profile

## match consensus seq to pdbseq
```{}
data_variability_with_resno <- NULL

for (pdb_entry in na.omit(unique(data_variability$PDB))) {
  print(pdb_entry)
  
  tryCatch({
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_variability_subset <- data_variability %>% filter(PDB == pdb_entry)
  
  consensus_sequence <- str_c(
    data_variability_subset %>% filter(consensus_res != "*") %>% pull(consensus_res),
    collapse = ""
  )
  
  alignment_results <- map.seq.to.pdb(consensus_sequence, pdb_object)
  
  ca_atoms <- pdb_object$atom %>%
    filter(
      type == "ATOM",
      elety == "CA",
      alt %in% c(NA, "A"),
      resid %in% toupper(AMINO_ACID_CODE)
    )
  
  resmap_variability <- create.resmap.variability(ca_atoms, data_variability_subset, alignment_results)

  if (is.null(data_variability_with_resno)) {
    data_variability_with_resno <- resmap_variability
  } else {
    # Otherwise, rbind the current data to the combined data frame
    data_variability_with_resno <- rbind(data_variability_with_resno, resmap_variability)
  }
  
  }, error = function(e) {
    warning("Failed on PDB entry: ", pdb_entry, "\nReason: ", e$message)
  })
}

data_variability_with_resno 
#data_variability_with_resno %>%
#  #select(PDB,resno,resid, consensus_res,var_counts,var_freq,n_seq,seq_length) %>%
#  filter(if_any(everything(), is.na))
```

```{, test}
for (pdb_entry in unique(data_predictions$PDB)[2]) {
  print(pdb_entry)
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_variability_subset <- data_variability %>% filter(PDB == pdb_entry, consensus_res != "*") %>% 
    mutate(Position = position) %>% select(!position)
  
  consensus_sequence <- data_variability_subset %>%
    pull(consensus_res) %>% 
    str_c(collapse = "")
  
  required_adjustment <- map.seqposno.to.resno(sequence_AA1 = consensus_sequence, 
                                               pdb_object = pdb_object,
                                               verbose = TRUE)
}
```

```{}
data_variability_subset
consensus_sequence

alignment_results <- map.seq.to.pdb(consensus_sequence, pdb_object, alignment_method = "overlap")
alignment_results
as.character(alignedPattern(alignment_results))
as.character(alignedSubject(alignment_results))
```


```{}
#old
data_predictions_with_variability <- full_join(
  data_predictions,
  data_variability_with_resno %>%
    filter(!is.na(resid), consensus_res.x != "-") %>%
    mutate(resid = aa123(resid),
           consensus_res = consensus_res.x) %>%
    select(proteinclass, protein, PDB, resno, resid, 
           consensus_res, position, var_counts, var_freq, n_seq,seq_length),
  by = c("proteinclass","protein","PDB","resno","resid"))

data_predictions_with_variability
```

```{r}
#new
data_predictions_with_variability <- full_join(
  data_predictions,
  aligned_resno %>% select(-chain) %>% na.omit(),
  by = c("proteinclass","protein","PDB","resno","resid"))

data_predictions_with_variability
```


```{r}
data_predictions_with_variability %>% filter(PDB == "3ZFJ") %>%
  select(proteinclass, protein, PDB, chain, resno, resid, consensus_res, seqposno) 

data_variability %>% filter(PDB == "3ZFJ")
data_predictions_longestchain %>% filter(PDB == "3ZFJ")
```

#### do actual alignment
maybe do this for all pdb sequences?
```{}
#aln_protein_names

mismatch_map <- data_variability %>% select(PDB, protein) %>% filter(PDB %in% mismatch_pdb_list) %>% distinct(PDB, .keep_all = TRUE)

for (pdb_entry in mismatch_map$PDB) {
  print(pdb_entry)
  protein_name <- mismatch_map$protein[mismatch_map$PDB == pdb_entry]
  matched_filename <- names(aln_protein_names)[aln_protein_names == protein_name]
  
  aa_file <- file.path(AA_fasfiles_path, matched_filename)
  #print(matched_file)
  
  aa_seqs <- RemoveGaps(readAAStringSet(aa_file)) 
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  longest_chain <- pdb_object$atom %>%
    filter(type == "ATOM",
           elety == "CA",
           alt %in% c(NA, "A"),
           resid %in% toupper(AMINO_ACID_CODE)) %>%
    add_count(chain, name = "chain_length") %>%
    arrange(desc(chain_length), chain) %>%
    filter(chain_length == max(chain_length)) %>%
    filter(chain == min(chain)) %>%
    pull(chain) %>%
    unique()
  
  ca_atoms <- pdb_object$atom %>%
    filter(
      type == "ATOM",
      elety == "CA",
      chain == longest_chain,
      alt %in% c(NA, "A"),
      resid %in% toupper(AMINO_ACID_CODE)
    )
  
  new_seq <- str_c(aa321(ca_atoms$resid), collapse = "")
  
  new_seq_set <- AAStringSet(setNames(new_seq, pdb_entry))
  
  combined_seqs <- c(new_seq_set, aa_seqs)
  
  aligned_combined <- AlignSeqs(combined_seqs, terminalGap=-1)
  
  aligned_new_seq <- aligned_combined[[pdb_entry]]
}
```

##### all pdbs to same protein together
```{}
for (protein_name in unique(mismatch_map$protein)[4]) {
  cat("\nProcessing protein:", protein_name, "\n")
  
  # Match the filename for the current protein
  matched_filename <- names(aln_protein_names)[aln_protein_names == protein_name]
  
  if (length(matched_filename) == 0) {
    warning("No matching .fas file for protein: ", protein_name)
    next
  }
  
  # Load existing alignment and remove gaps
  aa_file <- file.path(AA_fasfiles_path, matched_filename)
  aa_seqs <- RemoveGaps(readAAStringSet(aa_file))
  
  # Get PDB entries corresponding to this protein
  pdb_entries <- mismatch_map %>% filter(protein == protein_name) %>% pull(PDB)
  
  # Collect all sequences from PDB entries
  pdb_strings <- c()
  
  for (pdb_entry in pdb_entries) {
    print(pdb_entry)
    pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
    
    longest_chain <- pdb_object$atom %>%
      filter(type == "ATOM",
             elety == "CA",
             alt %in% c(NA, "A"),
             resid %in% toupper(AMINO_ACID_CODE)) %>%
      add_count(chain, name = "chain_length") %>%
      arrange(desc(chain_length), chain) %>%
      filter(chain_length == max(chain_length)) %>%
      slice_min(chain, with_ties = FALSE) %>%
      pull(chain)

    
    if (length(longest_chain) != 1) {
      warning("Ambiguous chain in PDB:", pdb_entry)
      next
    }
    
    ca_atoms <- pdb_object$atom %>%
      filter(type == "ATOM",
             elety == "CA",
             chain == longest_chain,
             alt %in% c(NA, "A"),
             resid %in% toupper(AMINO_ACID_CODE))
    
    new_seq <- str_c(aa321(ca_atoms$resid), collapse = "")
    names(new_seq) <- pdb_entry
    pdb_strings <- c(pdb_strings, new_seq)
  }
  
  pdb_stringset <- AAStringSet(pdb_strings, use.names = TRUE)

  # Combine and align
  combined_seqs <- c(pdb_stringset, aa_seqs)
  #aligned_combined <- AlignSeqs(combined_seqs,  gapOpening = -5, gapExtension = -2, terminalGap = -0.1, substitutionMatrix = "BLOSUM80")
  aligned_combined <- AlignSeqs(combined_seqs, terminalGap = -0.1)
  
  # Optionally: Extract only aligned PDB entries
  aligned_new_seqs <- aligned_combined[names(pdb_strings)]
  
  # Inspect or save
  print(aligned_new_seqs)
}

```



```{}
for (pdb_entry in mismatch_pdb_list[3]) {
  print(pdb_entry)
  
  tryCatch({
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_variability_subset <- data_variability %>% filter(PDB == pdb_entry)
  
  consensus_sequence <- str_c(
    data_variability_subset %>% filter(consensus_res != "*") %>% pull(consensus_res),
    collapse = ""
  )
  
  alignment_results <- map.seq.to.pdb(consensus_sequence, pdb_object)
  
  ca_atoms <- pdb_object$atom %>%
    filter(
      type == "ATOM",
      elety == "CA",
      alt %in% c(NA, "A"),
      resid %in% toupper(AMINO_ACID_CODE)
    )
  
  resmap_variability <- create.resmap.variability(ca_atoms, data_variability_subset, alignment_results)

  #if (is.null(data_variability_with_resno)) {
  #  data_variability_with_resno <- resmap_variability
  #} else {
  #  # Otherwise, rbind the current data to the combined data frame
  #  data_variability_with_resno <- rbind(data_variability_with_resno, resmap_variability)
  #}
  #
  #}, error = function(e) {
  #  warning("Failed on PDB entry: ", pdb_entry, "\nReason: ", e$message)
  })
}
```





## plot
### tracks
```{r,warning=FALSE, fig.width = 14}
pivot_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score", 
  "var_freq"
)

for (pdb_entry in na.omit(unique(data_predictions_with_variability$PDB))) {
  data_subset <- data_predictions_with_variability %>%    
    filter(PDB == pdb_entry)  
  
    longest_chain <- data_subset %>% 
      filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
    add_count(chain, name = "chain_length") %>%
    arrange(desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
    filter(chain_length == max(chain_length)) %>%  
    filter(chain == min(chain)) %>% 
    pull(chain) %>%
    unique()
    
    data_subset_long <- data_subset %>% 
      pivot_longer(cols = pivot_list, 
                   names_to = "Score_Type", values_to = "Score_Value") %>%
      group_by(Score_Type) %>%
      mutate(normalized_score = case_when(
        Score_Type == "BepiPred-2.0 probability" ~ Score_Value,
        TRUE ~ rescale(Score_Value, to = c(0, 1))
      )
      )
 
  
  plot <- data_subset_long %>% filter(chain == longest_chain | is.na(chain)) %>%
    ggplot(aes(x=resno, y=Score_Value, fill = normalized_score)) +
    geom_col() +
    
    facet_wrap(~Score_Type, ncol = 1, scales = "free_y") +
    
    scale_fill_gradient2(low = "blue", mid = "gray95", high = "red", 
                         midpoint = 0.5) +  # Now uses median per facet
    
    geom_text(aes(label = Score_Type, 
                  x = (min(resno) + (max(resno) - min(resno)) / 2)), 
              color = "black", y = Inf, vjust = 1.5, size = 4, family = "Arial", 
              check_overlap = TRUE) +
    
    labs(
      title = paste0(pdb_entry, " chain ", data_subset$chain[1], " (",data_subset$proteinname[1],")"), 
         x = "Residue number") +
    
    theme_minimal() +
    theme(
      strip.text = element_blank(),
      axis.ticks = element_blank(),  
      axis.ticks.x = element_blank(),  
      axis.ticks.y = element_blank(),
      #axis.text.x = element_blank(),  
      axis.text.y = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
  
  print(plot)
  
  ggsave(filename = paste0("output/figures/variability/", pdb_entry, "_", longest_chain,
                           "_track.png"),
         plot = plot, 
         width = 14, height = 10, dpi = 300,
         bg="white")
}
```

```{r}
data_predictions_with_variability_normalized <- data_predictions_with_variability %>%
  group_by(PDB) %>%
  mutate(
    across(
      c(`BepiPred-3.0 score` , `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, Epigraph_score), 
      rescale
    )
  ) %>%
  ungroup()

data_predictions_with_variability_normalized
```

### heatmap
```{, fig.width=14}
for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  
  protein_name <- unique(data_subset$protein)
  
  print(PDB_entry)
 
  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)
    
    data_subset_wide <- data_subset_chain %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
      #distinct() %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    bar_data <- data_subset_wide$var_freq
    names(bar_data) <- rownames(data_subset_wide)
    
    data_subset_wide <- data_subset_wide %>% select(-var_freq)
    
    heatmap_matrix <- t(data_subset_wide)
    
    top_annot <- HeatmapAnnotation(
      `Variability` = anno_barplot(
        bar_data, 
        gp = gpar(fill = "gray30"), 
        border = FALSE, 
        bar_width = 1, 
        ylim = c(0,1),
        axis_param = list(
          at = seq(0, 1, by = 0.2),
          labels = seq(0, 1, by = 0.2)
        )
      ),
      show_annotation_name = TRUE,
      annotation_name_side = "left",
      height = unit(4, "cm")
    )
    
    residues <- as.numeric(colnames(heatmap_matrix))
    
    colorbar_df <- get_residueindex_colors(residues, rev = FALSE)
    
    annotation_df <- data.frame(NGL_Color = as.character(residues))
    rownames(annotation_df) <- as.character(residues)
    
    annotation_colors <- list(
      NGL_Color = setNames(colorbar_df$color, as.character(residues))
    )
    
    bottom_annot <- HeatmapAnnotation(
      `Residue Index` = annotation_df$NGL_Color,
      col = list(`Residue Index` = annotation_colors$NGL_Color),
      annotation_name_side = "right",
      show_legend = FALSE,
      which = "column"
    )

    ht <- Heatmap(
      heatmap_matrix,
      name = "Score",
      cluster_rows = TRUE,
      cluster_columns = FALSE,
      column_title = paste0(PDB_entry, " chain ", chain_index, " (", protein_name, ")"),
      col = colorRampPalette(c("blue", "gray95", "red"))(50),
      top_annotation = top_annot,
      bottom_annotation = bottom_annot,
      show_column_names = FALSE
    )
    
    figure <- draw(ht, heatmap_legend_side = "right", annotation_legend_side = "right")
    
    print(figure)
    
    png(filename = paste0("output/figures/heatmap/variability/", PDB_entry, "_", chain_index, "_heatmap.png"),
        width = 14, height = 10, units = "in", res = 300, bg = "white")
    
    draw(ht, heatmap_legend_side = "right", annotation_legend_side = "right")
    
    dev.off()

  }
}
```


#### longestchain
```{r, fig.width = 14}
for (PDB_entry in na.omit(unique(data_predictions_with_variability_normalized$PDB))) {
  print(PDB_entry)
  
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  
  protein_name <- unique(data_subset$protein)
  
  longest_chain <- data_subset %>% 
    filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
    add_count(chain, name = "chain_length") %>%
    arrange(desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
    filter(chain_length == max(chain_length)) %>%  
    filter(chain == min(chain)) %>% 
    pull(chain) %>%
    unique()
  
  print(longest_chain)
  data_subset_chain <- data_subset %>% filter(chain == longest_chain)
  
  data_subset_wide <- data_subset_chain %>%
    select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
           `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
    #distinct() %>%
    column_to_rownames("resno") %>%
    as.data.frame()
  
  bar_data <- data_subset_wide$var_freq
  names(bar_data) <- rownames(data_subset_wide)
  
  data_subset_wide <- data_subset_wide %>% select(-var_freq)
  
  heatmap_matrix <- t(data_subset_wide)
  
  top_annot <- HeatmapAnnotation(
    `Variability` = anno_barplot(
      bar_data, 
      gp = gpar(fill = "gray30"), 
      border = FALSE, 
      bar_width = 1, 
      ylim = c(0,1),
      axis_param = list(
        at = seq(0, 1, by = 0.2),
        labels = seq(0, 1, by = 0.2)
      )
    ),
    show_annotation_name = TRUE,
    annotation_name_side = "left",
    height = unit(4, "cm")
  )
  
  residues <- as.numeric(colnames(heatmap_matrix))
  
  colorbar_df <- get_residueindex_colors(residues, rev = FALSE)
  
  annotation_df <- data.frame(NGL_Color = as.character(residues))
  rownames(annotation_df) <- as.character(residues)
  
  annotation_colors <- list(
    NGL_Color = setNames(colorbar_df$color, as.character(residues))
  )
  
  bottom_annot <- HeatmapAnnotation(
    `Residue Index` = annotation_df$NGL_Color,
    col = list(`Residue Index` = annotation_colors$NGL_Color),
    annotation_name_side = "right",
    show_legend = FALSE,
    which = "column"
  )
  
  ht <- Heatmap(
    heatmap_matrix,
    name = "Score",
    cluster_rows = TRUE,
    cluster_columns = FALSE,
    column_title = paste0(PDB_entry, " chain ", longest_chain, " (", protein_name, ")"),
    col = colorRampPalette(c("blue", "gray95", "red"))(50),
    top_annotation = top_annot,
    bottom_annotation = bottom_annot,
    show_column_names = FALSE
  )
  
  figure <- draw(ht, heatmap_legend_side = "right", annotation_legend_side = "right")
  
  print(figure)
  
  png(filename = paste0("output/figures/heatmap/variability/", PDB_entry, "_", longest_chain, "_heatmap.png"),
      width = 14, height = 10, units = "in", res = 300, bg = "white")
  
  dev.off()
  
  
}
```


#### clustered
```{, fig.width=14}
for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  
  protein_name <- unique(data_subset$proteinname)
  
  print(PDB_entry)
 
  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)
    
    data_subset_wide <- data_subset_chain %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
      distinct() %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    bar_data <- data_subset_wide$var_freq
    names(bar_data) <- rownames(data_subset_wide)
    
    data_subset_wide <- data_subset_wide %>% select(-var_freq, -`BepiPred-2.0 probability`)
    
    heatmap_matrix <- t(data_subset_wide)
    
    top_annot <- HeatmapAnnotation(
      `Variability` = anno_barplot(
        bar_data, 
        gp = gpar(fill = "gray30"), 
        border = FALSE, 
        bar_width = 1, 
        ylim = c(0,1),
        axis_param = list(
          at = seq(0, 1, by = 0.2),
          labels = seq(0, 1, by = 0.2)
        )
      ),
      show_annotation_name = TRUE,
      annotation_name_side = "left",
      height = unit(4, "cm")
    )
    
    residues <- as.numeric(colnames(heatmap_matrix))
    
    colorbar_df <- get_residueindex_colors(residues, rev = FALSE)
    
    annotation_df <- data.frame(NGL_Color = as.character(residues))
    rownames(annotation_df) <- as.character(residues)
    
    annotation_colors <- list(
      NGL_Color = setNames(colorbar_df$color, as.character(residues))
    )
    
    bottom_annot <- HeatmapAnnotation(
      `Residue Index` = annotation_df$NGL_Color,
      col = list(`Residue Index` = annotation_colors$NGL_Color),
      annotation_name_side = "right",
      show_legend = FALSE,
      which = "column"
    )

    ht <- Heatmap(
      heatmap_matrix,
      name = "Score",
      cluster_rows = TRUE,
      cluster_columns = TRUE,
      column_title = paste0(PDB_entry, " chain ", chain_index, " (", protein_name, ")"),
      col = colorRampPalette(c("blue", "gray95", "red"))(50),
      top_annotation = top_annot,
      bottom_annotation = bottom_annot,
      show_column_names = FALSE
    )
    
    figure <- draw(ht, heatmap_legend_side = "right", annotation_legend_side = "right")
    
    print(figure)
    
    #ggsave(filename = paste0("output/figures/heatmap/variability/", pdb_entry, "_", chain_index, "_heatmap.png"), 
    #       plot = figure, 
    #       width = 14, height = 10, dpi = 300,
    #       bg="white")
  }
}
```

```{, fig.with = 14}
maxdist <- 10

for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)[3]) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  
  protein_name <- unique(data_subset$proteinname)
  
  print(PDB_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
 
  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)
    
    data_subset_wide <- data_subset_chain %>%
      select(resno, `BepiPred-2.0 probability`, `BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
      distinct() %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    bar_data <- data_subset_wide$var_freq
    names(bar_data) <- rownames(data_subset_wide)
    
    # Remove the var_freq and 'BepiPred-2.0 probability' columns for the heatmap data
    data_subset_wide <- data_subset_wide %>% select(-var_freq, -`BepiPred-2.0 probability`)
    
    heatmap_matrix <- t(data_subset_wide)
    
    top_annot <- HeatmapAnnotation(
      `Variability` = anno_barplot(
        bar_data, 
        gp = gpar(fill = "gray30"), 
        border = FALSE, 
        bar_width = 1, 
        ylim = c(0, 1),
        axis_param = list(
          at = seq(0, 1, by = 0.2),
          labels = seq(0, 1, by = 0.2)
        )
      ),
      show_annotation_name = TRUE,
      annotation_name_side = "left",
      height = unit(4, "cm")
    )
    
    residues <- as.numeric(colnames(heatmap_matrix))
    
    resnos <- pdb_object$atom %>% filter(chain == chain_id, elety == "CA", type == "ATOM") %>% pull(resno)
    
    # Create the xyz distance matrix for residues
    ca.inds <- atom.select(pdb_object, chain = chain_index, elety = "CA", type = "ATOM")
    coords <- pdb_object$xyz[ca.inds$xyz]
    
    dist.mat <- dist.xyz(coords)
    
    rownames(dist.mat) <- as.character(resnos)
    colnames(dist.mat) <- as.character(resnos)
    
    shared <- intersect(resnos, residues)
    setdiff(resnos, residues)
    
    dist.mat <- dist.mat[shared,shared]
    
    #rownames(dist.mat) <- residues
    #colnames(dist.mat) <- residues
    
    # Perform hierarchical clustering on the rows
    row_clustering <- hclust(as.dist(dist.mat), method = "complete")  # You can change the method to "single", "average", etc.
    
    # Order the residues based on the clustering
    heatmap_matrix <- heatmap_matrix[, intersect(resnos, residues)]
    heatmap_matrix <- heatmap_matrix[, row_clustering$order]
    
    colorbar_df <- get_residueindex_colors(residues, rev = FALSE)
    
    annotation_df <- data.frame(NGL_Color = as.character(residues))
    rownames(annotation_df) <- as.character(residues)
    
    annotation_colors <- list(
      NGL_Color = setNames(colorbar_df$color, as.character(residues))
    )
    
    bottom_annot <- HeatmapAnnotation(
      `Residue Index` = annotation_df$NGL_Color,
      col = list(`Residue Index` = annotation_colors$NGL_Color),
      annotation_name_side = "right",
      show_legend = FALSE,
      which = "column"
    )

    # Plot heatmap with hierarchical clustering of rows
    ht <- Heatmap(
      heatmap_matrix,
      name = "Score",
      cluster_rows = row_clustering,  # Use pre-clustered rows (row_clustering)
      cluster_columns = TRUE,  # Optional: Cluster columns as well
      column_title = paste0(PDB_entry, " chain ", chain_index, " (", protein_name, ")"),
      col = colorRampPalette(c("blue", "gray95", "red"))(50),
      top_annotation = top_annot,
      bottom_annotation = bottom_annot,
      show_column_names = FALSE
    )
    
    figure <- draw(ht, heatmap_legend_side = "right", annotation_legend_side = "right")
    
    print(figure)
    
    # Optionally save the plot
    # ggsave(filename = paste0("output/figures/heatmap/variability/", pdb_entry, "_", chain_index, "_heatmap.png"), 
    #        plot = figure, 
    #        width = 14, height = 10, dpi = 300,
    #        bg="white")
  }
}

```

#### with domain annotation
```{r}
library(colorspace)
all_domain_descriptions <- data_interpro_filtered %>%
  arrange(Description) %>%
  pull(Description) %>%
  unique() 

#domain_palette <- qualitative_hcl(length(all_domain_descriptions), palette = "Viridis")
domain_palette <- viridis(length(all_domain_descriptions), option="D")

domain_description_colors <- setNames(
  domain_palette, 
  all_domain_descriptions
  )

show_col(domain_description_colors)

```



```{, fig.width = 14}
for (PDB_entry in na.omit(unique(data_predictions_with_variability_normalized$PDB))) {
  print(PDB_entry)
  
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  
  protein_name <- unique(data_subset$protein)
  
  longest_chain <- data_subset %>% 
    filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
    add_count(chain, name = "chain_length") %>%
    arrange(desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
    filter(chain_length == max(chain_length)) %>%  
    filter(chain == min(chain)) %>% 
    pull(chain) %>%
    unique()
  
  print(longest_chain)
  data_subset_chain <- data_subset %>% filter(chain == longest_chain)
  
  data_subset_wide <- data_subset_chain %>%
    select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
           `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
    #distinct() %>%
    column_to_rownames("resno") %>%
    as.data.frame()
  
  bar_data <- data_subset_wide$var_freq
  names(bar_data) <- rownames(data_subset_wide)
  
  data_subset_wide <- data_subset_wide %>% select(-var_freq)
  
  heatmap_matrix <- t(data_subset_wide)
  
  top_annot <- HeatmapAnnotation(
    `Variability` = anno_barplot(
      bar_data, 
      gp = gpar(fill = "gray30"), 
      border = FALSE, 
      bar_width = 1, 
      ylim = c(0,1),
      axis_param = list(
        at = seq(0, 1, by = 0.2),
        labels = seq(0, 1, by = 0.2)
      )
    ),
    show_annotation_name = TRUE,
    annotation_name_side = "left",
    height = unit(4, "cm")
  )
  
  residues <- as.numeric(colnames(heatmap_matrix))
  
  colorbar_df <- get_residueindex_colors(residues, rev = FALSE)
  
  annotation_df <- data.frame(NGL_Color = as.character(residues))
  rownames(annotation_df) <- as.character(residues)
  
  annotation_colors <- list(
    NGL_Color = setNames(colorbar_df$color, as.character(residues))
  )
  
  interpro_domains <- data_interpro_filtered %>%
    filter(Protein_accession == PDB_entry) %>%
    mutate(Start_resno = Start + data_subset$resno[1] - data_subset$seqposno[1],
           End_resno = End + data_subset$resno[1] - data_subset$seqposno[1]) %>%
    select(Start_resno, End_resno, Description) %>%
    pmap_dfr(function(Start_resno, End_resno, Description) {
      tibble(
        resno = seq(Start_resno, End_resno),
        Description = Description
      )
    })
  
  interpro_domains
  
  domain_vector <- rep(NA_character_, length(residues))
  names(domain_vector) <- as.character(residues)
  
  matching_indices <- match(interpro_domains$resno, residues)
  domain_vector[matching_indices[!is.na(matching_indices)]] <- interpro_domains$Description[!is.na(matching_indices)]
  
  bottom_annot <- HeatmapAnnotation(
  `Residue Index` = annotation_df$NGL_Color,
  `Domain Annotation` = domain_vector,
  col = list(
    `Residue Index` = annotation_colors$NGL_Color,
    `Domain Annotation` = domain_description_colors[na.omit(unique(domain_vector))]
  ),
  annotation_name_side = "right",
  show_legend = c(FALSE, TRUE),
  na_col = "white",
  which = "column"
)

  ht <- Heatmap(
    heatmap_matrix,
    name = "Score",
    cluster_rows = TRUE,
    cluster_columns = FALSE,
    column_title = paste0(PDB_entry, " chain ", longest_chain, " (", protein_name, ")"),
    col = colorRampPalette(c("blue", "gray95", "red"))(50),
    top_annotation = top_annot,
    bottom_annotation = bottom_annot,
    show_column_names = FALSE
  )
  
  png(filename = paste0("output/figures/heatmap/domain/", PDB_entry, "_", longest_chain, "_heatmap.png"),
      width = 14, height = 10, units = "in", res = 300, bg = "white")
  
  draw(ht, heatmap_legend_side = "right", annotation_legend_side = "right")
  
  dev.off()
  
}
```



```{r, fig.width=14}
for (PDB_entry in na.omit(unique(data_predictions_with_variability_normalized$PDB))) {
  print(PDB_entry)
  
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  
  protein_name <- unique(data_subset$protein)
  
  longest_chain <- data_subset %>% 
    filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
    add_count(chain, name = "chain_length") %>%
    arrange(desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
    filter(chain_length == max(chain_length)) %>%  
    filter(chain == min(chain)) %>% 
    pull(chain) %>%
    unique()
  
  print(longest_chain)
  data_subset_chain <- data_subset %>% filter(chain == longest_chain)
  
  data_subset_wide <- data_subset_chain %>%
    select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
           `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
    #distinct() %>%
    column_to_rownames("resno") %>%
    as.data.frame()
  
  bar_data <- data_subset_wide$var_freq
  names(bar_data) <- rownames(data_subset_wide)
  
  data_subset_wide <- data_subset_wide %>% select(-var_freq)
  
  heatmap_matrix <- t(data_subset_wide)
  
  top_annot <- HeatmapAnnotation(
    `Variability` = anno_barplot(
      bar_data, 
      gp = gpar(fill = "gray30"), 
      border = FALSE, 
      bar_width = 1, 
      ylim = c(0,1),
      axis_param = list(
        at = seq(0, 1, by = 0.2),
        labels = seq(0, 1, by = 0.2)
      )
    ),
    show_annotation_name = TRUE,
    annotation_name_side = "left",
    height = unit(4, "cm")
  )
  
  residues <- as.numeric(colnames(heatmap_matrix))
  
  colorbar_df <- get_residueindex_colors(residues, rev = FALSE)
  
  annotation_df <- data.frame(NGL_Color = as.character(residues))
  rownames(annotation_df) <- as.character(residues)
  
  annotation_colors <- list(
    NGL_Color = setNames(colorbar_df$color, as.character(residues))
  )
  
  interpro_domains <- data_interpro_filtered %>%
    filter(Protein_accession == PDB_entry) %>%
    mutate(Start_resno = Start + data_subset$resno[1] - data_subset$seqposno[1],
           End_resno = End + data_subset$resno[1] - data_subset$seqposno[1]) %>%
    select(Start_resno, End_resno, Description) %>%
    pmap_dfr(function(Start_resno, End_resno, Description) {
      tibble(
        resno = seq(Start_resno, End_resno),
        Description = Description
      )
    })
  
  interpro_domains
  
  domain_vector <- rep(NA_character_, length(residues))
  names(domain_vector) <- as.character(residues)
  
  matching_indices <- match(interpro_domains$resno, residues)
  domain_vector[matching_indices[!is.na(matching_indices)]] <- interpro_domains$Description[!is.na(matching_indices)]
  
  # Get unique non-NA domain descriptions
  unique_domains <- na.omit(unique(domain_vector))
  domain_colors <- domain_description_colors[unique_domains]
  
  # Get blocks for each domain
  domain_blocks <- split(names(domain_vector)[!is.na(domain_vector)], domain_vector[!is.na(domain_vector)])
  domain_blocks_ranges <- lapply(domain_blocks, function(res_indices) {
    indices <- which(names(domain_vector) %in% res_indices)
    range(indices)
  })
  
  # Create align_to list: each domain name maps to its residue indices (as numeric positions)
  domain_align_to <- lapply(domain_blocks_ranges, function(rng) seq(rng[1], rng[2]))
  
  annotation_domain <- anno_block(
    align_to = domain_align_to,
    panel_fun = function(index, label) {
      grid.rect(gp = gpar(fill = domain_colors[[label]], col = "black"))
      block_width <- length(index)
      # Scale font size: base 6 + 1.2 * sqrt(width), capped at 15
      font_size <- min(1 + 1.2 * sqrt(block_width), 15)
      grid.text(label, 0.5, 0.5, gp = gpar(col = "white", fontsize = font_size))
    },
    height = unit(1, "cm"), show_name = TRUE
  )
  
  # Final bottom annotation with anno_block and Residue Index
  bottom_annot <- HeatmapAnnotation(
    `Residue Index` = annotation_df$NGL_Color,
    `Domain hit` = annotation_domain,
    col = list(
      `Residue Index` = annotation_colors$NGL_Color
    ),
    annotation_name_side = "right",
    show_legend = c(FALSE, FALSE),
    na_col = "white",
    which = "column"
  )
  
  
  ht <- Heatmap(
    heatmap_matrix,
    name = "Score",
    cluster_rows = TRUE,
    cluster_columns = FALSE,
    column_title = paste0(PDB_entry, " chain ", longest_chain, " (", protein_name, ")"),
    col = colorRampPalette(c("blue", "gray95", "red"))(50),
    top_annotation = top_annot,
    bottom_annotation = bottom_annot,
    show_column_names = FALSE
  )
  
  png(filename = paste0("output/figures/heatmap/domain/", PDB_entry, "_", longest_chain, "_heatmap.png"),
      width = 14, height = 10, units = "in", res = 300, bg = "white")
  
  figure <- draw(ht, heatmap_legend_side = "right", annotation_legend_side = "right")
  
  dev.off()
  
  print(figure)
}
```

FIXME: Refactor complexheatmap code
```{,fig.width=14}
# Helper function: get longest chain from a data subset
get_longest_chain <- function(data_subset) {
  data_subset %>%
    filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
    add_count(chain, name = "chain_length") %>%
    arrange(desc(chain_length), chain) %>%
    filter(chain_length == max(chain_length)) %>%
    slice_min(chain) %>%
    pull(chain) %>%
    unique()
}

# Helper: create domain annotation vector and colors
get_domain_annotation <- function(residues, interpro_data, resno_shift, domain_colors) {
  domains <- interpro_data %>%
    mutate(
      Start_resno = Start + resno_shift,
      End_resno = End + resno_shift
    ) %>%
    select(Start_resno, End_resno, Description) %>%
    pmap_dfr(function(Start_resno, End_resno, Description) {
      tibble(resno = seq(Start_resno, End_resno), Description = Description)
    })


  domain_vector <- rep(NA_character_, length(residues))
  names(domain_vector) <- as.character(residues)
  matching_indices <- match(domains$resno, residues)
  domain_vector[matching_indices[!is.na(matching_indices)]] <- domains$Description[!is.na(matching_indices)]

  unique_domains <- na.omit(unique(domain_vector))
  domain_blocks <- split(names(domain_vector)[!is.na(domain_vector)], domain_vector[!is.na(domain_vector)])
  domain_blocks_ranges <- lapply(domain_blocks, function(res_indices) {
    indices <- which(names(domain_vector) %in% res_indices)
    range(indices)
  })
  domain_align_to <- lapply(domain_blocks_ranges, function(rng) seq(rng[1], rng[2]))

  annotation <- anno_block(
    align_to = domain_align_to,
    panel_fun = function(index, label) {
      grid.rect(gp = gpar(fill = domain_colors[[label]], col = "black"))
      block_width <- length(index)
      # Scale font size: base 6 + 1.2 * sqrt(width), capped at 15
      font_size <- min(6 + 1.2 * sqrt(block_width), 15)
      grid.text(label, 0.5, 0.5, gp = gpar(col = "white", fontsize = font_size))
    }
    ,
    height = unit(0.5, "cm")
  )

  list(annotation = annotation, domain_vector = domain_vector)
}

# Main loop
for (pdb in na.omit(unique(data_predictions_with_variability_normalized$PDB))) {
  message("Processing: ", pdb)

  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == pdb)
  protein_name <- unique(data_subset$protein)

  longest_chain <- get_longest_chain(data_subset)
  message("  Longest chain: ", longest_chain)

  chain_data <- data_subset %>% filter(chain == longest_chain)

  wide_data <- chain_data %>%
    select(resno, `BepiPred-2.0 probability`, `BepiPred-3.0 score`,
           `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
    column_to_rownames("resno")

  bar_data <- wide_data$var_freq
  wide_data <- wide_data %>% select(-var_freq)
  heatmap_matrix <- t(as.matrix(wide_data))
  residues <- as.numeric(colnames(heatmap_matrix))

  top_annot <- HeatmapAnnotation(
    `Variability` = anno_barplot(
      bar_data,
      gp = gpar(fill = "gray30"),
      border = FALSE,
      bar_width = 1,
      ylim = c(0, 1),
      axis_param = list(at = seq(0, 1, 0.2), labels = seq(0, 1, 0.2))
    ),
    show_annotation_name = TRUE,
    annotation_name_side = "left",
    height = unit(4, "cm")
  )

  colorbar_df <- get_residueindex_colors(residues, rev = FALSE)
  annotation_colors <- list(NGL_Color = setNames(colorbar_df$color, as.character(residues)))
  annotation_df <- data.frame(NGL_Color = as.character(residues), row.names = as.character(residues))

  interpro_data <- data_interpro_filtered %>% filter(Protein_accession == pdb)
  resno_shift <- chain_data$resno[1] - chain_data$seqposno[1]
  domain_annot <- get_domain_annotation(residues, interpro_data, resno_shift, domain_description_colors)

  bottom_annot <- HeatmapAnnotation(
    `Residue Index` = annotation_df$NGL_Color,
    `Domain Annotation` = domain_annot$annotation,
    col = list(`Residue Index` = annotation_colors$NGL_Color),
    annotation_name_side = "right",
    show_legend = c(FALSE, FALSE),
    na_col = "white",
    which = "column"
  )

  ht <- Heatmap(
    heatmap_matrix,
    name = "Score",
    cluster_rows = TRUE,
    cluster_columns = FALSE,
    column_title = paste0(pdb, " chain ", longest_chain, " (", protein_name, ")"),
    col = colorRampPalette(c("blue", "gray95", "red"))(50),
    top_annotation = top_annot,
    bottom_annotation = bottom_annot,
    show_column_names = FALSE
  )

  png(filename = paste0("output/figures/heatmap/domain/", PDB_entry, "_", longest_chain, "_heatmap.png"),
      width = 14, height = 10, units = "in", res = 300, bg = "white")
  
  draw(ht, heatmap_legend_side = "right", annotation_legend_side = "right")
  
  dev.off()
}


```


## Scatterplot
```{r, fig.width=14,warning=FALSE}
pivot_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score"
)

for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  
  protein_name <- unique(data_subset$protein)
  
  print(PDB_entry)
 
  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)
    
    data_subset_long <- data_subset_chain %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
      pivot_longer(cols = pivot_list, 
                   names_to = "Score_Type", values_to = "Score_Value") 
    
    plot <- data_subset_long %>% 
      ggplot(aes(x=var_freq,y=Score_Value)) +
      
      geom_smooth(
        aes(color = Score_Type, group = Score_Type),
        method = "lm",
        se = FALSE, linewidth = 2
      ) +
      stat_regline_equation(aes(label = ..rr.label.., group = Score_Type, color = Score_Type), label.x = 0.05, size = 5) +
      scale_color_manual(
        values = setNames(
          c(
            palette.colors()[7],
            palette.colors()[6],
            palette.colors()[4],
            palette.colors()[8]
          ),
          pivot_list[1:4]
        ),
        name = "Score_Type"
      ) +
      
      new_scale_color() +
      geom_point(aes(color = resno, shape = Score_Type), size=3) +
      scale_color_gradientn(
        colors = rainbow_ngl,
        limits = c(0, max(data_subset_long$resno)),
        name = "Residue number"
      ) +
      scale_x_continuous(
        limit = c(0,NA)
      ) +
      labs(
      title = paste0(PDB_entry, " chain ", chain_index, " (",data_subset$protein[1],")"),
      x = "Residue variability",
      y = "Epitope prediction score"
        ) +
      theme_bw() #+
      #guides(
      #  shape = guide_legend(
      #    override.aes = list(
      #      color = setNames(
      #        c(
      #          palette.colors()[7],
      #          palette.colors()[6],
      #          palette.colors()[4],
      #          palette.colors()[8]
      #        ),
      #        pivot_list[1:4]
      #      )
      #    )
      #  )
      #)
      
      print(plot)
      
      ggsave(filename = paste0("output/figures/scatterplot/merged/", PDB_entry, "_", chain_index, "_merged.png"), 
           plot = plot, 
           width = 14, height = 10, dpi = 300,
           bg="white")
  }
}
```

```{r, fig.width=14, warning=FALSE}
pivot_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score"
)

for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  
  protein_name <- unique(data_subset$protein)
  
  print(PDB_entry)
 
  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)
    
    data_subset_long <- data_subset_chain %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
      pivot_longer(cols = pivot_list, 
                   names_to = "Score_Type", values_to = "Score_Value") 
    
    plot <- data_subset_long %>% 
      ggplot(aes(x=var_freq,y=Score_Value)) +
      
      geom_smooth(
        aes(group = Score_Type),
        method = "lm",
        se = FALSE, size = 2, color="black"
      ) +
      scale_color_manual(
        values = setNames(
          c(
            palette.colors()[7],
            palette.colors()[6],
            palette.colors()[4],
            palette.colors()[8]
          ),
          pivot_list[1:4]
        ),
        name = "Score_Type"
      ) +
      new_scale_color() +
      geom_point(aes(color = resno), size=3) +
      scale_color_gradientn(
        colors = rainbow_ngl,
        limits = c(0, max(data_subset_long$resno)),
        name = "Residue number"
      ) +
      scale_x_continuous(
        limit = c(0,NA)
      ) +
      stat_regline_equation(aes(label = ..rr.label..), label.x = 0.05, size = 5) +
      facet_wrap(~Score_Type, nrow=2) +
        
      labs(
      title = paste0(PDB_entry, " chain ", data_subset$chain[1], " (",data_subset$protein[1],")"),
      x = "Residue variability",
      y = "Epitope prediction score"
        ) +
      theme_bw() #+
      #guides(
      #  shape = guide_legend(
      #    override.aes = list(
      #      color = setNames(
      #        c(
      #          palette.colors()[7],
      #          palette.colors()[6],
      #          palette.colors()[4],
      #          palette.colors()[8]
      #        ),
      #        pivot_list[1:4]
      #      )
      #    )
      #  )
      #)
      
      print(plot)
      
      ggsave(filename = paste0("output/figures/scatterplot/wrapped/", PDB_entry, "_", chain_index, "_wrapped.png"), 
           plot = plot, 
           width = 14, height = 10, dpi = 300,
           bg="white")
  }
}
```

```{r,fig.width=14}
pivot_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score"
)

for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  protein_name <- unique(data_subset$protein)
  
  print(PDB_entry)

  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)

    data_subset_long <- data_subset_chain %>%
      select(resno, `BepiPred-2.0 probability`, `BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
      pivot_longer(cols = pivot_list, 
                   names_to = "Score_Type", values_to = "Score_Value")

    # ➤ Compute mean score per Score_Type
    mean_df <- data_subset_long %>%
      group_by(Score_Type) %>%
      summarize(mean_value = mean(Score_Value, na.rm = TRUE))

    # ➤ Subset again for points: only rows with non-zero variability
    data_subset_long_nonzero_var <- data_subset_long %>% filter(var_freq > 0)
    
    if (nrow(data_subset_long_nonzero_var) == 0) {
      message(paste("Skipping", PDB_entry, chain_index, "- no non-zero variability"))
      next
    }

    plot <- ggplot(data_subset_long_nonzero_var, aes(x = var_freq, y = Score_Value)) +
      
      # Average horizontal line per Score_Type
      geom_hline(
        data = mean_df,
        aes(yintercept = mean_value),
        linetype = "dashed",size=1.5, color = "gray"
      ) +
      
      # Linear regression line (fit on full data)
      geom_smooth(
        data = data_subset_long,
        aes(group = Score_Type),
        method = "lm",
        se = FALSE, size = 2, color = "black"
      ) +

      scale_color_manual(
        values = setNames(
          c(
            palette.colors()[7],
            palette.colors()[6],
            palette.colors()[4],
            palette.colors()[8]
          ),
          pivot_list[1:4]
        ),
        name = "Score_Type"
      ) +
      new_scale_color() +
      
      # Plot only non-zero variability points
      geom_point(aes(color = resno), size = 3) +
      scale_color_gradientn(
        colors = rainbow_ngl,
        limits = c(0, max(data_subset_long$resno)),
        name = "Residue number"
      ) +
      scale_x_continuous(limits = c(0, NA)) +
      stat_regline_equation(aes(label = ..rr.label..), size = 5) +
      facet_wrap(~Score_Type, nrow = 2) +

      labs(
        title = paste0(PDB_entry, " chain ", data_subset$chain[1], " (", data_subset$protein[1], ")"),
        x = "Residue variability",
        y = "Epitope prediction score"
      ) +
      theme_bw()

    print(plot)

    ggsave(filename = paste0("output/figures/scatterplot/wrapped/", PDB_entry, "_", chain_index, "_wrapped.png"), 
           plot = plot, 
           width = 14, height = 10, dpi = 300,
           bg = "white")
  }
}

```

### violin
```{r, fig.width=14, warning=FALSE}
pivot_list <- c(
  "BepiPred-2.0 probability", 
  "BepiPred-3.0 score",
  "DiscoTope-3.0_calibrated_score", 
  "Epigraph_score"
)

for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  protein_name <- unique(data_subset$protein)
  
  print(PDB_entry)
  
  longest_chain <- data_subset %>% 
    filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
    add_count(chain, name = "chain_length") %>%
    arrange(desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
    filter(chain_length == max(chain_length)) %>%  
    filter(chain == min(chain)) %>% 
    pull(chain) %>%
    unique()

  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)

    # Prepare data
    data_subset_long_violin <- data_subset_chain %>%
      select(resno, all_of(pivot_list), var_freq) %>%
      pivot_longer(cols = pivot_list, names_to = "Score_Type", values_to = "Score_Value") %>%
      mutate(
        Variability_Group = ifelse(var_freq > 0, "Variable", "No variability")
      )

    # Skip if only one group exists
    if (length(unique(data_subset_long_violin$Variability_Group)) < 2) {
      message(paste("Skipping", PDB_entry, chain_index, "- only one variability group"))
      next
    }

    plot_violin <- ggplot(data_subset_long_violin, aes(x = Variability_Group, y = Score_Value, fill = Variability_Group)) +
      geom_violin(trim = FALSE, scale = "width", alpha = 0.6) +
      geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.8) +
      stat_compare_means(
        method = "wilcox.test",
        label = "p.signif",
        comparisons = list(c("No variability", "Variable")),
        hide.ns = TRUE
      ) +
      facet_wrap(~ Score_Type, nrow = 2) +
      scale_fill_manual(values = c("No variability" = "#999999", "Variable" = "#E69F00")) +
      labs(
        title = paste0(PDB_entry, " chain ", chain_index, " (", protein_name, ")"),
        x = "Residue variability",
        y = "Epitope prediction score"
      ) +
      theme_bw() +
      theme(legend.position = "none")

    print(plot_violin)

    ggsave(
      filename = paste0("output/figures/violin/", PDB_entry, "_", chain_index, "_violin.png"),
      plot = plot_violin,
      width = 14, height = 10, dpi = 300, bg = "white"
    )
  }
}



```



