---
title: "Streptococcus pneumoniae epitope visualization"
author: "Jan Paul van Meenen"
date: "`r Sys.Date()`"
output: html_document
---

#Packages
```{r}
source("requirements/packages.R")
```

#Functions
```{r}
function_files <- list.files(path = "functions", pattern = "\\.R$",full.names = TRUE)

for (function_file in function_files) {
  source(function_file)
}
```

# Plot all together
## Load in from 01
```{r}
data_BepiPred_2_with_resno <- read.csv("output/BepiPred2.0/data_BepiPred2.csv", check.names = FALSE)
data_BepiPred_3_with_resno <- read.csv("output/BepiPred3.0/data_BepiPred3.csv", check.names = FALSE)
data_Epigraph <- read.csv("output/EpiGraph/data_Epigraph.csv", check.names = FALSE)
data_DiscoTope <- read.csv("output/DiscoTope/data_DiscoTope.csv", check.names = FALSE)
```


## create joined predictions df
```{r}
data_predictions_sequencebased <- full_join(
  data_BepiPred_2_with_resno %>%
    mutate(
      `BepiPred-2.0 probability` = EpitopeProbability,
      `RSA NetsurfP` = RelativeSurfaceAccessilibity
    ) %>%
    select(
      proteinclass, protein, TIGR4_name, PDB, chain, resno, resid, x,y,z,b, seqresno, seqres_resid,
      `BepiPred-2.0 probability`,`RSA NetsurfP`
    ) 
  ,
  
  data_BepiPred_3_with_resno %>%
    select(
      proteinclass, protein, TIGR4_name, PDB, chain, resno, resid, x,y,z,b, seqresno, seqres_resid, 
      `BepiPred-3.0 score`,`BepiPred-3.0 linear epitope score`
    ) 
  ,
  by = c("proteinclass","protein","TIGR4_name","PDB","chain","resno","resid","seqres_resid","seqresno","x","y","z","b")
)

data_predictions_sequencebased
```


```{r}
data_predictions_structurebased <- full_join(
  data_Epigraph %>%
    mutate(
      Epigraph_score = Score
      ) %>%
    select(proteinclass,protein,TIGR4_name,PDB, chain, resno, resid, Epigraph_score, RSA)
  ,
  data_DiscoTope %>%
    mutate(
           `RSA Shrake-Rupley` = rsa,
           `DiscoTope-3.0_calibrated_score` = calibrated_score
    ) %>%
    select(
      proteinclass,protein,TIGR4_name,PDB, chain, resno, resid, `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, `RSA Shrake-Rupley`
      )
  , 
  by = c("proteinclass","protein","TIGR4_name","PDB","chain","resno","resid")
)
data_predictions_structurebased
```


```{r}
data_predictions <- full_join(
  data_predictions_sequencebased,
  data_predictions_structurebased %>%
    add_count(chain, name = "chain_length") %>%
    arrange(desc(chain_length), chain) %>%
    filter(chain_length == max(chain_length)) %>%
    filter(chain == min(chain)),
  by = c("proteinclass","protein","TIGR4_name","PDB","chain","resno","resid")
  )

data_predictions %>%
  filter(if_any(everything(), is.na))

#data_predictions <- data_predictions %>%
#  na.omit() 

data_predictions
```


## normalize
```{r}
data_predictions_normalized <- data_predictions %>%
  group_by(protein,PDB,chain) %>%
  mutate(
    across(
      c(#`BepiPred-2.0 probability`,
        `BepiPred-3.0 score` , `BepiPred-3.0 linear epitope score`, 
        `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, Epigraph_score), 
      rescale
    )
  ) %>%
  ungroup()

data_predictions_normalized
```

###track
```{, warning=FALSE}
track_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score"
)

for (pdb_entry in unique(data_predictions$PDB)) {
  data_predictions_subset <- data_predictions %>%
    filter(PDB == pdb_entry) 
  
  proteinname <- unique(data_predictions_subset$proteinname)
  
  for (chain_index in unique(data_predictions_subset$chain)) {
    
    data_predictions_subset_long <- data_predictions_subset %>%
      filter(chain == chain_index) %>%
      pivot_longer(cols = track_list, 
                   names_to = "Score_Type", values_to = "Score_Value") %>%
      group_by(Score_Type) %>%
      mutate(normalized_score = case_when(
        Score_Type == "BepiPred-2.0 probability" ~ Score_Value,
        TRUE ~ rescale(Score_Value, to = c(0, 1))
        )
      )
    
    figure <- data_predictions_subset_long %>%
      ggplot(aes(x=resno, y=Score_Value, fill = normalized_score)) +
      geom_col() +
      
      facet_wrap(~Score_Type, ncol = 1, scales = "free_y") +
      
      scale_fill_gradient2(low = "blue", mid = "gray95", high = "red", 
                           midpoint = 0.5) +  # Now uses median per facet
      
      geom_text(aes(label = Score_Type, 
                    x = (min(resno) + (max(resno) - min(resno)) / 2)), 
                color = "black", y = Inf, vjust = 1.5, size = 4, family = "Arial", 
                check_overlap = TRUE) +
      
      labs(title = paste0(pdb_entry, " chain ", chain_index, " (",proteinname,")")) +
      
      theme_minimal() +
      theme(
        strip.text = element_blank(),
        axis.ticks = element_blank(),  
        axis.ticks.x = element_blank(),  
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),  
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      )
    
    print(figure)
    
    ggsave(filename = paste0("output/figures/tracks/", pdb_entry, "_", chain_index, "_track.png"), 
           plot = figure, 
           width = 14, height = 10, dpi = 300,
           bg="white")
  }
}
```


###heatmap (pheatmap)
```{, fig.width = 14}
for (pdb_entry in unique(data_predictions_normalized$PDB)) {
  data_predictions_normalised_subset <- data_predictions_normalized %>% na.omit() %>%
    filter(PDB == pdb_entry) 
  
  proteinname <- unique(data_predictions_normalised_subset$proteinname)
  
  for (chain_index in unique(data_predictions_normalised_subset$chain)) {
    
    data_predictions_normalised_subset_wide <- data_predictions_normalised_subset %>%
      filter(chain == chain_index) %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score) %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    figure <- pheatmap(t(data_predictions_normalised_subset_wide), 
                       cluster_rows = TRUE,         
                       cluster_cols = FALSE,         
                       main = paste0(pdb_entry, " chain ", chain_index, " (",proteinname,")"), 
                       color = colorRampPalette(c("blue", "gray95", "red"))(50),
                       labels_col = "resno",
                       clustering_distance_rows = "correlation") 
    
    print(figure)
    
    ggsave(filename = paste0("output/figures/heatmap/", pdb_entry, "_", chain_index, "_heatmap.png"), 
           plot = figure, 
           width = 14, height = 10, dpi = 300,
           bg="white")
  }
}
```

```{, fig.width=14}
for (pdb_entry in unique(data_predictions_normalized$PDB)[2]) {
  data_predictions_normalised_subset <- data_predictions_normalized %>% na.omit() %>%
    filter(PDB == pdb_entry) 
  
  proteinname <- unique(data_predictions_normalised_subset$proteinname)
  
  for (chain_index in unique(data_predictions_normalised_subset$chain)) {
    
    data_predictions_normalised_subset_wide <- data_predictions_normalised_subset %>%
      filter(chain == chain_index) %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score) %>%
      distinct(resno, .keep_all = TRUE) %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    residues <- as.numeric(rownames(data_predictions_normalised_subset_wide))
    
    colorbar_df <- get_residueindex_colors_v2(residues)
    
    annotation_col <- data.frame(NGL_Color = residues)
    rownames(annotation_col) <- residues
    
    annotation_colors <- list(
      NGL_Color = setNames(colorbar_df$color, colorbar_df$resno)
    )
    
    figure <- pheatmap::pheatmap(t(data_predictions_normalised_subset_wide), 
                       cluster_rows = TRUE,         
                       cluster_cols = FALSE,         
                       main = paste0(pdb_entry, " chain ", chain_index, " (",proteinname,")"), 
                       color = colorRampPalette(c("blue", "gray95", "red"))(50),
                       labels_col = "resno",
                       clustering_distance_rows = "correlation",
                       annotation_col = annotation_col,
                       annotation_colors = annotation_colors,
                       annotation_legend = FALSE, 
                       annotation_names_col = FALSE
    ) 
    
    print(figure)
    
    ggsave(filename = paste0("output/figures/heatmap/", pdb_entry, "_", chain_index, "_heatmap.png"), 
           plot = figure, 
           width = 14, height = 10, dpi = 300,
           bg="white")
  }
}
```

### heatmap (complexheatmap)
```{r}
for (pdb_entry in unique(data_predictions_normalized$PDB)) {
  data_predictions_normalised_subset <- data_predictions_normalized %>% na.omit() %>%
    filter(PDB == pdb_entry) 
  
  proteinname <- unique(data_predictions_normalised_subset$proteinname)
  
  for (chain_index in unique(data_predictions_normalised_subset$chain)) {
    
    data_predictions_normalised_subset_wide <- data_predictions_normalised_subset %>%
      filter(chain == chain_index) %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score) %>%
      distinct(resno, .keep_all = TRUE) %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    residues <- as.numeric(rownames(data_predictions_normalised_subset_wide))
    
    colorbar_df <- get_residueindex_colors(residues, rev = FALSE)
    
    annotation_col <- data.frame(NGL_Color = factor(residues))
    
    annotation_colors <- list(
      NGL_Color = setNames(colorbar_df$color, residues)
    )
    
    figure <- ComplexHeatmap::pheatmap(
      t(data_predictions_normalised_subset_wide), 
      cluster_rows = TRUE,         
      cluster_cols = FALSE,         
      main = paste0(pdb_entry, " chain ", chain_index, " (", proteinname, ")"), 
      color = colorRampPalette(c("blue", "gray95", "red"))(50),
      #labels_col = rownames(data_predictions_normalised_subset_wide),
      clustering_distance_rows = "correlation",
      annotation_col = annotation_col,
      annotation_colors = annotation_colors,
      annotation_legend = FALSE, 
      annotation_names_col = FALSE
    )
    
    
    print(figure)
    
    #ggsave(filename = paste0("output/figures/heatmap/", pdb_entry, "_", chain_index, "_heatmap.png"), 
    #       plot = figure, 
    #       width = 14, height = 10, dpi = 300,
    #       bg="white")
  }
}
```


## 3D
### bepipred-3.0
```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_BepiPred_3_with_resno_subset <- data_BepiPred_3_with_resno %>% filter(Accession == pdb_entry) %>% na.omit()
  
  for (chain_index in unique(data_BepiPred_3_with_resno_subset$chain)) {
    data_BepiPred_3_with_resno_subset_chain <- data_BepiPred_3_with_resno_subset %>% filter(chain == chain_index)
    
    min_score <- min(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, na.rm = TRUE)
    median_score <- median(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, na.rm = TRUE)
    max_score <- max(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, na.rm = TRUE)
    
    
    low_to_median <- colorRampPalette(c("blue", "gray95"))(10)
    median_to_high <- colorRampPalette(c("gray95", "red"))(10)
    
    breaks <- seq(min_score, max_score, length.out = 21)  # 20 intervals
    colors <- c(low_to_median, median_to_high)  # 20 colors
    
    color_map <- data.frame(
      ProbabilityBin = cut(seq(min_score, max_score, length.out = 20), 
                           breaks = breaks, include.lowest = TRUE),
      color = colors
    ) %>%
      rbind(c(NA,"black"))
    
    predictions_df <- data_BepiPred_3_with_resno_subset_chain %>%
      mutate(ProbabilityBin = cut(`BepiPred-3.0 score`, 
                                  breaks = breaks, include.lowest = TRUE)
      ) %>%
      group_by(ProbabilityBin) %>% 
      summarize(resno_list = list(paste0(resno,":",chain_index)), .groups = "drop") %>%
      mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " "))) %>%
      left_join(color_map, by = "ProbabilityBin")
    
    view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index, sele_color_df = predictions_df)
    
    #print(view)
    
    saveWidget(view, file = paste0("output/BepiPred3.0/snapshot/score/",pdb_entry, "_", chain_index,"_bepipred3.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```


### discotope-3.0
```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  
  for (chain_index in unique(data_DiscoTope_subset$chain)) {
  data_DiscoTope_subset_chain <- data_DiscoTope_subset %>% filter(chain == chain_index)
  
  min_score <- min(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  median_score <- median(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  max_score <- max(data_DiscoTope_subset$calibrated_score, na.rm = TRUE)
  
  
  low_to_median <- colorRampPalette(c("blue", "gray95"))(10)
  median_to_high <- colorRampPalette(c("gray95", "red"))(10)
  
  breaks <- seq(min_score, max_score, length.out = 21)  # 20 intervals
  colors <- c(low_to_median, median_to_high)  # 20 colors
  
  color_map <- data.frame(
    ProbabilityBin = cut(seq(min_score, max_score, length.out = 20), 
                         breaks = breaks, include.lowest = TRUE),
    color = colors
  ) %>%
  rbind(c(NA,"black"))
  
  predictions_df <- data_DiscoTope_subset %>%
    mutate(ProbabilityBin = cut(calibrated_score, 
                                breaks = breaks, include.lowest = TRUE)
    ) %>%
    group_by(ProbabilityBin) %>% 
    summarize(resno_list = list(paste0(resno,":",chain_index)), .groups = "drop") %>%
    mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " "))) %>%
    left_join(color_map, by = "ProbabilityBin")
  
  view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index,sele_color_df = predictions_df)
  
  #print(view)
  
  saveWidget(view, file = paste0("output/DiscoTope/snapshot/score/",pdb_entry, "_", chain_index,"_discotope3.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```

### colour epitopes by resno
Colouring predicted epitopes by resno could help identify discontinous epitopes
```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_BepiPred_3_with_resno_subset <- data_BepiPred_3_with_resno %>% filter(Accession == pdb_entry) %>% na.omit()
  
  quantile_threshold <- 0.85  #0.8 == top 20%
  
  for (chain_index in unique(data_BepiPred_3_with_resno_subset$chain)) {
    data_BepiPred_3_with_resno_subset_chain <- data_BepiPred_3_with_resno_subset %>% filter(chain == chain_index)
    
    score_threshold <- quantile(data_BepiPred_3_with_resno_subset_chain$`BepiPred-3.0 score`, 
                                probs = quantile_threshold, na.rm = TRUE)
    
    predictions_df <- data_BepiPred_3_with_resno_subset_chain %>%
      mutate(color = ifelse(`BepiPred-3.0 score` >= score_threshold, "residueindex", "white")) %>%
      group_by(color) %>% 
      summarize(resno_list = list(paste0(resno, ":", chain_index)), .groups = "drop") %>%
      mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " ")))
    
    view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index, sele_color_df = predictions_df)
    
    #print(view)
    
    saveWidget(view, file = paste0("output/BepiPred3.0/snapshot/conformational/",pdb_entry, "_", chain_index,"_resid_bepipred3.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```

```{r}
for (pdb_entry in pdb_files_list) {
  print(pdb_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
  
  data_DiscoTope_subset <- data_DiscoTope %>% filter(pdb == pdb_entry)
  
  quantile_threshold <- 0.85  #0.8 == top 20%
  
  for (chain_index in unique(data_DiscoTope_subset$chain)) {
    data_DiscoTope_subset_chain <- data_DiscoTope_subset %>% filter(chain == chain_index)
    
    score_threshold <- quantile(data_DiscoTope_subset_chain$`DiscoTope-3.0_score`, 
                                probs = quantile_threshold, na.rm = TRUE)
    
    predictions_df <- data_DiscoTope_subset_chain %>%
      mutate(color = ifelse(`DiscoTope-3.0_score` >= score_threshold, "residueindex", "white")) %>%
      group_by(color) %>% 
      summarize(resno_list = list(paste0(resno, ":", chain_index)), .groups = "drop") %>%
      mutate(resno_string = sapply(resno_list, function(x) str_c(x, collapse = " ")))
    
    view <- prediction.viewer(pdb_entry, pdbs_path, chain = chain_index, sele_color_df = predictions_df)
    
    print(view)
    
    saveWidget(view, file = paste0("output/DiscoTope/snapshot/conformational/",pdb_entry, "_", chain_index,"_resid_discotope.html"), title=paste0(pdb_entry, " chain ", chain_index), selfcontained = TRUE)
  }
}
```


## Longest chain
### Getting the longest chain
```{r}
data_predictions_longestchain <- inner_join(
  data_predictions_structurebased %>%
    add_count(PDB, chain, name = "chain_length") %>%  # Add chain length count
    arrange(PDB,desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
    group_by(PDB) %>%
    filter(chain_length == max(chain_length)) %>%  
    filter(chain == min(chain)) %>% 
    ungroup() %>%  
    select(-chain_length),
  data_predictions_sequencebased,
  by = c("PDB", "chain", "resno","resid")
  ) %>%
  
  left_join(
    data_namesclasspdb
  )

data_predictions_longestchain
```

### PCA
```{r}
score_columns <- c("Epigraph_score", "DiscoTope-3.0_calibrated_score",
                   "BepiPred-2.0 probability", 
                   "BepiPred-3.0 score")

score_columns
```

```{r}
scaled_scores <- data_predictions_longestchain %>%
  select(all_of(score_columns)) %>%
  scale()
  
pca_result <- prcomp(t(scaled_scores),center = TRUE, scale. = TRUE)
```


```{r}
explained_variance <- data.frame(
  PC = paste0("PC", 1:length(pca_result$sdev)),
  Variance = (pca_result$sdev^2) / sum(pca_result$sdev^2) * 100
)

# Scree plot
ggplot(explained_variance, aes(x = PC, y = Variance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Scree Plot", x = "Principal Components", y = "Variance Explained (%)") +
  theme_minimal()
```

```{r}
library(ggrepel)

pca_scores <- as.data.frame(pca_result$x) %>%
  rownames_to_column(var = "score_type")

# Scatter plot of PC1 vs PC2
ggplot(pca_scores, aes(x = PC1, y = PC2, label = score_type)) +
  geom_point(alpha = 0.7, color = "blue") +
  geom_text_repel() +
  labs(title = "PCA of concatenated epitope prediction scores across different tools", x = "PC1", y = "PC2") +
  theme_minimal()
```



# Variability
matching variability consensus sequence to epitope predictions sequences
## variability_data_with_resno
```{r}
data_variability <- read.csv("output/data_variability.csv") #%>%
  #distinct(structure,chain,x,y,z, .keep_all = TRUE)

data_variability
```

### check the missing ones
Currently working with PDB entries only, not all proteins have PDB's associated with them, so some are missing
```{r}
sort(unique(data_predictions$protein))
sort(unique(data_variability$protein))

setdiff(
  tolower(unique(data_predictions$protein)),
  tolower(unique(data_variability$protein))
)

setdiff(
  tolower(unique(data_variability$protein)),
  tolower(unique(data_predictions$protein))
)
```

```{r}
data_predictions_with_variability <- full_join(
  data_predictions,
  data_variability %>% select(-chain),
  by = c("proteinclass","protein","PDB","resno","resid", "seqresno", "seqres_resid", "x", "y", "z", "b"))

data_predictions_with_variability
```


```{r}
data_predictions %>% filter(PDB == "3CX3") 
data_variability %>% filter(PDB == "3CX3")
```

### normalized
```{r}
data_predictions_with_variability_normalized <- data_predictions_with_variability %>%
  group_by(PDB) %>%
  mutate(
    across(
      c(`BepiPred-3.0 score` , `DiscoTope-3.0_score`, `DiscoTope-3.0_calibrated_score`, Epigraph_score), 
      rescale
    )
  ) %>%
  ungroup()

data_predictions_with_variability_normalized
```

## plot
### tracks
```{,warning=FALSE, fig.width = 14}
pivot_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score", 
  "var_freq"
)

for (pdb_entry in na.omit(unique(data_predictions_with_variability$PDB))) {
  data_subset <- data_predictions_with_variability %>%    
    filter(PDB == pdb_entry)  
  
    longest_chain <- data_subset %>% 
      filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
    add_count(chain, name = "chain_length") %>%
    arrange(desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
    filter(chain_length == max(chain_length)) %>%  
    filter(chain == min(chain)) %>% 
    pull(chain) %>%
    unique()
    
    data_subset_long <- data_subset %>% 
      pivot_longer(cols = pivot_list, 
                   names_to = "Score_Type", values_to = "Score_Value") %>%
      group_by(Score_Type) %>%
      mutate(normalized_score = case_when(
        Score_Type == "BepiPred-2.0 probability" ~ Score_Value,
        TRUE ~ rescale(Score_Value, to = c(0, 1))
      )
      )
 
  
  plot <- data_subset_long %>% filter(chain == longest_chain | is.na(chain)) %>%
    ggplot(aes(x=resno, y=Score_Value, fill = normalized_score)) +
    geom_col() +
    
    facet_wrap(~Score_Type, ncol = 1, scales = "free_y") +
    
    scale_fill_gradient2(low = "blue", mid = "gray95", high = "red", 
                         midpoint = 0.5) +  # Now uses median per facet
    
    geom_text(aes(label = Score_Type, 
                  x = (min(resno) + (max(resno) - min(resno)) / 2)), 
              color = "black", y = Inf, vjust = 1.5, size = 4, family = "Arial", 
              check_overlap = TRUE) +
    
    labs(
      title = paste0(pdb_entry, " chain ", data_subset$chain[1], " (",data_subset$proteinname[1],")"), 
         x = "Residue number") +
    
    theme_minimal() +
    theme(
      strip.text = element_blank(),
      axis.ticks = element_blank(),  
      axis.ticks.x = element_blank(),  
      axis.ticks.y = element_blank(),
      #axis.text.x = element_blank(),  
      axis.text.y = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
  
  print(plot)
  
  ggsave(filename = paste0("output/figures/variability/", pdb_entry, "_", longest_chain,
                           "_track.png"),
         plot = plot, 
         width = 14, height = 10, dpi = 300,
         bg="white")
}
```



### heatmap
```{, fig.width=14}
for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  
  protein_name <- unique(data_subset$protein)
  
  print(PDB_entry)
 
  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)
    
    data_subset_wide <- data_subset_chain %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
      #distinct() %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    bar_data <- data_subset_wide$var_freq
    names(bar_data) <- rownames(data_subset_wide)
    
    data_subset_wide <- data_subset_wide %>% select(-var_freq)
    
    heatmap_matrix <- t(data_subset_wide)
    
    top_annot <- HeatmapAnnotation(
      `Variability` = anno_barplot(
        bar_data, 
        gp = gpar(fill = "gray30"), 
        border = FALSE, 
        bar_width = 1, 
        ylim = c(0,1),
        axis_param = list(
          at = seq(0, 1, by = 0.2),
          labels = seq(0, 1, by = 0.2)
        )
      ),
      show_annotation_name = TRUE,
      annotation_name_side = "left",
      height = unit(4, "cm")
    )
    
    residues <- as.numeric(colnames(heatmap_matrix))
    
    colorbar_df <- get_residueindex_colors(residues, rev = FALSE)
    
    annotation_df <- data.frame(NGL_Color = as.character(residues))
    rownames(annotation_df) <- as.character(residues)
    
    annotation_colors <- list(
      NGL_Color = setNames(colorbar_df$color, as.character(residues))
    )
    
    bottom_annot <- HeatmapAnnotation(
      `Residue Index` = annotation_df$NGL_Color,
      col = list(`Residue Index` = annotation_colors$NGL_Color),
      annotation_name_side = "right",
      show_legend = FALSE,
      which = "column"
    )

    ht <- Heatmap(
      heatmap_matrix,
      name = "Score",
      cluster_rows = TRUE,
      cluster_columns = FALSE,
      column_title = paste0(PDB_entry, " chain ", chain_index, " (", protein_name, ")"),
      col = colorRampPalette(c("blue", "gray95", "red"))(50),
      top_annotation = top_annot,
      bottom_annotation = bottom_annot,
      show_column_names = FALSE
    )
    
    figure <- draw(ht, heatmap_legend_side = "right", annotation_legend_side = "right")
    
    print(figure)
    
    png(filename = paste0("output/figures/heatmap/variability/", PDB_entry, "_", chain_index, "_heatmap.png"),
        width = 14, height = 10, units = "in", res = 300, bg = "white")
    
    draw(ht, heatmap_legend_side = "right", annotation_legend_side = "right")
    
    dev.off()

  }
}
```


#### longestchain
```{r, fig.width = 14}
for (structure_entry in na.omit(unique(data_predictions_with_variability_normalized$structure))[37]) {
  print(structure_entry)
  
  data_subset <- data_predictions_with_variability_normalized %>% filter(structure == structure_entry)
  
  protein_name <- unique(data_subset$protein)
  
  longest_chain <- data_subset %>% 
    filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
    add_count(chain, name = "chain_length") %>%
    arrange(desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
    filter(chain_length == max(chain_length)) %>%  
    filter(chain == min(chain)) %>% 
    pull(chain) %>%
    unique()
  
  print(longest_chain)
  data_subset_chain <- data_subset %>% filter(chain == longest_chain)
  
  data_subset_wide <- data_subset_chain %>%
    select(seqresno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
           `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
    #distinct() %>%
    column_to_rownames("seqresno") %>%
    as.data.frame()
  
  bar_data <- data_subset_wide$var_freq
  names(bar_data) <- rownames(data_subset_wide)
  
  data_subset_wide <- data_subset_wide %>% select(-var_freq)
  
  heatmap_matrix <- t(data_subset_wide)
  
  top_annot <- HeatmapAnnotation(
    `Variability` = anno_barplot(
      bar_data, 
      gp = gpar(fill = "gray30"), 
      border = FALSE, 
      bar_width = 1, 
      ylim = c(0,1),
      axis_param = list(
        at = seq(0, 1, by = 0.2),
        labels = seq(0, 1, by = 0.2)
      )
    ),
    show_annotation_name = TRUE,
    annotation_name_side = "left",
    height = unit(4, "cm")
  )
  
  residues <- as.numeric(colnames(heatmap_matrix))
  
  colorbar_df <- get_residueindex_colors(residues, rev = FALSE)
  
  annotation_df <- data.frame(NGL_Color = as.character(residues))
  rownames(annotation_df) <- as.character(residues)
  
  annotation_colors <- list(
    NGL_Color = setNames(colorbar_df$color, as.character(residues))
  )
  
  bottom_annot <- HeatmapAnnotation(
    `Residue Index` = annotation_df$NGL_Color,
    col = list(`Residue Index` = annotation_colors$NGL_Color),
    annotation_name_side = "right",
    show_legend = FALSE,
    which = "column"
  )
  
  ht <- Heatmap(
    heatmap_matrix,
    name = "Score",
    cluster_rows = TRUE,
    cluster_columns = FALSE,
    column_title = paste0(structure_entry, " chain ", longest_chain, " (", protein_name, ")"),
    col = colorRampPalette(c("blue", "gray95", "red"))(50),
    top_annotation = top_annot,
    bottom_annotation = bottom_annot,
    show_column_names = FALSE
  )
  
  figure <- draw(ht, heatmap_legend_side = "right", annotation_legend_side = "right")
  
  print(figure)
  
  png(filename = paste0("output/figures/heatmap/variability/", structure_entry, "_", longest_chain, "_heatmap.png"),
      width = 14, height = 10, units = "in", res = 300, bg = "white")
  
  dev.off()
  
  
}
```


#### clustered
```{, fig.width=14}
for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  
  protein_name <- unique(data_subset$proteinname)
  
  print(PDB_entry)
 
  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)
    
    data_subset_wide <- data_subset_chain %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
      distinct() %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    bar_data <- data_subset_wide$var_freq
    names(bar_data) <- rownames(data_subset_wide)
    
    data_subset_wide <- data_subset_wide %>% select(-var_freq, -`BepiPred-2.0 probability`)
    
    heatmap_matrix <- t(data_subset_wide)
    
    top_annot <- HeatmapAnnotation(
      `Variability` = anno_barplot(
        bar_data, 
        gp = gpar(fill = "gray30"), 
        border = FALSE, 
        bar_width = 1, 
        ylim = c(0,1),
        axis_param = list(
          at = seq(0, 1, by = 0.2),
          labels = seq(0, 1, by = 0.2)
        )
      ),
      show_annotation_name = TRUE,
      annotation_name_side = "left",
      height = unit(4, "cm")
    )
    
    residues <- as.numeric(colnames(heatmap_matrix))
    
    colorbar_df <- get_residueindex_colors(residues, rev = FALSE)
    
    annotation_df <- data.frame(NGL_Color = as.character(residues))
    rownames(annotation_df) <- as.character(residues)
    
    annotation_colors <- list(
      NGL_Color = setNames(colorbar_df$color, as.character(residues))
    )
    
    bottom_annot <- HeatmapAnnotation(
      `Residue Index` = annotation_df$NGL_Color,
      col = list(`Residue Index` = annotation_colors$NGL_Color),
      annotation_name_side = "right",
      show_legend = FALSE,
      which = "column"
    )

    ht <- Heatmap(
      heatmap_matrix,
      name = "Score",
      cluster_rows = TRUE,
      cluster_columns = TRUE,
      column_title = paste0(PDB_entry, " chain ", chain_index, " (", protein_name, ")"),
      col = colorRampPalette(c("blue", "gray95", "red"))(50),
      top_annotation = top_annot,
      bottom_annotation = bottom_annot,
      show_column_names = FALSE
    )
    
    figure <- draw(ht, heatmap_legend_side = "right", annotation_legend_side = "right")
    
    print(figure)
    
    #ggsave(filename = paste0("output/figures/heatmap/variability/", pdb_entry, "_", chain_index, "_heatmap.png"), 
    #       plot = figure, 
    #       width = 14, height = 10, dpi = 300,
    #       bg="white")
  }
}
```

```{, fig.with = 14}
maxdist <- 10

for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)[3]) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  
  protein_name <- unique(data_subset$proteinname)
  
  print(PDB_entry)
  
  pdb_object <- read.pdb.custom(pdb_entry = pdb_entry, pdb_path = pdbs_path)
 
  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)
    
    data_subset_wide <- data_subset_chain %>%
      select(resno, `BepiPred-2.0 probability`, `BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
      distinct() %>%
      column_to_rownames("resno") %>%
      as.data.frame()
    
    bar_data <- data_subset_wide$var_freq
    names(bar_data) <- rownames(data_subset_wide)
    
    # Remove the var_freq and 'BepiPred-2.0 probability' columns for the heatmap data
    data_subset_wide <- data_subset_wide %>% select(-var_freq, -`BepiPred-2.0 probability`)
    
    heatmap_matrix <- t(data_subset_wide)
    
    top_annot <- HeatmapAnnotation(
      `Variability` = anno_barplot(
        bar_data, 
        gp = gpar(fill = "gray30"), 
        border = FALSE, 
        bar_width = 1, 
        ylim = c(0, 1),
        axis_param = list(
          at = seq(0, 1, by = 0.2),
          labels = seq(0, 1, by = 0.2)
        )
      ),
      show_annotation_name = TRUE,
      annotation_name_side = "left",
      height = unit(4, "cm")
    )
    
    residues <- as.numeric(colnames(heatmap_matrix))
    
    resnos <- pdb_object$atom %>% filter(chain == chain_id, elety == "CA", type == "ATOM") %>% pull(resno)
    
    # Create the xyz distance matrix for residues
    ca.inds <- atom.select(pdb_object, chain = chain_index, elety = "CA", type = "ATOM")
    coords <- pdb_object$xyz[ca.inds$xyz]
    
    dist.mat <- dist.xyz(coords)
    
    rownames(dist.mat) <- as.character(resnos)
    colnames(dist.mat) <- as.character(resnos)
    
    shared <- intersect(resnos, residues)
    setdiff(resnos, residues)
    
    dist.mat <- dist.mat[shared,shared]
    
    #rownames(dist.mat) <- residues
    #colnames(dist.mat) <- residues
    
    # Perform hierarchical clustering on the rows
    row_clustering <- hclust(as.dist(dist.mat), method = "complete")  # You can change the method to "single", "average", etc.
    
    # Order the residues based on the clustering
    heatmap_matrix <- heatmap_matrix[, intersect(resnos, residues)]
    heatmap_matrix <- heatmap_matrix[, row_clustering$order]
    
    colorbar_df <- get_residueindex_colors(residues, rev = FALSE)
    
    annotation_df <- data.frame(NGL_Color = as.character(residues))
    rownames(annotation_df) <- as.character(residues)
    
    annotation_colors <- list(
      NGL_Color = setNames(colorbar_df$color, as.character(residues))
    )
    
    bottom_annot <- HeatmapAnnotation(
      `Residue Index` = annotation_df$NGL_Color,
      col = list(`Residue Index` = annotation_colors$NGL_Color),
      annotation_name_side = "right",
      show_legend = FALSE,
      which = "column"
    )

    # Plot heatmap with hierarchical clustering of rows
    ht <- Heatmap(
      heatmap_matrix,
      name = "Score",
      cluster_rows = row_clustering,  # Use pre-clustered rows (row_clustering)
      cluster_columns = TRUE,  # Optional: Cluster columns as well
      column_title = paste0(PDB_entry, " chain ", chain_index, " (", protein_name, ")"),
      col = colorRampPalette(c("blue", "gray95", "red"))(50),
      top_annotation = top_annot,
      bottom_annotation = bottom_annot,
      show_column_names = FALSE
    )
    
    figure <- draw(ht, heatmap_legend_side = "right", annotation_legend_side = "right")
    
    print(figure)
    
    # Optionally save the plot
    # ggsave(filename = paste0("output/figures/heatmap/variability/", pdb_entry, "_", chain_index, "_heatmap.png"), 
    #        plot = figure, 
    #        width = 14, height = 10, dpi = 300,
    #        bg="white")
  }
}

```

#### with domain annotation
```{r}
library(colorspace)
all_domain_descriptions <- data_interpro_filtered %>%
  arrange(Description) %>%
  pull(Description) %>%
  unique() 

#domain_palette <- qualitative_hcl(length(all_domain_descriptions), palette = "Viridis")
domain_palette <- viridis(length(all_domain_descriptions), option="D")

domain_description_colors <- setNames(
  domain_palette, 
  all_domain_descriptions
  )

show_col(domain_description_colors)

```

```{r, fig.width=14}
for (PDB_entry in na.omit(unique(data_predictions_with_variability_normalized$PDB))) {
  print(PDB_entry)
  
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  
  protein_name <- unique(data_subset$protein)
  
  longest_chain <- data_subset %>% 
    filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
    add_count(chain, name = "chain_length") %>%
    arrange(desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
    filter(chain_length == max(chain_length)) %>%  
    filter(chain == min(chain)) %>% 
    pull(chain) %>%
    unique()
  
  print(longest_chain)
  data_subset_chain <- data_subset %>% filter(chain == longest_chain)
  
  data_subset_wide <- data_subset_chain %>%
    select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
           `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
    #distinct() %>%
    column_to_rownames("resno") %>%
    as.data.frame()
  
  bar_data <- data_subset_wide$var_freq
  names(bar_data) <- rownames(data_subset_wide)
  
  data_subset_wide <- data_subset_wide %>% select(-var_freq)
  
  heatmap_matrix <- t(data_subset_wide)
  
  top_annot <- HeatmapAnnotation(
    `Variability` = anno_barplot(
      bar_data, 
      gp = gpar(fill = "gray30"), 
      border = FALSE, 
      bar_width = 1, 
      ylim = c(0,1),
      axis_param = list(
        at = seq(0, 1, by = 0.2),
        labels = seq(0, 1, by = 0.2)
      )
    ),
    show_annotation_name = TRUE,
    annotation_name_side = "left",
    height = unit(4, "cm")
  )
  
  residues <- as.numeric(colnames(heatmap_matrix))
  
  colorbar_df <- get_residueindex_colors(residues, rev = FALSE)
  
  annotation_df <- data.frame(NGL_Color = as.character(residues))
  rownames(annotation_df) <- as.character(residues)
  
  annotation_colors <- list(
    NGL_Color = setNames(colorbar_df$color, as.character(residues))
  )
  
  rsa_vector <- data_subset_chain %>%
    filter(resno %in% residues) %>%
    arrange(match(resno, residues)) %>%
    pull(`RSA Shrake-Rupley`)
  
  names(rsa_vector) <- as.character(residues)
  
  rsa_colors <- circlize::colorRamp2(c(0, 1), c("#E69F00", "#56B4E9"))
  
  interpro_domains <- data_interpro_filtered %>%
    filter(Protein_accession == PDB_entry) %>%
    mutate(Start_resno = Start + data_subset$resno[1] - data_subset$seqposno[1],
           End_resno = End + data_subset$resno[1] - data_subset$seqposno[1]) %>%
    select(Start_resno, End_resno, Description) %>%
    pmap_dfr(function(Start_resno, End_resno, Description) {
      tibble(
        resno = seq(Start_resno, End_resno),
        Description = Description
      )
    })
  
  interpro_domains
  
  domain_vector <- rep(NA_character_, length(residues))
  names(domain_vector) <- as.character(residues)
  
  matching_indices <- match(interpro_domains$resno, residues)
  domain_vector[matching_indices[!is.na(matching_indices)]] <- interpro_domains$Description[!is.na(matching_indices)]
  
  # Get unique non-NA domain descriptions
  unique_domains <- na.omit(unique(domain_vector))
  domain_colors <- domain_description_colors[unique_domains]
  
  # Get blocks for each domain
  domain_blocks <- split(names(domain_vector)[!is.na(domain_vector)], domain_vector[!is.na(domain_vector)])
  domain_blocks_ranges <- lapply(domain_blocks, function(res_indices) {
    indices <- which(names(domain_vector) %in% res_indices)
    range(indices)
  })
  
  # Create align_to list: each domain name maps to its residue indices (as numeric positions)
  domain_align_to <- lapply(domain_blocks_ranges, function(rng) seq(rng[1], rng[2]))
  
  annotation_domain <- anno_block(
    align_to = domain_align_to,
    panel_fun = function(index, label) {
      grid.rect(gp = gpar(fill = domain_colors[[label]], col = "black"))
      block_width <- length(index)
      # Scale font size: base 6 + 1.2 * sqrt(width), capped at 15
      font_size <- min(1 + 1.2 * sqrt(block_width), 15)
      grid.text(label, 0.5, 0.5, gp = gpar(col = "white", fontsize = font_size))
    },
    height = unit(1, "cm"), show_name = TRUE
  )
  
  # Final bottom annotation with anno_block and Residue Index
  bottom_annot <- HeatmapAnnotation(
    `Residue Index` = annotation_df$NGL_Color,
    `RSA` = rsa_vector,
    `Domain hit` = annotation_domain,
    col = list(
      `Residue Index` = annotation_colors$NGL_Color,
      `RSA` = rsa_colors
    ),
    annotation_name_side = "right",
    show_legend = c(FALSE, TRUE, FALSE),
    na_col = "white",
    which = "column"
  )
  
  
  ht <- Heatmap(
    heatmap_matrix,
    name = "Score",
    cluster_rows = TRUE,
    cluster_columns = FALSE,
    column_title = paste0(PDB_entry, " chain ", longest_chain, " (", protein_name, ")"),
    col = colorRampPalette(c("blue", "gray95", "red"))(50),
    top_annotation = top_annot,
    bottom_annotation = bottom_annot,
    show_column_names = FALSE
  )
  
  png(filename = paste0("output/figures/heatmap/domain/", PDB_entry, "_", longest_chain, "_heatmap.png"),
      width = 14, height = 10, units = "in", res = 300, bg = "white")
  
  figure <- draw(ht, heatmap_legend_side = "right", annotation_legend_side = "right")
  
  dev.off()
  
  print(figure)
}
```

FIXME: Refactor complexheatmap code
```{,fig.width=14}
# Helper function: get longest chain from a data subset
get_longest_chain <- function(data_subset) {
  data_subset %>%
    filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
    add_count(chain, name = "chain_length") %>%
    arrange(desc(chain_length), chain) %>%
    filter(chain_length == max(chain_length)) %>%
    slice_min(chain) %>%
    pull(chain) %>%
    unique()
}

# Helper: create domain annotation vector and colors
get_domain_annotation <- function(residues, interpro_data, resno_shift, domain_colors) {
  domains <- interpro_data %>%
    mutate(
      Start_resno = Start + resno_shift,
      End_resno = End + resno_shift
    ) %>%
    select(Start_resno, End_resno, Description) %>%
    pmap_dfr(function(Start_resno, End_resno, Description) {
      tibble(resno = seq(Start_resno, End_resno), Description = Description)
    })


  domain_vector <- rep(NA_character_, length(residues))
  names(domain_vector) <- as.character(residues)
  matching_indices <- match(domains$resno, residues)
  domain_vector[matching_indices[!is.na(matching_indices)]] <- domains$Description[!is.na(matching_indices)]

  unique_domains <- na.omit(unique(domain_vector))
  domain_blocks <- split(names(domain_vector)[!is.na(domain_vector)], domain_vector[!is.na(domain_vector)])
  domain_blocks_ranges <- lapply(domain_blocks, function(res_indices) {
    indices <- which(names(domain_vector) %in% res_indices)
    range(indices)
  })
  domain_align_to <- lapply(domain_blocks_ranges, function(rng) seq(rng[1], rng[2]))

  annotation <- anno_block(
    align_to = domain_align_to,
    panel_fun = function(index, label) {
      grid.rect(gp = gpar(fill = domain_colors[[label]], col = "black"))
      block_width <- length(index)
      # Scale font size: base 6 + 1.2 * sqrt(width), capped at 15
      font_size <- min(6 + 1.2 * sqrt(block_width), 15)
      grid.text(label, 0.5, 0.5, gp = gpar(col = "white", fontsize = font_size))
    }
    ,
    height = unit(0.5, "cm")
  )

  list(annotation = annotation, domain_vector = domain_vector)
}

# Main loop
for (pdb in na.omit(unique(data_predictions_with_variability_normalized$PDB))) {
  message("Processing: ", pdb)

  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == pdb)
  protein_name <- unique(data_subset$protein)

  longest_chain <- get_longest_chain(data_subset)
  message("  Longest chain: ", longest_chain)

  chain_data <- data_subset %>% filter(chain == longest_chain)

  wide_data <- chain_data %>%
    select(resno, `BepiPred-2.0 probability`, `BepiPred-3.0 score`,
           `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
    column_to_rownames("resno")

  bar_data <- wide_data$var_freq
  wide_data <- wide_data %>% select(-var_freq)
  heatmap_matrix <- t(as.matrix(wide_data))
  residues <- as.numeric(colnames(heatmap_matrix))

  top_annot <- HeatmapAnnotation(
    `Variability` = anno_barplot(
      bar_data,
      gp = gpar(fill = "gray30"),
      border = FALSE,
      bar_width = 1,
      ylim = c(0, 1),
      axis_param = list(at = seq(0, 1, 0.2), labels = seq(0, 1, 0.2))
    ),
    show_annotation_name = TRUE,
    annotation_name_side = "left",
    height = unit(4, "cm")
  )

  colorbar_df <- get_residueindex_colors(residues, rev = FALSE)
  annotation_colors <- list(NGL_Color = setNames(colorbar_df$color, as.character(residues)))
  annotation_df <- data.frame(NGL_Color = as.character(residues), row.names = as.character(residues))

  interpro_data <- data_interpro_filtered %>% filter(Protein_accession == pdb)
  resno_shift <- chain_data$resno[1] - chain_data$seqposno[1]
  domain_annot <- get_domain_annotation(residues, interpro_data, resno_shift, domain_description_colors)

  bottom_annot <- HeatmapAnnotation(
    `Residue Index` = annotation_df$NGL_Color,
    `Domain Annotation` = domain_annot$annotation,
    col = list(`Residue Index` = annotation_colors$NGL_Color),
    annotation_name_side = "right",
    show_legend = c(FALSE, FALSE),
    na_col = "white",
    which = "column"
  )

  ht <- Heatmap(
    heatmap_matrix,
    name = "Score",
    cluster_rows = TRUE,
    cluster_columns = FALSE,
    column_title = paste0(pdb, " chain ", longest_chain, " (", protein_name, ")"),
    col = colorRampPalette(c("blue", "gray95", "red"))(50),
    top_annotation = top_annot,
    bottom_annotation = bottom_annot,
    show_column_names = FALSE
  )

  png(filename = paste0("output/figures/heatmap/domain/", PDB_entry, "_", longest_chain, "_heatmap.png"),
      width = 14, height = 10, units = "in", res = 300, bg = "white")
  
  draw(ht, heatmap_legend_side = "right", annotation_legend_side = "right")
  
  dev.off()
}


```


## Scatterplot
```{, fig.width=14,warning=FALSE}
pivot_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score"
)

for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  
  protein_name <- unique(data_subset$protein)
  
  print(PDB_entry)
 
  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)
    
    data_subset_long <- data_subset_chain %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
      pivot_longer(cols = pivot_list, 
                   names_to = "Score_Type", values_to = "Score_Value") 
    
    plot <- data_subset_long %>% 
      ggplot(aes(x=var_freq,y=Score_Value)) +
      
      geom_smooth(
        aes(color = Score_Type, group = Score_Type),
        method = "lm",
        se = FALSE, linewidth = 2
      ) +
      stat_regline_equation(aes(label = ..rr.label.., group = Score_Type, color = Score_Type), label.x = 0.05, size = 5) +
      scale_color_manual(
        values = setNames(
          c(
            palette.colors()[7],
            palette.colors()[6],
            palette.colors()[4],
            palette.colors()[8]
          ),
          pivot_list[1:4]
        ),
        name = "Score_Type"
      ) +
      
      new_scale_color() +
      geom_point(aes(color = resno, shape = Score_Type), size=3) +
      scale_color_gradientn(
        colors = rainbow_ngl,
        limits = c(0, max(data_subset_long$resno)),
        name = "Residue number"
      ) +
      scale_x_continuous(
        limit = c(0,NA)
      ) +
      labs(
      title = paste0(PDB_entry, " chain ", chain_index, " (",data_subset$protein[1],")"),
      x = "Residue variability",
      y = "Epitope prediction score"
        ) +
      theme_bw() #+
      #guides(
      #  shape = guide_legend(
      #    override.aes = list(
      #      color = setNames(
      #        c(
      #          palette.colors()[7],
      #          palette.colors()[6],
      #          palette.colors()[4],
      #          palette.colors()[8]
      #        ),
      #        pivot_list[1:4]
      #      )
      #    )
      #  )
      #)
      
      print(plot)
      
      ggsave(filename = paste0("output/figures/scatterplot/merged/", PDB_entry, "_", chain_index, "_merged.png"), 
           plot = plot, 
           width = 14, height = 10, dpi = 300,
           bg="white")
  }
}
```

```{, fig.width=14, warning=FALSE}
pivot_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score"
)

for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  
  protein_name <- unique(data_subset$protein)
  
  print(PDB_entry)
 
  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)
    
    data_subset_long <- data_subset_chain %>%
      select(resno,`BepiPred-2.0 probability`,`BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
      pivot_longer(cols = pivot_list, 
                   names_to = "Score_Type", values_to = "Score_Value") 
    
    plot <- data_subset_long %>% 
      ggplot(aes(x=var_freq,y=Score_Value)) +
      
      geom_smooth(
        aes(group = Score_Type),
        method = "lm",
        se = FALSE, size = 2, color="black"
      ) +
      scale_color_manual(
        values = setNames(
          c(
            palette.colors()[7],
            palette.colors()[6],
            palette.colors()[4],
            palette.colors()[8]
          ),
          pivot_list[1:4]
        ),
        name = "Score_Type"
      ) +
      new_scale_color() +
      geom_point(aes(color = resno), size=3) +
      scale_color_gradientn(
        colors = rainbow_ngl,
        limits = c(0, max(data_subset_long$resno)),
        name = "Residue number"
      ) +
      scale_x_continuous(
        limit = c(0,NA)
      ) +
      stat_regline_equation(aes(label = ..rr.label..), label.x = 0.05, size = 5) +
      facet_wrap(~Score_Type, nrow=2) +
        
      labs(
      title = paste0(PDB_entry, " chain ", data_subset$chain[1], " (",data_subset$protein[1],")"),
      x = "Residue variability",
      y = "Epitope prediction score"
        ) +
      theme_bw() #+
      #guides(
      #  shape = guide_legend(
      #    override.aes = list(
      #      color = setNames(
      #        c(
      #          palette.colors()[7],
      #          palette.colors()[6],
      #          palette.colors()[4],
      #          palette.colors()[8]
      #        ),
      #        pivot_list[1:4]
      #      )
      #    )
      #  )
      #)
      
      print(plot)
      
      ggsave(filename = paste0("output/figures/scatterplot/wrapped/", PDB_entry, "_", chain_index, "_wrapped.png"), 
           plot = plot, 
           width = 14, height = 10, dpi = 300,
           bg="white")
  }
}
```

```{r,fig.width=14}
pivot_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score"
)

min_freq <- 0.01

for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  protein_name <- unique(data_subset$protein)
  
  print(PDB_entry)

  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)

    data_subset_long <- data_subset_chain %>%
      select(resno, `BepiPred-2.0 probability`, `BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq) %>%
      pivot_longer(cols = pivot_list, 
                   names_to = "Score_Type", values_to = "Score_Value")

    mean_df <- data_subset_long %>%
      group_by(Score_Type) %>%
      summarize(mean_value = mean(Score_Value, na.rm = TRUE))

    data_subset_long_nonzero_var <- data_subset_long %>% filter(var_freq > min_freq)
    
    if (nrow(data_subset_long_nonzero_var) == 0) {
      message(paste("Skipping", PDB_entry, chain_index, "- no non-zero variability"))
      next
    }

    plot <- ggplot(data_subset_long_nonzero_var, aes(x = var_freq, y = Score_Value)) +
      
      # Average horizontal line per Score_Type
      geom_hline(
        data = mean_df,
        aes(yintercept = mean_value),
        linetype = "dashed",size=1.5, color = "gray"
      ) +
      
      geom_vline(
        xintercept = min_freq,
        linetype = "dashed",size=1.5, color = "gray"
      ) +
      
      # Linear regression line (fit on full data)
      geom_smooth(
        data = data_subset_long_nonzero_var,
        aes(group = Score_Type),
        method = "lm",
        se = FALSE, size = 2, color = "black"
      ) +

      scale_color_manual(
        values = setNames(
          c(
            palette.colors()[7],
            palette.colors()[6],
            palette.colors()[4],
            palette.colors()[8]
          ),
          pivot_list[1:4]
        ),
        name = "Score_Type"
      ) +
      new_scale_color() +
      
      stat_regline_equation(data = data_subset_long_nonzero_var, aes(label = ..rr.label..), size = 5) +
      facet_wrap(~Score_Type, nrow = 2) +
      
      geom_point(data = data_subset_long,aes(color = resno), size = 3) +
      scale_color_gradientn(
        colors = rainbow_ngl,
        limits = c(0, max(data_subset_long$resno)),
        name = "Residue number"
      ) +
      scale_x_continuous(limits = c(0, NA)) +

      labs(
        title = paste0(PDB_entry, " chain ", data_subset$chain[1], " (", data_subset$protein[1], ")"),
        x = "Residue variability",
        y = "Epitope prediction score"
      ) +
      theme_bw()

    print(plot)

    ggsave(filename = paste0("output/figures/scatterplot/wrapped/", PDB_entry, "_", chain_index, "_wrapped.png"), 
           plot = plot, 
           width = 14, height = 10, dpi = 300,
           bg = "white")
  }
}

```

#### vardist
```{r,fig.width=14}
pivot_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score"
)

for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  protein_name <- unique(data_subset$protein)
  
  print(PDB_entry)

  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)
    
    # Identify variable residues
    var_residues <- data_subset_chain %>% filter(var_freq > 0)
    
    if (nrow(var_residues) < 1) {
      message(paste("Skipping", PDB_entry, chain_index, "- no variable residues"))
      next
    }
    
    # Initialize distance column to Inf
    data_subset_chain <- data_subset_chain %>%
      mutate(min_distance_to_var = Inf,
             closest_var_freq = NA_real_)
    
    # Loop through each variable residue and update the minimum distance
    for (i in 1:nrow(var_residues)) {
      x_var <- var_residues$x[i]
      y_var <- var_residues$y[i]
      z_var <- var_residues$z[i]
      freq_var <- var_residues$var_freq[i]
      
      # Calculate distance from this variable residue to all residues
      distances <- sqrt((data_subset_chain$x - x_var)^2 +
                          (data_subset_chain$y - y_var)^2 +
                          (data_subset_chain$z - z_var)^2)
      
      # Find which distances are smaller than current min
      closer <- distances < data_subset_chain$min_distance_to_var
      
      # Update only where the new distance is shorter
      data_subset_chain$min_distance_to_var[closer] <- distances[closer]
      data_subset_chain$closest_var_freq[closer] <- freq_var
    }
    
    data_subset_long <- data_subset_chain %>%
      select(resno, `BepiPred-2.0 probability`, `BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq, min_distance_to_var, closest_var_freq) %>%
      pivot_longer(cols = pivot_list, 
                   names_to = "Score_Type", values_to = "Score_Value")
    
    
    data_subset_long_nonzero_dist <- data_subset_long %>% filter(min_distance_to_var > 0)
    
    mean_df <- data_subset_long %>%
      group_by(Score_Type) %>%
      summarize(mean_value = mean(Score_Value, na.rm = TRUE))
    
    plot <- data_subset_long %>% 
      ggplot(aes(x = min_distance_to_var, y = Score_Value)) +
      
      geom_hline(
        data = mean_df,
        aes(yintercept = mean_value),
        linetype = "dashed",size=1.5, color = "gray"
      ) +
      
      # Linear regression line (fit on full data)
      geom_smooth(
        data = data_subset_long_nonzero_dist,
        aes(group = Score_Type),
        method = "lm",
        se = FALSE, size = 2, color = "black"
      ) +

      scale_color_manual(
        values = setNames(
          c(
            palette.colors()[7],
            palette.colors()[6],
            palette.colors()[4],
            palette.colors()[8]
          ),
          pivot_list[1:4]
        ),
        name = "Score_Type"
      ) +
      new_scale_color() +
      
      # Plot only non-zero variability points
      geom_point(aes(color = resno, size = closest_var_freq), alpha = 0.5) +
      scale_color_gradientn(
        colors = rainbow_ngl,
        limits = c(0, max(data_subset_long$resno)),
        name = "Residue number"
      ) +
      scale_size_continuous(limits = c(0, NA),
                            name = "Variability closest variable residue") +
      scale_x_continuous(limits = c(0, NA)) +
      stat_regline_equation(aes(label = ..rr.label..), size = 5) +
      facet_wrap(~Score_Type, nrow = 2) +

      labs(
        title = paste0(PDB_entry, " chain ", data_subset$chain[1], " (", data_subset$protein[1], ")"),
        x = "CA distance to closest variable residue (A)",
        y = "Epitope prediction score"
      ) +
      theme_bw()

    print(plot)
    
    ggsave(filename = paste0("output/figures/scatterplot/wrapped/vardist/", PDB_entry, "_", chain_index, "_wrapped.png"), 
           plot = plot, 
           width = 14, height = 10, dpi = 300,
           bg = "white")

  }
}

```

#### freqdist
```{r,fig.width=14}
min_var_freq = 0.05
#max_distance_to_var = 10

pivot_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score"
)

for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  protein_name <- unique(data_subset$protein)
  
  print(PDB_entry)

  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)
    
    # Identify variable residues
    var_residues <- data_subset_chain %>% filter(var_freq > min_var_freq)
    
    if (nrow(var_residues) < 1) {
      message(paste("Skipping", PDB_entry, chain_index, "- no variable residues"))
      next
    }
    
    # Initialize distance column to Inf
    data_subset_chain <- data_subset_chain %>%
      mutate(min_distance_to_var = Inf,
             closest_var_freq = NA_real_)
    
    # Loop through each variable residue and update the minimum distance
    for (i in 1:nrow(var_residues)) {
      x_var <- var_residues$x[i]
      y_var <- var_residues$y[i]
      z_var <- var_residues$z[i]
      freq_var <- var_residues$var_freq[i]
      
      # Calculate distance from this variable residue to all residues
      distances <- sqrt((data_subset_chain$x - x_var)^2 +
                          (data_subset_chain$y - y_var)^2 +
                          (data_subset_chain$z - z_var)^2)
      
      # Find which distances are smaller than current min
      closer <- distances < data_subset_chain$min_distance_to_var
      
      # Update only where the new distance is shorter
      data_subset_chain$min_distance_to_var[closer] <- distances[closer]
      data_subset_chain$closest_var_freq[closer] <- freq_var
    }
    
    data_subset_long <- data_subset_chain %>%
      select(resno, `BepiPred-2.0 probability`, `BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq, min_distance_to_var, closest_var_freq) %>%
      pivot_longer(cols = pivot_list, 
                   names_to = "Score_Type", values_to = "Score_Value")
    
    
    data_subset_long_nonzero_dist <- data_subset_long %>% filter(min_distance_to_var > 0)
    
    mean_df <- data_subset_long %>%
      group_by(Score_Type) %>%
      summarize(mean_value = mean(Score_Value, na.rm = TRUE))
    
    plot <- data_subset_long %>% 
      ggplot(aes(x = closest_var_freq, y = Score_Value)) +
      
      geom_hline(
        data = mean_df,
        aes(yintercept = mean_value),
        linetype = "dashed",size=1.5, color = "gray"
      ) +
      
      # Linear regression line (fit on full data)
      geom_smooth(
        data = data_subset_long_nonzero_dist,
        aes(group = Score_Type),
        method = "lm",
        se = FALSE, size = 2, color = "black"
      ) +

      scale_color_manual(
        values = setNames(
          c(
            palette.colors()[7],
            palette.colors()[6],
            palette.colors()[4],
            palette.colors()[8]
          ),
          pivot_list[1:4]
        ),
        name = "Score_Type"
      ) +
      new_scale_color() +
      
      # Plot only non-zero variability points
      geom_point(aes(color = resno, size = -min_distance_to_var), alpha = 0.5) +
      scale_color_gradientn(
        colors = rainbow_ngl,
        limits = c(0, max(data_subset_long$resno)),
        name = "Residue number"
      ) +
      scale_x_continuous(limits = c(0, NA)) +
      scale_size_continuous(limits = c(-10, -3.8),
                            breaks = c(-10, -8, -6, -4, -3.8),
                            labels = c("10+", "8", "6", "4", "3.8"),
                            name = "Distance to variable residue (Å)") +
      stat_regline_equation(aes(label = ..rr.label..), size = 5) +
      facet_wrap(~Score_Type, nrow = 2) +

      labs(
        title = paste0(PDB_entry, " chain ", data_subset$chain[1], " (", data_subset$protein[1], ")"),
        x = "Variability of closest variable residue",
        y = "Epitope prediction score"
      ) +
      theme_bw()

    print(plot)
    
    ggsave(filename = paste0("output/figures/scatterplot/wrapped/freqdist/", PDB_entry, "_", chain_index, "_wrapped.png"), 
           plot = plot, 
           width = 14, height = 10, dpi = 300,
           bg = "white")

  }
}

```

```{r,fig.width=14}
min_var_freq = 0.05
max_distance_to_var = 10
min_RSA = 0.2

pivot_list <- c(
  "BepiPred-2.0 probability", 
  # "RSA NetsurfP", 
  "BepiPred-3.0 score",
  # "BepiPred-3.0 linear epitope score", 
  # "DiscoTope-3.0_score",
  "DiscoTope-3.0_calibrated_score", 
  # "RSA Shrake-Rupley", 
  "Epigraph_score"
)

for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  protein_name <- unique(data_subset$protein)
  
  print(PDB_entry)

  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)
    
    # Identify variable residues
    var_residues <- data_subset_chain %>% filter(var_freq > min_var_freq)
    
    if (nrow(var_residues) < 1) {
      message(paste("Skipping", PDB_entry, chain_index, "- no variable residues"))
      next
    }
    
    # Initialize distance column to Inf
    data_subset_chain <- data_subset_chain %>%
      mutate(min_distance_to_var = Inf,
             closest_var_freq = NA_real_)
    
    # Loop through each variable residue and update the minimum distance
    for (i in 1:nrow(var_residues)) {
      x_var <- var_residues$x[i]
      y_var <- var_residues$y[i]
      z_var <- var_residues$z[i]
      freq_var <- var_residues$var_freq[i]
      
      # Calculate distance from this variable residue to all residues
      distances <- sqrt((data_subset_chain$x - x_var)^2 +
                          (data_subset_chain$y - y_var)^2 +
                          (data_subset_chain$z - z_var)^2)
      
      # Find which distances are smaller than current min
      closer <- distances < data_subset_chain$min_distance_to_var
      
      # Update only where the new distance is shorter
      data_subset_chain$min_distance_to_var[closer] <- distances[closer]
      data_subset_chain$closest_var_freq[closer] <- freq_var
    }
    
    data_subset_long <- data_subset_chain %>%
      mutate(
        min_distance_to_var = case_when(
          min_distance_to_var == Inf ~ NA,
          TRUE ~ min_distance_to_var
        )
        ) %>%
      select(resno, `BepiPred-2.0 probability`, `BepiPred-3.0 score`, 
             `DiscoTope-3.0_calibrated_score`, Epigraph_score, var_freq, min_distance_to_var, closest_var_freq, RSA) %>%
      pivot_longer(cols = pivot_list, 
                   names_to = "Score_Type", values_to = "Score_Value")
    
    
    data_subset_long_nonzero_dist <- data_subset_long %>% 
      filter(closest_var_freq > min_var_freq,
             min_distance_to_var < max_distance_to_var,
             RSA > min_RSA)
    
    mean_df <- data_subset_long %>%
      filter(RSA > min_RSA) %>%
      group_by(Score_Type) %>%
      summarize(mean_value = mean(Score_Value, na.rm = TRUE))
    
    plot <- data_subset_long %>% 
      ggplot(aes(x = closest_var_freq, y = Score_Value)) +
      
      geom_hline(
        data = mean_df,
        aes(yintercept = mean_value),
        linetype = "dashed",size=1.5, color = "gray"
      ) +
      
      # Linear regression line (fit on full data)
      geom_smooth(
        data = data_subset_long_nonzero_dist,
        aes(group = Score_Type),
        method = "lm",
        se = FALSE, size = 2, color = "black"
      ) +

      scale_color_manual(
        values = setNames(
          c(
            palette.colors()[7],
            palette.colors()[6],
            palette.colors()[4],
            palette.colors()[8]
          ),
          pivot_list[1:4]
        ),
        name = "Score_Type"
      ) +
      new_scale_color() +
      
      # Plot only non-zero variability points
      geom_point(aes(color = resno, size = RSA), alpha = 0.5) +
      scale_color_gradientn(
        colors = rainbow_ngl,
        limits = c(0, max(data_subset_long$resno)),
        name = "Residue number"
      ) +
      scale_x_continuous(limits = c(0, NA)) +
      #scale_size_continuous(limits = c(-10, -3.8),
      #                      breaks = c(-10, -8, -6, -4, -3.8),
      #                      labels = c("10+", "8", "6", "4", "3.8"),
      #                      name = "Distance to variable residue (Å)") +
      stat_regline_equation(aes(label = ..rr.label..), size = 5) +
      facet_wrap(~Score_Type, nrow = 2) +

      labs(
        title = paste0(PDB_entry, " chain ", data_subset$chain[1], " (", data_subset$protein[1], ")"),
        x = "Variability of closest variable residue",
        y = "Epitope prediction score"
      ) +
      theme_bw()

    print(plot)
    
    #ggsave(filename = paste0("output/figures/scatterplot/wrapped/freqdist/", PDB_entry, "_", chain_index, "_wrapped.png"), 
    #       plot = plot, 
    #       width = 14, height = 10, dpi = 300,
    #       bg = "white")

  }
}

```


## violin
```{r, fig.width=14, warning=FALSE}
pivot_list <- c(
  "BepiPred-2.0 probability", 
  "BepiPred-3.0 score",
  "DiscoTope-3.0_calibrated_score", 
  "Epigraph_score"
)

min_freq <- 0.01

for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  protein_name <- unique(data_subset$protein)
  
  print(PDB_entry)
  
  longest_chain <- data_subset %>% 
    filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
    add_count(chain, name = "chain_length") %>%
    arrange(desc(chain_length), chain) %>%  # Sort by length (descending), then alphabetically
    filter(chain_length == max(chain_length)) %>%  
    filter(chain == min(chain)) %>% 
    pull(chain) %>%
    unique()

  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)

    # Prepare data
    data_subset_long_violin <- data_subset_chain %>%
      select(resno, all_of(pivot_list), var_freq) %>%
      pivot_longer(cols = pivot_list, names_to = "Score_Type", values_to = "Score_Value") %>%
      mutate(
        Variability_Group = ifelse(var_freq > min_freq, "Variable", "Not variable")
      )

    # Skip if only one group exists
    if (length(unique(data_subset_long_violin$Variability_Group)) < 2) {
      message(paste("Skipping", PDB_entry, chain_index, "- only one variability group"))
      next
    }

    plot_violin <- ggplot(data_subset_long_violin, aes(x = Variability_Group, y = Score_Value, fill = Variability_Group)) +
      geom_violin(trim = FALSE, scale = "width", alpha = 0.6) +
      geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.8) +
      stat_compare_means(
        method = "wilcox.test",
        label = "p.signif",
        comparisons = list(c("Not variable", "Variable")),
        hide.ns = TRUE
      ) +
      facet_wrap(~ Score_Type, nrow = 2) +
      scale_fill_manual(values = c("Not variable" = "#999999", "Variable" = "#E69F00")) +
      labs(
        title = paste0(PDB_entry, " chain ", chain_index, " (", protein_name, ")"),
        x = "Residue variability",
        y = "Epitope prediction score"
      ) +
      theme_bw() +
      theme(legend.position = "none")

    print(plot_violin)

    ggsave(
      filename = paste0("output/figures/violin/", PDB_entry, "_", chain_index, "_violin.png"),
      plot = plot_violin,
      width = 14, height = 10, dpi = 300, bg = "white"
    )
  }
}



```

### within 10A
```{r}
pivot_list <- c(
  "BepiPred-2.0 probability", 
  "BepiPred-3.0 score",
  "DiscoTope-3.0_calibrated_score", 
  "Epigraph_score"
)

score_label_map <- c(
        "BepiPred-2.0 probability" = "BepiPred 2.0",
        "BepiPred-3.0 score" = "BepiPred 3.0",
        "DiscoTope-3.0_calibrated_score" = "DiscoTope 3.0",
        "Epigraph_score" = "Epigraph"
      )

```

```{r, fig.width=14}
# Loop through each unique PDB entry
for (PDB_entry in unique(data_predictions_with_variability_normalized$PDB)) {
  data_subset <- data_predictions_with_variability_normalized %>% filter(PDB == PDB_entry)
  protein_name <- unique(data_subset$protein)
  
  print(PDB_entry)
  
  # Determine the longest chain
  longest_chain <- data_subset %>% 
    filter(resid %in% toupper(AMINO_ACID_CODE)) %>%
    add_count(chain, name = "chain_length") %>%
    arrange(desc(chain_length), chain) %>%
    filter(chain_length == max(chain_length)) %>%  
    filter(chain == min(chain)) %>% 
    pull(chain) %>%
    unique()
  
  # Loop through each chain
  for (chain_index in unique(data_subset$chain)) {
    print(chain_index)
    data_subset_chain <- data_subset %>% filter(chain == chain_index)
    
    # Identify residues with var_freq > 0
    var_residues <- data_subset_chain %>% filter(var_freq > 0)
    
    if (nrow(var_residues) < 1) {
      message(paste("Skipping", PDB_entry, chain_index, "- no variable residues"))
      next
    }
    
    # Initialize distance column to Inf
    data_subset_chain <- data_subset_chain %>%
      mutate(min_distance_to_var = Inf,
             closest_var_freq = NA_real_)
    
    # Loop through each variable residue and update the minimum distance
    for (i in 1:nrow(var_residues)) {
      x_var <- var_residues$x[i]
      y_var <- var_residues$y[i]
      z_var <- var_residues$z[i]
      freq_var <- var_residues$var_freq[i]
      
      # Calculate distance from this variable residue to all residues
      distances <- sqrt((data_subset_chain$x - x_var)^2 +
                          (data_subset_chain$y - y_var)^2 +
                          (data_subset_chain$z - z_var)^2)
      
      # Find which distances are smaller than current min
      closer <- distances < data_subset_chain$min_distance_to_var
      
      # Update only where the new distance is shorter
      data_subset_chain$min_distance_to_var[closer] <- distances[closer]
      data_subset_chain$closest_var_freq[closer] <- freq_var
    }
    
    data_subset_long_violin <- data_subset_chain %>%
      select(resno, all_of(pivot_list), var_freq, min_distance_to_var) %>%
      pivot_longer(cols = pivot_list, names_to = "Score_Type", values_to = "Score_Value") %>%
      mutate(
        Near_Var_Group = ifelse(min_distance_to_var <= 10, "Near variable", "Not near variable")
      )
    
    plot_violin <- ggplot(data_subset_long_violin, aes(x = Near_Var_Group, y = Score_Value, fill = Near_Var_Group)) +
      geom_violin(trim = FALSE, scale = "width", alpha = 0.6) +
      geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.8) +
      stat_compare_means(
        method = "wilcox.test",
        label = "p.signif",
        comparisons = list(c("Not near variable", "Near variable")),
        hide.ns = TRUE
      ) +
      facet_wrap(~ Score_Type, nrow = 2) +
      scale_fill_manual(values = c("Not near variable" = "#999999", "Near variable" = "#E69F00")) +
      labs(
        title = paste0(PDB_entry, " chain ", chain_index, " (", protein_name, ")"),
        x = "Residue proximity to variable position (<10 CA euclidean distance)",
        y = "Epitope prediction score"
      ) +
      theme_bw() +
      theme(legend.position = "none")
    
    print(plot_violin)
    
    ggsave(
      filename = paste0("output/figures/violin/vardist/", PDB_entry, "_", chain_index, "_violin.png"),
      plot = plot_violin,
      width = 14, height = 10, dpi = 300, bg = "white"
    )
  }
}
```


## summary statistics
```{r}
global_pLDDT_minimum <- 80

data_predictions_with_variability_normalized %>% group_by(proteinclass,protein,chain,structure) %>%
  filter(is.na(PDB)) %>%
  summarize(mean_b = mean(b)) %>%
  filter(!is.na(mean_b)) %>%
  arrange(mean_b)

low_AF_confidence_proteins <- data_predictions_with_variability_normalized %>%
  group_by(proteinclass,protein,chain,structure) %>%
  filter(is.na(PDB)) %>%
  summarize(mean_b = mean(b)) %>%
  filter(mean_b < global_pLDDT_minimum) %>% 
  pull(protein)

low_AF_confidence_proteins

high_confidence_proteins <- data_predictions_with_variability_normalized %>%
  filter(!(protein %in% low_AF_confidence_proteins))

high_confidence_proteins
```


### score vs var
```{r}
data_predictions_with_variability_normalized %>% filter(PDB == "1W9R")
```

```{r, fig.width=14}
library(ggrepel)
min_var_freq <- 0.05

high_confidence_proteins %>%
  pivot_longer(cols = pivot_list, 
               names_to = "Score_Type", values_to = "Score_Value") %>%
  group_by(proteinclass, structure, Score_Type) %>%
  mutate(score_label = score_label_map[Score_Type]) %>%
  filter(PDB == "2BIB") %>%
  #filter(RSA > 0.2) %>%
  #summarize(
  #  mean_score = mean(Score_Value, na.rm=TRUE),
  #  mean_variability = mean(var_freq, na.rm=TRUE),
  #  n_seq = mean(n_seq),
  #  seq_length = mean(seq_length)
  #) %>%
  #mutate(
  #  mean_variability = case_when(
  #    is.na(mean_variability) ~ 0,
  #    TRUE ~ mean_variability
  #  ),
  #  mean_score = case_when(
  #    is.na(mean_score) ~ 0,
  #    TRUE ~ mean_score
  #  )
  #) %>%
  #filter(mean_score != 0,
  #       mean_variability != 0) %>%
  ggplot(aes(x=var_freq, y=Score_Value)) +
  geom_point(aes(col = proteinclass)) +
  #geom_text_repel(aes(label = PDB, col = proteinclass, alpha = n_seq/seq_length), min.segment.length = 0.01, max.overlaps = 100) +
  
  geom_text(
    aes(x=(max(var_freq, na.rm = TRUE) - min(var_freq, na.rm = TRUE))/2,
        y=Inf, label = score_label), check_overlap = TRUE, vjust = 1, size = 5
            ) +
  
  geom_smooth(
        data = . %>% filter(var_freq > min_var_freq),
        #aes(group = structure, color = proteinclass),
        method = "lm",
        se = FALSE, size = 2, color="black"
      ) +
  
  #geom_smooth(
  #      data = . %>% filter(var_freq > min_var_freq),
  #      aes(group = structure, color = proteinclass),
  #      method = "lm",
  #      se = FALSE, size = 1
  #    ) +
  
  stat_regline_equation(
    data = . %>% filter(var_freq > min_var_freq),
    aes(label = ..rr.label..), size = 5, 
    label.x.npc = "right",   
    label.y.npc = "top",     
    hjust = 1,               
    vjust = 1                
  ) +
  
  scale_y_continuous(
    limits = c(0, 1.1),
    expand = c(0, 0), 
    breaks = seq(0,1, by = 0.25),
    labels = seq(0,1, by = 0.25)
  ) +
  
  labs(
    y= "Residue epitope score",
    x= "Residue variability"
  ) +
  scale_alpha_continuous(limits = c(0,NA)) +
  theme_bw() +
  facet_wrap(~Score_Type) +
  theme(
    panel.spacing = unit(1, "lines"), 
    #panel.grid = element_blank()
    
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    
    strip.background = element_blank(),
    strip.text = element_blank()
  )
```

```{r, fig.width=14}
library(ggrepel)
min_var_freq <- 0.05

high_confidence_proteins %>%
  pivot_longer(cols = pivot_list, 
               names_to = "Score_Type", values_to = "Score_Value") %>%
  group_by(proteinclass, structure, Score_Type) %>%
  mutate(score_label = score_label_map[Score_Type]) %>%
  #filter(RSA > 0.2) %>%
  #summarize(
  #  mean_score = mean(Score_Value, na.rm=TRUE),
  #  mean_variability = mean(var_freq, na.rm=TRUE),
  #  n_seq = mean(n_seq),
  #  seq_length = mean(seq_length)
  #) %>%
  #mutate(
  #  mean_variability = case_when(
  #    is.na(mean_variability) ~ 0,
  #    TRUE ~ mean_variability
  #  ),
  #  mean_score = case_when(
  #    is.na(mean_score) ~ 0,
  #    TRUE ~ mean_score
  #  )
  #) %>%
  #filter(mean_score != 0,
  #       mean_variability != 0) %>%
  ggplot(aes(x=var_freq, y=Score_Value)) +
  geom_point(aes(col = proteinclass)) +
  #geom_text_repel(aes(label = PDB, col = proteinclass, alpha = n_seq/seq_length), min.segment.length = 0.01, max.overlaps = 100) +
  
  geom_text(
    aes(x=(max(var_freq, na.rm = TRUE) - min(var_freq, na.rm = TRUE))/2,
        y=Inf, label = score_label), check_overlap = TRUE, vjust = 1, size = 5
            ) +
  
  geom_smooth(
        data = . %>% filter(var_freq > min_var_freq),
        #aes(group = structure, color = proteinclass),
        method = "lm",
        se = FALSE, size = 2, color="black"
      ) +
  
  geom_smooth(
        data = . %>% filter(var_freq > min_var_freq),
        aes(group = structure, color = proteinclass),
        method = "lm",
        se = FALSE, size = 1
      ) +
  
  stat_regline_equation(
    data = . %>% filter(var_freq > min_var_freq),
    aes(label = ..rr.label..), size = 5, 
    label.x.npc = "right",   
    label.y.npc = "top",     
    hjust = 1,               
    vjust = 1                
  ) +
  
  scale_y_continuous(
    limits = c(0, 1.1),
    expand = c(0, 0), 
    breaks = seq(0,1, by = 0.25),
    labels = seq(0,1, by = 0.25)
  ) +
  
  labs(
    y= "Residue epitope score",
    x= "Residue variability"
  ) +
  scale_alpha_continuous(limits = c(0,NA)) +
  theme_bw() +
  facet_wrap(~Score_Type) +
  theme(
    panel.spacing = unit(1, "lines"), 
    #panel.grid = element_blank()
    
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    
    strip.background = element_blank(),
    strip.text = element_blank()
  )
```


```{r}
high_confidence_proteins_with_seroprof <- high_confidence_proteins %>%
  rowwise() %>%
    mutate(
    seroprof_titers = data_variability_with_seroprof$seroprof_titers[
      match(structure, data_variability_with_seroprof$structure)
    ]
  )
```


```{r, fig.width=14}
library(ggrepel)
min_var_freq <- 0.05

high_confidence_proteins_with_seroprof %>%
  pivot_longer(cols = pivot_list, 
               names_to = "Score_Type", values_to = "Score_Value") %>%
  group_by(proteinclass, structure, Score_Type) %>%
  mutate(score_label = score_label_map[Score_Type]) %>%
  #filter(RSA > 0.2) %>%
  #summarize(
  #  mean_score = mean(Score_Value, na.rm=TRUE),
  #  mean_variability = mean(var_freq, na.rm=TRUE),
  #  n_seq = mean(n_seq),
  #  seq_length = mean(seq_length)
  #) %>%
  #mutate(
  #  mean_variability = case_when(
  #    is.na(mean_variability) ~ 0,
  #    TRUE ~ mean_variability
  #  ),
  #  mean_score = case_when(
  #    is.na(mean_score) ~ 0,
  #    TRUE ~ mean_score
  #  )
  #) %>%
  #filter(mean_score != 0,
  #       mean_variability != 0) %>%
  ggplot(aes(x=var_freq, y=Score_Value)) +
  geom_point(aes(col = seroprof_titers)) +
  #geom_text_repel(aes(label = PDB, col = proteinclass, alpha = n_seq/seq_length), min.segment.length = 0.01, max.overlaps = 100) +
  
  geom_text(
    aes(x=(max(var_freq, na.rm = TRUE) - min(var_freq, na.rm = TRUE))/2,
        y=Inf, label = score_label), check_overlap = TRUE, vjust = 1, size = 5
            ) +
  
  geom_smooth(
        data = . %>% filter(var_freq > min_var_freq),
        #aes(group = structure, color = proteinclass),
        method = "lm",
        se = FALSE, size = 2, color="black"
      ) +
  
  geom_smooth(
        data = . %>% filter(var_freq > min_var_freq),
        aes(group = structure, color = seroprof_titers),
        method = "lm",
        se = FALSE, size = 1
      ) +
  
  stat_regline_equation(
    data = . %>% filter(var_freq > min_var_freq),
    aes(label = ..rr.label..), size = 5, 
    label.x.npc = "right",   
    label.y.npc = "top",     
    hjust = 1,               
    vjust = 1                
  ) +
  
  scale_y_continuous(
    limits = c(0, 1.1),
    expand = c(0, 0), 
    breaks = seq(0,1, by = 0.25),
    labels = seq(0,1, by = 0.25)
  ) +
  
  labs(
    y= "Residue epitope score",
    x= "Residue variability"
  ) +
  scale_alpha_continuous(limits = c(0,NA)) +
  theme_bw() +
  facet_wrap(~Score_Type) +
  theme(
    panel.spacing = unit(1, "lines"), 
    #panel.grid = element_blank()
    
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    
    strip.background = element_blank(),
    strip.text = element_blank()
  )
```

```{r, fig.width=14}
library(ggrepel)

high_confidence_proteins_with_seroprof %>%
  pivot_longer(cols = pivot_list, 
               names_to = "Score_Type", values_to = "Score_Value") %>%
  group_by(proteinclass, structure, Score_Type) %>%
  #filter(RSA > 0.2) %>%
  summarize(
    mean_score = mean(Score_Value, na.rm=TRUE),
    mean_variability = mean(var_freq, na.rm=TRUE),
    n_seq = mean(n_seq),
    seq_length = mean(seq_length)
  ) %>%
  mutate(score_label = score_label_map[Score_Type]) %>%
  #mutate(
  #  mean_variability = case_when(
  #    is.na(mean_variability) ~ 0,
  #    TRUE ~ mean_variability
  #  ),
  #  mean_score = case_when(
  #    is.na(mean_score) ~ 0,
  #    TRUE ~ mean_score
  #  )
  #) %>%
  #filter(mean_score != 0,
  #       mean_variability != 0) %>%
  ggplot(aes(x=mean_variability, y=mean_score)) +
  geom_point(
    aes(col = proteinclass#, 
        #alpha = n_seq/seq_length
        )
    ) +
  #geom_text_repel(aes(label = PDB, col = proteinclass, alpha = n_seq/seq_length), min.segment.length = 0.01, max.overlaps = 100) +
  
  geom_text(
    aes(x=(max(mean_variability, na.rm = TRUE) - min(mean_variability, na.rm = TRUE))/2,
        y=Inf, label = score_label), check_overlap = TRUE, vjust = 1, size = 5
            ) +
  
  geom_smooth(
        data = . %>% filter(mean_variability > 0),
        #aes(group = structure, color = proteinclass),
        method = "lm",
        se = FALSE, size = 2, color="black"
      ) +
  
  geom_smooth(
        data = . %>% filter(mean_variability > 0),
        aes(col = proteinclass),
        method = "lm",
        se = FALSE, size = 1
      ) +
  
  stat_regline_equation(
    data = . %>% filter(mean_variability > 0),
    aes(label = ..rr.label..), size = 5, 
    label.x.npc = "right",   
    label.y.npc = "top",     
    hjust = 1,               
    vjust = 1                
  ) +
  
  labs(
    y= "Mean epitope score",
    x= "Mean variability"
  ) +
  
  scale_y_continuous(
    limits = c(0, 1.1),
    expand = c(0, 0), 
    breaks = seq(0,1, by = 0.25),
    labels = seq(0,1, by = 0.25)
  ) +
  theme_bw() +
  facet_wrap(~Score_Type) +
  theme(
    panel.spacing = unit(1, "lines"), 
    #panel.grid = element_blank()
    
    strip.background = element_blank(),
    strip.text = element_blank()
  )
```

### residue only
#### regression
```{r}
regression_results_all <- high_confidence_proteins_with_seroprof %>%
  filter(!(protein %in% low_AF_confidence_proteins)) %>%
  select(PDB,resno, all_of(pivot_list), var_freq, seroprof_titers) %>%
  pivot_longer(cols = pivot_list, 
                   names_to = "Score_Type", values_to = "Score_Value") %>%
  group_by(PDB, Score_Type) %>%
  na.omit() %>%
  group_modify(~ {
    model <- lm(Score_Value ~ var_freq, data = .x, na.action = "na.exclude")
    tidy(model) %>% mutate(
      r.squared = summary(model)$r.squared
    ) 
  }) #%>%
  #ungroup()

regression_results_all
```



#### binary classification
##### var only
```{r}
RSA_min <- 0.15
var_freq_min <- 0.05

class_results_residue <- high_confidence_proteins_with_seroprof %>%
  
  select(proteinclass,protein,PDB,resno, chain,all_of(pivot_list), var_freq, RSA, x, y, z, b, seqresno, seroprof_titers) %>%
  pivot_longer(cols = pivot_list, 
               names_to = "Score_Type", values_to = "Score_Value") %>%
  group_by(PDB, Score_Type) %>%
  mutate(
    var_class_00 = case_when(
      is.na(var_freq) ~ NA_character_,
      var_freq <= 0 ~ "nvar",
      TRUE ~ "var"
    ),
    var_class_05 = case_when(
      is.na(var_freq) ~ NA_character_,
      var_freq <= var_freq_min ~ "nvar",
      TRUE ~ "var"
    ),
    RSA_class = case_when(
      RSA < RSA_min ~ "nsur",
      TRUE ~ "sur"
    )
  ) %>%
  pivot_longer(
    cols = c(var_class_00, var_class_05),
    names_to = "var_class_type",
    values_to = "var_class"
  ) %>%
  ungroup() %>%
  #drop_na(Score_Value, var_class) %>%
  group_by(PDB, Score_Type, var_class_type) #%>%
  #nest() %>%
  #mutate(
  #  valid = map_lgl(data, ~ {
  #    vc <- .x$var_class
  #    g_sizes <- table(vc)
  #    length(g_sizes) == 2 && all(g_sizes >= 2)
  #  }),

  #  t_test = map2(data, valid, ~ if (.y) t.test(Score_Value ~ var_class, data = .x) else NULL),

  #  tidy_res = map(t_test, ~ if (!is.null(.x)) tidy(.x) else NULL)
  #) %>%
  #unnest(tidy_res)

class_results_residue
```

```{r, fig.width=14}
class_results_residue %>%
  drop_na(var_class) %>%
  filter(var_class_type == "var_class_05") %>%
  mutate(score_label = score_label_map[Score_Type]) %>%
  ggplot(aes(x= var_class, fill = var_class, y = Score_Value)) +
  geom_violin(
    aes(fill = var_class),
    trim = FALSE, scale = "width", alpha = 0.6, position = position_dodge(width = 0.8)
  ) +
  geom_boxplot(
    aes(fill = var_class),
    width = 0.1, outlier.shape = NA, alpha = 0.8, position = position_dodge(width = 0.8)
  ) +
  #geom_point(aes(color = proteinclass)) +
  #geom_path(aes(group = structure, color = proteinclass)) +
  stat_compare_means(
    comparisons = list(c("nvar", "var")),
    method = "wilcox",
    label = "p.signif",
    hide.ns = TRUE, label.y = 1.1
  ) +
  labs(
    x = "",
    y = "Residue epitope prediction score",
    title = "Not variable vs variable (wilcox.test)",
    fill = "Variability"
  ) +
  #facet_grid(var_class_type~proteinclass) +
  facet_grid(
    ~Score_Type, 
    labeller = labeller(
      Score_Type = score_label_map
    )
  ) +
  geom_text(
    aes(x=1.5,y=1.3, label = score_label), check_overlap = TRUE, size = 5
            ) +
  scale_fill_manual(
    values = c("nvar" = "#D55E00", "var" = "#56B4E9"),
    labels = c("nvar" = "not variable", "var" = "variable")
  ) +
  scale_x_discrete(
    labels = c(
      "nvar" = "not variable", 
      "var" = "variable"
    )
  ) +
  scale_y_continuous(
    limits = c(0, 1.35),
    expand = c(0, 0), 
    breaks = seq(0,1, by = 0.25),
    labels = seq(0,1, by = 0.25)
  ) +
  theme_bw() +
  theme(
    panel.spacing = unit(0, "lines"), 
    legend.position = "none",
    panel.grid = element_blank(), 
    #axis.title.x = ,
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),

    strip.background = element_blank(),
    strip.text = element_blank()
    )
```

```{r, fig.width = 14}
class_results_residue %>%
  drop_na(var_class) %>%
  filter(var_class_type == "var_class_05") %>%
  mutate(
    is_sur.var = factor(
      interaction(RSA_class, var_class),
      levels = c("nsur.nvar","nsur.var", "sur.nvar", "sur.var")
      )
    ) %>%
  mutate(score_label = score_label_map[Score_Type]) %>%
  ggplot(aes(x= is_sur.var, fill = is_sur.var, y = Score_Value)) +
  geom_violin(
    trim = FALSE, scale = "width", alpha = 0.6, position = position_dodge(width = 0.8)
  ) +
  geom_boxplot(
    width = 0.1, outlier.shape = NA, alpha = 0.8, position = position_dodge(width = 0.8)
  ) +
  #geom_point(aes(color = proteinclass)) +
  #geom_path(aes(group = structure, color = proteinclass)) +
  stat_compare_means(
    comparisons = combn(c("nsur.nvar","nsur.var", "sur.nvar", "sur.var"), 2, simplify = FALSE),
    method = "wilcox",
    label = "p.signif",
    hide.ns = FALSE,
    label.y = seq(1,1.6,by=0.1)
  ) +
  labs(
    x = "",
    y = "Residue epitope prediction score",
    title = "Not variable vs variable & surface vs non-surface (wilcox.test)",
    fill = "Accessibility % Variability"
  ) +
  #facet_grid(var_class_type~proteinclass) +
  facet_grid(~Score_Type) +
  scale_fill_manual(
    values =  c("nsur.nvar" = "#D55E00","nsur.var" = "#E69F00", "sur.nvar"  = "#0072B2", "sur.var" = "#56B4E9"),
    labels =  c("nsur.nvar" = "non-surface\nnot variable", 
                "nsur.var" = "non-surface\nvariable", 
                "sur.nvar" = "surface\nnot variable", 
                "sur.var" = "surface\nvariable")
  ) +
  scale_x_discrete(
    labels = c(
      "nsur.nvar" = "non-surface\nnot variable", 
      "nsur.var" = "non-surface\nvariable", 
      "sur.nvar" = "surface\nnot variable", 
      "sur.var" = "surface\nvariable"
    )
  ) +
  
  geom_text(
    aes(x=2.5,y=1.65, label = score_label), check_overlap = TRUE, size = 5
            ) +
  
  scale_y_continuous(
    limits = c(0, 1.75),
    expand = c(0, 0), 
    breaks = seq(0,1, by = 0.25),
    labels = seq(0,1, by = 0.25)
  ) +
  theme_bw() +
  theme(
    panel.spacing = unit(0, "lines"), 
    legend.position = "none",
    panel.grid = element_blank(), 
    #axis.title.x = ,
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),

    strip.background = element_blank(),
    strip.text = element_blank()
    )
```

```{r, fig.width = 14}
class_results_residue %>%
  drop_na(var_class) %>%
  filter(var_class_type == "var_class_05") %>%
  mutate(
    is_sur.var = factor(
      interaction(RSA_class, var_class),
      levels = c("nsur.nvar","nsur.var", "sur.nvar", "sur.var")
      )
    ) %>%
  mutate(score_label = score_label_map[Score_Type]) %>%
  ggplot(aes(x= is_sur.var, fill = is_sur.var, y = Score_Value)) +
  geom_violin(
    trim = FALSE, scale = "width", alpha = 0.6, position = position_dodge(width = 0.8)
  ) +
  geom_boxplot(
    width = 0.1, outlier.shape = NA, alpha = 0.8, position = position_dodge(width = 0.8)
  ) +
  #geom_point(aes(color = proteinclass)) +
  #geom_path(aes(group = structure, color = proteinclass)) +
  stat_compare_means(
    comparisons = combn(c("nsur.nvar","nsur.var", "sur.nvar", "sur.var"), 2, simplify = FALSE),
    method = "wilcox",
    label = "p.signif",
    hide.ns = FALSE,
    label.y = seq(1,1.6,by=0.1)
  ) +
  labs(
    x = "",
    y = "Residue epitope prediction score",
    title = "Not variable vs variable & surface vs non-surface (wilcox.test)",
    fill = "Accessibility % Variability"
  ) +
  #facet_grid(var_class_type~proteinclass) +
  facet_grid(proteinclass~score_label) +
  scale_fill_manual(
    values =  c("nsur.nvar" = "#D55E00","nsur.var" = "#E69F00", "sur.nvar"  = "#0072B2", "sur.var" = "#56B4E9"),
    labels =  c("nsur.nvar" = "non-surface\nnot variable", 
                "nsur.var" = "non-surface\nvariable", 
                "sur.nvar" = "surface\nnot variable", 
                "sur.var" = "surface\nvariable")
  ) +
  scale_x_discrete(
    labels = c(
      "nsur.nvar" = "non-surface\nnot variable", 
      "nsur.var" = "non-surface\nvariable", 
      "sur.nvar" = "surface\nnot variable", 
      "sur.var" = "surface\nvariable"
    )
  ) +
  
  #geom_text(
  #  aes(x=2.5,y=1.65, label = score_label), check_overlap = TRUE, size = 5
  #          ) +
  #
  scale_y_continuous(
    limits = c(0, 1.75),
    expand = c(0, 0), 
    breaks = seq(0,1, by = 0.25),
    labels = seq(0,1, by = 0.25)
  ) +
  theme_bw() +
  theme(
    panel.spacing = unit(0, "lines"), 
    legend.position = "none",
    panel.grid = element_blank(), 
    #axis.title.x = ,
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),

    #strip.background = element_blank(),
    #strip.text = element_blank()
    )
```

```{r, fig.width = 14}
class_results_residue %>%
  drop_na(var_class) %>%
  filter(var_class_type == "var_class_05") %>%
  mutate(
    is_sur.var = factor(
      interaction(RSA_class, var_class),
      levels = c("nsur.nvar","nsur.var", "sur.nvar", "sur.var")
      )
    ) %>%
  mutate(score_label = score_label_map[Score_Type]) %>%
  ggplot(aes(x= is_sur.var, fill = is_sur.var, y = Score_Value)) +
  geom_violin(
    trim = FALSE, scale = "width", alpha = 0.6, position = position_dodge(width = 0.8)
  ) +
  geom_boxplot(
    width = 0.1, outlier.shape = NA, alpha = 0.8, position = position_dodge(width = 0.8)
  ) +
  #geom_point(aes(color = proteinclass)) +
  #geom_path(aes(group = structure, color = proteinclass)) +
  stat_compare_means(
    comparisons = combn(c("nsur.nvar","nsur.var", "sur.nvar", "sur.var"), 2, simplify = FALSE),
    method = "wilcox",
    label = "p.signif",
    hide.ns = FALSE,
    label.y = seq(1,1.6,by=0.1)
  ) +
  labs(
    x = "",
    y = "Residue epitope prediction score",
    title = "Not variable vs variable & surface vs non-surface (wilcox.test)",
    fill = "Accessibility % Variability"
  ) +
  #facet_grid(var_class_type~proteinclass) +
  facet_grid(seroprof_titers~score_label) +
  scale_fill_manual(
    values =  c("nsur.nvar" = "#D55E00","nsur.var" = "#E69F00", "sur.nvar"  = "#0072B2", "sur.var" = "#56B4E9"),
    labels =  c("nsur.nvar" = "non-surface\nnot variable", 
                "nsur.var" = "non-surface\nvariable", 
                "sur.nvar" = "surface\nnot variable", 
                "sur.var" = "surface\nvariable")
  ) +
  scale_x_discrete(
    labels = c(
      "nsur.nvar" = "non-surface\nnot variable", 
      "nsur.var" = "non-surface\nvariable", 
      "sur.nvar" = "surface\nnot variable", 
      "sur.var" = "surface\nvariable"
    )
  ) +
  
  #geom_text(
  #  aes(x=2.5,y=1.65, label = score_label), check_overlap = TRUE, size = 5
  #          ) +
  #
  scale_y_continuous(
    limits = c(0, 1.75),
    expand = c(0, 0), 
    breaks = seq(0,1, by = 0.25),
    labels = seq(0,1, by = 0.25)
  ) +
  theme_bw() +
  theme(
    panel.spacing = unit(0, "lines"), 
    legend.position = "none",
    panel.grid = element_blank(), 
    #axis.title.x = ,
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),

    #strip.background = element_blank(),
    #strip.text = element_blank()
    )
```


##### means
```{}
class_results_residue_means <- class_results_residue %>%
  group_by(PDB,protein,proteinclass, Score_Type, var_class_type) %>%
  drop_na() %>%
  group_modify(function(.x, .y) {
    data_do <- .x
    
    # Ensure var_class has exactly 2 levels
    if (n_distinct(data_do$var_class) != 2) {
      return(tibble())
    }
    
    # Ensure both groups have at least 2 observations
    counts <- table(data_do$var_class)
    if (any(counts < 2)) {
      return(tibble())
    }

    # Run t-test
    test <- t.test(Score_Value ~ var_class, data = data_do)
    tidy_out <- tidy(test)

    # Rename group means using actual level names
    var_levels <- levels(factor(data_do$var_class))
    tidy_out <- tidy_out %>%
      rename_with(~ var_levels, c("estimate1", "estimate2"))

    tidy_out
  }) %>%
  pivot_longer(
    cols = c("not variable", "variable"),  # adjust if needed
    names_to = "mean_group",
    values_to = "mean_value"
  )

class_results_residue_means
```

```{, fig.width=14}
class_results_residue_means %>%
  ggplot(aes(x = Score_Type, fill = mean_group, y = mean_value, group = interaction(Score_Type, mean_group))) +
  geom_violin(
    trim = FALSE, scale = "width", alpha = 0.6, position = position_dodge(width = 0.8)
  ) +
  
  geom_line(
    aes(
      x = as.numeric(factor(Score_Type)) + ifelse(mean_group == "not variable", -0.2, 0.2),
      group = interaction(PDB, Score_Type),
      color = proteinclass,
      alpha = -log10(p.value)
    )
  ) +
  
  geom_boxplot(
    width = 0.1, outlier.shape = NA, alpha = 0.8, position = position_dodge(width = 0.8)
  ) +
  geom_point(
    aes(color = proteinclass),
    position = position_dodge(width = 0.8)
  ) +
  
  stat_compare_means(
    #aes(group = interaction(Score_Type, mean_group)),
    #comparisons = c("not variable", "variable"),
    method = "wilcox.test",
    label = "p.signif",
    hide.ns = FALSE
  ) +
  
  #geom_text(
  #  aes(
  #    x = Score_Type, 
  #    group = interaction(Score_Type, mean_group),
  #    label = case_when(
  #      p.value < 0.001 ~ "***",
  #      p.value < 0.01  ~ "**",
  #      p.value < 0.05  ~ "*",
  #      TRUE            ~ "ns"
  #    )
  #  )
  #) +
  
  theme_bw() +
  labs(
    x = "Prediction method",
    y = "Mean prediction score",
    title = "Mean prediction score: not variable vs variable (wilcox.test)"
  ) +
  scale_fill_manual(
    values = c(
      "not variable" = "#999999",
      "variable" = "#E69F00"
    )
  ) +
  
  scale_alpha_continuous(
    limits = c(0,2),
    breaks = c(0, -log10(0.05), 2)
  ) +
  
  facet_wrap(~var_class_type, nrow=2) 

  
```

```{, fig.width=14}
class_results_residue_means %>%
  ggplot(aes(x = Score_Type, fill = mean_group, y = mean_value, group = interaction(Score_Type, mean_group))) +
  geom_violin(
    trim = FALSE, scale = "width", alpha = 0.6, position = position_dodge(width = 0.8)
  ) +
  
  geom_line(
    aes(
      x = as.numeric(factor(Score_Type)) + ifelse(mean_group == "not variable", -0.2, 0.2),
      group = interaction(PDB, Score_Type),
      color = proteinclass,
      alpha = -log10(p.value)
    )
  ) +
  
  geom_boxplot(
    width = 0.1, outlier.shape = NA, alpha = 0.8, position = position_dodge(width = 0.8)
  ) +
  geom_point(
    aes(color = proteinclass),
    position = position_dodge(width = 0.8)
  ) +
  
  stat_compare_means(
    #aes(group = interaction(Score_Type, mean_group)),
    #comparisons = c("not variable", "variable"),
    method = "wilcox.test",
    label = "p.signif",
    hide.ns = FALSE
  ) +
  
  #geom_text(
  #  aes(
  #    x = Score_Type, 
  #    group = interaction(Score_Type, mean_group),
  #    label = case_when(
  #      p.value < 0.001 ~ "***",
  #      p.value < 0.01  ~ "**",
  #      p.value < 0.05  ~ "*",
  #      TRUE            ~ "ns"
  #    )
  #  )
  #) +

  labs(
    x = "Prediction method",
    y = "Mean prediction score",
    title = "Mean prediction score: not variable vs variable (wilcox.test)"
  ) +
  scale_fill_manual(
    values = c(
      "not variable" = "#999999",
      "variable" = "#E69F00"
    )
  ) +
  
  scale_alpha_continuous(
    limits = c(0,2),
    breaks = c(0, -log10(0.05), 2)
  ) +
  facet_grid(var_class_type~proteinclass) +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )

  
```


## 10A radius
```{r}
get_min_distance_to_var <- function(df_sub, min_var_freq = 0) {
  # Separate variable residues and all residues
  var_res <- df_sub %>% filter(var_freq > min_var_freq)
  all_res <- df_sub
  
  # If no variable residues, return Inf and NA for all
  if (nrow(var_res) == 0) {
    all_res <- all_res %>%
      mutate(min_distance_to_var = NA_real_, closest_var_freq = NA_real_)
    return(all_res)
  }
  
  # Coordinates matrix
  coords_all <- as.matrix(all_res[, c("x", "y", "z")])
  coords_var <- as.matrix(var_res[, c("x", "y", "z")])
  
  # Find nearest variable residue for each residue
  nn_res <- RANN::nn2(coords_var, coords_all, k = 1)
  
  # nn_res$nn.dists is the distance; nn_res$nn.idx the index of closest var residue
  all_res <- all_res %>%
    mutate(
      min_distance_to_var = nn_res$nn.dists,
      closest_var_freq = var_res$var_freq[nn_res$nn.idx]
    )
  
  return(all_res)
}
```

```{r}
results_all_radius <- class_results_residue %>%
  drop_na(x,y,z) %>%
  group_by(protein, PDB, chain, Score_Type, var_class_type) %>%
  group_modify(~ get_min_distance_to_var(.x)) %>%
  ungroup()

results_all_radius
```

```{}
max_residue_distance <- 10

class_results_radius_means <- results_all_radius %>%
  mutate(
    distance_variable_00 = case_when(
      is.na(closest_var_freq) ~ NA_character_,
      closest_var_freq >= 0 & min_distance_to_var <= max_residue_distance ~ "neighbor",
      TRUE ~ "nneighbor"
    ),
    distance_variable_05 = case_when(
      is.na(closest_var_freq) ~ NA_character_,
      closest_var_freq >= 0.05 & min_distance_to_var <= max_residue_distance ~ "neighbor",
      TRUE ~ "nneighbor"
    )
  ) %>%
  
  pivot_longer(
    cols = c(distance_variable_00, distance_variable_05),
    names_to = "near_variable_type",
    values_to = "variable_distance"
  ) %>%
  mutate(
    variable_distance = factor(
      variable_distance,
      levels = c("nneighbor", "neighbor")
    )
  ) %>%
  group_by(PDB,protein,proteinclass, Score_Type, near_variable_type) %>%
  drop_na(proteinclass, protein, Score_Type, Score_Value, near_variable_type, variable_distance) %>%
  group_modify(function(.x, .y) {
    data_do <- .x
    
    # Ensure var_class has exactly 2 levels
    if (n_distinct(data_do$near_variable) != 2) {
      return(tibble())
    }
    
    # Ensure both groups have at least 2 observations
    counts <- table(data_do$near_variable)
    if (any(counts < 2)) {
      return(tibble())
    }

    # Run t-test
    test <- t.test(Score_Value ~ near_variable, data = data_do)
    tidy_out <- tidy(test)

    # Rename group means using actual level names
    var_levels <- levels(factor(data_do$near_variable))
    tidy_out <- tidy_out %>%
      rename_with(~ var_levels, c("estimate1", "estimate2"))

    tidy_out
  }) %>%
  
  
  #do({
  #  data_do <- .
  #  test <- t.test(Score_Value ~ near_variable, data = data_do)
  #  tidy_out <- tidy(test)
  #  
  #  var_levels <- levels(data_do$near_variable)
  #  
  #  tidy_out <- tidy_out %>%
  #    rename_with(~var_levels, c("estimate1", "estimate2"))
  #  tidy_out
  #}) %>%
  pivot_longer(
    cols = c("no", "yes"),
    names_to = "near_variable", values_to = "group_mean"
  ) %>%
  mutate(
    near_variable = factor(
      near_variable, 
      levels = c("no", "yes")
    )
  ) 

class_results_radius_means
```

```{r, fig.width=14}
class_results_radius_means %>%
  ggplot(aes(x = Score_Type, fill = near_variable, y = group_mean, group = interaction(Score_Type, near_variable))) +
  geom_violin(
    trim = FALSE, scale = "width", alpha = 0.6, position = position_dodge(width = 0.8)
  ) +
  
  geom_line(
    aes(
      x = as.numeric(factor(Score_Type)) + ifelse(near_variable == "no", -0.2, 0.2),
      group = interaction(PDB, Score_Type),
      color = proteinclass
    )
  ) +
  
  geom_boxplot(
    width = 0.1, outlier.shape = NA, alpha = 0.8, position = position_dodge(width = 0.8)
  ) +
  geom_point(
    aes(color = proteinclass),
    position = position_dodge(width = 0.8)
    ) +
  
  stat_compare_means(
    #aes(group = interaction(Score_Type, mean_group)),
    #comparisons = c("not variable", "variable"),
    method = "wilcox.test",
    label = "p.signif",
    hide.ns = FALSE
  ) +
  
  labs(
    x = "Prediction method",
    y = "Mean prediction score",
    title = "Mean prediction score: not variable vs near variable (t.test)",
    subtitle = paste0("Within ", max_residue_distance, " A"),
  ) +
  scale_fill_manual(
    values = c(
      "no" = "#999999",
      "yes" = "#E69F00"
    )
  ) +
  
  theme_bw() +
  facet_wrap(~near_variable_type, nrow=2)
  
  
 
```

```{r, fig.width=14}
class_results_radius_means %>%
  ggplot(aes(x = Score_Type, fill = near_variable, y = group_mean, group = interaction(Score_Type, near_variable))) +
  geom_violin(
    trim = FALSE, scale = "width", alpha = 0.6, position = position_dodge(width = 0.8)
  ) +
  
  geom_line(
    aes(
      x = as.numeric(factor(Score_Type)) + ifelse(near_variable == "no", -0.2, 0.2),
      group = interaction(PDB, Score_Type),
      color = proteinclass
    )
  ) +
  
  geom_boxplot(
    width = 0.1, outlier.shape = NA, alpha = 0.8, position = position_dodge(width = 0.8)
  ) +
  geom_point(
    aes(color = proteinclass),
    position = position_dodge(width = 0.8)
    ) +
  
  stat_compare_means(
    #aes(group = interaction(Score_Type, near_variable)),
    #comparisons = c("not variable", "variable"),
    method = "wilcox.test",
    label = "p.signif",
    hide.ns = FALSE
  ) +
  
  labs(
    x = "Prediction method",
    y = "Mean prediction score",
    title = "Mean prediction score: not variable vs near variable (t.test)",
    subtitle = paste0("Within ", max_residue_distance, " A"),
  ) +
  scale_fill_manual(
    values = c(
      "no" = "#999999",
      "yes" = "#E69F00"
    )
  ) +
  
  facet_grid(near_variable_type~proteinclass) +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
  
```

## mixed effects
```{r}
library(MuMIn)
library(broom.mixed)
library(lmerTest)
library(lme4)

mixed_effects <- data_predictions_with_variability_normalized %>% 
  pivot_longer(cols = pivot_list, 
               names_to = "Score_Type", values_to = "Score_Value") %>%
  group_by(Score_Type) %>%
  group_modify(~ {
    model <- lmer(Score_Value ~ RSA + var_freq + (1 | proteinclass/protein/PDB), data = .x)
    print(colnames(summary(model)$coefficients))
    r2 <- r.squaredGLMM(model)
    
    # Extract fixed effects with p-values from lmerTest summary
    coef_df <- summary(model)$coefficients %>%
      as.data.frame() %>%
      tibble::rownames_to_column("term") %>%
      select(
        term,
        estimate = Estimate,
        std.error = `Std. Error`,
        statistic = `t value`,
        p.value = `Pr(>|t|)`
      )
    
    tibble(
      marginal_r2 = r2[1],
      conditional_r2 = r2[2],
      term_data = list(coef_df)
    )
  }) %>%
  unnest(term_data)

colnames(mixed_effects)
mixed_effects
```

```{r, fig.width=14}
mixed_effects %>%
  #filter(effect == "fixed") %>%
  ggplot(aes(x=term, y=estimate)) +
  geom_point(
    #aes(size=-log10(p.value))
    ) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
  facet_wrap(~ Score_Type) +
  theme_bw() +
  labs(title = "Fixed Effect Estimates per Score Type",
       y = "Estimate", x = "Term") +
  
  scale_size_continuous(
        limits = c(0, 3),
        breaks = c(-log10(0.1),-log10(0.05), 2, 3),  # Use -log10(0.05) as a significant break
        labels = c("0.1","0.05", "0.01", "0.001"),
        name = "p-value"
      ) +
  
  
  geom_text(
    data = mixed_effects %>%
      distinct(Score_Type, marginal_r2, conditional_r2) %>%
      mutate(
        label = paste0("Marginal R²: ", round(marginal_r2, 3), "\n",
                       "Conditional R²: ", round(conditional_r2, 3)),
        x = Inf,
        y = Inf
      ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust = 1.1,
    vjust = 1.1,
    size = 5
  ) +
  
  geom_text(
    aes(
      label = case_when(
        p.value < 0.001 ~ "***",
        p.value < 0.01  ~ "**",
        p.value < 0.05  ~ "*",
        TRUE            ~ "ns"
      ), 
      vjust = -(std.error + 0.2) 
      
    )
  )

```

### radius
```{r}
mixed_effects_radius <- results_all_radius %>%
  mutate(
    min_distance_to_var = case_when(
      min_distance_to_var == Inf ~ NA,
      TRUE ~ min_distance_to_var
    )
  ) %>%
  pivot_longer(cols = pivot_list, 
               names_to = "Score_Type", values_to = "Score_Value") %>%
  group_by(Score_Type) %>%
  group_modify(~ {
    model <- lmer(
      Score_Value ~ RSA + (1 | proteinclass/protein/PDB) + closest_var_freq*min_distance_to_var,
      data = .x
    )
    print(colnames(summary(model)$coefficients))
    r2 <- r.squaredGLMM(model)
    
    # Extract fixed effects with p-values from lmerTest summary
    coef_df <- summary(model)$coefficients %>%
      as.data.frame() %>%
      tibble::rownames_to_column("term") %>%
      select(
        term,
        estimate = Estimate,
        std.error = `Std. Error`,
        statistic = `t value`,
        p.value = `Pr(>|t|)`
      )
    
    tibble(
      marginal_r2 = r2[1],
      conditional_r2 = r2[2],
      term_data = list(coef_df)
    )
  }) %>%
  unnest(term_data)
  

mixed_effects_radius
```

```{r, fig.width=14}
mixed_effects_radius %>%
  #filter(effect == "fixed") %>%
  mutate(
    term = factor(
      term,
      levels = c("(Intercept)","RSA", "closest_var_freq", "min_distance_to_var", "closest_var_freq:min_distance_to_var")
    )
  ) %>%
  ggplot(aes(x=term, y=estimate)) +
  geom_point(
    #aes(size=-log10(p.value))
    ) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
  facet_wrap(~ Score_Type) +
  theme_bw() +
  labs(title = "Fixed Effect Estimates per Score Type",
       y = "Estimate", x = "Term") +
  
  scale_size_continuous(
        limits = c(0, 3),
        breaks = c(-log10(0.1),-log10(0.05), 2, 3),  # Use -log10(0.05) as a significant break
        labels = c("0.1","0.05", "0.01", "0.001"),
        name = "p-value"
      ) +
  
  
  geom_text(
    data = mixed_effects_radius %>%
      distinct(Score_Type, marginal_r2, conditional_r2) %>%
      mutate(
        label = paste0("Marginal R²: ", round(marginal_r2, 3), "\n",
                       "Conditional R²: ", round(conditional_r2, 3)),
        x = Inf,
        y = Inf
      ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust = 1.1,
    vjust = 1.1,
    size = 5
  ) +
  
  geom_text(
    aes(
      label = case_when(
        p.value < 0.001 ~ "***",
        p.value < 0.01  ~ "**",
        p.value < 0.05  ~ "*",
        TRUE            ~ "ns"
      ), 
      vjust = -(std.error + 0.2) 
      
    )
  ) +
  
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
```

## New classification
```{r}
min_var_freq = 0.05
max_distance_to_var = 10
min_RSA = 0.2

results_all_radius_classification <- results_all_radius %>%
  drop_na(var_freq, x, y, z) %>%
  group_by(protein, PDB, chain) %>%
  group_modify(~ get_min_distance_to_var(.x, min_var_freq)) %>%
  
  mutate(
    min_distance_to_var = case_when(
      min_distance_to_var == Inf ~ NA,
      TRUE ~ min_distance_to_var
    ),
    
    distance_variable_00 = case_when(
      is.na(closest_var_freq) ~ NA_character_,
      closest_var_freq >= 0 & min_distance_to_var <= max_residue_distance ~ "neighbor",
      TRUE ~ "nneighbor"
    ),
    distance_variable_05 = case_when(
      is.na(closest_var_freq) ~ NA_character_,
      closest_var_freq >= 0.05 & min_distance_to_var <= max_residue_distance ~ "neighbor",
      TRUE ~ "nneighbor"
    )
  ) %>%
  pivot_longer(
    cols = c(distance_variable_00, distance_variable_05),
    names_to = "near_variable_type",
    values_to = "variable_distance"
  ) %>%
  mutate(
    variable_distance = factor(
      variable_distance,
      levels = c("nneighbor", "neighbor")
    )
  ) 

results_all_radius_classification
```

```{r, fig.width=14}
results_all_radius_classification %>%
  drop_na(RSA_class, variable_distance) %>%
  filter(near_variable_type == "distance_variable_05") %>%
  mutate(
    is_sur.nb = factor(
      interaction(RSA_class, variable_distance),
      levels = c("nsur.nneighbor","nsur.neighbor", "sur.nneighbor", "sur.neighbor")
      )
    ) %>%
  mutate(score_label = score_label_map[Score_Type]) %>%
  ggplot(aes(x= is_sur.nb, fill = is_sur.nb, y = Score_Value)) +
  geom_violin(
    trim = FALSE, scale = "width", alpha = 0.6, position = position_dodge(width = 0.8)
  ) +
  geom_boxplot(
    width = 0.1, outlier.shape = NA, alpha = 0.8, position = position_dodge(width = 0.8)
  ) +
  #geom_point(aes(color = proteinclass)) +
  #geom_path(aes(group = structure, color = proteinclass)) +
  stat_compare_means(
    comparisons = combn(c("nsur.nneighbor","nsur.neighbor", "sur.nneighbor", "sur.neighbor"), 2, simplify = FALSE),
    method = "wilcox",
    label = "p.signif",
    hide.ns = FALSE,
    label.y = seq(1,1.6,by=0.1)
  ) +
  labs(
    x = "",
    y = "Residue epitope prediction score",
    title = "Not variable residue neighbor vs variable residue neighbor & surface vs non-surface (wilcox.test)",
    fill = "Accessibility % Variability"
  ) +
  #facet_grid(var_class_type~proteinclass) +
  facet_grid(~score_label) +
  scale_fill_manual(
    values =  c("nsur.nneighbor" = "#D55E00","nsur.neighbor" = "#E69F00", "sur.nneighbor"  = "#0072B2", "sur.neighbor" = "#56B4E9"),
    labels =  c("nsur.nneighbor" = "non-surface\nnon-neighbor", 
                "nsur.neighbor" = "non-surface\nneighbor", 
                "sur.nneighbor" = "surface\nnon-neighbor", 
                "sur.neighbor" = "surface\nneighbor")
  ) +
  scale_x_discrete(
    labels = c("nsur.nneighbor" = "non-surface\nnon-neighbor", 
                "nsur.neighbor" = "non-surface\nneighbor", 
                "sur.nneighbor" = "surface\nnon-neighbor", 
                "sur.neighbor" = "surface\nneighbor")
  ) +
  
  geom_text(
    aes(x=2.5,y=1.65, label = score_label), check_overlap = TRUE, size = 5
            ) +
  
  scale_y_continuous(
    limits = c(0, 1.75),
    expand = c(0, 0), 
    breaks = seq(0,1, by = 0.25),
    labels = seq(0,1, by = 0.25)
  ) +
  theme_bw() +
  theme(
    panel.spacing = unit(0, "lines"), 
    legend.position = "none",
    panel.grid = element_blank(), 
    #axis.title.x = ,
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),

    strip.background = element_blank(),
    strip.text = element_blank()
    )
```

```{r, fig.width=14}
results_all_radius_classification %>%
  drop_na(RSA_class, variable_distance) %>%
  filter(near_variable_type == "distance_variable_05") %>%
  mutate(
    is_sur.nb = factor(
      interaction(RSA_class, variable_distance),
      levels = c("nsur.nneighbor","nsur.neighbor", "sur.nneighbor", "sur.neighbor")
      )
    ) %>%
  mutate(score_label = score_label_map[Score_Type]) %>%
  ggplot(aes(x= is_sur.nb, fill = is_sur.nb, y = Score_Value)) +
  geom_violin(
    trim = FALSE, scale = "width", alpha = 0.6, position = position_dodge(width = 0.8)
  ) +
  geom_boxplot(
    width = 0.1, outlier.shape = NA, alpha = 0.8, position = position_dodge(width = 0.8)
  ) +
  #geom_point(aes(color = proteinclass)) +
  #geom_path(aes(group = structure, color = proteinclass)) +
  stat_compare_means(
    comparisons = combn(c("nsur.nneighbor","nsur.neighbor", "sur.nneighbor", "sur.neighbor"), 2, simplify = FALSE),
    method = "wilcox",
    label = "p.signif",
    hide.ns = FALSE,
    label.y = seq(1,1.6,by=0.1)
  ) +
  labs(
    x = "",
    y = "Residue epitope prediction score",
    title = "Not variable residue neighbor vs variable residue neighbor & surface vs non-surface (wilcox.test)",
    fill = "Accessibility % Variability"
  ) +
  #facet_grid(var_class_type~proteinclass) +
  facet_grid(proteinclass~score_label) +
  scale_fill_manual(
    values =  c("nsur.nneighbor" = "#D55E00","nsur.neighbor" = "#E69F00", "sur.nneighbor"  = "#0072B2", "sur.neighbor" = "#56B4E9"),
    labels =  c("nsur.nneighbor" = "non-surface\nnon-neighbor", 
                "nsur.neighbor" = "non-surface\nneighbor", 
                "sur.nneighbor" = "surface\nnon-neighbor", 
                "sur.neighbor" = "surface\nneighbor")
  ) +
  scale_x_discrete(
    labels = c("nsur.nneighbor" = "non-surface\nnon-neighbor", 
                "nsur.neighbor" = "non-surface\nneighbor", 
                "sur.nneighbor" = "surface\nnon-neighbor", 
                "sur.neighbor" = "surface\nneighbor")
  ) +
  
  #geom_text(
  #  aes(x=2.5,y=1.65, label = score_label), check_overlap = TRUE, size = 5
  #          ) +
  
  scale_y_continuous(
    limits = c(0, 1.75),
    expand = c(0, 0), 
    breaks = seq(0,1, by = 0.25),
    labels = seq(0,1, by = 0.25)
  ) +
  theme_bw() +
  theme(
    panel.spacing = unit(0, "lines"), 
    legend.position = "none",
    panel.grid = element_blank(), 
    #axis.title.x = ,
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),

    #strip.background = element_blank(),
    #strip.text = element_blank()
    )
```

```{r, fig.width=14}
results_all_radius_classification %>%
  drop_na(RSA_class, variable_distance) %>%
  filter(near_variable_type == "distance_variable_05") %>%
  mutate(
    is_sur.nb = factor(
      interaction(RSA_class, variable_distance),
      levels = c("nsur.nneighbor","nsur.neighbor", "sur.nneighbor", "sur.neighbor")
      )
    ) %>%
  mutate(score_label = score_label_map[Score_Type]) %>%
  ggplot(aes(x= is_sur.nb, fill = is_sur.nb, y = Score_Value)) +
  geom_violin(
    trim = FALSE, scale = "width", alpha = 0.6, position = position_dodge(width = 0.8)
  ) +
  geom_boxplot(
    width = 0.1, outlier.shape = NA, alpha = 0.8, position = position_dodge(width = 0.8)
  ) +
  #geom_point(aes(color = proteinclass)) +
  #geom_path(aes(group = structure, color = proteinclass)) +
  stat_compare_means(
    comparisons = combn(c("nsur.nneighbor","nsur.neighbor", "sur.nneighbor", "sur.neighbor"), 2, simplify = FALSE),
    method = "wilcox",
    label = "p.signif",
    hide.ns = FALSE,
    label.y = seq(1,1.6,by=0.1)
  ) +
  labs(
    x = "",
    y = "Residue epitope prediction score",
    title = "Not variable residue neighbor vs variable residue neighbor & surface vs non-surface (wilcox.test)",
    fill = "Accessibility % Variability"
  ) +
  #facet_grid(var_class_type~proteinclass) +
  facet_grid(seroprof_titers~score_label) +
  scale_fill_manual(
    values =  c("nsur.nneighbor" = "#D55E00","nsur.neighbor" = "#E69F00", "sur.nneighbor"  = "#0072B2", "sur.neighbor" = "#56B4E9"),
    labels =  c("nsur.nneighbor" = "non-surface\nnon-neighbor", 
                "nsur.neighbor" = "non-surface\nneighbor", 
                "sur.nneighbor" = "surface\nnon-neighbor", 
                "sur.neighbor" = "surface\nneighbor")
  ) +
  scale_x_discrete(
    labels = c("nsur.nneighbor" = "non-surface\nnon-neighbor", 
                "nsur.neighbor" = "non-surface\nneighbor", 
                "sur.nneighbor" = "surface\nnon-neighbor", 
                "sur.neighbor" = "surface\nneighbor")
  ) +
  
  #geom_text(
  #  aes(x=2.5,y=1.65, label = score_label), check_overlap = TRUE, size = 5
  #          ) +
  
  scale_y_continuous(
    limits = c(0, 1.75),
    expand = c(0, 0), 
    breaks = seq(0,1, by = 0.25),
    labels = seq(0,1, by = 0.25)
  ) +
  theme_bw() +
  theme(
    panel.spacing = unit(0, "lines"), 
    legend.position = "none",
    panel.grid = element_blank(), 
    #axis.title.x = ,
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),

    #strip.background = element_blank(),
    #strip.text = element_blank()
    )
```

```{, fig.width=14}
p_adjust_method <- "BH"

results_all_radius_classification_means %>%
  filter(Score_Type != "BepiPred-2.0 probability") %>%
  
  ggplot(aes(x = is_sur.var, y = Score_Value)) +
  
  #geom_path(aes(group = PDB, col = proteinclass)) +
  
  geom_violin(
    aes(fill = is_sur.var),
    trim = FALSE, scale = "width", alpha = 0.6, show.legend = FALSE
  ) +
  
  geom_boxplot(
    width = 0.1, outlier.shape = NA, alpha = 0.8
  ) +
  
  geom_point(
    aes(color = proteinclass)
    ) +
  
  stat_compare_means(
    aes(group = PDB),  
    method = "wilcox",
    label = "p.signif",
    comparisons = combn(levels(results_all_radius_classification_means$is_sur.var), 2, simplify = FALSE),
    p.adjust.method = p_adjust_method
  ) +
  
  theme_bw() +
  
  facet_grid(Score_Type~proteinclass, scales = "free_y") +
  
  labs(
    title = "Mean score comparisons per Score Type",
    subtitle = paste0(
      "minimum variability: ", min_var_freq,
      ", maximum distance to variable residue: ", max_distance_to_var, " A",
      ", minimum RSA: ", min_RSA,
      ", correction: ", p_adjust_method
    ),
    y = "Mean Score", 
    x = "comparison(is surface, is near variable)"
  ) +
  
  scale_fill_manual(
    values = c("no.no" = "#D55E00","no.yes" = "#E69F00", "yes.no" = "#0072B2", "yes.yes" = "#56B4E9")
  ) +
  
  theme(
    panel.spacing = unit(0, "lines"),
    #strip.background = element_blank(),
    strip.placement = "outside",
    panel.border = element_blank(),
    axis.line = element_line()
  ) 

```

```{r, figure.width=14}
prediction %>%
  drop_na(RSA_class, variable_distance) %>%
  filter(near_variable_type == "distance_variable_05", var_class_type == "var_class_05",
         protein %in% ordered_proteins) %>%
  
```


```{r}
setdiff(
  ordered_proteins,
  results_all_radius_classification %>% filter(protein %in% ordered_proteins) %>% pull(protein) %>% unique
)
```

```{r, fig.width=14}
data_predictions_with_variability_normalized_with_seroprof <- data_predictions_with_variability_normalized %>%
  left_join(distinct(data_variability_with_seroprof %>% select(protein, seroprof_titers)), by = "protein") 
  
```

```{r, fig.width=14}
lowest_plot <- data_predictions_with_variability_normalized_with_seroprof %>%
  filter(protein %in% ordered_proteins) %>%
  mutate(protein = factor(protein, levels = ordered_proteins)) %>%
  pivot_longer(cols = pivot_list, 
               names_to = "Score_Type", values_to = "Score_Value") %>%
  group_by(proteinclass, structure, Score_Type) %>%
  group_by(seroprof_titers,proteinclass, protein, Score_Type) %>%
  distinct() %>%
  summarize(
    mean_score = mean(Score_Value, na.rm=TRUE),
    .groups = "drop"
  ) %>%
  mutate(score_label = score_label_map[Score_Type]) %>%
  filter(score_label != "BepiPred 2.0") %>%
  
  ggplot(aes(x = protein, y = mean_score)) +
  
  geom_rect(
    aes(
      xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = seroprof_titers
    ),
    inherit.aes = FALSE, alpha = 0.3,
    data = data_variability_with_seroprof %>% distinct(seroprof_titers),
    show.legend = FALSE  # Exclude facet-based color fill from the legend
  ) +
  
  scale_fill_brewer(palette = "RdYlBu") +
  
  new_scale_fill() +
  
  geom_col(aes(fill = proteinclass)) +
  
  facet_grid(
    score_label~seroprof_titers,
    scales = "free_x",
    space = "free_x",
    labeller = label_wrap_gen(width = 10)
  ) +
  
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 1), breaks = seq(0, 1, by = 0.25) ) +
  
  theme_bw() +
  
  labs(y = "Mean Normalized Variability", x = NULL) +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 16), # Adjust facet label text size
    panel.background = element_rect(fill = "white", colour = "white"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.spacing = unit(0, "lines"),
    axis.line = element_line(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
  
lowest_plot
```

```{r, fig.width = 14}
plot_grid(
  upper_plot + 
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      plot.margin = margin(t = 5, r = 5, b = 1.5, l = 5)
    )
  ,
  lower_plot + 
    theme(
        axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      strip.text = element_blank(),
      strip.background = element_blank(),
      legend.position = "none",
      plot.margin = margin(t = 0, r = 5, b = 1.5, l = 5) 
    ),
  lowest_plot + 
    theme(
      strip.text.x = element_blank(),
      strip.background.x = element_blank(),
      legend.position = "none",
      plot.margin = margin(t = 0, r = 5, b = 5, l = 5) 
    ),
  ncol = 1,
  align = "v",
  axis = "rl",
  rel_heights = c(1, 1, 2)  # adjust height ratio as needed
)
```

